### 2020-11-17 tex/LitReview/LitReview.tex
initial commit from desktop.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\noindent \citet{yuan2017}: Disruption Management for the Real-Time Home Caregiver Scheduling and Routing Problem

\section{Differences}
\begin{itemize}[leftmargin=*, itemsep=-0.1em]
	\item Paper covers updating an existing schedule when an disruption occurs. The disruption in this case is when a new service is added to the schedule. However, in our case there could be a variety of reasons why a disruption may occur, such as a delay in a service or carer illness.
	\item This paper does \emph{not} consider double services or synchronised (dependent) services.
	\item The paper assumes that all carers start and end their routes at the Home Health Care Centre, rather than their own homes.
	\item The paper also states that each patient is visited by a carer exactly once, which is not always the case in our problem, as a patient may require multiple visits in a single day.
	\item 	
\end{itemize}

\bibliographystyle{plainnat}
\bibliography{includes/bibliography}

\end{document}

### 2020-11-17 tex/LitReview/includes/preamble.tex
initial commit from desktop.

%---- PREAMBLE ----%

% Allows separate files to be used in the main file
\usepackage{subfiles}

% Algorithms
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\renewcommand{\algorithmicrequire}{\textbf{Input:}}
\renewcommand{\algorithmicensure}{\textbf{Output:}}
\newcommand{\algorithmicbreak}{\State \textbf{break}}
\newcommand{\Break}{\algorithmicbreak}
\renewcommand{\algorithmicreturn}{\State \textbf{return}}
\newcommand{\algorithmicto}{\textbf{ to }}
\newcommand{\To}{\algorithmicto}
\newcommand{\algorithmicrun}{\State \textbf{call }}
\newcommand{\Run}{\algorithmicrun}
\newcommand{\algorithmicoutput}{\textbf{output }}
\newcommand{\Output}{\algorithmicoutput}
\newcommand{\algorithmictrue}{\textbf{true}}
\newcommand{\True}{\algorithmictrue}
\newcommand{\algorithmicfalse}{\textbf{false}}
\newcommand{\False}{\algorithmicfalse}
\newcommand{\algorithmicand}{\textbf{ and }}
\newcommand{\AAnd}{\algorithmicand}
\newcommand{\algorithmicnot}{\textbf{not}}
\newcommand{\Not}{\algorithmicnot}
\newcommand{\algorithmicor}{\textbf{ or }}
\newcommand{\Or}{\algorithmicor}

% Contains advanced math extensions
\usepackage{amsmath}

% Introduces the *proof* environment and the \theoremstyle command
\usepackage{amsthm}

% Adds new symbols to be used in math mode, e.g. \mathbb
\usepackage{amssymb}

% To declare multiple authors
%\usepackage{authblk}

% Provides extra comands as well as optimisation for producing tables
\usepackage{booktabs}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}

% Allows customisation of appearance and placements for figures/tables etc.
\usepackage{caption}

\usepackage{comment}

% Adds support for arbitrarily-deep nested lists
\usepackage[inline]{enumitem}

%Support for changing size of font in table footnotes
%\usepackage{etoolbox}

% Improves the interface for defining floating objects such as figures/tables
\usepackage{float}

%\usepackage{fullpage}

% For easy management of document margins and the document page size
\usepackage[a4paper]{geometry}

% Allows insertion of graphic files within a document
\usepackage{graphicx}

% Manage links within the document or to any URL when you compile in PDF
\usepackage[colorlinks]{hyperref} 
\usepackage[dvipsnames]{xcolor}
%Tikz colours, used in tikz figures only
\colorlet{tBlue}{RoyalBlue!35!Cerulean} %tikz color
\colorlet{tRed}{Red} %tikz color
\definecolor{tGreen}{HTML}{569909} %tikz color
\definecolor{tOrange}{HTML}{FA7602} %tikz color
\definecolor{tLightGreen1}{HTML}{C1E685} %tikz color
\definecolor{tLightOrange1}{HTML}{FFCD4F} %tikz color
\colorlet{tLightGreen}{LimeGreen!70!OliveGreen!45!White}
\colorlet{tLightOrange}{Dandelion!65!White}
\definecolor{tLightPink}{HTML}{FFD4EB} %tikz color
\definecolor{tLightBlue}{HTML}{CEF0FF} %tikz color

%Text colours
\colorlet{myRed}{Red!50!OrangeRed}
\definecolor{myOrange}{HTML}{FA7602}
\definecolor{myGreen}{HTML}{569909}
\definecolor{myAqua}{HTML}{00B1BA} %02BEB8
\definecolor{myBlue}{HTML}{0095FF} %00B3FF
\colorlet{myPurple}{Orchid} 
\colorlet{myPink}{Rhodamine!65!Lavender}
\colorlet{myGray}{Gray!90!White}

\hypersetup{
		linkcolor=myPurple,
		citecolor=myAqua,
		urlcolor=black
}

% The document class `elsarticle` uses the \AtBeginDocument comman to define the colour of all link types as blue, so we use the \AtBeginDocument command again to overwrite the colour settings to the colors we choose. REMEMBER TO REMOVE THIS BEFORE SUBMITTING.
%\AtBeginDocument{% 
%	\hypersetup{
%		linkcolor=myPurple,
%		citecolor=myAqua,
%		urlcolor=black
%	}
%} %end of \AtBeginDocument

\newcommand{\intro}[1]{{\color{myOrange}#1}} % \intro{<text>}, makes <text> orange, use to highlight things to do, urgent, or revisit in introduction section

\newcommand{\ahc}[1]{{\color{myBlue}#1}} % \ahc{<text>}, makes <text> blue, use to highlight things to do, urgent, or revisit in AHC section

\newcommand{\heur}[1]{{\color{myPink}#1}} % \scspp{<text>}, makes <text> pink, use to highlight things to do, urgent, or revisit in SCSPP section

\newcommand{\ea}[1]{{\color{myRed}#1}} % \ea{<text>}, makes <text> red, use to highlight things to do, urgent, or revisit in EA section

\newcommand{\cmsa}[1]{{\color{myGreen}#1}} % \cmsa{<text>}, makes <text> green, use to highlight things to do, urgent, or revisit in CMSA section

\newcommand{\conc}[1]{{\color{myPurple}#1}} % \conc{<text>}, makes <text> purple, use to highlight things to do, urgent, or revisit in Conclusion section

\newcommand{\note}[1]{{\color{myPurple}#1}} % \note{<text>}, makes <text> turquoise, use to highlight things to do, urgent, or revisit in any section

\newcommand{\alert}[1]{{\color{myRed}#1}} % \alert{<text>}, makes <text> red, use to highlight things to do, urgent, or revisit.

\newcommand{\done}[1]{{\color{myGray}#1}} % \done{<text>}, makes <text> gray, use to highlight things that are not required or should be ignored.

\newcommand{\ialert}[1]{{\color{myRed}\item#1}} % \ialert{<text>}, makes <text> a red bullet point, use for bullet points of things to do, urgent, or revisit.

\newcommand{\idone}[1]{{\color{myGray}\item#1}} % \idone{<text>}, makes <text> a gray bullet point, use for bullet points that have been completed, that are not required or should be ignored.

\renewcommand{\idone}[1]{} % hides all \idone bullet points

% Successor of amsmath
\usepackage{mathtools}

\usepackage{multirow}

% No indentation, space between paragraphs
%\usepackage{parskip}

\usepackage[round]{natbib}

%Include standalone .tex files
\usepackage{standalone}

% Define multiple floats (figures/tables) within one environment with individual captions 1a, 1b etc
\usepackage{subcaption}

%\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{tabularx}

\usepackage{threeparttable}
%\appto\TPTnoteSettings{\footnotesize} % Change font size of table footnotes to footnotesize

\usepackage{tikz}
\usetikzlibrary{shapes.geometric}

\usepackage{wrapfig}

% Need \usepackage{mathtools} to use floor/ceil below.
% Example: \floor*{\frac{x}{2}}, \ceil*{\frac{x}{2}}
% The asterisk resizes floor/ceil brackets.
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

%Theorem style
%\theoremstyle{plain}% default
\newtheorem{theorem}{Theorem}
%\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}

%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
%\newtheorem{proposition}{Proposition}
%\newtheorem{exmp}{Example}[section]

### 2020-11-17 tex/SocialCare/SocialCare.tex
initial commit from desktop.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\section{Constraints}
\begin{itemize}[leftmargin=*, itemsep=-0.1em]
	\item Maximum number of carers per patient
	\item PEG feed
	\item Determine the final objective, or allow for multiple objective (e.g. reduce costs, reduce travel time, reduce tardiness etc.)
	\item Maximum distance from depot (home), 6.7 miles?
	\item Maximum number of dependent jobs/precendence relations
	\item Determine the hard and soft constraints - current assuming that all patients must be visited by qualified carers, and tardiness is allowed as a result
	\item Total number of patients and total number of carers in each location/area
	\item How often are new patients added to the system/schedule?
	\item Carers updating patients files after each service, has this time been included in the service time?
	\item Number of patients in each location that require a carer that speaks the same language and is qualified for the job
	\item PPE removal and application between services
	\item Can patients refuse `treatment' if carer is late? (non-urgent service, e.g. running errands or going outdoors)
	\item Preferences, e.g. allergies, male or female carer/patient, languages, personal preferences (some carers may not wish to work with one another on double services) 
	\item Noted that carers are assigned different patients over longer periods of time to maintain professional relationships between carers and patients
	\item Maximum travel distance for individual carer's routes?
	\item Length of day for carers (half day/full days set or can schedule be flexible)?
	\item Lunch breaks?
	\item Shadowing carers
	\item Parking availability at each patient's household, or does walking time also need to be included?
	\item Are there distinct job types?
\end{itemize}

\section{Rerouting and Rescheduling}
\begin{itemize}[leftmargin=*, itemsep=-0.em]
	\item In what scenarios does and r\&r need to occur?
	\item Addition of new patient (job) to the schedule (for example, emergency revisit)
	\item Cancellation of job
	\item Scenario where a patient that originally only required one carer now requires multiple carers
	\item Change in skill requirement of a patient, where the originally scheduled carer does not have the required skill level
	\item Patient emergency that requires the carer to remain with the patient; thus remaining jobs in the carer's route will need to be serviced by one or more other carers
	\item Carer unwell at the beginning of the day, so all jobs assigned to that carer will need to be reassigned to other carer's working that day	
\end{itemize}

\end{document}

### 2020-12-01 tex/LitReview/LitReview.tex
added more information.

@@
 \begin{document}
-\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
+\title{Literature: On Disruption Management and Reactive Algorithms for Routing and Scheduling Problems}
 %\author{Asyl L. Hawa}
@@
 
-\noindent \citet{yuan2017}: Disruption Management for the Real-Time Home Caregiver Scheduling and Routing Problem
+\section{Disruption Management for the RT-HCSRP}
+\noindent \citet{yuan2017}: Disruption Management for the Real-Time Home Caregiver Scheduling and Routing Problem.
+The problem involves changing a given schedule of the HCSRP (Home Caregiver Scheduling and Routing Problem) when an unexpected event occurs, which in this paper is when a new service request arises.
+The RT-HCSRP (Real-Time) consists in creating an updated schedule and route for the carers to the patients based on the original schedule; that is, a feasible solution to the original HCSRP).
+Thus, the input for the RT-HCSRP is a schedule for the HCSRP with the new additional service added to a carer's route in some arbitrary position.
+Then, a tabu search procedure is used to optimise the solution.
+The goal is to minimise the negative effect of unexpected events on patients, carers, and companies.
+Deviation measurements are formed independently for patients, carers, and companies, which are quantitatively measured and taken as minimisation objectives in a weighted objective function.
 
-\section{Differences}
-\begin{itemize}[leftmargin=*, itemsep=-0.1em]
-	\item Paper covers updating an existing schedule when an disruption occurs.
+\subsection{Differences}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item The paper covers updating an existing schedule when an disruption occurs.
 The disruption in this case is when a new service is added to the schedule.
 However, in our case there could be a variety of reasons why a disruption may occur, such as a delay in a service or carer illness.
-	\item This paper does \emph{not} consider double services or synchronised (dependent) services.
+	\item The paper does \emph{not} consider double services or synchronised (dependent) services.
 	\item The paper assumes that all carers start and end their routes at the Home Health Care Centre, rather than their own homes.
-	\item The paper also states that each patient is visited by a carer exactly once, which is not always the case in our problem, as a patient may require multiple visits in a single day.
+	\item The paper also states that each patient is visited by a carer \emph{exactly once}, which is not always the case in our problem, as a patient may require multiple visits in a single day.	
+	\item The paper assumes that there is an additional carer available at the depot (HHC centre) that has the highest possible skill level (i.e. is able to do any type of job).
+\end{itemize}
+
+\subsection{Model and Solution Approach}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item In the paper, a new patient is added to the original plan (i.e.
+the solution to the original HCSRP) to generate an initial solution for the RT-HCSRP.
+This is done by assigning the new patient (job/service) to the carer who can do the job with the least increase in the objective function value.
+This new solution with the additional job is considered the initial input solution, and the tabu search procedure is applied to this solution.
+	\item When an unexpected event occurs (the addition of a new service), at time $T_0$, each of the patients (jobs) are in one of four states: 
+	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
+		\item The job has been serviced (is completed); 
+		\item The job has not yet been serviced; 
+		\item The job is currently in service (a carer is at the job); and
+		\item A newly occuring job (new job being added to the schedule).
+	\end{enumerate}
+	\item Only (ii)--(iv) are considered because jobs in (i) have already been fulfilled.
+	\item Carers are also in one of three states regarding their locations at time $T_0$:
+	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
+		\item Carers in transit, travelling to the next job;
+		\item Carers currently at a job; and
+		\item Carers at the HHC centre (depot).
+	\end{enumerate}
+	\item The planning period (for the new schedule) is between time $T_0$ (the time of the disruption) and the end of the day.
+	\item Locations of carers at a job or the first jobs to be serviced after $T_0$ in the original plan are referred to as \emph{critical nodes} ($CN$).
+	\item Critical nodes in the node set $CN$ are the starting nodes (locations) of carers in the updated plan.
+	\item The paper outlines three different \emph{deviation measurements} for patients (jobs), carers, and the companies respectively, and the objective function minimises the weighted sum of the deviation measurements.
+	\item The solution approach is using tabu search (TS).
+The procedure starts from an initial solution, and in each iteration the neighbour set of a solution is generated using neighbourhood structures: remove a job from a route visited by a carer, and reinsert the job into the route of another carer.
+	\item The solution in the neighbour set that has best quality (least solution evaluation value) replaces the current solution, and the tabu list is updated.
+	\item The process continues until a stopping criteria is satisfied (time limit).
+	\item Each solution $s$ is characterised by an \emph{attribute set}, $B(s) = \{(i,k) : \text{job $i$ is perfomed by carer $k$}\}$, which is recorded in the tabu list.
+	\item A neighbour of a solution is obtained by replacing some attributes of $B(s)$ with other attributes.
+	\item Then, when a job $i$ is removed from a carer $k$'s route, the attribute $(i,k)$ is assigned a tabu status, as well as a tabu duration $\theta$. 
+	\item Aspiration criterion :the tabu status of an attribute can be revoked if accepting the attribute leads to a solution with better quality (less objective functon value) than the current best solution that contains the attribute.
+The OF value of the current best solution that contains an attribute is called the \emph{aspiration level} of the attribute.
+	\item To guide the search into less explored solution space, if a solution $s'$ is a non-improving neighbourhood w.r.t.
+solution $s$, then $s'$ is penalised by a term proportional to the addition frequency of its attributes.
+\end{itemize}
+
+\subsection{Modifications}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item In our problem, we may have many other scenarios where a disruption and rescheduling can occur, not just the addition of a new service, and to create an initial solution for the R\&R we it may involve reassigning multiple services to other carers, not just one new service as with \citep{yuan2017}.
+	\item We also need to consider double services, e.g. if the disruption occurs on a double/synchronised service, as this can affect multiple carers' routes.
+\end{itemize}
+
+\section{Disruption Management of the VRP with Breakdown}
+\noindent \citet{mu2011}: Disruption management of the vehicle routing problem with vehicle breakdown.
+The problem arises when a vehicle breaks down on a route, and a new solution needs to be created quickly such that the remaining customers on the vehicle's route are reassigned to other vehicles, whilst also minimising costs.
+The paper introduces two tabu search algorithms to solve the problem.
+The disruption in this problem, when a vehicle breaks down, corresponds to the scenario in our problem where a carer must stay with a patient (stay at a job) due to an emergency, and so all remaining jobs in the carer's route will need to be reassigned to other carers.
+
+\subsection{Differences}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item There are no double services or synchronised deliveries (jobs) to customers.
+	\item There are no time windows for vehicle deliveries.
+	\item There is an extra vehicle (EV) available at the depot, however it may not be possible to have an extra carer on stand-by as this will incur extra costs.
+	\item All vehicles start and end at a central depot, whereas in our problem carers start and end their routes at their own homes.
+	\item There are no route distance nor distance radius constraints.
+	\item It is assumed that the goods being delivered by the vehicle are all the same, and each customer is expecting the same goods; thus any vehicle can deliver to any customer.
+Obviously, this is not the case in our problem, as each job requires a carer with the appropriate skill level.
+	\item This problem also deals with the issue of vehicles being empty (i.e. running out of goods to deliver to customers). As the carers offer a service in our problem, we do not have this constraint.
+	\item Each customer only needs to be visited once by one vehicle; again, it may be that a patient needs to be visited multiple times in a day by different carers at different times (note that these are just considered as individual/separate jobs in our model).
+\end{itemize}
+
+\subsection{Model and Solution Approach}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item The initial solution for the new plan is based on the original plan to save computing time.
+	\item There are two ways of creating an initial solution: either with an extra vehicle (EV) or without an EV.
+	\item If an EV must be used, the initial solution is formed as follows: the vehicles that are not broken-down continue on their routes as in the original plan, and are not affected by any changes.
+Note that the next customer that each of these vehicles is to serve at the time of the disruption is referred to as the `starting point' of these vehicles' routes in the new plan.
+An EV is sent out from the depot to resume the route of the broken-down vehicle, starting at the customer that the broken vehicle was on the way to serve (the starting point of the EV is the customer that was next in the original vehicle's route before it broke down).
+	\item When an EV is not required, an insertion algorithm is used to find an initial solution: the vehicles that are not broken-down continue on their routes as in the original plan (as above), and then the unserved customers in the disrupted route are inserted into the closest routes, according to their distance to the midpoint between the starting point and the depot.
+If an insertion violates the capacity constraint, then the customer is inserted into the next closest route.
+If there is no feasible insertion route, then the customer is simply inserted into the closest route, regardless of infeasibility.
+	\item TS is used, starting from an initial solution (which can be infeasible or feasible), moving to the best neighbouring solution until a time limit is reached.
+	\item Neighbourhood search is based on a relocation process: removing a customer from one route and inserting it into another route (same as \citet{yuan2017}).
+In each iteration, \emph{all} possible moves are tried for all customers, and the move that gives the least cost (i.e.
+the best of all neighbouring solutions) is chosen as the next move, provided it is not in the tabu list.
+	\item To explore more of the search space, customers that have been moved frequently are penalised.
+	\item To intensify the search space, each route is improve using two-opt and relocation after every $x$ iterations of TS for $y$ iterations; this is also performed every time a new best feasible solution is found.
+\end{itemize}
+
+\subsection{Modifications}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
 	\item 

### 2020-12-01 tex/SocialCare/SocialCare.tex
added more information.

@@
 \section{Constraints}
-\begin{itemize}[leftmargin=*, itemsep=-0.1em]
+\begin{itemize}[label=\textcolor{myRed}{\textbullet},leftmargin=*, itemsep=-0.1em]
 	\item Maximum number of carers per patient
@@
 	\item Are there distinct job types?
+	\item Will there always be enough carers for urgent services, i.e. if there are 10 patients that must be visited at 8am in the morning, are there always 10 carers available at that time?
+	\item Are there other carers ``on call'' that are able to work as-and-when if an emergency occurs, or is it that the carers working that day are the only carers available to do all the services.
+	\item Are there specific jobs that \emph{must} be done on time for the patient's wellbeing?
+(In contrast, there may be jobs where although tardiness will be penalised, it does not have a detrimental effect on the patient).
+This could include, for example, services where medication must be administered at a particular time.
+	\item Do schedules change daily or are services fixed daily/weekly?
+	\item In what situations will a new schedule/R\&R need to occur?
+	\item Are skills in levels? I.e. if a carer can do a PEG feed, can they do other tasks too?
 \end{itemize}
@@
 \section{Rerouting and Rescheduling}
-\begin{itemize}[leftmargin=*, itemsep=-0.em]
-	\item In what scenarios does and r\&r need to occur?
-	\item Addition of new patient (job) to the schedule (for example, emergency revisit)
-	\item Cancellation of job
-	\item Scenario where a patient that originally only required one carer now requires multiple carers
-	\item Change in skill requirement of a patient, where the originally scheduled carer does not have the required skill level
-	\item Patient emergency that requires the carer to remain with the patient; thus remaining jobs in the carer's route will need to be serviced by one or more other carers
-	\item Carer unwell at the beginning of the day, so all jobs assigned to that carer will need to be reassigned to other carer's working that day	
+\noindent In what scenarios does an R\&R need to occur?\vspace{3mm}
+
+\noindent \textbf{1. Addition of new job $j$ (either a new patient or an unexpected necessary revisit to a patient that has already been seen earlier in the day) to the schedule}
+\begin{itemize}[label=\textcolor{myRed}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item This would involve rescheulding one or more carers' routes to insert the new job $j$ in the correct time frame in a carer's route. 
+	\item If $j$ is a double service, then the job will have to be inserted into two different carers' routes in the same time frame, and the carers must be sufficiently skilled to perform the service.
+\end{itemize}
+
+\noindent \textbf{2. Patient emergency that requires the carer to remain with the patient (stay at job $j$)}
+\begin{itemize}[label=\textcolor{myOrange}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item Remaining jobs in the carer's route will need to be serviced by one or more other carers.
+	\item Note that at some point the carer may be able to resume work, and so another schedule may have to be created to reinsert the carer back into the plan.
+	\item If $j$ is a double service with carers $i_1$ and $i_2$, then it may be that only one carer needs to stay with the patient, whilst the other carer can resume their usual route.
+Then, the problem is to decide which carer should stay with the patient, as that carer's route will be affected by the time delay, and so the remaining jobs on that carer's route will need to be reassigned to other carers.
+Factors in this decision could include the skill level of the carers (it may be more beneficial for the carer with the higher skill level to resume their work), the locations of the remaining jobs of each carer's routes, whether one or more of the remaining jobs in their routes are double services/synchronised services (it may be easier for the carer assigned to a double service later on to resume their work, as rescheduling a double service is more difficult as it affects other carers), and the preferences of the remaining jobs and the carers.
+	\item However, it may be the case that both carers $i_1$ and $i_2$ must remain with the patient, and so all remaining jobs in both carers' routes will need to be reassigned to other carers.
+	\item Otherwise, if $j$ is a single service, then the remaining jobs in that carer's route will need to be reassigned to other carers.
+	\item If one or more of the remaining jobs in the route is/are double services, then the aim should be to reassign the job to another carer that has the same (or better) skill level as the original carer assigned the job, so that the other carer already assigned to that double service job can remain on the job and does not have to be reassigned because the total skill level between the two carers is insufficient.
+\end{itemize}
+
+\noindent \textbf{3. Cancellation of a job $j$}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item Carer $i$ can simply go to the next job $j+1$ on the route, although there will be a significantly longer waiting time for carer $i$ at job $j+1$. 
+	\item The same occurs if $j$ is a double service, carers $i_1$ and $i_2$ can simply skip job $j$ and go to the next job on their respective routes.
+\end{itemize}
+
+\noindent \textbf{4. Scenario where a job $j$ that originally only required one carer $i_1$ now requires multiple carers}
+\begin{itemize}[label=\textcolor{myGreen}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item The job $j$ may be able to remain in $i_1$'s route, provided another carer $i_2$ can be found that has the required skills in combination with carer $i_1$ to perform the service $j$ at the requested time.
+However, this may also delay the services in carer $i_2$'s route.
+	\item Furthermore, if there are no other carers that can perform the service $j$ with the carer $i_1$ (the original carer assigned $j$), then it may be that the job $j$ is unassigned from carer $i_1$, and $j$ is reassigned to two different carers.
+	\item In this scenario, carer $i_1$ would then be `free' during the time slot where $j$ was originally scheduled, and so it may be possible that carer $i_1$ is able to take over the job in the corresponding time slot of one of the other two carers that is now attending to $j$ (provided $i_1$ is skilled to do the job).
+If $i_1$ cannot perform a job for one of the other two carers, then other carers must be assigned those jobs, and so a larger rescheduling and rerouting must occur.
+\end{itemize}
+
+\noindent \textbf{5. Change in skill requirement of a job $j$, where the originally scheduled carer $i_1$ does not have the required skill level}
+\begin{itemize}[label=\textcolor{myBlue}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item Ideally in this situation, we would try to find another carer $i_2$ that has the required skill to do job $j$, and is also assigned a job $k$ at the same time as job $j$ with a lower skill requirement such that carer $i_1$ is skilled to perform job $k$.
+Then, we would be able to simply swap jobs $j$ and $k$ between carers $i_1$ and $i_2$.
+	\item However, this does not take into consideration the distance of the jobs from the carers' previous jobs in their respective routes, i.e.
+the distance from job $j-1$ to job $k$ and the distance from job $k-1$ to job $j$.
+\end{itemize}
+
+\noindent \textbf{6. Carer unwell at the beginning of the day/during the day/personal emergency}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item All (remaining) jobs assigned to that carer will need to be reassigned to other carers working that day.
+\end{itemize}
+
+\noindent \textbf{7. Carer requiring assistance from another carer}
+\begin{itemize}[label=\textcolor{myPink}{\textbullet},leftmargin=*, itemsep=-0.1em]
+	\item Another carer must be rerouted to the service to help the original carer on the job. 
+	\item Need to ensure that the carer chosen has the skill to assist on the job, and also select a carer that is closest to the location of the job.
+	\item This will disrupt the schedule of the rerouted carer, and so this will also have to be addressed.
 \end{itemize}

### 2020-12-01 tex/Pres/main.tex
presentation files.

\documentclass[usenames,dvipsnames]{beamer}
\input{includes/preamble.tex}

\begin{document}

\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
\date{}
\author{A. L. Hawa, C. Lamas-Fernandez, A. Martinez-Sykora, C. Currie, T. Monks, T. Chalklen}
\titlegraphic{\includegraphics[height=1.2cm]{figures/sotonabicare2.png}}
\maketitle

%\section{Introduction}

\begin{frame}{Task}
To create an efficient route and schedule for each carer such that all patients (jobs) are visited at the correct times by suitable carers.
\small
\begin{multicols}{2}
	\begin{itemize}
		\item Assign jobs to sufficiently skilled carer(s)
		\item Assign jobs to preferred carers(s)
		\item Assign carers to preferred jobs
		\item Workload balance %($\overline{B}$)
		\item Minimise waiting times %($\overline{W}$)
		\item Minimise tardiness %($\overline{Z}$)
		\item Minimise overtime %($\overline{O}$)
		\item Minimise travel time %($\overline{T}$)
		\item Minimise travel distance*
		%\item Assign jobs and carers according to preferences ($\overline{P}$)
	\end{itemize}	
\end{multicols}	
\end{frame}

\begin{frame}{Current Model and Approach}
%\begin{multicols}{2}
	\begin{itemize}
		\item Weighted objective function 
		\item Produce multiple schedules depending on the desired objective
		\item Currently implemented in C++ and Python
	\end{itemize}	
%\end{multicols}	
\end{frame}

\begin{frame}{Data}
This is required from each location:
\footnotesize
\begin{multicols}{2}
	\begin{itemize}
		\item Total number of patients/carers in each location
		\item Max travel distance (route/radius)
		\item Skills of carers
		\item Time window of each job
		\item Patient preferences: language, gender of carers, personal
		\item Carer preferences: gender of patient, allergies, personal
		\item Length of day for carers (full/half/flexible)
		\item Number of DS/synchronised services
		\item Distinct job types?
		\item Carers ``on call''
		\item Strength of preferences
	\end{itemize}	
\end{multicols}	
\end{frame}

\begin{frame}{Queries}
\scriptsize
%\begin{multicols}{2}
	\begin{itemize}
		\item \note{Desired objective(s)}
		\item Are there skill ``levels'' for carers?
		\item Distinct job types or customised per job
		\item Jobs that \emph{cannot} start late
		\item Lunch breaks
		\item Parking availability at patient's home
		\item PPE/COVID-19 conditions
		\item Frequency of schedule formation
		\item Database/software currently used
		\item Hard and soft constraints
		\item Any additional constraints?
		\item Who will be using the application?
		\item Additional carers available?
	\end{itemize}	
%\end{multicols}
\end{frame}

\begin{frame}{Rerouting and Rescheduling (R\&R)}
In the event of an unexpected disruption, the original plan may need to be updated to accommodate any changes that occur.

\note{Which of these do we need to address?}
\small
\begin{itemize}
	\item Cancellation of a service
	\item Addition of a new service (single or double)
	\item Modification of a service (single or double)
	\item Patient emergency, carer remains with patient
	\item Increase in skill level of service
	\item Increase in number of carers required for a service
	\item Carer unwell at beginning/during the day or personal emergency
	\item Carer requiring urgent assistance from another carer
\end{itemize}	
\end{frame}

\begin{frame}{Meeting}
\small
\begin{itemize}
	\item EasyLog only has details of time/length of job
	\item Preferences/carer skills/client needs are known by staff only
	\item No jobs between 2-4pm
	\item Day in two shifts: 7am-2pm, 4pm-10pm
	\item ``Ideal rotas'' on EasyLog
	\item Availability of extra carers (e.g. part-time, can help with issues)
	\item TW 15mins before and after start time
	\item QR code carer clock in/out, not accurate
	\item Larger gaps between jobs for new carers, e.g. +10mins travel time
	\item Continuity of care
\end{itemize}

\end{frame}

\begin{frame}{Literature}

\end{frame}

\end{document}

### 2021-01-07 tex/Pres/main.tex
WIP commit.

@@
 \input{includes/preamble.tex}
+\setbeamertemplate{caption}{\raggedright\insertcaption\par}
 
@@
 
+\begin{frame}{Comparison}
+\small
+\begin{multicols}{3}	
+\begin{block}{Hampshire}
+	\begin{itemize}
+		\item Carers = 15
+		\item Clients = 34
+		\item Jobs = 480
+	\end{itemize}
+\end{block}
+\begin{block}{Monmouth}
+	\begin{itemize}
+		\item Carers = 23
+		\item Clients = 47
+		\item Jobs = 735
+	\end{itemize}
+\end{block}
+\begin{block}{Aldershot}
+	\begin{itemize}
+		\item Carers = 9
+		\item Clients = 19
+		\item Jobs = 184
+	\end{itemize}
+\end{block}
+\end{multicols}	
+\end{frame}
+
+\begin{frame}{Comparison: Hampshire Time and Distance}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/Hampshire_time_comparison_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/Hampshire_dist_comparison_abi}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison: Monmouth Time and Distance}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/Monmouth_time_comparison_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/Monmouth_dist_comparison_abi}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison: Aldershot Time and Distance}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/Aldershot_time_comparison_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/Aldershot_dist_comparison_abi}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison: Hampshire Peak/Off-Peak Times}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/Hampshire_peaktime_comparison_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/Hampshire_offpeaktime_comparison_abi}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison: Monmouth Peak/Off-Peak Times}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/Monmouth_peaktime_comparison_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/Monmouth_offpeaktime_comparison_abi}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison: Aldershot Peak/Off-Peak Times}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/Aldershot_peaktime_comparison_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/Aldershot_offpeaktime_comparison_abi}
+	\end{figure}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[width=0.5\textwidth]{figures/Hampshire_timeofday_jobs_abi}
+	\end{figure}
+	\vspace*{-0.9cm}
+	\begin{columns}[T]
+		\begin{column}{0.5\textwidth}
+			\begin{figure}
+				\includegraphics[width=\textwidth]{figures/Monmouth_timeofday_jobs_abi}
+			\end{figure}
+		\end{column}%
+		\begin{column}{0.5\textwidth}
+			\begin{figure}
+				\includegraphics[width=\textwidth]{figures/Aldershot_timeofday_jobs_abi}
+			\end{figure}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}
+	\begin{figure}
+		\includegraphics[width=0.5\textwidth]{figures/Hampshire_shift_jobs_abi}
+	\end{figure}
+	\vspace*{-0.8cm}
+	\begin{columns}[T]
+		\begin{column}{0.5\textwidth}
+			\begin{figure}
+				\includegraphics[width=\textwidth]{figures/Monmouth_shift_jobs_abi}
+			\end{figure}
+		\end{column}%
+		\begin{column}{0.5\textwidth}
+			\begin{figure}
+				\includegraphics[width=\textwidth]{figures/Aldershot_shift_jobs_abi}
+			\end{figure}
+		\end{column}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Comparison: Total Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+	\midrule
+	02-Nov && 5640.00 & \gre{5452.46} && 6285.00 & \gre{5851.16} && 1480.00 & \gre{1449.90}\\
+	03-Nov && 5460.00 & \gre{5282.68} && 6030.00 & \gre{5758.70} && 1080.00 & \gre{1015.74}\\
+	04-Nov && 5160.00 & \gre{4996.38} && 6135.00 & \gre{6044.82} && 1355.00 & \gre{1321.32}\\
+	05-Nov && 5265.00 & \gre{5089.72} && 7200.00 & \gre{6955.46} && 2000.00 & \gre{1905.00}\\
+	06-Nov && 4875.00 & \gre{4617.36} && 7455.00 & \gre{7100.96} && 1470.00 & \gre{1345.28}\\
+	07-Nov && 3780.00 & \gre{3616.35} && 4845.00 & \gre{4574.79} && 1065.00 & \gre{1005.92}\\
+	08-Nov && 3720.00 & \gre{3608.01} && 4470.00 & \gre{4380.78} && 1050.00 & \gre{1006.87}\\
+	\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison: Total Travel Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+	\midrule
+	02-Nov && 475.00 & \gre{232.28} && 722.55 & \gre{687.00} && 192.77 & \gre{124.81}\\
+	03-Nov && 489.88 & \gre{227.02} && \red{728.62} & 745.88 && 166.05 & \gre{119.21}\\
+	04-Nov && 490.77 & \gre{242.58} && \red{769.78} & 895.18 && 178.28 & \gre{119.79}\\
+	05-Nov && 490.42 & \gre{271.44} && 798.20 & \gre{687.87} && 211.58 & \gre{122.75}\\
+	06-Nov && 479.50 & \gre{243.76} && \red{864.52} & 920.79 && 152.97 & \gre{108.11}\\
+	07-Nov && 452.67 & \gre{243.81} && \red{694.07} & 779.27 && 167.70 & \gre{108.08}\\
+	08-Nov && 405.41 & \gre{223.68} && \red{666.17} & 802.96 && 176.48 & \gre{119.98}\\
+	\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison NEW: Total Travel Time (minutes)}
+	\scriptsize
+	\begin{table}
+		%\caption{Comparison}
+		\renewcommand{\arraystretch}{1.5}
+		\begin{tabular}{ccrrcrrcrr}\toprule
+			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+			\midrule
+			02-Nov && 372.88 & \gre{232.28} && \red{674.46} & 687.00 && 153.05 & \gre{124.81}\\
+			03-Nov && 381.58 & \gre{227.02} && \red{675.89} & 745.88 && 129.47 & \gre{119.21}\\
+			04-Nov && 384.52 & \gre{242.58} && \red{722.42} & 895.18 && 141.19 & \gre{119.79}\\
+			05-Nov && 389.95 & \gre{271.44} && 745.86 & \gre{687.87} && 168.99 & \gre{122.7}5\\
+			06-Nov && 378.22 & \gre{243.76} && \red{824.72} & 920.79 && 123.08 & \gre{108.11}\\
+			07-Nov && 356.83 & \gre{243.81} && \red{661.50} & 779.27 && 133.91 & \gre{108.08}\\
+			08-Nov && 324.72 & \gre{223.68} && \red{636.07} & 802.96 && 142.60 & \gre{119.98}\\
+			\bottomrule
+		\end{tabular}
+	\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison: Total Waiting Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+	\midrule
+	02-Nov && \red{1595.00} & 1650.18 && 1407.45 & \gre{1009.16} && \red{222.23} & 260.09\\
+	03-Nov && \red{1580.12} & 1665.65 && 1026.38 & \gre{737.81} && 73.95 & \gre{56.53}\\
+	04-Nov && \red{1189.23} & 1273.80 && 775.22 & \gre{559.64} && \red{36.72} & 61.53\\
+	05-Nov && \red{1159.58} & 1203.28 && 1661.80 & \gre{1527.59} && 338.42 & \gre{332.25}\\
+	06-Nov && 1065.50 & \gre{1043.60} && 1505.48 & \gre{1095.16} && 402.03 & \gre{322.17}\\
+	07-Nov && \red{957.33} & 1002.54 && 790.93 & \gre{435.51} && \red{212.30} & 212.84\\
+	08-Nov && \red{1019.59} & 1089.32 && 548.83 & \gre{322.82} && \red{143.52} & 156.89\\
+	\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison NEW: Total Waiting Time (minutes)}
+	\scriptsize
+	\begin{table}
+		%\caption{Comparison}
+		\renewcommand{\arraystretch}{1.5}
+		\begin{tabular}{ccrrcrrcrr}\toprule
+			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+			\midrule
+			02-Nov && 1697.12 & \gre{1650.18} && 1455.54 & \gre{1009.16} && 261.95 & \gre{260.09}\\
+			03-Nov && 1688.42 & \gre{1665.65} && 1079.11 & \gre{737.81} && 110.53 & \gre{56.53}\\
+			04-Nov && 1295.48 & \gre{1273.80} && 822.58 & \gre{559.64} && 73.81 & \gre{61.53}\\
+			05-Nov && 1260.05 & \gre{1203.28} && 1714.14 & \gre{1527.59} && 381.01 & \gre{332.25}\\
+			06-Nov && 1166.78 & \gre{1043.60} && 1545.28 & \gre{1095.16} && 431.92 & \gre{322.17}\\
+			07-Nov && 1053.17 & \gre{1002.54} && 823.50 & \gre{435.51} && 246.09 & \gre{212.84}\\
+			08-Nov && 1100.28 & \gre{1089.32} && 578.93 & \gre{322.82} && 177.41 & \gre{156.89}\\
+			\bottomrule
+		\end{tabular}
+	\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison: Total Travel Distance (km)}
+	\scriptsize
+	\begin{table}
+		%\caption{Comparison}
+		\renewcommand{\arraystretch}{1.5}
+		\begin{tabular}{ccrrcrrcrr}\toprule
+			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+			\midrule
+			02-Nov && 288.26 & \gre{161.16} && \red{497.03} & 569.09 && 127.14 & \gre{100.00}\\
+			03-Nov && 307.63 & \gre{153.31} && \red{502.23} & 662.55 && 106.41 & \gre{99.03}\\
+			04-Nov && 301.65 & \gre{165.69} && \red{550.60} & 822.70 && 108.11 & \gre{90.21}\\
+			05-Nov && 302.08 & \gre{193.55} && \red{561.90} & 578.20 && 129.57 & \gre{92.00}\\
+			06-Nov && 301.39 & \gre{167.34} && \red{639.99} & 831.14 && 98.11 & \gre{83.70}\\
+			07-Nov && 291.58 & \gre{176.95} && \red{503.89} & 662.54 && 110.56 & \gre{82.77}\\
+			08-Nov && 250.77 & \gre{160.10} && \red{489.51} & 676.26 && 113.47 & \gre{91.53}\\
+			\bottomrule
+		\end{tabular}
+	\end{table}%
+\end{frame}
+
+\begin{comment}
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}[T]
+		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_time_info_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/4_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/5_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/6_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/7_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_workload_abi}%
+		\includegraphics[width=0.5\linewidth]{figures/8_Nov_20_Hampshire_workload_dst}
+	\end{figure}
+\end{frame}
+\end{comment}
+
+\begin{comment}
+%SAVE TEMPLATE
+\begin{frame}{Comparison: Total Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+\midrule
+02-Nov && 5640.00 & 5452.46 && 6285.00 & 5851.16 && 1480.00 & 1449.90\\
+03-Nov && 5460.00 & 5282.68 && 6030.00 & 5758.70 && 1080.00 & 1015.74\\
+04-Nov && 5160.00 & 4996.38 && 6135.00 & 6044.82 && 1355.00 & 1321.32\\
+05-Nov && 5265.00 & 5089.72 && 7200.00 & 6955.46 && 2000.00 & 1905.00\\
+06-Nov && 4875.00 & 4617.36 && 7455.00 & 7100.96 && 1470.00 & 1345.28\\
+07-Nov && 3780.00 & 3616.35 && 4845.00 & 4574.79 && 1065.00 & 1005.92\\
+08-Nov && 3720.00 & 3608.01 && 4470.00 & 4380.78 && 1050.00 & 1006.87\\
+\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison: Total Travel Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+\midrule
+02-Nov && 475.00 & 232.28 && 722.55 & 687.00 && 192.77 & 124.81\\
+03-Nov && 489.88 & 227.02 && 728.62 & 745.88 && 166.05 & 119.21\\
+04-Nov && 490.77 & 242.58 && 769.78 & 895.18 && 178.28 & 119.79\\
+05-Nov && 490.42 & 271.44 && 798.20 & 687.87 && 211.58 & 122.75\\
+06-Nov && 479.50 & 243.76 && 864.52 & 920.79 && 152.97 & 108.11\\
+07-Nov && 452.67 & 243.81 && 694.07 & 779.27 && 167.70 & 108.08\\
+08-Nov && 405.41 & 223.68 && 666.17 & 802.96 && 176.48 & 119.98\\
+\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison: Total Waiting Time (minutes)}
+	\scriptsize
+	\begin{table}
+		%\caption{Comparison}
+		\renewcommand{\arraystretch}{1.5}
+		\begin{tabular}{ccrrcrrcrr}\toprule
+			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+			\midrule
+			02-Nov && 1595.00 & 1650.18 && 1407.45 & 1009.16 && 222.23 & 260.09\\
+			03-Nov && 1580.12 & 1665.65 && 1026.38 & 737.81 && 73.95 & 56.53\\
+			04-Nov && 1189.23 & 1273.80 && 775.22 & 559.64 && 36.72 & 61.53\\
+			05-Nov && 1159.58 & 1203.28 && 1661.80 & 1527.59 && 338.42 & 332.25\\
+			06-Nov && 1065.50 & 1043.60 && 1505.48 & 1095.16 && 402.03 & 322.17\\
+			07-Nov && 957.33 & 1002.54 && 790.93 & 435.51 && 212.30 & 212.84\\
+			08-Nov && 1019.59 & 1089.32 && 548.83 & 322.82 && 143.52 & 156.89\\
+			\bottomrule
+		\end{tabular}
+	\end{table}%
+\end{frame}
+\end{comment}
+
+\begin{comment}
+\begin{frame}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/Hampshire_timeofday_jobs_abi}
+\end{figure}
+\vspace*{-0.9cm}
+\begin{columns}[T]
+\begin{column}{0.5\textwidth}
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/Monmouth_timeofday_jobs_abi}
+\end{figure}
+\end{column}%
+\begin{column}{0.5\textwidth}
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/Aldershot_timeofday_jobs_abi}
+\end{figure}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{figures/Hampshire_shift_jobs_abi}
+\end{figure}
+\vspace*{-0.8cm}
+\begin{columns}[T]
+\begin{column}{0.5\textwidth}
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/Monmouth_shift_jobs_abi}
+\end{figure}
+\end{column}%
+\begin{column}{0.5\textwidth}
+\begin{figure}
+\includegraphics[width=\textwidth]{figures/Aldershot_shift_jobs_abi}
+\end{figure}
+\end{column}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Comparison}
+\begin{figure}[T]
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_abi}%
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_time_info_dst}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_abi}%
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Hampshire_workload_dst}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Comparisonx}
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_abi}%
+\includegraphics[width=0.5\linewidth]{figures/3_Nov_20_Hampshire_workload_dst}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+\begin{figure}[B]
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Monmouth_time_info_abi}%
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Monmouth_workload_abi}
+\end{figure}
+\end{frame}
+
+\begin{frame}{Comparison}
+\begin{figure}
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Aldershot_time_info_abi}%
+\includegraphics[width=0.5\linewidth]{figures/2_Nov_20_Aldershot_workload_abi}
+\end{figure}
+\end{frame}
+
+\begin{frame}
+\begin{columns}
+\column{0.5\textwidth}
+\begin{minipage}[c][0.4\textheight][c]{\linewidth}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Hampshire_timeofday_jobs_abi}
+\end{minipage}
+\begin{minipage}[c][0.4\textheight][c]{\linewidth}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Monmouth_timeofday_jobs_abi}
+\end{minipage}
+\column{0.5\textwidth}
+\begin{minipage}[c][0.4\textheight][c]{\linewidth}
+\begin{enumerate}
+\item First.
+\item Second.
+\item Third.
+\end{enumerate}
+\end{minipage}
+\begin{minipage}[c][0.4\textheight][c]{\linewidth}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/Aldershot_timeofday_jobs_abi}
+\end{minipage}
+\end{columns}
+\end{frame}
+
+\begin{frame}
+\scriptsize
+\begin{multicols}{3}
+\begin{table}
+\setlength{\tabcolsep}{3pt}
+\renewcommand{\arraystretch}{1.5}
+%		\caption{Hampshire}
+\begin{tabular}{ccc}\toprule
+Date & Abicare & DST \\
+\midrule
+02-Nov & 5640.00 & 5452.46\\
+03-Nov & 5460.00 & 5282.68\\
+04-Nov & 5160.00 & 4996.38\\
+05-Nov & 5265.00 & 5089.72\\
+06-Nov & 4875.00 & 4617.36 \\
+07-Nov & 3780.00 & 3616.35\\
+08-Nov & 3720.00 & 3608.01\\
+\bottomrule
+\end{tabular}
+\end{table}%
+\vspace{2mm}
+\begin{table}
+\setlength{\tabcolsep}{3pt}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccc}\toprule
+Date & Abicare & DST \\
+\midrule
+02-Nov & 6285.00 & 5851.16\\
+03-Nov & 6030.00 & 5758.70\\
+04-Nov & 6135.00 & 6044.82\\
+05-Nov & 7200.00 & 6955.46\\
+06-Nov & 7455.00 & 7100.96 \\
+07-Nov & 4845.00 & 4574.79\\
+08-Nov & 4470.00 & 4380.78\\
+\bottomrule
+\end{tabular}
+\end{table}%
+\vspace{2mm}
+\begin{table}
+\setlength{\tabcolsep}{3pt}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccc}\toprule
+Date & Abicare & DST \\
+\midrule
+02-Nov & 1480.00 & 1449.90\\
+03-Nov & 1080.00 & 1015.74\\
+04-Nov & 1355.00 & 1321.32\\
+05-Nov & 2000.00 & 1905.00\\
+06-Nov & 1470.00 & 1345.28 \\
+07-Nov & 1065.00 & 1005.92\\
+08-Nov & 1050.00 & 1006.87\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{multicols}
+\end{frame}
+\end{comment}
+
 \end{document}

### 2021-01-07 code/c++/GRASPMSVS/GRASPMSVS/constructive.c

@@ -1,0 +1,2926 @@
+#include <stdlib.h> // To have min in linux
+#include "Python.h"
+#include "constructive.h"
+#include "grasp.h"
+
+// #ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+// #    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+// #else
+// #  define MODULE_API
+// #endif
+
+int main(int argc, char const* argv[]) {
+	struct INSTANCE inst = generate_instance();
+	struct INSTANCE* ip = &inst;
+	ip->verbose = 4;
+	return main_with_output(ip, NULL, NULL);
+}
+
+MODULE_API int python_entry(int nJobs_data, int nNurses_data, int nSkills_data, int verbose_data, float MAX_TIME_SECONDS, double* od_data, double* nurse_travel_from_depot_data,
+	double* nurse_travel_to_depot_data, int* nurseWorkingTimes_data, int* jobTimeInfo_data, int* jobRequirements_data, int* nurseSkills_data, int* solMatrixPointer, int* doubleService_data,
+	int* dependsOn_data, int* mk_mind_data, int* mk_maxd_data, int* capabilityOfDoubleServices, double* prefScore, double* algorithmOptions_data, int randomSeed) {
+
+	// printf("Inside C function. Random seed = %d.\n", randomSeed);
+
+	double eps = 1e-6;
+	int printInputData = (int)algorithmOptions_data[99] + eps;
+
+	int qm = (int)(algorithmOptions_data[0] + eps); // printf("Quality measure	= %d\n\n", qm);
+	int do_gap_not_precedence = (int)(algorithmOptions_data[12] + eps); // printf("Use gaps for dependent jobs (rather than precedence)	= %d\n", do_gap_not_precedence);
+
+	if (qm == 0 && do_gap_not_precedence == 0)
+		printf("WARNING: Solving with AIT H quality measure but with PRECEDENCE rather than GAP.\n");
+
+	// printf("Quality measure	= %d\n\n", (int) (algorithmOptions_data[0] + eps));
+	// printf("do_twopt 		= useTwoOpt = (int) ip->algorithmOptions[1] + eps = %d\n", (int) (algorithmOptions_data[1] + eps));
[…]

### 2021-01-07 code/c++/GRASPVNS/constructive.h

@@ -21,9 +21,11 @@ struct INSTANCE {
 	int nSkills; //Number of skills
 	int verbose; // Additional information
 	int isFeasible; // Feasibility (deprecated)
-	float objTravel; // Objective travel
 	float objTime; // Objective time
+	float objTravel; // Objective travel time
+	float objWaiting; // Objective waiting time, NEW 07/01/2021
 	float objTardiness; // Objective tardiness
+	float objShortestDay; // Objective shortest day, NEW 07/01/2021
 	float objLongestDay; // Objective longest day
 	float objOvertime; // Objective overtime
 	float totPref; // Total preference score
@@ -116,6 +118,10 @@ double sol_quality_light(struct INSTANCE* ip);
 double sol_quality_optimised(struct INSTANCE* ip, int n1, int n2);
 double sol_quality(struct INSTANCE* ip, int report);
 int synchronise_jobs(struct INSTANCE* ip);
+// void mins_to_time(double time, int* hours, int* minutes, int* seconds);
+int* mins_to_time(double time);
+// double mins_to_minsecs(double time);
+int* mins_to_minsecs(double time);
 double obj_from_times(struct INSTANCE* ip, int report);
 double alternative_quality(struct INSTANCE* ip, int report);
 int random_integer(int min_val, int max_val);

### 2021-01-07 code/screpo/constructive_wrapper.py

@@ -16,8 +16,9 @@ from joblib import Parallel, delayed
 # Own modules:
 import mankowska_data as Mk
 import python_resources as pr
-import instance_handler as hhc
-import convert_dict_inst as cdi
+# import instance_handler as hhc
+import instance_handler_idict as hhc
+# import convert_dict_inst as cdi
 
 def main():
     # New main function.
@@ -29,58 +30,64 @@ def main():
     # max_time_seconds = 5
     # verbose_level = 1
     create_python_plots = True
-    create_html_website = False
-    results_file_name = '08_Nov_Salisbury_results.txt'
+    # create_html_website = False
+    create_html_website = True
+    input_filename = 'all_inst_Aldershot.p'
+    # results_filename = '02_Nov_Hampshire_results.txt'
+    # results_filename = 'test_02_Nov_Hampshire_results.txt'
     # run_in_parallel = False
     # parallel_workers = 1 # Only works if previous is true
     # random_seed = -1 # Only if run_in_parallel = False. -1 is a true random seed, otherwise set to this value
-    random_seed = 13826 # NOTE: used for testing only.
+    random_seed = 13027 # NOTE: used for testing only. #35807 600seconds 13027
 
     options_vector = hhc.default_options_vector() 
-    options_vector[1] = 1.0 # Two-opt active
-    options_vector[3] = 0.0 # Nurse order change active (neighbourhood in local search)
-    options_vector[4] = 0.5 # -   GRASP: Delta low
-    options_vector[5] = 0.48 # -   GRASP: Delta range
-    options_vector[6] = 1.0 # Nurse order change active (In GRASP, between calls)
-    options_vector[7] = 1.0 # deprecated
-    options_vector[8] = 20.0 # Solutions in pool
-    options_vector[9] = 3.0 # PR_STRATEGY
-    options_vector[10] = 2.0 # RCL Strategy
-    options_vector[11] = 2.0 # PR_DIRECTION
-    options_vector[12] = 0 # -   Use gap (1) or precedence (0)
-    options_vector[50] = 0 # 1 if tardiness and overtime are infeasible, 0 if feasible
-    options_vector[51] = -1/3*60 # alpha_1 Travel time
-    options_vector[52] = 0.0 # alpha_2 Waiting time
-    options_vector[53] = -1/3*60 # alpha_3 Tardiness
-    options_vector[54] = 0.0 # alpha_4 Overtime
[…]

### 2021-01-07 code/screpo/convert_dict_inst.py

@@ -11,7 +11,7 @@ import sys
 import glob
 import pickle
 import math
-import requests # For website
+import requests # For osrm
 import numpy as np
 import pandas as pd
 import geopandas
@@ -20,9 +20,20 @@ from scipy import stats
 
 # Import python file in project
 import tools_and_scripts.class_cpo_df as ccd
+# from instance_handler import geopoint
 
 # Display all rows and columns of dataframes in command prompt:
-# pd.set_option("display.max_rows", None, "display.max_columns", None)
+pd.set_option("display.max_rows", None, "display.max_columns", None)
+
+class geopoint1(object):
+    def __init__(self, latitude, longitude):
+        self.lat = latitude
+        self.long = longitude
+    def latlong(self):
+        return([self.lat, self.long])
+    def longlat(self):
+        return([self.long, self.lat])
+### --- End class geopoint  --- ###
 
 def convert_dict_inst(inst, idict):
     # This function creates an instance 'inst' of the class 'INSTANCE', and initialises inst with data from our idict dictionary instance created by analyse_mileage.py.
@@ -38,6 +49,8 @@ def convert_dict_inst(inst, idict):
     for i in range(inst.nNurses): #range(len(idict['rota']['start']))
         inst.nurseWorkingTimes[i][0] = idict['rota']['start'][i]
         inst.nurseWorkingTimes[i][1] = idict['rota']['finish'][i]
+        # inst.nurseWorkingTimes[i][0] = idict['rota']['home_start'][i]
+        # inst.nurseWorkingTimes[i][1] = idict['rota']['home_finish'][i]
         # We dont have max working time, so we need to calculate the duration of each carer's shift (total minutes that each carer works during the day)
         maxWorkingTime = idict['rota']['finish'][i] - idict['rota']['start'][i]
         inst.nurseWorkingTimes[i][2] = maxWorkingTime
@@ -53,8 +66,8 @@ def convert_dict_inst(inst, idict):
     inst.jobSkillsRequired = np.ones((inst.nJobs, inst.nSkills), dtype=np.int32) # Note: this will only work if nSkills > 0.
[…]

### 2021-01-07 code/screpo/instance_handler.py

@@ -11,6 +11,7 @@ import subprocess
 from math import sqrt
 import pickle
 import osrm
+import requests # for osrm
 
 import mankowska_data as Mk
 import convert_dict_inst as cdi
@@ -25,14 +26,20 @@ def create_solve_inst(idict, options_vector, random_seed):
     
     # i = INSTANCE()
     inst = INSTANCE()
+    # inst.init_job_and_nurse_objects()
 
     if mkVNS: # For some reason an error occurs (cannot call inst.solve()) without this if statement, so just left it in.
         inst = Mk.generate_Mk(inst, matfile=file_to_run, filetype='large_vns')
     else:
         inst = cdi.convert_dict_inst(inst, idict)
+    inst.init_job_and_nurse_objects()
+    inst = cdi.initialise_nurse_job(inst, idict)
     # exit(-1)
     inst.algorithmOptions = options_vector
-    inst.algorithmOptions[0] = 1.0 # MANKOWSKA
+    # inst.algorithmOptions[0] = 0.0 # ait_h
+    inst.algorithmOptions[0] = 1.0 # mankowska
+    # inst.algorithmOptions[0] = 5.0 # workload_balance
+    # inst.algorithmOptions[0] = 6.0 # paper
     saved_od_data = inst.od[0][0]
     inst.solve(randomSeed=random_seed, printAllCallData=False)
     inst.solve = []
@@ -216,6 +223,7 @@ class JOB(object):
         self.dependsOn = -1
         self.minimumGap = 0
         self.maximumGap = 0
+        self.location = [] # NOTE: NEW, 27/12/2020 ALH
         # Calculated (solution)
         self.assignedNurse = -1
         self.positionInSchedule = 0
@@ -238,7 +246,7 @@ class NURSE(object):
         self.preferences = []
         self.preferredJobs = []
         self.jobsToAvoid = []
-
+        self.postcode = 'Unknown' # NOTE: NEW, 27/12/2020 ALH
         # Calculated (solution)
         self.jobsServed = 0
         self.startTime = 0
@@ -283,7 +291,7 @@ class INSTANCE(object):
         self.capabilityOfDoubleServices = []
 
         # Post-processed solution:
-        self.nurseRoute = []
+        self.nurseRoute = [] # In post_process_solution, this will become equivalent to allNurseRoutes, with dimensions nNurses x nPositions(jobs), and so nurseRoute[nurse][position] = job.
         self.nurseWaitingTime = []
         self.nurseServiceTime = []
         self.nurseTravelTime = []
[…]

### 2021-01-07 code/screpo/src/C/grasp.c

@@ -1087,8 +1087,8 @@ void rcl_pick(int** bestIndices, int* RCL_seeds, int rcl_size, int* cNurse, int*
 	// This function randomly picks an element from the RCL, which returns to us the indicies for the job-nurse assignment (cNurse and cJob).
 	// Picks one random element from bestIndices. The index of bestIndices comes from RCL_seeds. The elements considered of RCL_seeds are only the first rcl_size elements. Output is saved in cNurse and cJob.
 
-	// int rd_int = pick_integer(rcl_size);
-	int rd_int = 0; //ISSUE, THIS SHOULD BE SELECTING A RANDOM INTEGER BETWEEN 0 AND rcl_size, SHOULD NOT ALWAYS BE 0 AS THIS MEANS THAT THE ELEMENT PICKED WILL ALWAYS BE THE 0TH ELEMENT OF RCL_seeds!
+	 int rd_int = pick_integer(rcl_size);
+//	int rd_int = 0; //ISSUE, THIS SHOULD BE SELECTING A RANDOM INTEGER BETWEEN 0 AND rcl_size, SHOULD NOT ALWAYS BE 0 AS THIS MEANS THAT THE ELEMENT PICKED WILL ALWAYS BE THE 0TH ELEMENT OF RCL_seeds!
 
 	int el = RCL_seeds[rd_int]; //el = element, (should be) selected at random from RCL_seeds.
 
@@ -1339,8 +1339,9 @@ int find_arc_destination(int source_nurse, int source_job, struct INSTANCE* ip) 
 	// If there is a job after source_job in source_nurse's route, then this function returns that job, otherwise it returns -1 (which means that there are no other jobs in source_nurse's route,
 	// source_job is the last job in source_nurse's route and so the nurse ends up returning to the depot).
 
-	int c_pos = ip->solMatrix[source_nurse][source_job]; //c_pos position of job in nurse's route
-	int n_pos = c_pos++; /// Increment is AFTER, so n_pos = c_pos, and THEN c_pos is increased by one, should this be ++c_pos or just c_pos + 1 instead as we're trying to find the NEXT position in the route?
+	int c_pos = ip->solMatrix[source_nurse][source_job]; //c_pos position of job in nurse's route, 'current position'
+	/*int n_pos = c_pos++;*/ /// 'next position'. Increment is AFTER, so n_pos = c_pos, and THEN c_pos is increased by one, should this be ++c_pos or just c_pos + 1 instead as we're trying to find the NEXT position in the route?
[…]

### 2021-01-07 code/screpo/tools_and_scripts/abicare_data.py

@@ -1,0 +1,206 @@
+#--------------------#
+# fc_analyse_mileage.py
+# 17/12/2020
+# 
+#--------------------#
+
+import math
+import pickle
+import datetime
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+
+# fc_analyse_mileage.py: full codepoint analyse mileage.
+
+class POSTCODE_FINDER():
+    """
+    docstring
+    """
+    df = []
+    filename = []
+    def __init__(self, filename=r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\full_codepoint.csv'):
+        self.filename = filename
+        # self.df = pd.read_csv(self.filename)
+        # self.df['post'] = self.df['Postcode 1'].str.replace(' ', '')
+        # self.df['post'] = self.df['post'].str.lower()
+        # self.n_adr_filename = r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\full_codepoint.csv'
+        self.df = pd.read_csv(self.filename)
+        self.df['PC'] = self.df['PC'].str.replace('"', '')
+        self.df['PC'] = self.df['PC'].str.replace(' ', '') # Changed from self.adr_df['PC'].str.replace(' ', '') to self.adr_df['post'].str.replace(' ', '')
+        self.df['PC'] = self.df['PC'].str.lower()
+
+    def find_postcode_latlon(self, postcode): # Returns Latitude and Longitude of given postcode.
+        dfview = self.df[self.df['PC'] == postcode]
+        if dfview.empty:
+            return [None, None]
+        else:
+            # Need to convert eastings and northings to lat and lon
+            return [dfview.iloc[0]['EA'], dfview.iloc[0]['NO']]
+    
+    def find_postcode_n_addresses(self, postcode): # Returns number of registered properties at a given postcode.
+        dfview = self.df[self.df['PC'] == postcode]
+        if dfview.empty:
+            return -1
+        else:
+            return dfview.iloc[0]['RP']
+
+def time_dif_to_minutes(endtime, starttime, round=True):
+    mins = (endtime - starttime).seconds/60
+    if round:
+        return np.round(mins, 0)
+    else:
+        return mins
+
+def inst_dicts_to_dfs(inst):
+    inst['rota'] = pd.DataFrame(inst['rota'])
+    inst['tasks'] = pd.DataFrame(inst['tasks'])
[…]

### 2021-01-07 code/screpo/tools_and_scripts/analyse_mileage.py

@@ -16,7 +16,7 @@ class POSTCODE_FINDER():
         self.df = pd.read_csv(self.filename)
         self.df['post'] = self.df['Postcode 1'].str.replace(' ', '')
         self.df['post'] = self.df['post'].str.lower()
-        self.n_adr_filename = r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\salisbury_full_codepoint.csv'
+        self.n_adr_filename = r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\full_codepoint.csv'
         self.adr_df = pd.read_csv(self.n_adr_filename)
         self.adr_df['post'] = self.adr_df['PC'].str.replace('"', '')
         self.adr_df['post'] = self.adr_df['post'].str.replace(' ', '') # Changed from self.adr_df['PC'].str.replace(' ', '') to self.adr_df['post'].str.replace(' ', '')
@@ -85,7 +85,7 @@ days_in_df.sort() # Sort the dates in chronological order
 all_instances = []
 largeprint = True
 timewindow_interval = datetime.timedelta(minutes = 15) # Timewindow generated as plus-minus these minutes of the start date, change this value 30, 15, etc.
-only_areas = ['Salisbury']
+only_areas = ['Hampshire']
 min_addresses = 10
 largeprintstr = ''
 
@@ -212,7 +212,7 @@ for a_name in u_areas: # For each area in the list of Areas
     
         inst['txtsummary'] = largeprintstr
         all_instances.append(inst_dicts_to_dfs(inst).copy())
-        pickle.dump(all_instances, open('all_inst_salisbury.p', 'wb'))
+        pickle.dump(all_instances, open('all_inst_hampshire.p', 'wb'))
 
 #print('Last instance:')
 #print(print_inst(all_instances[-1]))

### 2021-01-07 code/screpo/tools_and_scripts/class_cpo_df.py

@@ -1,21 +1,25 @@
+#--------------------#
+# class_cpo_df.py
+# 16/12/2020
+# class CPO_DF(): CodepointOpen Dataframe class
+# This class opens all csv files in codepo_gb folder, and goes through all files to find postcodes for the given CC/DC codes.
+# These dataframes are put into a dictionary, which are then put into a list and merged together into a single dataframe, df.
+# As codepoint open only contains eastings and northings, geopandas is used to convert these to longitudes and latitudes, which are added as columns to df.
+# This class is used to obtain the longitudes and latitudes of any postcode in the given CC/DC areas, directly from codepoint open rather than using postcodelookup and fullcodepoint files.
+# Note: codepoint open does not contain the number of addresses per postcode (i.e. 'RP' column).
+# Used in convert_dict_inst.pu and compre_abicare_osrm.py.
+#--------------------#
+
 import os
 import glob
-import pickle
-import datetime
 import numpy as np
 import pandas as pd
 import geopandas
-import matplotlib.pyplot as plt
-from scipy import stats
 
 class CPO_DF(): # Codepoint Open DataFrame
-    """
-    docstring
-    """
-    # df = []
     def __init__(self, foldername=r'data\codepo_gb'):
+    # def __init__(self, foldername=r'..\data\codepo_gb'):
         self.foldername = foldername # CP_Open_Folder = r'..\data\codepo_gb'
-        # self.foldername = r'..\data\codepo_gb'
         all_cpopen_csvs = glob.glob(os.path.join(self.foldername, 'data\\CSV', '*.csv'))
 
         # Get the headers in a list:
@@ -23,9 +27,9 @@ class CPO_DF(): # Codepoint Open DataFrame
             raw_headers = f.readline()
             cp_open_headers = raw_headers.split(',') # cp_open_headers:  ['PC', 'PQ', 'EA', 'NO', 'CY', 'RH', 'LH', 'CC', 'DC', 'WC\n'], it's a list of strings.
         
-        # We are interested in the areas 'Wiltshire', 'Hampshire', and 'Monmouthshire', which we search with the codes below (county code 'CC', district code 'DC')
-        utas = [['DC', 'E06000054'], ['CC', 'E10000014'], ['DC', 'W06000021']]
-        # utas = [['DC', 'E06000054']]
[…]

### 2021-01-07 code/screpo/tools_and_scripts/class_cpo_df1.py

@@ -1,0 +1,98 @@
+#--------------------#
+# class_cpo_df.py
+# 16/12/2020
+# class CPO_DF(): CodepointOpen Dataframe class
+# This class opens all csv files in codepo_gb folder, and goes through all files to find postcodes for the given CC/DC codes.
+# These dataframes are put into a dictionary, which are then put into a list and merged together into a single dataframe, df.
+# As codepoint open only contains eastings and northings, geopandas is used to convert these to longitudes and latitudes, which are added as columns to df.
+# This class is used to obtain the longitudes and latitudes of any postcode in the given CC/DC areas, directly from codepoint open rather than using postcodelookup and fullcodepoint files.
+# Note: codepoint open does not contain the number of addresses per postcode (i.e. 'RP' column).
+# Used in convert_dict_inst.pu and compre_abicare_osrm.py.
+#--------------------#
+
+import os
+import glob
+import numpy as np
+import pandas as pd
+import geopandas
+
+class CPO_DF1(): # Codepoint Open DataFrame
+    # def __init__(self, foldername=r'data\codepo_gb'):
+    def __init__(self, foldername=r'..\data\codepo_gb'):
+        self.foldername = foldername # CP_Open_Folder = r'..\data\codepo_gb'
+        all_cpopen_csvs = glob.glob(os.path.join(self.foldername, 'data\\CSV', '*.csv'))
+
+        # Get the headers in a list:
+        with open(os.path.join(self.foldername, 'Doc', 'Code-Point_Open_Column_Headers.csv')) as f:
+            raw_headers = f.readline()
+            cp_open_headers = raw_headers.split(',') # cp_open_headers:  ['PC', 'PQ', 'EA', 'NO', 'CY', 'RH', 'LH', 'CC', 'DC', 'WC\n'], it's a list of strings.
+        
+        # We are interested in the areas 'Wiltshire', 'Hampshire', 'Berkshire/Buckinghamshire', and 'Monmouthshire', which we search with the codes below (CC county code, DC district code).
+        # Wiltshire: DC E06000054, Hampshire: CC E10000014, Berkshire/Buckinghamshire: DC E06000060, Monmouthshire: DC W06000019, W06000020 and W06000021, W06000022
+        utas = [['DC', 'E06000054'], ['CC', 'E10000014'], ['DC', 'E06000060'], ['DC', 'W06000019'], ['DC', 'W06000020'], ['DC', 'W06000021'], ['DC', 'W06000022']]
+
+        count = 0
+        dict_dfs = {}
[…]

### 2021-01-07 code/screpo/tools_and_scripts/compare_abicare_dst.py

@@ -1,0 +1,467 @@
+# Import modules
+import os
+import sys
+import glob
+import math
+import folium
+import pickle
+import requests # For website
+import geopandas
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+
+
+def inst_dicts_to_dfs(inst):
+    inst['stats'] = pd.DataFrame(inst['stats'])
+    inst['abi'] = pd.DataFrame(inst['abi'])
+    inst['dst'] = pd.DataFrame(inst['dst'])
+    return inst
+### --- End inst_dicts_to_dfs --- ###
+
+def timemins_to_string(mins):
+    minsRound = math.floor(mins)
+    # print('minsRound: ', minsRound)
+    hours = mins // 60
+    # print('hours:', hours)
+    minutes = (minsRound % 60)
+    # print('minutes: ', minutes)
+    seconds = (mins - minsRound) * 60
+    seconds = round(seconds)
+    # print('seconds: ', seconds)
+    return('{0:0>2}:{1:0>2}:{2:0>2}'.format(int(hours), int(minutes), int(seconds)))
+### --- End def timemins_to_string --- ###
+
+def create_df(filename, a_name=''):
+    # filename = '../outputs/results/HC.csv' #File from Abicare
+    # a_name = 'Hampshire'
+    df = pd.read_csv(filename)
+    # df['Date'] = pd.to_datetime(df['Date'], format='%d-%b-%y %I:%M %p')
+
+    inst = {
+            'name' : a_name,
+            'all_dates' : [],
+            'nNurses' : 0,
+            'nJobs' : 0,
+            'nClients' : 0,
+            'stats' : {'nurses' : [], 'jobs' : [], 'clients' : [], 'servicetime' : []},
+            'abi' : {'ttime' : [], 'ttravel' : [], 'ttravelosrm' : [], 'twaiting' : [], 'twaitingosrm' : [], 'dist' : [], 'distjobs' : [], 'distjobsosrm' : [], 'aith': [], 'mk' : [], 'wb' : [], 'paper' : []},
+            'dst' : {'ttime' : [], 'ttravel' : [], 'twaiting' : [], 'dist' : [], 'distjobs' : [], 'aith': [], 'mk' : [], 'wb' : [], 'paper' : [], 'algtime' : [], 'iterations' : []},
[…]

### 2021-01-07 code/screpo/tools_and_scripts/fc_analyse_mileage.py

@@ -1,0 +1,367 @@
+#--------------------#
+# fc_analyse_mileage.py
+# 17/12/2020
+# 
+#--------------------#
+
+import math
+import pickle
+import datetime
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+
+# fc_analyse_mileage.py: full codepoint analyse mileage.
+
+class POSTCODE_FINDER():
+    """
+    docstring
+    """
+    df = []
+    filename = []
+    def __init__(self, filename=r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\full_codepoint.csv'):
+        self.filename = filename
+        # self.df = pd.read_csv(self.filename)
+        # self.df['post'] = self.df['Postcode 1'].str.replace(' ', '')
+        # self.df['post'] = self.df['post'].str.lower()
+        # self.n_adr_filename = r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\full_codepoint.csv'
+        self.df = pd.read_csv(self.filename)
+        self.df['PC'] = self.df['PC'].str.replace('"', '')
+        self.df['PC'] = self.df['PC'].str.replace(' ', '') # Changed from self.adr_df['PC'].str.replace(' ', '') to self.adr_df['post'].str.replace(' ', '')
+        self.df['PC'] = self.df['PC'].str.lower()
+
+    def find_postcode_eastnorth(self, postcode): # Returns eastings and northings of given postcode
+        dfview = self.df[self.df['PC'] == postcode]
+        if dfview.empty:
+            return [None, None]
+        else:
+            # Need to convert eastings and northings to lat and lon
+            return [dfview.iloc[0]['EA'], dfview.iloc[0]['NO']]
+    
+    def find_postcode_n_addresses(self, postcode): # Returns number of registered properties at a given postcode.
+        dfview = self.df[self.df['PC'] == postcode]
+        if dfview.empty:
+            return -1
+        else:
+            return dfview.iloc[0]['RP']
+
+def time_dif_to_minutes(endtime, starttime, round=True):
+    mins = (endtime - starttime).seconds/60
+    if round:
+        return np.round(mins, 0)
+    else:
+        return mins
+
+def inst_dicts_to_dfs(inst):
+    inst['rota'] = pd.DataFrame(inst['rota'])
[…]

### 2021-01-07 docs/constructive.c

@@ -1,0 +1,3011 @@
+#include <stdlib.h> // To have min in linux
+#include "Python.h"
+#include "constructive.h"
+#include "grasp.h"
+
+// #ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+// #    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+// #else
+// #  define MODULE_API
+// #endif
+
+int main(int argc, char const* argv[]){
+    struct INSTANCE inst = generate_instance();
+    struct INSTANCE* ip = &inst;
+    ip->verbose = 4;
+    return main_with_output(ip, NULL, NULL);
+}
+
+MODULE_API int python_entry(int nJobs_data, int nNurses_data, int nSkills_data, int verbose_data, float MAX_TIME_SECONDS, double* od_data, double* nurse_travel_from_depot_data,
+                            double* nurse_travel_to_depot_data, int* nurseWorkingTimes_data, int* jobTimeInfo_data, int* jobRequirements_data, int* nurseSkills_data, int* solMatrixPointer,
+                            int* doubleService_data,
+                            int* dependsOn_data, int* mk_mind_data, int* mk_maxd_data, int* capabilityOfDoubleServices, double* prefScore, double* algorithmOptions_data, int randomSeed){
+
+    // printf("Inside C function. Random seed = %d.\n", randomSeed);
+
+    double eps = 1e-6;
+    int printInputData = (int) algorithmOptions_data[99] + eps;
+
+    int qm = (int) (algorithmOptions_data[0] + eps); // printf("Quality measure	= %d\n\n", qm);
+    int do_gap_not_precedence = (int) (algorithmOptions_data[12] + eps); // printf("Use gaps for dependent jobs (rather than precedence)	= %d\n", do_gap_not_precedence);
+
+    if(qm==0 && do_gap_not_precedence==0)
+        printf("WARNING: Solving with AIT H quality measure but with PRECEDENCE rather than GAP.\n");
+
+    // printf("Quality measure	= %d\n\n", (int) (algorithmOptions_data[0] + eps));
+    // printf("do_twopt 		= useTwoOpt = (int) ip->algorithmOptions[1] + eps = %d\n", (int) (algorithmOptions_data[1] + eps));
[…]

### 2021-01-07 docs/constructive.h

@@ -1,0 +1,148 @@
+#include<stdio.h>
+#include <stdlib.h>
+#include <time.h>
+
+#ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+#    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+#else
+#  define MODULE_API
+#endif
+
+
+#define bigM 9999999;
+
+struct INSTANCE {
+	int nJobs; // Number of jobs
+	int nNurses; // Number of nurses
+	int nSkills; //Number of skills
+	int verbose; // Additional information
+	int isFeasible; // Feasibility (deprecated)
+	float objTravel; // Objective travel
+	float objTime; // Objective time
+	float objTardiness; // Objective tardiness
+	float objLongestDay; // Objective longest day
+	float objOvertime; // Objective overtime
+	float totPref; // Total preference score
+	int quality_measure; // Temp. Allow to choose (Mk, Ait h., etc.) Ideally we want a vector with weights! This should be set ONLY in "instance_from_python" (read_instance.c). The value is hardcoded at the moment.
+	float MAX_TIME_SECONDS; // Maximum time limit in seconds
+	// double **od_cost;
+	int** nurseWorkingTimes; // 2D array, size = nNurses x 3. For each nurse i (row): column[0] = start time, column[1] = finish time, column[2] = max working time. (column[2] not used?)
+	int** solMatrix; // 2D array, size = nNurses x nJobs. For each nurse i (row), the element in solMatrix[i][j] gives the position in nurse i's route where nurse i visits job j.
+	// E.g. if solMatrix[2][4] = 3 -> the third job in nurse 2's route is job 4. If solMatrix[i][j] = -1, then job j is not in nurse i's route.
+	int** jobTimeInfo; // 2D array, size = nJobs x 3. For each job i (row): column[0] = start time window, column[1] = finish time window, column[2] = length of job.
+	int** jobRequirements; // 2D array, size = nJobs x nSkills. Skill 0 means headcount needed. (deprecated)
+	int** nurseSkills; // 2D array, size = nNurses x nSkills. Skill 0 is always 1 (one worker). (deprecated)
+	int** nurseSkilled; // 2D array, size = nNurses x nJobs. For each nurse i (row) and each job j (column): nurseSkilled[i][j] = 1 if nurse i can do job j, = 0 otherwise.
[…]

### 2021-01-07 docs/grasp.h

@@ -1,0 +1,29 @@
+#include "math.h"
+
+
+int GRASP(struct INSTANCE * ip);
+void randomised_constructive(struct INSTANCE * ip, int randomness, double delta, int rcl_strategy);
+void grasp_ls(struct INSTANCE * ip, int LS_ITERS);
+void generate_rcl(double delta, int * rcl_size, int * RCL_seeds, double * rankValues, int rv_size, double bestRank, double worstRank, int strategy);
+void rcl_pick(int ** bestIndices, int * RCL_seeds, int rcl_size, int * cNurse, int * cJob);
+void best_index_pick(int ** bestIndices, int GRASP_param, int * cNurse, int * cJob);
+int pick_integer(int max_int);
+void update_top_values(int GRASP_param, int ** bestIndices, double * bestValues, double newValue, int bestNurse, int bestJob);
+void clean_solution_from_struct(struct INSTANCE * ip);
+int identify_early_insert(struct INSTANCE * ip, struct INSTANCE * guiding, int guiding_nurse, int guiding_nurse_pos);
+void find_best_worst_pool_qualities(double * pool_best, double * pool_worst, struct INSTANCE ** pool, double * pool_quality, int pool_size);
+double solution_dissimilarity(struct INSTANCE * input1, struct INSTANCE * input2);
+int find_arc_destination(int source_nurse, int source_job, struct INSTANCE * ip);
+void print_and_recalculate_pool_contents(struct INSTANCE ** pool, double * pool_quality,  int solutions_in_pool);
+void calculate_and_print_dissimilartiy_matrix(struct INSTANCE ** pool, int solutions_in_pool);
+double forward_path_relinking(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output);
+double forward_and_backward_path_relinking(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output1);
+double backward_path_relinking(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output);
+double directed_path_relinking(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output, int direction);
+double path_relinking(struct INSTANCE * ip, struct INSTANCE * guiding);
+void get_nurse_and_position_of_job(struct INSTANCE * ip, int job, int * nurse, int * position);
[…]

### 2021-04-20 tex/Pres/main.tex
updated commit.

@@
 \date{}
-\author{A. L. Hawa, C. Lamas-Fernandez, A. Martinez-Sykora, C. Currie, T. Monks, T. Chalklen}
+%\author{A. L. Hawa, C. Lamas-Fernandez, A. Martinez-Sykora, C. Currie, T. Monks, T. Chalklen}
 \titlegraphic{\includegraphics[height=1.2cm]{figures/sotonabicare2.png}}
@@
 
+\begin{comment}
 \begin{frame}{Queries}
@@
 \end{frame}
+\end{comment}
 
-\begin{frame}{Comparison}
+\begin{frame}{Comparison: Abicare vs DST}
 \small
@@
 	\begin{itemize}
+		\setlength{\itemindent}{-.1in}
 		\item Carers = 15
@@
 	\begin{itemize}
+		\setlength{\itemindent}{-.1in}
 		\item Carers = 23
@@
 	\begin{itemize}
+		\setlength{\itemindent}{-.1in}
 		\item Carers = 9
@@
 			\midrule
-	02-Nov && 475.00 & \gre{232.28} && 722.55 & \gre{687.00} && 192.77 & \gre{124.81}\\
-	03-Nov && 489.88 & \gre{227.02} && \red{728.62} & 745.88 && 166.05 & \gre{119.21}\\
-	04-Nov && 490.77 & \gre{242.58} && \red{769.78} & 895.18 && 178.28 & \gre{119.79}\\
-	05-Nov && 490.42 & \gre{271.44} && 798.20 & \gre{687.87} && 211.58 & \gre{122.75}\\
-	06-Nov && 479.50 & \gre{243.76} && \red{864.52} & 920.79 && 152.97 & \gre{108.11}\\
-	07-Nov && 452.67 & \gre{243.81} && \red{694.07} & 779.27 && 167.70 & \gre{108.08}\\
-	08-Nov && 405.41 & \gre{223.68} && \red{666.17} & 802.96 && 176.48 & \gre{119.98}\\
-	\bottomrule
-\end{tabular}
-\end{table}%
-\end{frame}
-
-\begin{frame}{Comparison NEW: Total Travel Time (minutes)}
-	\scriptsize
-	\begin{table}
-		%\caption{Comparison}
-		\renewcommand{\arraystretch}{1.5}
-		\begin{tabular}{ccrrcrrcrr}\toprule
-			&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
-			\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
-			\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
-			\midrule
 			02-Nov && 372.88 & \gre{232.28} && \red{674.46} & 687.00 && 153.05 & \gre{124.81}\\
@@
 			04-Nov && 384.52 & \gre{242.58} && \red{722.42} & 895.18 && 141.19 & \gre{119.79}\\
-			05-Nov && 389.95 & \gre{271.44} && 745.86 & \gre{687.87} && 168.99 & \gre{122.7}5\\
+			05-Nov && 389.95 & \gre{271.44} && 745.86 & \gre{687.87} && 168.99 & \gre{122.75}\\
 			06-Nov && 378.22 & \gre{243.76} && \red{824.72} & 920.79 && 123.08 & \gre{108.11}\\
@@
 \begin{frame}{Comparison: Total Waiting Time (minutes)}
-\scriptsize
-\begin{table}
-%\caption{Comparison}
-\renewcommand{\arraystretch}{1.5}
-\begin{tabular}{ccrrcrrcrr}\toprule
-	&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
-	\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
-	\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
-	\midrule
-	02-Nov && \red{1595.00} & 1650.18 && 1407.45 & \gre{1009.16} && \red{222.23} & 260.09\\
-	03-Nov && \red{1580.12} & 1665.65 && 1026.38 & \gre{737.81} && 73.95 & \gre{56.53}\\
-	04-Nov && \red{1189.23} & 1273.80 && 775.22 & \gre{559.64} && \red{36.72} & 61.53\\
-	05-Nov && \red{1159.58} & 1203.28 && 1661.80 & \gre{1527.59} && 338.42 & \gre{332.25}\\
-	06-Nov && 1065.50 & \gre{1043.60} && 1505.48 & \gre{1095.16} && 402.03 & \gre{322.17}\\
-	07-Nov && \red{957.33} & 1002.54 && 790.93 & \gre{435.51} && \red{212.30} & 212.84\\
-	08-Nov && \red{1019.59} & 1089.32 && 548.83 & \gre{322.82} && \red{143.52} & 156.89\\
-	\bottomrule
-\end{tabular}
-\end{table}%
-\end{frame}
-
-\begin{frame}{Comparison NEW: Total Waiting Time (minutes)}
 	\scriptsize
@@
 \end{frame}
+
+%----------------------------------- FIGURES AND PLOTS ------------------------------------%
+
+%=================================== HAMPSHIRE 03 NOV ===================================%
+
+\begin{frame} % HAMPSHIRE 03-NOV nNurses = 11, nJobs = 70, nClients = 28
+	\begin{columns}
+		\column{0.75\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\hspace{35mm}
+			\textbf{Hampshire: 03-Nov-2020}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.9\linewidth]{figures/03HampshireAbi}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.9\linewidth]{figures/03HampshireDST}
+		\end{minipage}
+		\column{0.25\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			%
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\scriptsize
+			\textbf{Abicare:}
+			\begin{itemize}
+				\setlength{\itemindent}{-.2in}
+				\aitem Total: 91:00:00
+				\aitem Travel: 06:21:35
+				\aitem Waiting: 28:08:25
+				\aitem Distance: 307.63km
+			\end{itemize}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\scriptsize
+			\textbf{DST:}
+			\begin{itemize}
+				\setlength{\itemindent}{-.2in}
+				\ditem Total: 88:02:41
+				\ditem Travel: 03:47:01
+				\ditem Waiting: 27:45:39
+				\ditem Distance: 153.31km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame} % PLOTS HAMPSHIRE 03-NOV nNurses = 11, nJobs = 70, nClients = 28
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/3_Nov_20_Hampshire_time_info_abiosrm}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/3_Nov_20_Hampshire_workload_abiosrm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/3_Nov_20_Hampshire_time_info_dst}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/3_Nov_20_Hampshire_workload_dst}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+%=================================== HAMPSHIRE 06 NOV ===================================%
+
+\begin{frame} % HAMPSHIRE 06-NOV nNurses = 11, nJobs = 71, nClients = 27
+	\begin{columns}
+		\column{0.75\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\hspace{35mm}
+			\textbf{Hampshire: 06-Nov-2020}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/06HampshireAbi}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/06HampshireDST}
+		\end{minipage}
+		\column{0.25\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			%
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\scriptsize
+			\textbf{Abicare:}
+			\begin{itemize}
+				\setlength{\itemindent}{-.2in}
+				\aitem Total: 81:15:00
+				\aitem Travel: 06:18:13
+				\aitem Waiting: 19:26:47
+				\aitem Distance: 301.39km
+			\end{itemize}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\scriptsize
+			\textbf{DST:}
+			\begin{itemize}
+				\setlength{\itemindent}{-.2in}
+				\ditem Total: 76:57:22
+				\ditem Travel: 04:03:45
+				\ditem Waiting: 17:23:26
+				\ditem Distance: 167.34km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame} % PLOTS HAMPSHIRE 06-NOV nNurses = 11, nJobs = 71, nClients = 27
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Hampshire_time_info_abiosrm}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Hampshire_workload_abiosrm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Hampshire_time_info_dst}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Hampshire_workload_dst}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+%=================================== MONMOUTH 02 NOV ===================================%
+
+\begin{frame} % MONMOUTH 02-NOV nNurses = 13, nJobs = 102, nClients = 37
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\hspace{35mm}
+			\textbf{Monmouth:}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/02MonmouthAbi}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{Abicare:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\aitem Total: 104:45:00
+				\aitem Travel: 11:14:28
+				\aitem Waiting: 24:15:32
+				\aitem Distance: 497.03km
+			\end{itemize}
+		\vspace{10mm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\textbf{02-Nov-2020}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/02MonmouthDST}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{DST:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\ditem Total: 97:31:10
+				\ditem Travel: 11:27:00
+				\ditem Waiting: 16:49:10
+				\ditem Distance: 569.09km
+			\end{itemize}
+			\vspace{10mm}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame} % PLOTS MONMOUTH 02-NOV nNurses = 13, nJobs = 102, nClients = 37
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/2_Nov_20_Monmouth_time_info_abiosrm}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/2_Nov_20_Monmouth_workload_abiosrm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/2_Nov_20_Monmouth_time_info_dst}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/2_Nov_20_Monmouth_workload_dst}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+%=================================== MONMOUTH 05 NOV ===================================%
+
+\begin{frame} % MONMOUTH 05-NOV nNurses = 16, nJobs = 116, nClients = 40
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\hspace{35mm}
+			\textbf{Monmouth:}
+		\end{minipage}
+		\begin{minipage}[c][0.65\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/05MonmouthAbi}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{Abicare:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\aitem Total: 120:00:00
+				\aitem Travel: 12:25:52
+				\aitem Waiting: 28:34:08
+				\aitem Distance: 561.90km
+			\end{itemize}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\textbf{05-Nov-2020}
+		\end{minipage}
+		\begin{minipage}[c][0.65\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/05MonmouthDST}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{DST:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\ditem Total: 115:55:27
+				\ditem Travel: 11:27:52
+				\ditem Waiting: 25:27:35
+				\ditem Distance: 578.20km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame} % PLOTS MONMOUTH 05-NOV nNurses = 16, nJobs = 116, nClients = 40
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Monmouth_time_info_abiosrm}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Monmouth_workload_abiosrm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Monmouth_time_info_dst}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Monmouth_workload_dst}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+%=================================== ALDERSHOT 05 NOV ===================================%
+
+\begin{frame} % ALDERSHOT 05-NOV nNurses = 5, nJobs = 30, nClients = 16
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\hspace{37mm}
+			\textbf{Aldershot:}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.8\linewidth]{figures/05AldershotAbi2}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{Abicare:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\aitem Total: 33:20:00
+				\aitem Travel: 02:48:59
+				\aitem Waiting: 06:21:01
+				\aitem Distance: 129.57km
+			\end{itemize}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\textbf{05-Nov-2020}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.8\linewidth]{figures/05AldershotDST2}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{DST:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\ditem Total: 31:45:00
+				\ditem Travel: 02:02:25
+				\ditem Waiting: 05:32:15
+				\ditem Distance: 92.00km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame} % PLOTS ALDERSHOT 05-NOV nNurses = 5, nJobs = 30, nClients = 16
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Aldershot_time_info_abiosrm}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Aldershot_workload_abiosrm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Aldershot_time_info_dst}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/5_Nov_20_Aldershot_workload_dst}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+%=================================== ALDERSHOT 06 NOV ===================================%
+
+\begin{frame} % ALDERSHOT 06-NOV nNurses = 5, nJobs = 27, nClients = 15
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\hspace{37mm}
+			\textbf{Aldershot:}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/06AldershotAbi}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{Abicare:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\aitem Total: 24:30:00
+				\aitem Travel: 02:03:05
+				\aitem Waiting: 07:11:55
+				\aitem Distance: 98.11km
+			\end{itemize}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\textbf{06-Nov-2020}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/06AldershotDST}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{DST:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\ditem Total: 22:25:16
+				\ditem Travel: 01:48:06
+				\ditem Waiting: 05:22:10
+				\ditem Distance: 83.70km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame} % ALDERSHOT 06-NOV nNurses = 5, nJobs = 27, nClients = 15
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Aldershot_time_info_abiosrm}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Aldershot_workload_abiosrm}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Aldershot_time_info_dst}
+		\end{minipage}
+		\begin{minipage}[c][0.45\textheight][c]{\linewidth}
+			\vspace{5mm}
+			\centering
+			\includegraphics[width=1\linewidth]{figures/6_Nov_20_Aldershot_workload_dst}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+%==========================================================================================
+
+\begin{frame}{Issues}
+	\begin{itemize}
+		\aitem Missing miles/estimated time for some tasks
+		\aitem No travel data (miles/estimated time) between carer's home and first/last tasks
+		\aitem Some clients have tasks back-to-back that are given separately - why?
+		\aitem All other areas cannot be studied due to missing postcodes
+	\end{itemize}
+\end{frame}
+
+%==========================================================================================
 
@@
 	\end{figure}
+\end{frame}
+\end{comment}
+
+\begin{comment}
+\begin{frame}{Comparison OLD: Total Travel Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+\midrule
+02-Nov && 475.00 & \gre{232.28} && 722.55 & \gre{687.00} && 192.77 & \gre{124.81}\\
+03-Nov && 489.88 & \gre{227.02} && \red{728.62} & 745.88 && 166.05 & \gre{119.21}\\
+04-Nov && 490.77 & \gre{242.58} && \red{769.78} & 895.18 && 178.28 & \gre{119.79}\\
+05-Nov && 490.42 & \gre{271.44} && 798.20 & \gre{687.87} && 211.58 & \gre{122.75}\\
+06-Nov && 479.50 & \gre{243.76} && \red{864.52} & 920.79 && 152.97 & \gre{108.11}\\
+07-Nov && 452.67 & \gre{243.81} && \red{694.07} & 779.27 && 167.70 & \gre{108.08}\\
+08-Nov && 405.41 & \gre{223.68} && \red{666.17} & 802.96 && 176.48 & \gre{119.98}\\
+\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{frame}{Comparison OLD: Total Waiting Time (minutes)}
+\scriptsize
+\begin{table}
+%\caption{Comparison}
+\renewcommand{\arraystretch}{1.5}
+\begin{tabular}{ccrrcrrcrr}\toprule
+&& \multicolumn{2}{c}{Hampshire} && \multicolumn{2}{c}{Monmouth} && \multicolumn{2}{c}{Aldershot} \\
+\cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10}
+\multicolumn{1}{c}{Date} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} && \multicolumn{1}{c}{Abicare} & \multicolumn{1}{c}{DST} \\
+\midrule
+02-Nov && \red{1595.00} & 1650.18 && 1407.45 & \gre{1009.16} && \red{222.23} & 260.09\\
+03-Nov && \red{1580.12} & 1665.65 && 1026.38 & \gre{737.81} && 73.95 & \gre{56.53}\\
+04-Nov && \red{1189.23} & 1273.80 && 775.22 & \gre{559.64} && \red{36.72} & 61.53\\
+05-Nov && \red{1159.58} & 1203.28 && 1661.80 & \gre{1527.59} && 338.42 & \gre{332.25}\\
+06-Nov && 1065.50 & \gre{1043.60} && 1505.48 & \gre{1095.16} && 402.03 & \gre{322.17}\\
+07-Nov && \red{957.33} & 1002.54 && 790.93 & \gre{435.51} && \red{212.30} & 212.84\\
+08-Nov && \red{1019.59} & 1089.32 && 548.83 & \gre{322.82} && \red{143.52} & 156.89\\
+\bottomrule
+\end{tabular}
+\end{table}%
+\end{frame}
+
+\begin{figure}
+\centering
+\begin{subfigure}[h]{0.67\textwidth}
+\vspace{3mm}
+\includegraphics[width=1\linewidth]{figures/03HampshireAbi}
+\caption{Hampshire 03-Nov: Abicare}
+\label{fig:03hampabi} 
+\end{subfigure}
+%\vspace{-10mm}
+\begin{subfigure}[h]{0.67\textwidth}
+\includegraphics[width=1\linewidth]{figures/03HampshireDST}
+\caption{Hampshire 03-Nov: DST}
+\label{fig:03hampdst}
+\end{subfigure}
+\end{figure}
+
+\begin{frame}
+\begin{columns}
+\column{0.75\textwidth}
+\begin{minipage}[c][0.47\textheight][c]{\linewidth}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/03HampshireAbi}
+\end{minipage}
+\begin{minipage}[c][0.47\textheight][c]{\linewidth}
+\centering
+\includegraphics[width=0.9\linewidth]{figures/03HampshireDST}
+\end{minipage}
+\column{0.25\textwidth}
+\begin{minipage}[c][0.47\textheight][c]{\linewidth}
+\scriptsize
+Hampshire 03Nov: Abicare
+\begin{itemize}
+\setlength{\itemindent}{-.2in}
+\item Total: 91:00:00
+\item Travel: 08:09:53
+\item Waiting: 26:20:07
+\end{itemize}
+\end{minipage}
+\begin{minipage}[c][0.47\textheight][c]{\linewidth}
+\scriptsize
+Hampshire 03-Nov: DST
+\begin{itemize}
+\setlength{\itemindent}{-.2in}
+\item Total: 88:02:41
+\item Travel: 03:47:01
+\item Waiting: 27:45:39
+\end{itemize}
+\end{minipage}
+\end{columns}
 \end{frame}

### 2021-04-20 code/screpo/dst_solver.py

@@ -1,0 +1,275 @@
+import ctypes
+import numpy as np
+from numpy.ctypeslib import ndpointer
+import matplotlib.pyplot as plt
+import copy
+import time
+# import random
+import os
+
+from joblib import Parallel, delayed
+
+# import tkinter
+# from tkinter import messagebox
+# import tkMessageBox # Python 2.x version
+
+# Own modules:
+import mankowska_data as Mk
+import python_resources as pr
+import instance_handler as hhc
+
+# NOTE: THIS NEEDS TO BE CHANGED TO BE SIMILAR TO CONSTRUCTIVE_WRAPPER.PY
+
+def solver(opt_obj, instance_file, options_vector=[]):
+    # Solving parameters:
+    
+    quality_measure = opt_obj.quality_measure
+    ds_skill_type = opt_obj.ds_skill_type
+    max_time_seconds = opt_obj.max_time_seconds
+    verbose_level = opt_obj.verbose_level
+    create_python_plots = opt_obj.create_python_plots
+    create_html_website = opt_obj.create_html_website
+    results_file_name = opt_obj.results_file_name
+    run_in_parallel = opt_obj.run_in_parallel
+    parallel_workers = opt_obj.parallel_workers
+    random_seed = opt_obj.random_seed
+    file_idx = opt_obj.file_idx
+    # Set the default options:
+
+    if len(options_vector) < 1:
+        print('ERROR!')
+        # exit(-32423)
+
+
+
+    # For our own format, load driving matrix from disk. If not, it is saved after the first execution.
+    # File name should be: self.name + '_' + 'driving' + '_times.npy',
+    # where self.name = instance_file.split('.')[0]
+    load_from_disk = opt_obj.load_from_disk
+
+    # Choosing file to run. All files in file_list will be run in a loop.
+    # Need to set instance_type, directory and a file list (even if only with one file)
+
+    instance_type = opt_obj.instance_type
+    mankowska_vns = False
+    if instance_type == 'mankowska-vns':
+        instance_type = 'mankowska'
+        mankowska_vns = True
+    elif instance_type == 'mankowska-mat':
+        instance_type = 'mankowska'
+        mankowska_vns = True
+    
+    ###########################################################################################
+    ###########################################################################################
+    ###########################################################################################
[…]

### 2021-04-20 code/screpo/test_request.py

@@ -1,25 +1,51 @@
 import requests
 import pandas as pd
-# server = r'localhost:5000'
+server = r'localhost:5000'
 # coord1 = r'-1.813821,51.077959'
 # coord2 = r'-1.447691138804643,50.96380612421809'
 # coord1 = r'-1.1265538656195921,51.25167783293762'
 # coord2 = r'-1.0542636173681512,51.29435018372664'
+coord1 = r'-1.0932713683349489,51.26244013391559'
+coord2 = r'-1.1223542469709538,51.2517453438589'
+coord3 = r'-1.1108495601022277,51.247216458376776'
+coord4 = r'-1.11904,51.27989'
+coord5 = r'-1.04517,51.26668'
 # str_call = 'http://' + server + '/route/v1/driving/' + coord1 + ';' + coord2 + ';' + coord1 + '?overview=full&geometries=geojson'
-str_call = 'http://localhost:5000/route/v1/driving/-1.0932713683349489,51.26244013391559;-1.1223542469709538,51.2517453438589;-1.1108495601022277,51.247216458376776;-1.1066389846881866,51.27818905207848;-1.1195947760811282,51.27919535669177;-1.0993976532712253,51.262109743031516;-1.0932713683349489,51.26244013391559?overview=full&geometries=geojson'
-# str_call = 'http://' + server + '/table/v1/driving/' + coord1 + ';' + coord2
+# str_call = 'http://' + server + '/route/v1/driving/' + coord1 + ';' + coord3 + '?overview=full&geometries=geojson'
+# str_call = 'http://localhost:5000/route/v1/driving/-1.0932713683349489,51.26244013391559;-1.1223542469709538,51.2517453438589;-1.1108495601022277,51.247216458376776;-1.1066389846881866,51.27818905207848;-1.1195947760811282,51.27919535669177;-1.0993976532712253,51.262109743031516;-1.0932713683349489,51.26244013391559?overview=full&geometries=geojson'
+# str_call = 'http://localhost:5000/table/v1/driving/-1.0932713683349489,51.26244013391559;-1.1223542469709538,51.2517453438589;-1.1108495601022277,51.247216458376776;-1.11904,51.27989;-1.04517,51.26668?annotations=distance,duration'
[…]

### 2021-04-20 code/screpo/tools_and_scripts/abicare_solution_quality.py

@@ -15,17 +15,18 @@ from scipy import stats
 # Display all rows and columns of dataframes in command prompt:
 pd.set_option("display.max_rows", None, "display.max_columns", None)
 
-# Instance format from analyse_mileage.py (in this file, this is referred to as 'idict')
+# Instance format from fc_analyse_mileage.py (in this file, this is referred to as 'idict')
 # inst = {
-# 'name' : a_name + '_' + day.replace('-', '_'),
-# 'area' : a_name,
-# 'date' : day,
-# 'stats' : {'ncarers' : 0, 'ntasks' : 0, 'ttraveltime' : 0, 'ttravelmiles' : 0, 'tservicetime' : 0, 'tgaptime' : 0, 'twaitingtime' : 0, 'totaltime' : 0},
-# 'txtsummary' : '',
-# 'rota' : {'carer' : [], 'postcode' : [], 'num_addresses' : [], 'eastings': [], 'northings' : [], 'start' : [], 'finish' : [], 'shift' : [], 'home_start' : [], 'home_finish' : [], 'num_tasks' : [], 'travel_time' : [], 'wait_time' : [], 'service_time' : []},
-# 'tasks' : {'client' : [], 'postcode': [], 'num_addresses': [], 'eastings': [], 'northings' : [], 'duration' : [], 'miles' : [], 'metres' : [], 'esttime' : [], 'start' : [], 'end' : [], 'tw_start' : [], 'tw_end' : []},
-# 'routes' : []
-# }
+#         'name' : a_name + '_' + day.replace('-', '_'),
+#         'fname' : day.replace('-', '_') + '_' + a_name, # filename
+#         'area' : a_name,
+#         'date' : day,
+#         'stats' : {'ncarers' : 0, 'ntasks' : 0, 'ttraveltime' : 0, 'ttravelmiles' : 0, 'tservicetime' : 0, 'tgaptime' : 0, 'twaitingtime' : 0, 'totaltime' : 0},
+#         'txtsummary' : '',
+#         'rota' : {'carer' : [], 'postcode' : [], 'num_addr' : [], 'eastings': [], 'northings' : [], 'start' : [], 'finish' : [], 'shift' : [], 'home_start' : [], 'home_finish' : [], 'num_tasks' : [], 'travel_time' : [], 'wait_time' : [], 'service_time' : []},
[…]

### 2021-04-20 code/screpo/tools_and_scripts/class_cpo_df.py

@@ -17,7 +17,7 @@ import pandas as pd
 import geopandas
 
 class CPO_DF(): # Codepoint Open DataFrame
-    def __init__(self, foldername=r'data\codepo_gb'):
+    def __init__(self, foldername=r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\codepo_gb'):
     # def __init__(self, foldername=r'..\data\codepo_gb'):
         self.foldername = foldername # CP_Open_Folder = r'..\data\codepo_gb'
         all_cpopen_csvs = glob.glob(os.path.join(self.foldername, 'data\\CSV', '*.csv'))

### 2021-04-20 code/screpo/tools_and_scripts/class_cpo_df1.py

@@ -18,7 +18,7 @@ import geopandas
 
 class CPO_DF1(): # Codepoint Open DataFrame
     # def __init__(self, foldername=r'data\codepo_gb'):
-    def __init__(self, foldername=r'..\data\codepo_gb'):
+    def __init__(self, foldername=r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\codepo_gb'):
         self.foldername = foldername # CP_Open_Folder = r'..\data\codepo_gb'
         all_cpopen_csvs = glob.glob(os.path.join(self.foldername, 'data\\CSV', '*.csv'))
 

### 2021-04-20 code/screpo/tools_and_scripts/compare_abicare_dst.py

@@ -46,9 +46,9 @@ def create_df(filename, a_name=''):
             'nJobs' : 0,
             'nClients' : 0,
             'stats' : {'nurses' : [], 'jobs' : [], 'clients' : [], 'servicetime' : []},
-            'abi' : {'ttime' : [], 'ttravel' : [], 'ttravelosrm' : [], 'twaiting' : [], 'twaitingosrm' : [], 'dist' : [], 'distjobs' : [], 'distjobsosrm' : [], 'aith': [], 'mk' : [], 'wb' : [], 'paper' : []},
+            'abi' : {'ttime' : [], 'ttravel' : [], 'ttravelosrm' : [], 'twaiting' : [], 'twaitingosrm' : [], 'dist' : [], 'distjobs' : [], 'distjobsosrm' : [], 'aith': [], 'mk' : [], 'wb' : [], 'paper' : [], 'aithosrm': [], 'mkosrm' : [], 'wbosrm' : [], 'paperosrm' : []},
             'dst' : {'ttime' : [], 'ttravel' : [], 'twaiting' : [], 'dist' : [], 'distjobs' : [], 'aith': [], 'mk' : [], 'wb' : [], 'paper' : [], 'algtime' : [], 'iterations' : []},
-            'best' : {'ttime' : [], 'ttravel' : [], 'ttravelosrm' : [], 'twaiting' : [], 'twaitingosrm' : [], 'dist' : [], 'distjobs' : [], 'distjobsosrm' : [], 'aith' : [], 'mk' : [], 'wb' : [], 'paper' : []}
+            'best' : {'ttime' : [], 'ttravel' : [], 'ttravelosrm' : [], 'twaiting' : [], 'twaitingosrm' : [], 'dist' : [], 'distjobs' : [], 'distjobsosrm' : [], 'aith' : [], 'mk' : [], 'wb' : [], 'paper' : [], 'aithosrm': [], 'mkosrm' : [], 'wbosrm' : [], 'paperosrm' : []}
             }
 
     inst['nNurses'] = df.iloc[0]['nNurses']
@@ -82,6 +82,11 @@ def create_df(filename, a_name=''):
         inst['abi']['wb'].append(day_df.iloc[0]['WB Abi'])
         inst['abi']['paper'].append(day_df.iloc[0]['Paper Abi'])
 
+        inst['abi']['aithosrm'].append(day_df.iloc[0]['AitH AbiOsrm'])
[…]

### 2021-04-20 code/screpo/tools_and_scripts/compare_abicare_osrm.py

@@ -25,7 +25,7 @@ import matplotlib.pyplot as plt
 from scipy import stats
 
 # Import python file in project
-import class_cpo_df1 as ccd
+import class_cpo_df as ccd
 
 # Display all rows and columns of dataframes in command prompt:
 pd.set_option("display.max_rows", None, "display.max_columns", None)
@@ -35,7 +35,7 @@ def compare_all_inst_time_dist(filename, timeofday='all'):
     area_name = all_instances[0]['area']
     # print(a_name)
     # exit(-1)
-    cpo_inst = ccd.CPO_DF1()
+    cpo_inst = ccd.CPO_DF()
 
     abitime_all = []
     abidist_all = []
@@ -557,6 +557,8 @@ def calc_abicare_quality(idict):
 
     travel_time = idict['stats']['ttraveltime']
     waiting_time = idict['stats']['twaitingtime']
+    travel_timeosrm = idict['stats']['ttraveltimeosrm']
+    waiting_timeosrm = idict['stats']['twaitingtimeosrm']
     # print('travel_time ', idict['stats']['ttraveltime'])
     
     # Determine the shortest and longest shifts of all the nurses (in minutes)
@@ -571,53 +573,71 @@ def calc_abicare_quality(idict):
 
     # Ait H. quality:
     # ait_quality = -1*(0.3*totalTravel + ip->totPref);
-    ait_quality = -1*((0.3*travel_time) + 0)
+    ait_qualityabi = -1*((0.3*travel_time) + 0)
+    ait_qualityosrm = -1*((0.3*travel_timeosrm) + 0)
     # print('ait_quality: ', ait_quality)
 
     # Mankowska quality:
     # mk_quality = -1*(totalTravel + mk_allowed_tardiness + mk_max_tardiness)/3;
-    mk_quality = -1*(travel_time + 0 + 0)/3
+    mk_qualityabi = -1*(travel_time + 0 + 0)/3
+    mk_qualityosrm = -1*(travel_timeosrm + 0 + 0)/3
     # print('mk_quality: ', mk_quality)
 
     # Workload Balance quality:
     # wb_quality = -1*(0.3*totalTravel + ip->totPref) - 0.1*dayDiff + 0.1*minSpareTime;
-    wb_quality = -1*((0.3*travel_time) + 0) - (0.1*day_diff) + (0.1*0)
+    wb_qualityabi = -1*((0.3*travel_time) + 0) - (0.1*day_diff) + (0.1*0)
[…]

### 2021-04-20 code/screpo/tools_and_scripts/cpofc_analyse_mileage.py

@@ -1,0 +1,476 @@
+#--------------------#
+# cpofc_analyse_mileage.py
+# 08/01/2021
+# 
+#--------------------#
+
+# Import modules
+import os
+import sys
+import glob
+import math
+import folium
+import pickle
+import datetime
+import requests # For website
+import geopandas
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+
+# Import python file in project
+import class_cpo_df as ccd
+
+# cpofc_analyse_mileage.py: codepoint open and full codepoint analyse mileage.
+
+class POSTCODE_FINDER():
+    """
+    docstring
+    """
+    df = []
+    filename = []
+    def __init__(self, filename=r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\postcode\full_codepoint.csv'):
+        self.filename = filename
+        self.df = pd.read_csv(self.filename)
+        self.df['PC'] = self.df['PC'].str.replace('"', '')
+        self.df['PC'] = self.df['PC'].str.replace(' ', '') 
+        self.df['PC'] = self.df['PC'].str.lower()
+
+    def find_postcode_eastnorth(self, postcode): # Returns eastings and northings of given postcode
+        dfview = self.df[self.df['PC'] == postcode]
+        if dfview.empty:
+            return [None, None]
+        else:
+            # Need to convert eastings and northings to lat and lon
+            return [dfview.iloc[0]['EA'], dfview.iloc[0]['NO']]
+    
+    def find_postcode_n_addresses(self, postcode): # Returns number of registered properties at a given postcode.
+        dfview = self.df[self.df['PC'] == postcode]
+        if dfview.empty:
+            return -1
+        else:
+            return dfview.iloc[0]['RP']
+### --- End class POSTCODE_FINDER --- ###
+
+def time_dif_to_minutes(endtime, starttime, round=True):
+    mins = (endtime - starttime).seconds/60
+    if round:
+        return np.round(mins, 0)
+    else:
+        return mins
+### --- End time_dif_to_minutes --- ###
+
+def inst_dicts_to_dfs(inst):
+    inst['rota'] = pd.DataFrame(inst['rota'])
+    inst['tasks'] = pd.DataFrame(inst['tasks'])
+    return inst
+### --- End inst_dicts_to_dfs --- ###
+
+def miles_to_metres(distmiles):
[…]

### 2021-05-25 tex/Pres180521/main.tex
new presentation for abicare using new dataset.

--- 2021-04-20 tex/Pres/main.tex
@@
 
+\begin{frame}{Queries}
+	\scriptsize
+	%\begin{multicols}{2}
+	\begin{itemize}
+		\item Size of the dataset provided
+		\item Incomplete/incorrect postcodes (e.g. CU14 9DD, NP13)
+		\item Travel to/from carer's homes to be included?
+		\item Carers' locations between shifts (return home or elsewhere)
+		\item Areas have multiple locations (e.g. AndoverSalisbury, AldershotHampshire)
+		\item Clients' names in Carer Availability dataset
+		\item Missing details for some clients/carers
+		\item Duration of some tasks is 0 minutes
+	\end{itemize}	
+	%\end{multicols}
+\end{frame}
+
+\begin{frame}{Clients' Names in Carer Availability Dataset}
+	\begin{figure}
+		\includegraphics[width=1\linewidth]{figures/ClientNames}%
+	\end{figure}
+\end{frame}
+
+\begin{frame}{Incorrect Postcodes and Multiple Locations}
+\begin{columns}
+	\column{0.5\textwidth}
+	\begin{minipage}[c][1\textheight][c]{\linewidth}
+		\includegraphics[width=1\linewidth]{figures/IncorrectPostcodes}
+	\end{minipage}
+	\column{0.5\textwidth}
+	\begin{minipage}[c][1\textheight][c]{\linewidth}
+		\includegraphics[width=0.8\linewidth]{figures/MultipleAreas}
+	\end{minipage}
+\end{columns}
+\end{frame}
+
+\begin{frame}{Comparison: Salisbury vs Hampshire}
+	\small
+	\begin{multicols}{2}	
+		\begin{block}{Salisbury}
+			\begin{itemize}
+%				\setlength{\itemindent}{-.1in}
+				\item Carers = 35
+				\item Shifts = 70
+				\item Jobs = 14
+			\end{itemize}
+		\end{block}
+		\begin{block}{Hampshire}
+			\begin{itemize}
+%				\setlength{\itemindent}{-.1in}
+				\item Carers = 21
+				\item Shifts = 24
+				\item Jobs = 68
+			\end{itemize}
+		\end{block}
+	\end{multicols}	
+\end{frame}
+
+% Salisbury original vs including nursesto/from home. nNurses = 35, nJobs = 14, nShifts = 70
+\begin{frame} 
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\textbf{Salisbury}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.7\linewidth]{figures/ogSalisbury}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{Without home travel:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\setlength\itemsep{-0.2em}
+				\aitem Total: 38:24:18
+				\aitem Travel: 01:28:22
+				\aitem Waiting: 01:56:56
+				\aitem Distance: 246.53km
+			\end{itemize}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.7\linewidth]{figures/nurseSalisbury}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{With home travel:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\setlength\itemsep{-0.2em}
+				\ditem Total: 39:14:03
+				\ditem Travel: 02:01:38
+				\ditem Waiting: 02:13:25
+				\ditem Distance: 171.67km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Comparison: Salisbury with and without carer home travel}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/og_inst_Salisbury_2021-03-01_time_info_dst}%
+		\includegraphics[width=0.5\linewidth]{figures/nurse_inst_Salisbury_2021-03-01_time_info_dst}
+	\end{figure}
+\end{frame}
+
+% Hampshire original vs including nursesto/from home. nNurses = 21, nJobs = 68, nShifts = 24
+\begin{frame} 
+	\begin{columns}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+			\textbf{Hampshire}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.9\linewidth]{figures/ogHampshire}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{Without home travel:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\setlength\itemsep{-0.2em}
+				\aitem Total: 50:02:53
+				\aitem Travel: 05:30:20
+				\aitem Waiting: 01:47:33
+				\aitem Distance: 345.15km
+			\end{itemize}
+		\end{minipage}
+		\column{0.5\textwidth}
+		\begin{minipage}[c][0.05\textheight][c]{\linewidth}
+		\end{minipage}
+		\begin{minipage}[c][0.7\textheight][c]{\linewidth}
+			\centering
+			\includegraphics[width=0.9\linewidth]{figures/nurseHampshire}
+		\end{minipage}
+		\begin{minipage}[c][0.2\textheight][c]{\linewidth}
+			\scriptsize
+			\hspace{17mm}\textbf{With home travel:}
+			\begin{itemize}
+				\setlength{\itemindent}{0.5in}
+				\setlength\itemsep{-0.2em}
+				\ditem Total: 53:07:04
+				\ditem Travel: 06:02:27
+				\ditem Waiting: 04:19:37
+				\ditem Distance: 363.78km
+			\end{itemize}
+		\end{minipage}
+	\end{columns}
+\end{frame}
+
+\begin{frame}{Comparison: Hampshire with and without carer home travel}
+	\begin{figure}
+		\includegraphics[width=0.5\linewidth]{figures/og_inst_Hampshire_2021-03-01_time_info_dst}%
+		\includegraphics[width=0.5\linewidth]{figures/nurse_inst_Hampshire_2021-03-01_time_info_dst}
+	\end{figure}
+\end{frame}
+
+\begin{comment}
 \begin{frame}{Task}
@@
 
-\begin{comment}
 \begin{frame}{Queries}
@@
 \end{frame}
-\end{comment}
 
@@
 \end{frame}
-
+\end{comment}
+
+\begin{comment}
 \begin{frame}{Comparison: Hampshire Time and Distance}
@@
 %==========================================================================================
-
+\end{comment}
 \begin{comment}

### 2021-06-28 tex/Documentation/Documentation.tex
Documentation for python program abicare.

--- 2020-11-17 tex/LitReview/LitReview.tex
@@
 \begin{document}
-\title{Increasing Productivity in the Domiciliary Care Sector Through Automated Routing and Scheduling of Carers}
+\title{Documentation: Abicare DST}
 %\author{Asyl L. Hawa}
@@
 
-\noindent \citet{yuan2017}: Disruption Management for the Real-Time Home Caregiver Scheduling and Routing Problem
+\section{Disruption Management for the RT-HCSRP}
 
-\section{Differences}
-\begin{itemize}[leftmargin=*, itemsep=-0.1em]
-	\item Paper covers updating an existing schedule when an disruption occurs.
-The disruption in this case is when a new service is added to the schedule.
-However, in our case there could be a variety of reasons why a disruption may occur, such as a delay in a service or carer illness.
-	\item This paper does \emph{not} consider double services or synchronised (dependent) services.
-	\item The paper assumes that all carers start and end their routes at the Home Health Care Centre, rather than their own homes.
-	\item The paper also states that each patient is visited by a carer exactly once, which is not always the case in our problem, as a patient may require multiple visits in a single day.
+\subsection{Differences}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
 	\item 
@@
 
-\bibliographystyle{plainnat}
-\bibliography{includes/bibliography}
+\subsection{Model and Solution Approach}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item One
+	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
+		\item 
+	\end{enumerate}
+	\item Two
+	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
+		\item
+	\end{enumerate}
+	\item 
+\end{itemize}
+
+\subsection{Modifications}
+\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item 
+\end{itemize}
+
+\section{Disruption Management of the VRP with Breakdown}
+
+\subsection{Differences}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item 
+\end{itemize}
+
+\subsection{Model and Solution Approach}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item 
+\end{itemize}
+
+\subsection{Modifications}
+\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
+	\item 
+\end{itemize}
 

### 2021-07-07 tex/Documentation/Documentation.tex
updated documentation.

@@
 \input{includes/preamble.tex}
+\usepackage{parskip}
 
 \begin{document}
-\title{Documentation: Abicare DST}
+\title{\vspace{-20mm}Documentation: Abicare DST\vspace{-15mm}}
 %\author{Asyl L. Hawa}
@@
 
-\section{Disruption Management for the RT-HCSRP}
-
-\subsection{Differences}
-\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
+\section{About the Program}
+\begin{itemize}[leftmargin=*]
+	\item Takes in an input file containing details for carers and clients availabilities and locations, and uses the information to produce a schedule and route for each carer to visit the clients.
+	\item The algorithm used to determine a solution is a Greedy Randomised Adaptive Search Procedure (GRASP) combined with a Variable Neighbourhood Search (VNS) method, which is implemented in C.
+The pre- and post-processing of information is implemented in Python.
+	\item This program is designed for use in social care to improve schedules/routes of carers to clients.
 \end{itemize}
 
-\subsection{Model and Solution Approach}
-\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item One
-	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
-		\item 
-	\end{enumerate}
-	\item Two
-	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
-		\item
-	\end{enumerate}
-	\item 
+\section{Requirements}
+\subsection{Code-Point Open}
+The program uses Code-Point Open to obtain location data for carers and clients.
+Code-Point Open is available to download for free at https://www.ordnancesurvey.co.uk/business-government/products/code-point-open.
+The data format to download should be a CSV.
+Once the download is complete, extract the zip file and save the resulting \texttt{codepo\_gb} folder into the same directory as the program file.
+\begin{itemize}[leftmargin=*]
+	\item Code-Point Open is used to obtain exact locations from postcodes in the form of coordinates (latitude and longitude) from a free database provided by Ordnance Survey.
+	\item Code-Point Open is a necessary requirement for the execution of the program.
 \end{itemize}
 
-\subsection{Modifications}
-\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
+\subsection{OSRM}
+\begin{itemize}[leftmargin=*]
+	\item The Open Source Routing Machine (OSRM) is an open-source router designed for use with data from the OpenStreetMap project.
+	\item OSRM is used to obtain distances and travel times between coordinates.
+	\item OSRM is a necessary requirement for the execution of the program.
 \end{itemize}
 
-\section{Disruption Management of the VRP with Breakdown}
-
-\subsection{Differences}
-\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
+\section{User Input Variables}
+When the program is executed, a prompt will appear in which user input values can be provided for a number of parameters:
+\begin{itemize}[leftmargin=*]
+	\item Area name: the full name of the area to be scheduled and routed (default: Hampshire)
+	\item Time window interval: the number of minutes before and after the expected start time of a job in which a carer can arrive and start the job (default: 15)
+	\item Workload balance: a coefficient used to balance the workload for each carer (default: 1)
+	\item Quality measure: the measure used to calculate the objective value (default: `Default')
+	\item Time limit: the desired length of time (in seconds) for the algorithm to run (default: 60)
+	\item Create website: the creation of an html website which displays the solution on a map (default: `Yes')
+	\item Create plots: the creation of two plots showing the time information as a whole and the workload distribution for each carer (default: `Yes')
+	\item Select file: the input file containing the carer and client details (.xlsx)
 \end{itemize}
 
-\subsection{Model and Solution Approach}
-\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
-\end{itemize}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\linewidth]{figures/userinputscreenshot}%
+\end{figure}
 
-\subsection{Modifications}
-\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
+\section{Quality Measure Objective Functions}
+The four quality measures, `Default', `Ait H', `Mankowska', and `Workload Balance', have different objective functions, calculated as follows:
+\subsection{Default}
+\begin{equation}
+	\begin{split}
+		\text{quality} &= (-1\times \text{total travel time}) + (-1 \times \text{total waiting time}) \\
+		&+ (-5 \times \text{allowed tardiness}) + (-5 \times \text{total overtime}) \\
+		&+ (\text{wb coefficient} \times \text{min spare time}) + \text{total preference score}
+	\end{split}
+\end{equation}
+\noindent where the variable `wb coefficient' is the Workload Balance value provided by the user.
+
+\subsection{Ait Haddadene}
+\begin{equation}
+	\text{quality} = -1\times ((0.3 \times \text{total travel time}) + \text{total preference score}) \\
+\end{equation}
+
+\subsection{Mankowska}
+\begin{equation}
+	\text{quality} = -1\times \frac{\text{total travel time} + \text{allowed tardiness} + \text{max tardiness}}{3} \\
+\end{equation}
+
+\subsection{Workload Balance}
+\begin{equation}
+	\begin{split}
+		\text{quality} &= -1 \times ((0.3 \times \text{total travel time}) + \text{total preference score}) \\
+		& + (-0.1 \times \text{length of day}) + (0.1 \times \text{min spare time})
+	\end{split}
+\end{equation}
+\noindent where the variable `length of day' is the difference between the longest and shortest days in the solution.
+
+\section{Output Files}
+The program produces five output files: a text file, a CSV file, two image files containing plots, and a web file.
+\subsection{Results File}
+The results file is a text file (.txt) named as \texttt{area\_date\_results.txt}, where the area and date correspond to the area provided in the user input and the date for the schedule respectively.
+
+The results file contains the following information:
+\begin{enumerate}
+	\item Date: the current date and time
+	\item Quality: the final quality of the schedule and route produced by the program
+	\item Measure: the quality measure used to calculate the quality of the solution
+	\item Carers: the number of carers in the schedule
+	\item Jobs: the number of jobs in the schedule
+	\item Total time: the total amount of time for all carers in the schedule, including service time, travel time (excluding travel to and from carers' homes), waiting time, and overtime
+	\item Total service time: the total amount of time for all jobs
+	\item Total travel time: the total amount of time spent travelling for all carers (excluding travel to and from carers' homes)
+	\item Total waiting time: the total amount of waiting time for all carers
+	\item Total tardiness: the total amount of tardiness for all jobs
+	\item Total overtime: the total amount of overtime for all carers
+	\item Total mileage: the total distance travelled by all carers (excluding travel to and from carers' homes)
+	\item Total cost: the total cost incurred by all carers as a combination of travel costs and mileage costs (excluding travel to and from carers' homes)
+	\item Elapsed time: the total running time of the program
+\end{enumerate}
+
+\subsection{Solution File}
+The solution file is a CSV file (.csv) named as \texttt{area\_date\_solution.csv}, where the area and date correspond to the area provided in the user input and the date for the schedule respectively.
+
+The solution file contains the initial information provided for the clients, along with the following information for each job:
+\begin{enumerate}
+	\item carer\_id: the carer that is assigned to the job
+	\item arrive\_job: the time at which the carer arrives at the job location
+	\item start\_job: the time at which the carer starts the job
+	\item depart\_job: the time at which the carer leaves the job location
+	\item travel\_time: the time taken for the carer to travel to the job location from the previous job location (no travel time is provided for travel to the first job in a carer's route)
+	\item waiting\_time: the amount of time between the arrival of the carer at the job location and the start of the job
+	\item tardiness: the amount of time between the end of the job's time window and the start of the job 
+\end{enumerate}
+
+\subsection{Image Files}
+There are two image files (.png) containing plots produced by the program.
+The files are named \texttt{area\_date\_time\_info.png} and \texttt{area\_date\_workload.png}, where the area and date correspond to the area provided in the user input and the date for the schedule respectively.
+
+The \texttt{area\_date\_time\_info.png} file contains a pie chart depicting the percentage of total time spent on travel, waiting, and service.
+
+The \texttt{area\_date\_workload.png} file provides a stacked bar chart depicting the time spent on travel, waiting, and service for each carer.
+
+\subsection{Web File}
+The final file produced is a web file (.html) which provides a visualisation of the schedule and route on a map.
+The file is named \texttt{area\_date.html}, where the area and date correspond to the area provided in the user input and the date for the schedule respectively.
+
+\section{Possible Errors}
+\begin{itemize}[leftmargin=*]
+	\item If there is an incorrect postcode or missing postcode in the input file, an error will occur as the corresponding coordinates for the postcode will not be present in the Code-Point Open database.
+	\item In this case, the postcode should be replaced with a neighbouring valid postcode.
 \end{itemize}

### 2021-07-22 tex/Paper/appendix_randomly_generated_data.tex
EJOR paper HHCRSP.

%!TEX root=./main_paper.tex
\section{Randomly Generated Instances}\label{appendix:RandomlyGeneratedData}

The generation of random data was done by creating a list of available staff and a list of tasks that required to be performed by the qualified staff. 
For each of these, we define a number of properties that are important for our model. After identifying the structure that they follow in practice, we selected a number of values they could take and assigned a probability of occurring to each of these values, somewhat mimicking what happens in real life.

The instances used in this article were generated with the parameters described in Table~\ref{Appendix:Tab:InstanceDescription}, where each combination of parameters led to one instance, a total of 216. The exact instances used are available here: \url{https://github.com/c-lamas/instances_hhcrsp}.

\begin{table}[htbp!]
	\caption{Parameters used for generating the instance set. One instance was generated for each combination of parameters.}
	\centering
	\begin{tabular}{|c|c|}
		\hline
		\textbf{Parameter} & \textbf{Values} \\ \hline
		Number of nurses   &  4, 6, 8\\ \hline
		Skill mix   &  Low, Medium, High				   \\ \hline
		Proportion of jobs with time window & 10\%, 20\%\\ \hline
		Time window sizes & Small, Large, Mixed \\ \hline
		Proportion of double services & 10\%, 20\%\\ \hline
		Proportion of jobs with interdependencies & 10\%, 20\%\\ \hline
	\end{tabular}\label{Appendix:Tab:InstanceDescription}
\end{table}

The skill mix represented the diversity of skills in the workforce, as well as the broader or narrower range of tasks needed by patients. To mimic real settings, the workforce is organised in different bands. Each bad has at least one skill, and the probability of having further skills is higher in higher bands. The particular configuration of each skill mix used is described in Table~\ref{Appendix:Tab:SkillMixConfiguration}. The nurses were spread evenly across the bands, with more nurses in the lowest band if the total number of staff was non-divisible by the number of bands used.

\begin{table}[htbp!]
	\centering
	\caption{Details of the parameters used for creating instances with different skill requirements.}
	\begin{tabular}{|l|l|l|l|}
	\hline
		\textbf{Parameter} & \textbf{Low} & \textbf{Medium} & \textbf{High} \\
		\hline
		Number of skills   &  2 & 5 & 10\\ \hline
		Service times   &  Skill 1: 10 & Skill 1: 10 &Skills 1-3: 10\\
						&  Skill 2: 20 & Skills 2-3: 20 & Skills 4-6: 20 \\
						&  			   & Skill 4: 30 & Skills 7-9: 30 \\
						&  			   & Skill 5: 45 & Skill 10: 45 \\ \hline
		Number of bands & 1 & 2 & 3 \\ \hline
		Minimum skills & Band 1: 1 & Band 1: 1 & Band 1: 1 \\
		in each band  & 			 & Band 2: 2 & Band 2: 2 \\
										& 			 &  & Band 3: 3 \\ \hline
		Probability of having & Band 1: 0 & Band 1: 0.1 & Band 1: 0.1 \\
		an additional skill  & 			 & Band 2: 0.5 & Band 2: 0.2 \\
										& 			 &  & Band 3: 0.5 \\ \hline
		Probability of single & 1 Skill: 80\% & 1 Skill: 80\% & 1 Skill: 80\% \\		
		services requiring 	 & 				 & 2 Skills: 20\% & 2 Skills: 15\% \\		
		one or more skills   & 				 & & 3 Skills: 5\% \\ \hline
	\end{tabular}\label{Appendix:Tab:SkillMixConfiguration}
\end{table}

We allowed for two types of nurse shifts (4 hours, 20\% or 8 hours, 80\%), which could start at time 0 (80\%) or after 2 hours (20\%).

The number of jobs was selected to be between 9 and 11 per nurse available, a ratio similar to that in the teams we talked to. The jobs themselves were randomly assigned a location with coordinates in the square $[0, 100]^2$ and the distances calculated between them were Euclidean, with 10\% occurring at repeated locations. Dependent jobs were chosen to have a gap between them of 2, 4 or 6 hours, with equal probability.

Time windows were considered either small, large or mixed, with the following characteristics:
\begin{itemize}
	\item Small: 60 minutes duration
	\item Large: 180 minutes duration
	\item Mixed: Durations of 60 min (30\%), 120 min (40\%) or 180 min (30\%)
\end{itemize}
The start times of the time windows (in minutes) and their probabilities of occurrence were: 0 (30\%), 60 (15\%), 360 (10\%), 480 (15\%) or 540 (30\%).

The combination of skills available and skill requirements, different shifts, time windows and working hours can result in some realisations being infeasible. When those were detected, the instance was discarded and a new one was generated; as real teams would always configure the roster in a way that it can carry out the tasks they need.

% \correction{Dependency and DSs}

### 2021-07-22 tex/Paper/bib_section.tex
EJOR paper HHCRSP.

\begin{thebibliography}{}

\bibitem[{Ait Haddadene} et~al., 2016]{AitHaddadene2016}
{Ait Haddadene}, S.~R., Labadie, N., and Prodhon, C. (2016).
\newblock {A GRASP × ILS for the vehicle routing problem with time windows,
  synchronization and precedence constraints}.
\newblock {\em Expert Systems with Applications}, 66:274--294.

\bibitem[Bachouch et~al., 2011]{Bachouch2011}
Bachouch, R.~B., Guinet, A., and Hajri-Gabouj, S. (2011).
\newblock {A Decision-Making Tool for Home Health Care Nurses' Planning}.
\newblock {\em Supply Chain Forum: An International Journal}, 12(1):14--20.

\bibitem[Begur et~al., 1997]{Begur1997}
Begur, S.~V., Miller, D.~M., and Weaver, J.~R. (1997).
\newblock {An Integrated Spatial DSS for Scheduling and Routing
  Home-Health-Care Nurses}.
\newblock {\em Interfaces}, 27(4):35--48.

\bibitem[Bertels and Fahle, 2006]{BERTELS20062866}
Bertels, S. and Fahle, T. (2006).
\newblock {A hybrid setup for a hybrid scenario: combining heuristics for the
  home health care problem}.
\newblock {\em Computers {\&} Operations Research}, 33(10):2866--2890.

\bibitem[Bredstr{\"{o}}m and R{\"{o}}nnqvist, 2008]{Bredstrom2008}
Bredstr{\"{o}}m, D. and R{\"{o}}nnqvist, M. (2008).
\newblock {Combined vehicle routing and scheduling with temporal precedence and
  synchronization constraints}.
\newblock {\em European Journal of Operational Research}, 191(1):19--29.

\bibitem[Ciss{\'{e}} et~al., 2017]{Cisse2017}
Ciss{\'{e}}, M., Yal{\c{c}}ındağ, S., Kergosien, Y., Şahin, E., Lent{\'{e}},
  C., and Matta, A. (2017).
\newblock {OR problems related to Home Health Care: A review of relevant
  routing and scheduling problems}.

\bibitem[Clarke and Wright, 1964]{Clarke1964}
Clarke, G. and Wright, J.~W. (1964).
\newblock {Scheduling of Vehicles from a Central Depot to a Number of Delivery
  Points}.
\newblock {\em Operations Research}, 12(4):568--581.

\bibitem[Decerle et~al., 2018]{Decerle2018}
Decerle, J., Grunder, O., {Hajjam El Hassani}, A., and Barakat, O. (2018).
\newblock {A memetic algorithm for a home health care routing and scheduling
  problem}.
\newblock {\em Operations Research for Health Care}, 16:59--71.

\bibitem[Drexl, 2012]{Drexl2012}
Drexl, M. (2012).
\newblock {Synchronization in Vehicle Routing—A Survey of VRPs with Multiple
  Synchronization Constraints}.
\newblock {\em Transportation Science}, 46(3):297--316.

\bibitem[Eveborn et~al., 2006]{Eveborn2006}
Eveborn, P., Flisberg, P., and R{\"{o}}nnqvist, M. (2006).
\newblock {Laps Care—an operational system for staff planning of home care}.
\newblock {\em European Journal of Operational Research}, 171(3):962--976.

\bibitem[Eveborn et~al., 2009]{Eveborn2009a}
Eveborn, P., R{\"{o}}nnqvist, M., Einarsd{\'{o}}lttir, H., Eklund, M.,
  Lid{\'{e}}n, K., and Almroth, M. (2009).
\newblock {Operations research improves quality and efficiency in home care}.
\newblock {\em Interfaces}, 39(1):18--34.

\bibitem[Feo and Resende, 1989]{FEO198967}
Feo, T.~A. and Resende, M. G.~C. (1989).
\newblock {A probabilistic heuristic for a computationally difficult set
  covering problem}.
\newblock {\em Operations Research Letters}, 8(2):67--71.

\bibitem[Fikar and Hirsch, 2015]{Fikar2015}
Fikar, C. and Hirsch, P. (2015).
\newblock {A matheuristic for routing real-world home service transport systems
  facilitating walking}.
\newblock {\em Journal of Cleaner Production}, 105:300--310.

\bibitem[Fikar and Hirsch, 2017]{Fikar2017}
Fikar, C. and Hirsch, P. (2017).
\newblock {Home health care routing and scheduling: A review}.
\newblock {\em Computers {\&} Operations Research}, 77:86--95.

\bibitem[Hansen et~al., 2010]{hansen2010variable}
Hansen, P., Mladenovi{\'{c}}, N., and P{\'{e}}rez, J. A.~M. (2010).
\newblock {Variable neighbourhood search: methods and applications}.
\newblock {\em Annals of Operations Research}, 175(1):367--407.

\bibitem[Laguna and Marti, 1999]{laguna1999grasp}
Laguna, M. and Marti, R. (1999).
\newblock {GRASP and path relinking for 2-layer straight line crossing
  minimization}.
\newblock {\em INFORMS Journal on Computing}, 11(1):44--52.

\bibitem[L{\'{o}}pez-Ib{\'{a}}{\~{n}}ez et~al., 2016]{Lopez-Ibanez2016}
L{\'{o}}pez-Ib{\'{a}}{\~{n}}ez, M., Dubois-Lacoste, J., {P{\'{e}}rez
  C{\'{a}}ceres}, L., Birattari, M., and St{\"{u}}tzle, T. (2016).
\newblock {The irace package: Iterated racing for automatic algorithm
  configuration}.
\newblock {\em Operations Research Perspectives}, 3:43--58.

\bibitem[Mankowska et~al., 2014]{Mankowska2014}
Mankowska, D.~S., Meisel, F., and Bierwirth, C. (2014).
\newblock {The home health care routing and scheduling problem with
  interdependent services}.
\newblock {\em Health Care Management Science}, 17(1):15--30.

\bibitem[Mladenovi{\'{c}} and Hansen, 1997]{Mladenovic1997}
Mladenovi{\'{c}}, N. and Hansen, P. (1997).
\newblock {Variable neighborhood search}.
\newblock {\em Computers {\&} Operations Research}, 24(11):1097--1100.

\bibitem[Mosquera et~al., 2018]{Mosquera2018}
Mosquera, F., Smet, P., and {Vanden Berghe}, G. (2018).
\newblock {Flexible home care scheduling}.
\newblock {\em Omega}.

\bibitem[{NHS England}, 2019]{england2019nhs}
{NHS England} (2019).
\newblock {The NHS long term plan}.

\bibitem[{NHS Providers}, 2018]{NHSProviders2018}
{NHS Providers} (2018).
\newblock {Community services: taking centre stage}.

\bibitem[OECD, 2019]{OECDHealthataGlance2019}
OECD (2019).
\newblock {\em {Health at a Glance 2019}}.

\bibitem[Rasmussen et~al., 2012]{Rasmussen2012}
Rasmussen, M.~S., Justesen, T., Dohn, A., and Larsen, J. (2012).
\newblock {The Home Care Crew Scheduling Problem: Preference-based visit
  clustering and temporal dependencies}.
\newblock {\em European Journal of Operational Research}, 219(3):598--610.

\bibitem[Resende and Ribeiro, 2016]{resende2016optimization}
Resende, M. G.~C. and Ribeiro, C.~C. (2016).
\newblock {\em {Optimization by GRASP}}.
\newblock Springer.

\bibitem[Rest and Hirsch, 2016]{Rest2016}
Rest, K.-D. and Hirsch, P. (2016).
\newblock {Daily scheduling of home health care services using time-dependent
  public transport}.
\newblock {\em Flexible Services and Manufacturing Journal}, 28(3):495--525.

\bibitem[Robertson et~al., 2017]{robertson2017understanding}
Robertson, R., Wenzel, L., Thompson, J., and Charles, A. (2017).
\newblock {Understanding NHS financial pressures}.
\newblock {\em The King's Fund}.

\bibitem[Tarricone and Tsouros, 2008]{tarricone2008home}
Tarricone, R. and Tsouros, A.~D. (2008).
\newblock {\em {Home care in Europe: the solid facts}}.
\newblock WHO Regional Office Europe.

\end{thebibliography}

### 2021-07-22 tex/Paper/combined_table_aith_results.tex
EJOR paper HHCRSP.

\begin{table}[htbp!]
\centering
    \caption{Performance of the best and average results found by GRASP DST compared with the three algorithms presented in \cite{AitHaddadene2016} (ILS, \emph{ah\_ils}, basic GRASP,  \emph{ah\_grasp}, and GRASPxILS, \emph{ah\_gi}) for the three instance groups \emph{G1}, \emph{G2} and \emph{G3}.}
    \begin{tabular}{cccccc}
    \label{tab:group_table_aith_combined}\\\hline

Result  &   Group   &   \emph{ah\_ils}  &   \emph{ah\_grasp}    &   \emph{ah\_gi}   &   \emph{GRASP}    \\  \hline
\multicolumn{6}{l}{\emph{Best}} \\                                              
    &   \emph{G1}   &   38.16   &   \textbf{38.04}  &   \textbf{38.04}  &   \textbf{38.04}  \\  
    &   \emph{G2}   &   -97.83  &   -97.54  &   -99.57  &   \textbf{-99.69} \\  
    &   \emph{G3}   &   -251.35 &   -231.54 &   -245.42 &   \textbf{-251.55}    \\  
\multicolumn{6}{l}{\emph{Avg.}} \\                                              
    &   \emph{G1}   &   42.27   &   38.39   &   38.05   &   \textbf{38.04}  \\  
    &   \emph{G2}   &   -91.85  &   -94.04  &   -97.91  &   \textbf{-99.19} \\  
    &   \emph{G3}   &   -240.82 &   -226.79 &   -239.44 &   \textbf{-247.01}    \\  
    \hline
    \end{tabular}
\end{table}

### 2021-07-22 tex/Paper/main.tex
EJOR paper HHCRSP.

\documentclass[a4paper,11pt]{elsarticle}

\input{includes/preamble.tex}

\biboptions{authoryear}

\journal{European Journal of Operational Research}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

\title{A decision support tool for the Home Health Care Routing and Scheduling Problem}

\author[sbs]{Carlos Lamas-Fernandez$^{*,}$}
\author[exeter]{Thomas Monks}
\author[sbs]{Antonio Martinez-Sykora}

\address[sbs]{Southampton Business School, University of Southampton, Southampton SO17 1BJ, UK}
\address[exeter]{University of Exeter Medical School, University of Exeter, Exeter, UK}

\begin{abstract}
%% Text of abstract
The Home Health Care Routing and Scheduling Problem (HHCRSP) consists in assigning routes and schedules to nurses that deliver health care services to patients at their homes. A varied range of hard and soft constraints such as time windows, synchronisation of nurses, skill mix or patient and nurse preferences make this problem particularly difficult to solve manually. However, this is a challenge faced daily by many community nursing teams.

In this work we present a decision support tool (DST) that aims to help planners solve this problem in an automatic and efficient way. Our DST is based on a Greedy Randomised Adaptive Search Procedure (GRASP) with Path Relinking that solves a HHCRSP and can provide high quality solutions for the problem in terms of travel and waiting time, preference matching and workload balance.

%The DST is also linked with open source maps and an interactive reporting system that allows to visualise the solution of the HHCRSP problem in an intuitive way, which can help planners both save time and achieve more efficient plans for their everyday operations. \correction{We might need to remove here the DST stuff if not mentioned in the paper (maps, etc.)}

We compare the performance of the GRASP algorithm with state-of-the-art approaches for similar problems and show that it offers competitive results, while providing a wider range of possibilities when modelling real problems. For instance, the computational results show that considering waiting times in the objective function can reduce total operation times up to 19\% in comparison with alternative modelling approaches that minimise travel time only.

% Our preliminary results working with local home health care teams in the south of England suggest that this is a promising approach given its flexibility to include practical constraints and its speed. Furthermore, we expect that its availability and independence of commercial solvers will help reaching more teams in the future.
\end{abstract}

\begin{keyword}
OR in health services \sep Metaheuristics \sep Routing \sep Scheduling \sep GRASP
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

{\footnotesize
\noindent $^*$ Corresponding author\\
\noindent E-mail addresses: \text{C.Lamas-Fernandez@soton.ac.uk} (C. Lamas-Fernandez), \text{T.M.W.Monks@exeter.ac.uk} (T. Monks), \text{A.Martinez-Sykora@soton.ac.uk} (A. Martinez-Sykora)
}

\newpage

\section{Introduction}\label{seq:intro}

Home health care refers to a form of care where a health professional visits a patient in their own home to perform a health-related task \citep{Cisse2017}. In the European context, with an ever-ageing population, delivering health care at home is increasingly seen as a necessary alternative to traditional forms of care \citep{tarricone2008home}.
This view is shared by many health providers around the globe as managing long term conditions, frail patients and end-of-life care among others might be done more efficiently in the patient's own home. For example, in the OECD countries, the care at home for patients with long term conditions increased by 6\% \citep{OECDHealthataGlance2019} and in the UK the National Health Service (NHS) stated in their long term plan that in the future more care should be delivered to patients in their own homes \cite{england2019nhs}.

A key role in the health service to deliver this care are District Nurses (DN), specialised nursing teams that travel to patient's homes and perform a variety of tasks, ranging from insulin injections to dressing wounds. Nevertheless, DN numbers have fallen by 44\% \cite{NHSProviders2018} between 2010 and 2018 in the NHS. On the other hand, while there is no official data on demand, the workforce perception is that demand has risen significantly, putting the service under pressure \citep{robertson2017understanding}. There is, therefore, a strong need for efficiency within the community nursing service, to better utilise the available resources.

One of the first tasks that the DN team lead faces on a working day is to distribute the work for the team. In the DN teams we visited in the city of Southampton, UK, this entailed assigning a number of home health care visits (typically over a hundred) to the nurses available on the team, attending to hard requirements (skill matching, time windows, job dependencies) and soft ones (minimising travel and waiting times, maintaining workload balance, matching carer-patient preferences). Despite the abundant literature related to similar problems, the planning is currently performed manually in a lengthy process for the planner that potentially leads to suboptimal schedules. Manual planning has three further weaknesses. First there are inconsistencies between the different plans of a nurse, and between plans of different nurses. Although a single DN may become highly skilled in planning overtime that knowledge is lost when they leave the organisation. Second, a poor planning results in turn in a poor utilisation of the workforce, which might include excessive travel and waiting times, unbalanced workloads. Third, it can also affect the quality of care, leading to missed time windows or lack of preference matching for patients.

In an effort to automate and improve this process, we model the planning of home health care visits as a rich vehicle routing problem, with constraints and objectives drawn from our interaction with community nursing teams. This problem is often known in the literature as Home Health Care Routing and Scheduling Problem (HHCRSP). In our case, the schedule takes place over a single day, as it is the preferred approach in practice. This is motivated by the last-minute changes due to staff availability, patient next-day referrals and the nature of continuous review of care, which often prompts to scheduling the next visit right after reviewing their condition on the previous visit. A formal description of the problem is provided in Section~\ref{seq:model description}.

We propose a Decision Support Tool (DST) that can produce high quality daily routes and schedules for a DN team, according to a flexible objective function. The tool requires as input a patient requirement list and the available staff and then runs an optimisation algorithm based on a weighted objective function. The optimisation is based on a Greedy Randomised Adaptive Search Procedure (GRASP) with Path Relinking, which is detailed in Section~\ref{seq:DSS}.

The main contribution of this article is to present a DST for the HHCRSP that is designed to be used in real community nursing settings. It tackles the most common constraints found in literature (e.g. time windows or skills) but also others often found in practice (e.g. multiple depots or workload balance). We provide extensive computational results (Section~\ref{seq:Computational Experiments}) that show that our algorithm is competitive with state-of-the-art approaches and is flexible to allow real-world constraints and objectives; but also provide insight on the algorithm components that help achieve high quality routes and schedules.

\section{Literature review}
The generalised growth in demand and availability of home health care services of recent years has resulted in a growth in academic interest in the related scheduling problem. While the problem might be addressed by drawing from the extensive vehicle routing problem (VRP) literature (see, for example, \cite{Drexl2012} for a comprehensive review VRPs and synchronisation constraints), HHCRSP has developed its own body of literature. In particular, there are two recent review articles, \cite{Cisse2017} and \cite{Fikar2017}. We refer the reader to them for a broader understanding of what it is meant by HHCRSP and different solving approaches. Home Health Care and Home Care services across the world operate in different ways and, as such, the related scheduling problems have different objectives and constraints. For example, it might be possible to speed up the visits to some patients \citep{Mosquera2018}, to allow nurses to share a minibus for part of their route, while walking the rest of it \citep{Fikar2015}, for nurses to use public transport \citep{Rest2016}, or to allow time preferences for patients \citep{Decerle2018}. % This variety makes difficult comparing approaches and translating the findings from one application to another.

In this section we review only articles that tackle problems that are similar to those occurring in the service we model, or that are especially relevant for our research, particularly those that tackle the synchronisation of patient visits.

% Exact methods
Several articles propose exact methods. These are usually based on a Mixed Integer Linear Program (MILP) formulation that is solved with the help of a commercial solver \citep{Bredstrom2008,Bachouch2011, Mankowska2014, AitHaddadene2016}. Typically, these formulations are three index formulations that involve two types of variables: $x_{ijk}$ binary variables that indicate if carer $k$ traverses the arc $i$ and $j$; and $t_{ik}$ real variables that determine at what time carer $k$ starts performing job $i$. These formulations are limited in practice and can only solve to optimality small problems (\eg 45 patients in \cite{AitHaddadene2016} or 10 in \cite{Mankowska2014}). In order to address this, some researchers propose to decompose the problem in a column generation scheme. For example, \cite{Rasmussen2012} tackles the HHCRSP problem in Denmark. They propose a branch-and-price scheme where generalised synchronisation constraints are relaxed and dealt with in the branching. % \correction{Only Rasmussen?}
 % \correction{Cite relevant papers here}. The ILP formulations work well for small problems, but are often not practical to solve for problems with a significant number of patients. %around 50

Often, in order to tackle larger instances, the MILPs are solved in a restricted manner and, while they do not guarantee optimality, they might provide feasible solutions suitable for day to day planning. This is often referred to as a matheuristic algorithm and can be done by reducing the size of the original problem by fixing some aspects and solving it again. One example of this technique is given in \cite{Rasmussen2012} that also proposes clustering strategies to reduce the size of the problem, which allows them to solve problems with up to $150$ patients, although no longer proving optimality. Another is \cite{Bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time.

Another approach, which is the one we use in this article, is to propose a metaheuristic algorithm independent of a mathematical solver. \cite{Mankowska2014} and \cite{AitHaddadene2016} propose an Adaptive Large Neighbourhood search and a hybrid GRASP combined with Iterated Local Search respectively. One advantage of the use of heuristic algorithms is that they do not require availability of a (often commercial) MILP solver, which is unlikely to be available to health care professionals.

Finally, we look at the literature that discusses the software implementation of the solution approaches in practice. One of the earliest works is due to \cite{Begur1997}, who developed a decision support system based on the Clarke and Wright algorithm \citep{Clarke1964} for vehicle routing and then relied on integration with a Geographic Information System (GIS), an interactive user interface and a human planner to alter the routes as necessary. The Laps Care system \citep{Eveborn2006, Eveborn2009a} is a Swedish software for the HHCRSP, based on a repeated matching algorithm has been shown to be successful in saving planning time, minimising travel times and even reducing the number of missed visits. The authors estimated annual savings in 2009 of up to €120 M across teams in Norway, Finland and Sweden \citep{Eveborn2009a}. Another software is Parpap \citep{BERTELS20062866}, which proposes various approaches combining constraint programming and metaheuristics. A common claim in all these implementation-oriented works is that the software needs to be flexible enough to allow for real-world constraints to be implemented and offers interaction with mapping software.

%%%%%%%%%%%%%%%%% MODEL FOR HHCRSP %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{A model for health home care services}\label{seq:model description}

In this section we describe the notation and give a formal description of the HHCRSP problem tackled by our DST.

We consider a list of nurses $N = \{1,\dots,n\}$ that are available for the day. Each nurse $i$ might have a different starting time, $st(i)$, and end time, $en(i)$, of their working day. During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$, a subset of which, $DS \subset J$ are double services, \ie, they require two nurses to be present to be performed. Note that jobs refer to individual tasks that a nurse needs to perform and are not necessarily identified with unique patients or locations, meaning that the same patient might have more than one job on the list.

A schedule (not necessarily feasible) $\mathcal{S}$, is made of a pair   $\mathcal{S} = \{\mathcal{R}, \mathcal{O}\}$, where $\mathcal{O}$ is a permutation of the vector $o = (1,\dots,n)$ representing the order of the nurses in the solution and $\mathcal{R}$ is a set of $n$ routes, $\mathcal{R} = \{r_1,\dots,r_n\}$, one per nurse. Each route $r_i$ is itself an ordered list of $m_i$ jobs, $r_i = \{k^i_0, k^i_1,\dots,k^i_{m_i}, k^i_0\}$, that are the jobs assigned to nurse $i \in N$ and starts and ends at the depot of that nurse denoted by, $k^i_0$. Of course, a route might be empty, meaning that that particular nurse is not assigned any jobs and $m_i = 0$ and the route only contains the depot, $r_i = \{k^i_0, k^i_0\}$.
After a schedule is generated, for each job $j \in J$ we denote by $t_j$ its starting time. Note that this refers to the time at which the job starts, rather than the time at which the nurse (or nurses) performing it arrive to the job location, which might be different.

In the rest of the section, we formally define the constraints and objectives that commonly apply in practice and relate to the skill mix of the workforce, the preferences of some assignments, travel times and synchronisation, visits with two staff members and quality of the schedule.

\subsection{Skill mix}
Some jobs might require a specialised nurse to be performed, and nurses might only perform jobs that they are qualified to do. To model this, we define a function $S: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise. 
Naturally, for double service jobs, no single nurse is skilled to perform them $S_1(i,j) = 0, \forall i \in N, \forall j \in DS$. To account for this, we define a second function $S_2: N\times N\times J \rightarrow \{0,1\}$, that takes value $S_2(i_1,i_2,j) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform job $j$ and $0$ otherwise.

Using these functions, we can write constraints regarding job assignment and skill mix.
All individual service jobs must be assigned to a skilled nurse, thus in any feasible solution it must hold that if $j \in r_i$ then $S_1(i,j) = 1$. Similarly, all double service jobs must be assigned to a pair of skilled nurses so if $j \in r_i \cap r_{i'}$ then it must hold that $S_2(i, i', j) = 1$.

A relevant modelling consideration relates to the necessary skill mix for simultaneous double services or, in other words, how the function $S_2$ is created in practice. We distinguish the following cases:
\begin{itemize}
    \item \textbf{Duplicated}. In this case it is required for all staff members visiting the patient to have all necessary skills to perform the service.
    \item \textbf{Shared}. This is the case when it is sufficient that the combination of skills between the staff members present is enough to perform the service. In particular, it could happen that only one staff member is qualified for all services and the other for none. An example of this would be visits that require a second person to provide support and ensure safety, for example with patients that have a record of behaving inappropriately towards staff.
    \item \textbf{Strictly shared}. In this case each member of the service does not need to be qualified for all services but should be attending the visit to perform at least one service. This is the approach used in \cite{Mankowska2014} and \cite{AitHaddadene2016}.
    \item \textbf{Shared-Duplicated}. This case is analogous to \emph{Shared}, but it allows some skills to be required by more than one of the staff members, hence it generalises the previous cases.
\end{itemize}

In our DST, the default approach is \emph{Shared-Duplicated}, as it allows for the flexibility that is observed in real operations. The data is introduced as a list of skills and, if some skills need to be present in more than one staff member, they can simply appear duplicated on the list. We have, however, also implemented the Strictly Shared approach, in order to solve the instances from the literature.

% \subsubsection{Hierarchical skills and downgrading}
It is not uncommon in the literature to model the skills of the workforce as hierarchical, \ie each nurse is assigned a skill level and it is assumed that they possess all the skills that lower levels have, and some more. The model might then introduce a penalty or limit for high skilled nurses performing too many low skilled jobs, known as downgrading \cite[see, \eg][]{Fikar2015}. In the DN teams we observed, skills were not hierarchical, neither were any such limits in place. Nevertheless, nurses pointed out that high skilled nurses were typically assigned only the high skilled jobs, as they might need to perform other jobs (\eg attending meetings or other administrative tasks). While this is not something that we consider explicitly, a similar objective may be achieved through setting negative preferences between high skilled nurses and low skilled jobs.

\subsection{Preferences}
It is not uncommon in practice that district nurses do their planning using preferences. These might arise from multiple reasons, such as a preserving continuity of care, gender preference, patients and nurses speaking the same language, previous conflicts between a patient and a staff member, etc. We model this with a preference function, $P: N\times J \rightarrow \mathbb{R}$ such that $P(i,j)$ represents the preference score of assigning job $i$ to staff member $j$. The function $P$ should assign a higher (positive) score to preferred patient-nurse assignments and a lower (negative) score to assignments that are better avoided. We give an indication on how to choose the specific values of the function in Section~\ref{seq:adequate_weights}.

\subsection{Travel times}
To describe travel times, for each nurse we define $N$ functions $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, where $k^i_0$ refers to the start and end location of the nurse $i$. Each nurse might start from a different depot (typically, their own home), a fact that is accounted for in these functions. Note that with this definition we allow to have different travel times between jobs for different nurses, and this can be used to model, for example, different means of transport. In our observations, however, all nurses used their own vehicles, so we assume that $T_i$ returns driving times for all nurses and, in particular that:
\begin{equation}
    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2),\ \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J
\end{equation}

Note that this function can be substituted by a more sophisticated one without any changes to the algorithms described here.

% For simplicity, we also assume this in our experiments, although a more complex function could be provided without modifying the DST.

\subsection{Temporal constraints}\label{seq:sub temporal constraints}
Each job $j$ has a service time $s_j$ and a time window that is defined by its earliest start time, $e_j$ and the latest start time $l_j$.

It is a hard constraint that jobs cannot start before their earliest start time,
\begin{align}
    t_j \geq e_j, & \forall j \in J
\end{align}

However, they might start late. The tardiness with which a job $j$ starts, $z_j$ can be simply calculated by:
\begin{equation}
    z_j = \max(0, t_j - l_j)
\end{equation}

Note that tardiness here refers only to the difference between the job start and its original time window. Jobs that have dependencies or require two nurses might be further restricted temporally. This might happen in practice, for example, in situations where a certain medication must be administered a few hours after the first dose. If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$, that depends on the start time of $j_1$, $[e^{(j_1)}_{j_2} + t_{j_1}, l^{(j_1)}_{j_2} + t_{j_1}]$, then, the same time window constraints hold:
\begin{equation}
    t_{j_2} \geq e^{(j_1)}_{j_2} + t_{j_1}, % & \forall j \in J
\end{equation}
\begin{equation}
    t_{j_2} \leq l^{(j_1)}_{j_2} + t_{j_1}, % & \forall j \in J
\end{equation}
Similarly, $j_2$ will impose a time window on $j_1$, and in a valid instance of the problem it must hold that:
\begin{equation}
    e^{(j_1)}_{j_2} = -e^{(j_2)}_{j_1}
\end{equation}
and
\begin{equation}
    l^{(j_1)}_{j_2} = -l^{(j_2)}_{j_1}
\end{equation}

Then, we can define the \emph{gap tardiness} of job $j_2$, $\tilde{z}_{j_2}$ as:
\begin{equation}
    \tilde{z}_{j_2} = \max(0, t_{j_2} - (l^{(j_1)}_{j_2} + t_{j_1})]
\end{equation}

Since some jobs have time windows or might need to be synchronised, waiting times might occur. We denote by $w_{ij}$ the waiting time of nurse $i$ at job $j$. Note that $w_{ij} = 0$ if $j \notin r_i$ and, otherwise, $w_{ij}$ represents the difference between $t_j$ and the arrival time of nurse $i$ to job $j$.% , denoted by $a(i,j)$.

The arrival time $a(i,j)$ of the nurse $i$ at the job $j \in r_i$ is calculated as follows:

\begin{equation}
    a(i,j) = st(i) + \sum_{h \in 0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1})
\end{equation}
Where $j'$ denotes the position of job $j$ in the route $r_i$, \ie $j = k^i_{j'} \in r_i$. Note that there is not waiting or service times at the depot, \ie $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 \ \forall i \in \mathcal{N}$.

In practice, waiting times might occur for a variety of reasons. In order to maintain simplicity in our model, we only consider three reasons for which a nurse $i$ might have waiting time $w_{ij} \geq 0$ at the job $j \in r_i$:
\begin{itemize}
    \item Respecting a time window: 
    \begin{align}
        w_{ij} \geq e_i - a(i,j) & \forall i \in N, \forall j \in J
    \end{align}
    \item Waiting for a second nurse $i'$ to arrive for a double service
    \begin{align}
        w_{i j} \geq t_j - a(i',j) & \forall j \in DS,\ such\ that\ j \in r_{i} \cap r_{i'}
    \end{align}
    \item Waiting for the minimum gap of a synchronised service. If the job $j$ depends on $j'$, then:
    \begin{equation}\label{eq:waiting_for_synchronised}
        w_{ij} \geq e^{(j')}_{j} - a(i,j)
    \end{equation}
\end{itemize}

Further, we assume that waiting times are never larger than strictly necessary. To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based  in the order in which they appear in the solution. Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, and such that $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$ and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$. Then, we enforce that:
%  $i_1 < i_2$, \ie $j_1$ appears earlier in the vector $\mathcal{O}$ than $j_2$. Then, we enforce that:
\begin{equation}\label{eq:earliest_possible_start_assumption}
    w_{i_1 j_1} = \max \{ a(i,j_1), e_{j_1} \}
\end{equation}

and

\begin{equation}\label{eq:earliest_possible_start_assumption2}
    t_{j_1} = \max \{ a(i,j_1), e_{j_1} \}
\end{equation}

Note that this implies that the order in which nurse routes appear in a solution may have an impact in the solution quality. To mitigate this effect, in Section~\ref{seq:DSS} we describe strategies to search different orderings of nurses when solving the problem.

\subsection{Objective function}
% The objective function is based on the quality of a solution, which consists on a few measures of interest that are added together with different weights. Performance measures to be maximised have positive weights, while those to be minimised have negative weights.
We consider a weighted objective function that includes waiting time, travel time, tardiness, maximum tardiness, overtime, workload balance and preferences. The measures to be maximised have positive weights, while those to be minimised have negative weights.

\paragraph{Total waiting time}\ \\
The total waiting time in a solution $\bar{W}$, calculated as:
\begin{equation}
     \bar{W} = \sum_{i \in \mathcal{N}} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}
\end{equation}

\paragraph{Total Travel time}\ \\

If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
\begin{equation}
    TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1})
\end{equation}
The total travel time in $\mathcal{S}$, $\bar{T}$ is calculated as:
\begin{equation}
     \bar{T} = \sum_{i \in \mathcal{N}}  TR(r_i)
\end{equation}

\paragraph{Total Tardiness}\ \\
The total tardiness in a solution, is simply the sum of the tardiness across all jobs,
\begin{equation}
    \bar{Z} = \sum_{j \in \mathcal{J}} z_j
\end{equation}

\paragraph{Maximum Tardiness}\ \\
The maximum tardiness in a solution, $Z_{max}$, is given by:
\begin{equation}
    Z_{max} = \max_{j \in \mathcal{J}} z_j
\end{equation}

\paragraph{Total Overtime}\ \\
The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$. If we denote the arrival of a nurse $i$ to its end point by $arr(i)$,
\begin{equation}
	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k
\end{equation}
Then, the total overtime can be calculated as
\begin{equation}
     \bar{O} = \sum_{i \in \mathcal{N}} \max \{0, arr(i) - en(i)\}
\end{equation}

\paragraph{Workload Balance}\ \\

In order to produce practical solutions, maintaining a balance in the workload is fundamental. To achieve this, we quantify the minimum spare time on any nurse route in the schedule, $\bar{B}$. If $\bar{B}$ is minimised, we favour solutions in which the lowest spare time of any given nurse is as large as possible, hence removing imbalance when possible.

\begin{equation}
    \bar{B} = \min_{i \in \mathcal{N}} \left\{\max \{0, en(i) - arr(i)\} \right\}
\end{equation}

\paragraph{Total Preference Score}\ \\

The total preference score in the solution, $\bar{P}$, is computed as:
\begin{equation}
    \bar{P} = \sum_{i \in \mathcal{N}} \sum_{j \in r_i \subset \mathcal{R}} P(i,j)
\end{equation}

Finally, the full quality of the solution, $q(\mathcal{S})$ is then obtained as:
\begin{equation}\label{eq:solution_quality}
    q(\mathcal{S}) = \alpha_1 \bar{T} + % Travel time 
         \alpha_2 \bar{W} + % Waiting time 
         \alpha_3 \bar{Z} + % Tardiness 
         \alpha_4 \bar{O} + % Overtime 
         \alpha_5 \bar{B} + % Workload balance 
         \alpha_6 \bar{P} + % Preference score 
         \alpha_7 Z_{max} 
\end{equation}

Note that some of the terms considered in the quality function \eqref{eq:solution_quality} might be considered infeasible in some settings (\eg overtime and tardiness). For this reason, the objective function that is used during the optimisation, $Q(\mathcal{S})$, is defined as follows:

\begin{equation}\label{eq:real_obj_function}
  \max Q(\mathcal{S}) =
  \begin{cases}
   q(\ensuremath{\mathcal{S}}) & \text{if \ensuremath{I(\mathcal{S}) \leq 0}} \\
   q(\ensuremath{\mathcal{S}}) + \ensuremath{M_1 I(\mathcal{S})} + \ensuremath{M_2} & \text{if \ensuremath{I(\mathcal{S}) > 0}} \\
  \end{cases}
\end{equation}

where $I(\mathcal{S})$ represents the value of the infeasibility in the solution and hence solutions where $I(\mathcal{S}) > 0$ are considered infeasible. The measures that are included in the calculation of $I(\mathcal{S})$ might vary depending on the team being modelled. Typically, they include the total gap tardiness, $\tilde{Z}$:
\begin{equation}
    \tilde{Z} = \sum_{j \in \mathcal{J}} \tilde{z}_j
\end{equation}
 and often the overtime in the solution, $\bar{O}$ and the tardiness in the solution, $\bar{Z}$. For example, if no overtime or gap tardiness is allowed, $I(\mathcal{S})$ might be defined as $I(\mathcal{S}) = \bar{O} + \tilde{Z}$.

The terms $M_1$ and $M_2$ represent an arbitrary large value, which is used to guide the algorithm towards feasible solutions. We include $M_1$ to drive the search towards solutions with smaller infeasibility values and $M_2$ to prevent accepting infeasible solutions when the value of $I(\mathcal{S})$ is very small. In our experiments we have used $M_1 = M_2 = -10^3$.

\subsubsection{Setting adequate weights}\label{seq:adequate_weights}

Having different measures on the objective function, some of which might represent quantities of different nature (\eg waiting time and preference of assignment), raises the question on how the weights should be chosen. While the way to do this is better assessed by practitioners, we provide a simple guidance on how to calibrate these weights using time as a reference.
As a base measure, we use the total time used to provide care to patients, \ie the sum of travel time and waiting time (service time is not included, as it does not depend on the scheduling). As such, their weights are set to one, $\alpha_1 = \alpha_2 = 1$.
Now, the remaining weights are set according to their relative importance to this base measure. This means, answering the question: How many minutes of total time would we be willing to sacrifice to improve this other measure by one minute? The answer to this question indicates the weight that should be assigned to that measure.
The only exception is the preference score, for which we set a weight of one, $\alpha_3 = 1$, as their importance in the objective is determined by the value of the assignment. The individual value of the preference of each assignment should be determined by answering the question: How many minutes of overall service are we willing to sacrifice so this assignment is maintained (or avoided)? If an assignment is to be avoided, it should have a negative value.

The weights we propose for the Decision Support Tool are listed in Table~\ref{tab:adequate_weights}.

\begin{table}[htbp!]
    \centering
    \caption{Weights proposed for the objective function of the DST}
    \begin{tabular}{ccc}
    \hline
    \textbf{Weight}  & \textbf{Value} & \textbf{Description} \\
    \hline
    $\alpha_1$      &   -1   &   Travel time \\
    $\alpha_2$      &   -1   &   Waiting time \\
    $\alpha_3$      &   -15   &   Tardiness \\
    $\alpha_4$      &   -15   &   Overtime \\   
    $\alpha_5$      &   $\frac{1}{2}$   &   Workload balance \\       
    $\alpha_6$      &   1   &   Preference score \\    
    $\alpha_7$      &   0   &   Maximum tardiness\\    
    \hline
    \end{tabular}
    \label{tab:adequate_weights}
\end{table}

Note that the weight for workload balance is smaller than one, meaning that the workload will be balanced by one minute if that costs less than one minute to the overall schedule. This value might be adjusted depending on the number of nurses on the teams, as the impact on the total schedule might be easier to absorb if more nurses are present.

\section{Decision Support Tool}\label{seq:DSS}

The problem described in the previous section aims to capture all the real-world constraints we observed in practice, but also to include a high degree of flexibility in the modelling. This occurs for example by the introduction of preferences (which can account for a number of features, as continuity of care, for example) or weights in the objective function. The idea behind this is that community nursing teams might use this tool as a decision support tool and produce a few schedules with perhaps slightly different objectives or preference assignments. The main component of the tool is a Greedy Randomised Adaptive Search Procedure (GRASP) optimisation algorithm, which aims to provide efficient solutions. As noted by other researchers \cite[\eg][]{BERTELS20062866, Rasmussen2012, Mankowska2014}, the problem presented in Section~\ref{seq:model description} and its similar versions are NP-Hard.

\subsection{GRASP}

The GRASP framework was originally proposed by \cite{FEO198967} and has been widely used and extended since then \citep{resende2016optimization}. In its basic form, it consists in generating greedy solutions with a controlled degree of randomness and then improving them with a local search. The general layout of our implementation of the GRASP algorithm is depicted in Figure~\ref{fig:grasp_diagram}.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=0.7\textwidth]{im/scheme}
    % \includegraphics{}
    \caption{Scheme of the proposed GRASP algorithm.}
    \label{fig:grasp_diagram}
\end{figure}

Solutions are generated iteratively by the randomised constructive (described in Section~\ref{subseq:constructive}) and improved with local search (described in Section~\ref{subseq:local_search}). Good solutions are kept in a solution pool if they are different enough from the ones currently in the pool. In order to speed the convergence of the algorithm, the solutions of the pool are used to perform path relinking (see Section~\ref{subseq:PR}). After each iteration we check the stopping criterion. In our implementation this was a time limit. The order of the nurse list is randomly generated for each iteration, as this might have an impact on waiting times.

\subsection{Randomised constructive}\label{subseq:constructive}

The Randomised Constructive Algorithm (RCA) is used to create initial solutions for each iteration of the GRASP algorithm. It works by ranking the quality of all job-nurse assignments by the value of the objective function \eqref{eq:real_obj_function}, which might also be evaluated in partial solutions. These rankings are sorted by quality, and the highest quality ones are stored in a reduced candidate list (RCL). Then, one of the elements of the list is randomly selected and added to the solution. There are a number of ways in which an RCL can be created, relating to which is the `cut-off' value of the objective value that decides if an assignment enters the RCL or not. In our implementation we have tested two strategies, both depending on the hyper-parameter $\delta$. Let $B$ ($W$) be the best (worst) ranking found in a given iteration of the RCA, the strategies to find the cut-off value, $C$, are as follows:
\begin{itemize}
     \item[1)] Strategy 1:
     \begin{equation}
         C = B + \delta (B - W)
     \end{equation}

     \item[2)] Strategy 2:
     \begin{equation}
         C = B - \delta B
     \end{equation}
 \end{itemize} 
The parameter $\delta$ is randomly selected in each iteration, sampled from the interval $\delta \in [\texttt{grasp\_dl}, \texttt{grasp\_dl} + \texttt{grasp\_dr}]$.
The values of \texttt{grasp\_dl} + \texttt{grasp\_dr}, together with the type of strategy to use, \texttt{rcl\_strategy}, are hyper-parameters of the algorithm and are optimised in Section~\ref{seq:parameter_tuning}. The RCA has two phases. In the first part, each nurse is assigned a single job using this procedure. This ensures that all nurses will have an initial route, regardless of how far their starting location is, and helps constructing more balanced solutions. In the second part, jobs are ranked and assigned to the best position in any of the nurse routes, until no more jobs are left to assign.

\subsection{Local search}\label{subseq:local_search}
After each iteration of the constructive algorithm a local search procedure is applied to the resulting solution to improve its quality. The local search consists in a first improvement Variable Neighbourhood Search (VNS) framework \citep{Mladenovic1997}, a general metaheuristic algorithm that has proven to be successful in many application areas, including routing problems \cite{hansen2010variable}, due to its capacity to include a range of different local search movements. VNS takes as input an initial solution and an ordered neighbourhood structure. It searches the first neighbourhood, moves to the first improving solution found and then searches the neighbourhood again. If no improvement is found, it searches the next neighbourhood on the list, always returning to the first neighbourhood if an improvement is found. If there is not any improving solution in the last neighbourhood of the list, the algorithm stops and returns the latest solution.

The quality of the VNS is directly influenced by the neighbourhoods used and their order. We have implemented the following neighbourhoods:
% \correction{Review names and add descriptions}
\begin{itemize}
    \item NE01 Moving a job: This neighbourhood involves removing one service from the route of one nurse and inserting it in the best possible position of another (or the same) nurse.
    \item NE02 Exchange two services: Two jobs from different nurse routes are removed and reinserted in the corresponding position of the other route.
    % \correction{Check code for real ordering and numbers here}
    \item NE03 Nurse order change: This neighbourhood focuses on exploring different orderings of the nurse list, $\mathcal{O}$. It searches over exchanges between pairs of nurses in the list. 
    This can have an effect in the way waiting times are calculated (as detailed in Section~\ref{seq:sub temporal constraints}) and improve the overall quality.
    \item NE04 Remove and reinsert linked services: Two services that are dependent on each other (synchronised or double services) are removed simultaneously from the solution and reinserted in the best possible positions.
    \item NE05 2OPT: This is the classical vehicle routing neighbourhood and operates in a single route. Subsets of each route are removed and reinserted in a reversed order.
\end{itemize}

While some neighbourhoods proved key to the performance of the GRASP algorithm in our initial testing, the benefit of 2OPT and the nurse order change was not immediately evident. For this reason, in Section~\ref{seq:parameter_tuning} we analyse whether they are helpful to the overall performance of the algorithm or not.

\subsection{Path Relinking}\label{subseq:PR}
% \subsection{Description}
The basic idea behind Path Relinking (PR) \citep{laguna1999grasp,resende2016optimization} is to use two existing solutions to generate a third one, which can potentially have higher quality than both of them.
This is done by transforming one of them (called starting solution) step by step, by using local search movements, in the other solution (called guiding solution).

In our implementation this is done by means of two local search movements, re-arranging the position of a nurse in the nurse order vector and removing and re-inserting a job in a different position. Each of these movements is done such that they result in the nurse or job matching as closely as possible the position seen in the guiding solution. The intermediate solution of highest quality is then saved, and local search is applied on it. Note that, this intermediate solution is often not a local optimum (unlike the starting and the guiding solutions), so applying local search may lead to high quality solutions. The pseudo-code of the routine is described in Algorithm~\ref{alg::path_relinking}.

\input{pr_algorithm.tex}

Note that double services are removed and reinserted simultaneously (see line~\ref{alg:pr:for_insertion}). In the case that some jobs that are due to be inserted in a position later than the current length of the route, they are instead inserted in the last possible position. This might cause the solutions not to be equal after applying $n + m$ movements, so if this happens another pass is performed (line~\ref{alg:pr:main_while}).

During the GRASP algorithm, good solutions are kept into a pool, and the relinking occurs between a newly generated solution and one or more solutions from the pool.
The number of solutions in the pool is controlled by the parameter \texttt{pool\_size}. We determine a suitable value for this parameter in our experiments (see Section~\ref{seq:Computational Experiments}).

During the algorithm, solutions that have higher quality than the worst solution currently in the pool will enter it, unless they are too similar to any existing solution with better quality. To further ensure diversity, if a solution enters the pool, it will do so by replacing the solution most similar to it from those that have lowest quality. To assess if a solution can enter the pool and which solution it will replace, we use the dissimilarity function, $d(\mathcal{S}_1, \mathcal{S}_2)$, which is defined as follows. Let schedule $\mathcal{S}_1$ be the pair $\mathcal{S}_1 = \{\mathcal{R}_1, \mathcal{O}_1\}$ and schedule $\mathcal{S}_2 = \{\mathcal{R}_2, \mathcal{O}_2\}$.
\begin{equation}
    d(\mathcal{S}_1, \mathcal{S}_2) = d_o(\mathcal{O}_1, \mathcal{O}_2) + d_a(\mathcal{R}_1, \mathcal{R}_2) + d_r(\mathcal{R}_1, \mathcal{R}_2)
\end{equation}

This is a combination of three measures, the dissimilarity of the nurse order $d_o(\mathcal{O}_1, \mathcal{O}_2)$, of the nurse-job assignments, $d_a(\mathcal{R}_1, \mathcal{R}_2)$, and the differences in the routes themselves, $d_r(\mathcal{R}_1, \mathcal{R}_2)$.

The distance between the solution orderings is simply calculated as the Hamming distance between the ordering vectors, \ie the number of different digits in them:
\begin{equation}
    d_o(\mathcal{O}_1, \mathcal{O}_2) = |\{i\in\{1,\dots,N\} : \mathcal{O}_1(i) \neq  \mathcal{O}_2(i)\}|
\end{equation}
where $\mathcal{O}_1(i)$ and $\mathcal{O}_2(i)$ represent the $i^{th}$ element in the nurse order vector of schedules $\mathcal{S}_1$ and $\mathcal{S}_2$ respectively.

The dissimilarity in the assignments is calculated as the number of services that are assigned to different nurses in the solutions. If we denote by $\mathcal{R}_1 = \{
r^{(1)}_1, r^{(1)}_2, \dots, r^{(1)}_n\}$ and $\mathcal{R}_2 = \{r^{(2)}_1, r^{(2)}_2, \dots, r^{(2)}_n\}$ the routes in schedules $\mathcal{S}_1$ and $\mathcal{S}_2$ respectively, then:

\begin{equation}
    d_a(\mathcal{R}_1, \mathcal{R}_2) = |\{j\in\{1,\dots,J\} : \nexists i \in N such\ that\ j \in r^{(1)}_i \cap r^{(2)}_i\}|
\end{equation}

Finally, the dissimilarities in the routes are captured by looking at the presence of different arcs in the solution. Let $A(\mathcal{R}_1)$ and $A(\mathcal{R}_2)$ be the sets of arcs present in the routes of schedules $\mathcal{S}_1$ and $\mathcal{S}_2$ respectively, including those to and from the depots, then:
\begin{equation}
    d_r(\mathcal{R}_1, \mathcal{R}_2) = |A(\mathcal{R}_1)| - |A(\mathcal{R}_1) \cap A(\mathcal{R}_2)|
\end{equation}
Note that this function is not symmetrical when $A(\mathcal{R}_1) \neq A(\mathcal{R}_2)$, although this may only happen if one schedule has a different number of empty nurse routes, which is a rare case in practice. Nevertheless, in this case the dissimilarity value will be at least $3$ (two arcs and one assignment), and the difference between the two cases is bounded by $||A(\mathcal{R}_1)| - |A(\mathcal{R}_2)||$, which is at most $1$ for each difference in an empty route. 
Including the route dissimilarity is particularly important to differentiate the cases where solutions might have nurses with similar profiles (in terms of starting locations, shift times and skills). If the nurses are simply being interchanged and doing the same routes, the solutions will have a positive value for ordering and assignment dissimilarities, but no arc dissimilarity. Nevertheless, if they have genuinely different routes, even if they are interchanging their jobs, the solution will have a high route dissimilarity too.

In our experiments setting a dissimilarity threshold of $5$ proved sufficient to maintain diversity in the pool of solutions.

\subsubsection{Type of Path Relinking}
Path relinking is performed at the end of the algorithm between all pairs of solutions in the pool. However, it is also possible to perform path relinking at intermediate steps of the algorithm too, each time a new solution is generated. Since this is potentially very time consuming, we have we have investigated the following three strategies for relinking intermediate solutions:
\begin{itemize}
    \item PR01 Perform path relinking between every generated solution and another solution selected randomly from the pool.
    \item PR02 Perform path relinking between every new solution and all solutions in the pool.
    \item PR03 Do not perform path relinking during the algorithm.
\end{itemize}

Finally, another important parameter of PR is the direction. In forward path relinking, the quality of the starting solution is less or equal than the quality of the guiding solution. In backward path relinking, the quality of the guiding solution is lower. In order to determine the choice of direction, we use the parameter \texttt{pr\_direction}, which represents the choice of direction of path relinking in each iteration: backward, forward, backward and forward or a random pick between them. For relinking at the end of the algorithm we always use both, forward path relinking and backward path relinking, between all pairs of solutions.

Both, a suitable strategy and direction for PR are determined in Section~\ref{seq:parameter_tuning}.
% \end{itemize}

\section{Computational experiments}\label{seq:Computational Experiments}
\subsection{Testing data}
For our experiments, we use three datasets. Two are from two similar problems in the literature, presented in \cite{Mankowska2014} and \cite{AitHaddadene2016}, and are used to evaluate the algorithm performance in Section~\ref{seq:algorithm_performance}. The third dataset is composed of randomly generated instances based on road distances in the city of Southampton. This dataset is used for tuning the algorithm parameters in Section~\ref{seq:parameter_tuning} and to explore different weights of the objective function in Section~\ref{seq:exploring_obj_function}. Details on the instances and the process for generating them are detailed in \ref{appendix:RandomlyGeneratedData}.

\subsection{Parameter tuning}\label{seq:parameter_tuning}
The GRASP algorithm is made up of several components, for each of which we have identified various potential configurations. %For example, we can choose whether to use or not some of the neighbourhoods described in Section~\ref{subseq:local_search}.
In order to select the best possible configuration automatically and in a statistically sound manner, we used the software package \texttt{irace} \citep{Lopez-Ibanez2016} with a budget of 1000 runs. Table~\ref{tab:parameters} shows a list of the hyper-parameters and solving strategies tested. In the first column we show the name of the hyper-parameter, in the second column a brief description, in the third column the type of variable we used for testing it, in the fourth column the values that were tested and in the fifth column the best value found by \texttt{irace}.

% "optparams=-do_twopt=0 -no_change_ls=1 -no_change_grasp=1 -pr_strategy=1 -pr_direction=1 -sols_in_pool=19 -grasp_dl=0.25 -grasp_dr=0.81 -rcl_strategy=2"
\begin{table}[h]
    \centering
    \caption{Parameters tested for configuring the GRASP algorithm, along with their type, values tested and the best configuration found by \texttt{irace}. Variable types used are binary (B), Categorical (C), Integer (I) and Real (R).}
    \begin{tabular}{cp{5.5cm}ccc}
    \hline
    \textbf{Parameter}& \textbf{Explanation}  & \textbf{Type} & \textbf{Tested} & \textbf{Best} \\
    \hline
    NE03                    & Use a local search neighbourhood to change nurse order & B & Yes, No               &  Yes  \\ % no_change_ls
    NE05                    &   Use the two opt neighbourhood & B & Yes, No                           & No   \\ % do_twopt
    NurseChange             &   Change nurse order between GRASP iterations & B & Yes, No                        &  Yes  \\ % no_change_grasp
    \texttt{pr\_strategy}   &  Path relinking strategy, as listed in Section~\ref{subseq:PR}& C & 1, 2, 3   &   1 \\ %  
    \texttt{pr\_direction}  &   Path relinking direction: forward (1), backward (2), forward and backward (3), chosen randomly at each iteration (4).& C & 1, 2, 3, 4       &   1 \\ % performPathRelinking
    \texttt{pool\_size}     &  Number of solutions in pool & I & [1,100]                                            & 19    \\ % sols_in_pool
    \texttt{grasp\_dl}              & Minimum value of $\delta$ &   R & [0,0.5]                                     & 0.25  \\ % grasp_dl
    \texttt{grasp\_dr}              & Range of $\delta$ &   R & [0,1]                                               & 0.81\\ % grasp_dr
    \texttt{rcl\_strategy}          & Strategy for generating the reduced candidate list & C & 1,2          & 2 \\ % rcl_strategy    
    \hline
    \end{tabular}
    \label{tab:parameters}
\end{table}

For the remainder of the section we use the parameters found by \texttt{irace} in all our experiments.

\subsection{Variations in the literature}\label{seq:literature data}
While the problems reported by \cite{Mankowska2014} and \cite{AitHaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to considered.

\cite{Mankowska2014} describes the following objective function:
\begin{equation}
    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max}
\end{equation}
where $Dist$ denotes the distance travelled by all nurses, $Travel$ denotes the total tardiness and $Travel^{max}$ denotes the maximum tardiness. For \cite{Mankowska2014} tardiness refers to the time between the end of its time window and the actual start time of a job. Tardiness is not allowed for interdependent jobs, where the dependency gap is a hard constraint. Further, the synchronisation is simplified for jobs with two skills, where the skill with the lower index must always be performed earlier than the skill with the largest index.
Tardiness does not include overtime for the workers either, which is not mentioned in the model, so we have not set a limit for working hours of staff when solving these instances.

The lambda values are weights to determine the importance of each factor, and the values they chose for their experiments are $\lambda_1 = \lambda_2 = \lambda_3 = \frac{1}{3}$.

In \cite{AitHaddadene2016} tardiness is not allowed, but preferences are taken into account, so their objective function is as follows:
\begin{equation}
    \min Dist + Pref
\end{equation}
Where $Dist$ represents the cost of travelling for all nurses and $Pref$ the preference score.
In Table~\ref{tab:adapted_obj_values} we specify the weights used in the objective function \eqref{eq:solution_quality} in order to represent the objective functions of these algorithms.

\begin{table}[htbp!]
    \centering
    \caption{Weights used in the objective function to match those used in the methods from  \cite{AitHaddadene2016} (AH) and \cite{Mankowska2014} (MK). A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, \ie they contribute to $I(S)$.}
    \begin{tabular}{ccc}
    \hline
    \textbf{Weight}  & \textbf{Values for AH} & \textbf{Values for MK} \\
    \hline
    $\alpha_1$      &   $-0.3\times60$          &  $-\frac{1}{3}\times60$   \\
    $\alpha_2$      &   $0$                     &  $0$                      \\
    $\alpha_3$      &   -                       &  $-\frac{1}{3}\times60$   \\
    $\alpha_4$      &   -                       &  $0$   \\   
    $\alpha_5$      &   $0$                     &  $0$                      \\       
    $\alpha_6$      &   $-1$                    &  $0$                      \\    
    $\alpha_7$      &   -                       &  $-\frac{1}{3}\times60$             \\    
    \hline
    \end{tabular}
    \label{tab:adapted_obj_values}
\end{table}

Note that we multiply time measures by $60$, since we consider our objective function in minutes, so it is easier to understand by final users of the tool. Furthermore, since both articles present minimisation problems, we switch the sign of the solutions we report in Section~\ref{seq:algorithm_performance} for ease of comparison.

Another difference between \cite{AitHaddadene2016} and our model is that for time dependent services, we define a precedence between them, while they define a gap. That means that our constraints precisely determine which job should be executed first, while \cite{AitHaddadene2016} leaves this decision to the algorithm, specifying only the minimum and maximum amount of time that might separate their start times. This was accounted for in our algorithm by considering the dependency only at runtime and choosing the most convenient gap (positive or negative) in each objective function evaluation, depending on what the arrival times were.

\subsection{Algorithm performance}\label{seq:algorithm_performance}

While the problem description of our DST is slightly different from the examples seen in Section~\ref{seq:literature data}, we test the performance of GRASP algorithm, with the required modifications, in the instances provided by \cite{AitHaddadene2016} and \cite{Mankowska2014}.

For all the instances, we perform 5 runs and report the best and the average result. The full computational results and solving times are listed in the Appendix tables \ref{tab:mk_results}, \ref{tab:aith_results} and \ref{tab:longer_experiments}.

In Table~\ref{tab:group_table_aith_combined} we present a summary of the results found by our DST for the three instance groups presented in \cite{AitHaddadene2016}. The results are the average of the best and average results found for the instances of the three instance groups \emph{G1}, \emph{G2} and \emph{G3}. We have allowed a time of 5 seconds for instances in the group \emph{G1}, 5 minutes for those in \emph{G2} and three hours for the instances in \emph{G3}. See \ref{appendix:allowing_extra_time} for a note on how solution quality varies when increasing solving time.

\input{combined_table_aith_results.tex}

These results suggest that our DST can find better objective values within reasonable planning times. Further, we are able to replicate the 23 optimal solutions provided in the paper. The average results show that GRASP is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.

Similarly, in Table~\ref{tab:group_table_mk_avg_best} we present the comparison of our DST against the performance of the AVNS algorithm presented in \cite{Mankowska2014}.

\begin{table}
	\centering
	\caption{Performance of the GRASP algorithm (average of best and average results) compared to the AVNS algorithm presented in \cite{Mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}}
	\begin{tabular}{c|c|ccc}
		\hline
		        &  \emph{mk\_avns}    &   \multicolumn{3}{c}{GRASP} \\
		Group   &   Result    &   Best        &   Avg         &   Time limit\\ \hline
		\emph{A}    &   \textbf{225.6}  &   226.41  &   226.41      &   5 s\\
		\emph{B}    &   475.1   &   \textbf{408.25} &   408.31      &   30 s\\
		\emph{C}    &   713.6   &   \textbf{628.15} &   637.59      &   1 min\\
		\emph{D}    &   930.3   &   \textbf{798.8}  &   814.53      &   5 min\\
		\emph{E}    &   1064.7  &   \textbf{880.06} &   901         &   15 min\\
		\emph{F}    &   1588    &   \textbf{1405.68}&   1491.09     &   30 min\\
		\emph{G}    &   2161.2  &   \textbf{2042.1} &   2341.89     &   3 h\\

		\hline
	\end{tabular}
	\label{tab:group_table_mk_avg_best}
\end{table}

GRASP clearly outperforms AVNS in the majority of instances, especially on those of larger sizes.
Further, our algorithm is able to replicate most of the proven optimal solutions (9 of the 10 instances for which \cite{Mankowska2014} provides an optimal solution). In the instance group A, where the AVNS finds all optimal solutions for the problem, GRASP fails to replicate one of these solutions, hence increasing the average. This is one of the cases where the assumption that a start time of a job should be as early as possible for the first nurse arriving \eqref{eq:earliest_possible_start_assumption2} proves to be too restrictive and removes the optimal solution from the search space of the metaheuristic. Nevertheless, the loss in quality is under $5\%$ and the route itself found by the two algorithms is the same (but not the scheduled start times). While this effect might still occur in larger instances, it is not evident from our results, suggesting that the efficiency gains seem to outweigh the potential decrease in quality.

\subsection{Trade-offs in solution quality}\label{seq:exploring_obj_function}
While the previous section sheds some light on the algorithmic performance of the GRASP algorithm and, in general, of our ability to solve efficiently HHCRSP problems of size similar to those faced by district nursing teams in the UK, it does not provide any insight on the actual applicability of these solutions.
In this section we focus on the instances from \cite{AitHaddadene2016} that are solved to optimality (with up to 45 patients) and analyse how these solutions would look like from the perspective of a team of nurses undertaking the job. This means asking questions like: Is the workload fairly distributed? How long does it take to complete the shift? Are patients seeing their preferred carers (or vice versa)?

In order to answer these questions, we examine closely the values of a few key indicators of the quality of a solution: \emph{Total time}, \emph{Travel time}, \emph{Waiting time}, \emph{Preferences} and \emph{Longest Day}, which is the time it takes to perform the longest shift in the solution. We compare the indicators found by using the original objective function (listed in Table~\ref{tab:adapted_obj_values}) with those found by using our proposed objective function (the values listed in Table~\ref{tab:adequate_weights}, but with preferences being negative and time quantities scaled by $-0.3\times60$, in order to scale them to the data). The result of this comparison is shown in Table~\ref{tab:comparison_of_qualities}.

\input{"quality_comparison_table.tex"}

The comparison clearly indicates that there is a significant amount of waiting time that can be reduced by rearranging the routes, and it is often not considered in the literature. We show that these waiting times can be reduced, sometimes entirely, and that these reductions have a non-negligible impact in the total time of operation, generating on average a $10.1\%$ reduction, which can be up to $19.7\%$ of the total time in some instances. Further, these gains in time can be spread out evenly, and it is often possible to achieve a reduction of the workload for those staff members that are already doing the most, reducing up to $11.9\%$ their shift time. Nonetheless, these gains come at the expense of a lower matching of preferences (on average, up by $22.9\%$) and travel time (on average, up by $12.7\%$). For this example, we have considered equal weights for travel and waiting times, but it would be up to individual teams to determine appropriate costs for these, considering mileage costs and environmental factors when weighing travel time.

\section{Conclusion}\label{seq:Conclusion}

In this article we propose a Decision Support Tool for the Home Health Care Routing and Scheduling Problem that is daily faced by district nursing teams. We propose a model that includes the most common constraints observed in practice (skills, time windows, synchronisation, double services) and objectives (travel time, waiting time, assignment preferences, workload balance, tardiness, overtime). This comprehensive approach provides flexibility in the modelling of the working environment in order to be applicable in general real settings. This allows for modelling a variety of situations, for example, teams with large workloads might find that overtime and tardiness are acceptable but should be penalised, while other teams with less workload might find those solutions infeasible and prefer to weight more patient preferences instead.
At the core of the DST is a GRASP algorithm with Path Relinking that is capable to solve the underlying NP-Hard problem successfully for the usual instance sizes observed in practice. We show that the DST is competitive and outperforms other methods in the state-of-the-art literature, and it is able to identify most of the optimal solutions known for the problem. 
Furthermore, we show that current literature ignores route aspects that are important to decision makers, such as waiting time. By including waiting times, we can reduce total operation times by up to 19\% when compared to optimal solutions in the literature that minimise travel times only.

\section{Funding}

This article presents independent research partially supported by the National Institute for Health Research (NIHR) Collaboration for Leadership in Applied Health Research and Care (CLAHRC), NIHR Applied Research Collaboration (ARC) and Solent NHS Trust. The views expressed in this publication are those of the author(s) and not necessarily those of the National Health Service, the NIHR or the Department of Health and Social Care. The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.

%% References with bibTeX database:
% \bibliographystyle{model1-num-names}
% \bibliographystyle{apalike}

% \bibliography{library}

\input{bib_section.tex}

\newpage

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections

\pagenumbering{Roman}

\appendix

\input{"appendix_randomly_generated_data.tex"}

\section{Result tables}\label{a:Result_Tables}

Table~\ref{tab:mk_results} shows the results for the instances presented in \cite{Mankowska2014}:
\input{"mk_table.tex"}

Table~\ref{tab:aith_results} shows the results for the instances presented in \cite{AitHaddadene2016}. We have allowed a time of 5 seconds for the instances with $18$ patients, 5 minutes for the instances with $45$ patients and one hour for the instances with $73$ patients.

\input{"aith_table.tex"}

\section{Solving time and solution quality}\label{appendix:allowing_extra_time}
In order to understand better how GRASP performs over time, we performed an extra set of experiments, where GRASP was allowed one, two and three hours on the largest instances from \cite{AitHaddadene2016}. We compared these results to their ILS algorithm, which obtains the best results on large instances. They are shown in Table~\ref{tab:longer_experiments}.
{\footnotesize
\begin{longtable}{c|cc|cc|cc|cc}
\caption{Performance of the GRASP algorithm on the larger instances with more time compared with that of the ILS algorithm from \cite{AitHaddadene2016} (AH ILS)}
\label{tab:longer_experiments}\\
\hline
\textbf{Instance}  & \multicolumn{2}{c|}{\textbf{AH ILS}} & \multicolumn{2}{c|}{\textbf{GRASP 1h}} & \multicolumn{2}{c|}{\textbf{GRASP 2h}} & \multicolumn{2}{c}{\textbf{GRASP 3h}}\\
    & \emph{Best} & \emph{Avg} & \emph{Best} & \emph{Avg} & \emph{Best} & \emph{Avg} & \emph{Best} & \emph{Avg}\\
\hline                                              
    \hline																		
        \emph{73-16-s-2a}   	&	-203.87	&	-188.67	&	-197.40	&	-190.13	&	-207.15	&	-197.71	&	\textbf{-211.72}	&	-200.66	    \\	
        \emph{73-16-s-3}    	&	-203.33	&	-199.93	&	-198.90	&	-195.50	&	-199.37	&	-195.71	&	\textbf{-210.53}	&	-201.72	    \\	
        \emph{73-16-s-2b}   	&	\textbf{-203.87}	&	-188.67	&	-193.73	&	-189.97	&	-203.62	&	-195.57	&	-201.24	&	-197.88	    \\	
        \emph{73-16-m-3a}   	&	\textbf{-202.06}	&	-186.24	&	-197.35	&	-190.74	&	-198.59	&	-195.68	&	-200.41	&	-198.68	    \\	
        \emph{73-16-m-2}    	&	-215.03	&	-205.47	&	-208.36	&	-202.17	&	-213.54	&	-209.12	&	\textbf{-215.7}	&	-210.58	    \\	
        \emph{73-16-m-3b}   	&	\textbf{-221.56}	&	-214.23	&	-210.07	&	-208.09	&	-220.20	&	-213.75	&	-213.76	&	-211.30	    \\	
        \emph{73-16-l-2}    	&	-310.24	&	-303.66	&	-306.13	&	-302.92	&	-308.29	&	-306.80	&	\textbf{-314.53}	&	-310.25	    \\	
        \emph{73-16-l-3}    	&	-311.5	&	-302.23	&	-313.08	&	-306.22	&	-311.65	&	-309.18	&	\textbf{-314.15}	&	-310.65	    \\	
        \emph{73-16-l-4}    	&	-315.24	&	-305.66	&	-310.31	&	-306.78	&	\textbf{-316.3}	&	-311.01	&	-313.76	&	-311.94	    \\	
        \emph{73-16-l-5}    	&	\textbf{-326.79}	&	-313.46	&	-312.32	&	-307.27	&	-315.54	&	-313.38	&	-319.76	&	-316.46	    \\	\hline
\emph{Avg.}	&	-251.35	&	-240.82	&	-244.77	&	-239.98	&	-249.42	&	-244.79	&	\textbf{-251.55}	&	-247.01	    \\

\hline
\end{longtable}
}

These results suggest that allowing more time improves the performance of GRASP, to a quality similar to that of the ILS algorithm. Note that the current practice in district nursing teams is to plan work one day in advance, so times of up to three hours for running the algorithm are feasible in a practical setting. The average results also improve with more time, and are often better than the averages of ILS, suggesting GRASP is more stable,

\end{document}

### 2021-09-09 tex/Paper/appendix_randomly_generated_data.tex
files for paper.

@@
 %!TEX root=./main_paper.tex
-\section{Randomly Generated Instances}\label{appendix:RandomlyGeneratedData}
-
-The generation of random data was done by creating a list of available staff and a list of tasks that required to be performed by the qualified staff. 
+\section{Randomly Generated Instances}
+\label{appendix:randomlygenerateddata}
+\noindent The generation of random data was done by creating a list of available staff and a list of tasks that required to be performed by the qualified staff. 
 For each of these, we define a number of properties that are important for our model.
@@
 
-The instances used in this article were generated with the parameters described in Table~\ref{Appendix:Tab:InstanceDescription}, where each combination of parameters led to one instance, a total of 216.
+The instances used in this article were generated with the parameters described in Table~\ref{appendix:table:instancedescription}, where each combination of parameters led to one instance, a total of 216.
 The exact instances used are available here: \url{https://github.com/c-lamas/instances_hhcrsp}.
@@
 		Proportion of jobs with interdependencies & 10\%, 20\%\\ \hline
-	\end{tabular}\label{Appendix:Tab:InstanceDescription}
+	\end{tabular}\label{appendix:table:instancedescription}
 \end{table}
 
-The skill mix represented the diversity of skills in the workforce, as well as the broader or narrower range of tasks needed by patients.
+\noindent The skill mix represented the diversity of skills in the workforce, as well as the broader or narrower range of tasks needed by patients.
 To mimic real settings, the workforce is organised in different bands.
 Each bad has at least one skill, and the probability of having further skills is higher in higher bands.
-The particular configuration of each skill mix used is described in Table~\ref{Appendix:Tab:SkillMixConfiguration}.
+The particular configuration of each skill mix used is described in Table~\ref{appendix:table:skillmixconfiguration}.
 The nurses were spread evenly across the bands, with more nurses in the lowest band if the total number of staff was non-divisible by the number of bands used.
@@
 		one or more skills   & 				 & & 3 Skills: 5\% \\ \hline
-	\end{tabular}\label{Appendix:Tab:SkillMixConfiguration}
+	\end{tabular}\label{appendix:table:skillmixconfiguration}
 \end{table}

### 2021-09-09 tex/Paper/main.tex
main file for paper.

@@
-\documentclass[a4paper,11pt]{elsarticle}
-
+\documentclass[a4paper,11pt,authoryear]{elsarticle}
 \input{includes/preamble.tex}
 
-\biboptions{authoryear}
-
-\journal{European Journal of Operational Research}
-
+\begin{comment}
+\begin{itemize}
+    \item Change all terms `nurse(s)' to `carer(s)' and `patient(s)' to `client(s)'
+\end{itemize}
+\end{comment}
 \begin{document}
@@
 
-%% Title, authors and addresses
-
 \title{A decision support tool for the Home Health Care Routing and Scheduling Problem}
 
-\author[sbs]{Carlos Lamas-Fernandez$^{*,}$}
-\author[exeter]{Thomas Monks}
-\author[sbs]{Antonio Martinez-Sykora}
+\author[1]{Carlos Lamas-Fernandez\corref{cor1}}
+\ead{C.Lamas-Fernandez@soton.ac.uk}
+\cortext[cor1]{Corresponding author}
+\address[1]{Southampton Business School, University of Southampton, Southampton SO17 1BJ, UK}
 
-\address[sbs]{Southampton Business School, University of Southampton, Southampton SO17 1BJ, UK}
-\address[exeter]{University of Exeter Medical School, University of Exeter, Exeter, UK}
+\author[2]{Thomas Monks}
+\ead{T.M.W.Monks@exeter.ac.uk}
+\address[2]{University of Exeter Medical School, University of Exeter, Exeter, UK}
+\author[1]{Antonio Martinez-Sykora}
+\ead{A.Martinez-Sykora@soton.ac.uk}
 
 \begin{abstract}
-%% Text of abstract
 The Home Health Care Routing and Scheduling Problem (HHCRSP) consists in assigning routes and schedules to nurses that deliver health care services to patients at their homes.
@@
 OR in health services \sep Metaheuristics \sep Routing \sep Scheduling \sep GRASP
-%% keywords here, in the form: keyword \sep keyword
+\end{keyword}
 
-%% MSC codes here, in the form: \MSC code \sep code
-%% or \MSC[2008] code \sep code (2000 is the default)
-
-\end{keyword}
+\newpageafter{abstract}
 
@@
 
-{\footnotesize
-\noindent $^*$ Corresponding author\\
-\noindent E-mail addresses: \text{C.Lamas-Fernandez@soton.ac.uk} (C. Lamas-Fernandez), \text{T.M.W.Monks@exeter.ac.uk} (T. Monks), \text{A.Martinez-Sykora@soton.ac.uk} (A. Martinez-Sykora)
-}
+%%
+%% Start line numbering here if you want
+%%
+%% \linenumbers
 
-\newpage
-
-\section{Introduction}\label{seq:intro}
-
-Home health care refers to a form of care where a health professional visits a patient in their own home to perform a health-related task \citep{Cisse2017}.
-In the European context, with an ever-ageing population, delivering health care at home is increasingly seen as a necessary alternative to traditional forms of care \citep{tarricone2008home}.
+%----------------------------------------------------------------------------------%
+\section{Introduction}
+\label{sec:intro}
+\noindent Home health care refers to a form of care where a health professional visits a patient in their own home to perform a health-related task \citep{cisse2017}.
+In the European context, with an ever-ageing population, delivering health care at home is increasingly seen as a necessary alternative to traditional forms of care \citep{tarricone2008}.
 This view is shared by many health providers around the globe as managing long term conditions, frail patients and end-of-life care among others might be done more efficiently in the patient's own home.
-For example, in the OECD countries, the care at home for patients with long term conditions increased by 6\% \citep{OECDHealthataGlance2019} and in the UK the National Health Service (NHS) stated in their long term plan that in the future more care should be delivered to patients in their own homes \cite{england2019nhs}.
+For example, in the OECD countries, the care at home for patients with long term conditions increased by 6\% \citep{oecd2019} and in the UK the National Health Service (NHS) stated in their long term plan that in the future more care should be delivered to patients in their own homes \citep{nhs2019}.
 
 A key role in the health service to deliver this care are District Nurses (DN), specialised nursing teams that travel to patient's homes and perform a variety of tasks, ranging from insulin injections to dressing wounds.
-Nevertheless, DN numbers have fallen by 44\% \cite{NHSProviders2018} between 2010 and 2018 in the NHS.
-On the other hand, while there is no official data on demand, the workforce perception is that demand has risen significantly, putting the service under pressure \citep{robertson2017understanding}.
+Nevertheless, DN numbers have fallen by 44\% \citep{nhs2018} between 2010 and 2018 in the NHS.
+On the other hand, while there is no official data on demand, the workforce perception is that demand has risen significantly, putting the service under pressure \citep{robertson2017}.
 There is, therefore, a strong need for efficiency within the community nursing service, to better utilise the available resources.
@@
 This is motivated by the last-minute changes due to staff availability, patient next-day referrals and the nature of continuous review of care, which often prompts to scheduling the next visit right after reviewing their condition on the previous visit.
-A formal description of the problem is provided in Section~\ref{seq:model description}.
+A formal description of the problem is provided in Section~\ref{sec:model}.
 
@@
 The tool requires as input a patient requirement list and the available staff and then runs an optimisation algorithm based on a weighted objective function.
-The optimisation is based on a Greedy Randomised Adaptive Search Procedure (GRASP) with Path Relinking, which is detailed in Section~\ref{seq:DSS}.
+The optimisation is based on a Greedy Randomised Adaptive Search Procedure (GRASP) with Path Relinking, which is detailed in Section~\ref{sec:dst}.
 
@@
 multiple depots or workload balance).
-We provide extensive computational results (Section~\ref{seq:Computational Experiments}) that show that our algorithm is competitive with state-of-the-art approaches and is flexible to allow real-world constraints and objectives; but also provide insight on the algorithm components that help achieve high quality routes and schedules.
+We provide extensive computational results (Section~\ref{sec:computationalexperiments}) that show that our algorithm is competitive with state-of-the-art approaches and is flexible to allow real-world constraints and objectives; but also provide insight on the algorithm components that help achieve high quality routes and schedules.
 
+%----------------------------------------------------------------------------------%
 \section{Literature review}
-The generalised growth in demand and availability of home health care services of recent years has resulted in a growth in academic interest in the related scheduling problem.
-While the problem might be addressed by drawing from the extensive vehicle routing problem (VRP) literature (see, for example, \cite{Drexl2012} for a comprehensive review VRPs and synchronisation constraints), HHCRSP has developed its own body of literature.
-In particular, there are two recent review articles, \cite{Cisse2017} and \cite{Fikar2017}.
-We refer the reader to them for a broader understanding of what it is meant by HHCRSP and different solving approaches.
+\label{sec:litreview}
+\noindent The generalised growth in demand and availability of home health care services of recent years has resulted in a growth in academic interest in the related scheduling problem.
+While the problem might be addressed by drawing from the extensive vehicle routing problem (VRP) literature (see, for example, \cite{drexl2012} for a comprehensive review of VRPs and synchronisation constraints), HHCRSP has developed its own body of literature.
+In particular, there are two recent review articles, \cite{cisse2017} and \cite{fikar2017}.
+We refer the reader to them for a broader understanding of what is meant by HHCRSP and different solving approaches.
 Home Health Care and Home Care services across the world operate in different ways and, as such, the related scheduling problems have different objectives and constraints.
-For example, it might be possible to speed up the visits to some patients \citep{Mosquera2018}, to allow nurses to share a minibus for part of their route, while walking the rest of it \citep{Fikar2015}, for nurses to use public transport \citep{Rest2016}, or to allow time preferences for patients \citep{Decerle2018}.
+For example, it might be possible to speed up the visits to some patients \citep{mosquera2019}, to allow nurses to share a minibus for part of their route, while walking the rest of it \citep{fikar2015}, for nurses to use public transport \citep{rest2016}, or to allow time preferences for patients \citep{decerle2018}.
 % This variety makes difficult comparing approaches and translating the findings from one application to another.
@@
 Several articles propose exact methods.
-These are usually based on a Mixed Integer Linear Program (MILP) formulation that is solved with the help of a commercial solver \citep{Bredstrom2008,Bachouch2011, Mankowska2014, AitHaddadene2016}.
+These are usually based on a Mixed Integer Linear Program (MILP) formulation that is solved with the help of a commercial solver \citep{bredstrom2008,bachouch2011, mankowska2014, aithaddadene2016}.
 Typically, these formulations are three index formulations that involve two types of variables: $x_{ijk}$ binary variables that indicate if carer $k$ traverses the arc $i$ and $j$; and $t_{ik}$ real variables that determine at what time carer $k$ starts performing job $i$.
-These formulations are limited in practice and can only solve to optimality small problems (\eg 45 patients in \cite{AitHaddadene2016} or 10 in \cite{Mankowska2014}).
+These formulations are limited in practice and can only solve to optimality small problems (e.g.
+45 patients in \cite{aithaddadene2016} or 10 in \cite{mankowska2014}).
 In order to address this, some researchers propose to decompose the problem in a column generation scheme.
-For example, \cite{Rasmussen2012} tackles the HHCRSP problem in Denmark.
+For example, \cite{rasmussen2012} tackles the HHCRSP problem in Denmark.
 They propose a branch-and-price scheme where generalised synchronisation constraints are relaxed and dealt with in the branching.
@@
 This is often referred to as a matheuristic algorithm and can be done by reducing the size of the original problem by fixing some aspects and solving it again.
-One example of this technique is given in \cite{Rasmussen2012} that also proposes clustering strategies to reduce the size of the problem, which allows them to solve problems with up to $150$ patients, although no longer proving optimality.
-Another is \cite{Bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time.
+One example of this technique is given in \cite{rasmussen2012} that also proposes clustering strategies to reduce the size of the problem, which allows them to solve problems with up to $150$ patients, although no longer proving optimality.
+Another is \cite{bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time.
 
 Another approach, which is the one we use in this article, is to propose a metaheuristic algorithm independent of a mathematical solver.
-\cite{Mankowska2014} and \cite{AitHaddadene2016} propose an Adaptive Large Neighbourhood search and a hybrid GRASP combined with Iterated Local Search respectively.
+\cite{mankowska2014} and \cite{aithaddadene2016} propose an Adaptive Large Neighbourhood search and a hybrid GRASP combined with Iterated Local Search respectively.
 One advantage of the use of heuristic algorithms is that they do not require availability of a (often commercial) MILP solver, which is unlikely to be available to health care professionals.
@@
 Finally, we look at the literature that discusses the software implementation of the solution approaches in practice.
-One of the earliest works is due to \cite{Begur1997}, who developed a decision support system based on the Clarke and Wright algorithm \citep{Clarke1964} for vehicle routing and then relied on integration with a Geographic Information System (GIS), an interactive user interface and a human planner to alter the routes as necessary.
-The Laps Care system \citep{Eveborn2006, Eveborn2009a} is a Swedish software for the HHCRSP, based on a repeated matching algorithm has been shown to be successful in saving planning time, minimising travel times and even reducing the number of missed visits.
-The authors estimated annual savings in 2009 of up to €120 M across teams in Norway, Finland and Sweden \citep{Eveborn2009a}.
-Another software is Parpap \citep{BERTELS20062866}, which proposes various approaches combining constraint programming and metaheuristics.
+One of the earliest works is due to \cite{begur1997}, who developed a decision support system based on the Clarke and Wright algorithm \citep{clarke1964} for vehicle routing and then relied on integration with a Geographic Information System (GIS), an interactive user interface and a human planner to alter the routes as necessary.
+The Laps Care system \citep{eveborn2006, eveborn2009} is a Swedish software for the HHCRSP, based on a repeated matching algorithm has been shown to be successful in saving planning time, minimising travel times and even reducing the number of missed visits.
+The authors estimated annual savings in 2009 of up to \euro 120 million across teams in Norway, Finland and Sweden \citep{eveborn2009}.
+Another software is Parpap \citep{bertels2006}, which proposes various approaches combining constraint programming and metaheuristics.
 A common claim in all these implementation-oriented works is that the software needs to be flexible enough to allow for real-world constraints to be implemented and offers interaction with mapping software.
 
-%%%%%%%%%%%%%%%%% MODEL FOR HHCRSP %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{A model for health home care services}\label{seq:model description}
-
-In this section we describe the notation and give a formal description of the HHCRSP problem tackled by our DST.
+%----------------------------------------------------------------------------------%
+\section{A model for health home care services}
+\label{sec:model}
+\noindent In this section we describe the notation and give a formal description of the HHCRSP problem tackled by our DST.
 
@@
 Each nurse $i$ might have a different starting time, $st(i)$, and end time, $en(i)$, of their working day.
-During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$, a subset of which, $DS \subset J$ are double services, \ie, they require two nurses to be present to be performed.
+During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$, a subset of which, $DS \subset J$ are double services, i.e.
+they require two nurses to be present to be performed.
 Note that jobs refer to individual tasks that a nurse needs to perform and are not necessarily identified with unique patients or locations, meaning that the same patient might have more than one job on the list.
@@
 
+%----------------------------------------------------------------------------------%
 \subsection{Skill mix}
-Some jobs might require a specialised nurse to be performed, and nurses might only perform jobs that they are qualified to do.
-To model this, we define a function $S: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise.
+\label{sub:skillmix}
+\noindent Some jobs might require a specialised nurse to be performed, and nurses might only perform jobs that they are qualified to do.
+To model this, we define a function $S_1: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise.
 Naturally, for double service jobs, no single nurse is skilled to perform them $S_1(i,j) = 0, \forall i \in N, \forall j \in DS$.
@@
 In this case each member of the service does not need to be qualified for all services but should be attending the visit to perform at least one service.
-This is the approach used in \cite{Mankowska2014} and \cite{AitHaddadene2016}.
+This is the approach used in \cite{mankowska2014} and \cite{aithaddadene2016}.
     \item \textbf{Shared-Duplicated}. This case is analogous to \emph{Shared}, but it allows some skills to be required by more than one of the staff members, hence it generalises the previous cases.
@@
 
-In our DST, the default approach is \emph{Shared-Duplicated}, as it allows for the flexibility that is observed in real operations.
+\noindent In our DST, the default approach is \emph{Shared-Duplicated}, as it allows for the flexibility that is observed in real operations.
 The data is introduced as a list of skills and, if some skills need to be present in more than one staff member, they can simply appear duplicated on the list.
@@
 
-% \subsubsection{Hierarchical skills and downgrading}
-It is not uncommon in the literature to model the skills of the workforce as hierarchical, \ie each nurse is assigned a skill level and it is assumed that they possess all the skills that lower levels have, and some more.
-The model might then introduce a penalty or limit for high skilled nurses performing too many low skilled jobs, known as downgrading \cite[see, \eg][]{Fikar2015}.
+It is not uncommon in the literature to model the skills of the workforce as hierarchical, i.e.
+each nurse is assigned a skill level and it is assumed that they possess all the skills that lower levels have, and some more.
+The model might then introduce a penalty or limit for high skilled nurses performing too many low skilled jobs, known as downgrading \cite[see, for example,][]{fikar2015}.
 In the DN teams we observed, skills were not hierarchical, neither were any such limits in place.
-Nevertheless, nurses pointed out that high skilled nurses were typically assigned only the high skilled jobs, as they might need to perform other jobs (\eg attending meetings or other administrative tasks).
+Nevertheless, nurses pointed out that high skilled nurses were typically assigned only the high skilled jobs, as they might need to perform other jobs (e.g.
+attending meetings or other administrative tasks).
 While this is not something that we consider explicitly, a similar objective may be achieved through setting negative preferences between high skilled nurses and low skilled jobs.
 
+%----------------------------------------------------------------------------------%
 \subsection{Preferences}
-It is not uncommon in practice that district nurses do their planning using preferences.
-These might arise from multiple reasons, such as a preserving continuity of care, gender preference, patients and nurses speaking the same language, previous conflicts between a patient and a staff member, etc.
-We model this with a preference function, $P: N\times J \rightarrow \mathbb{R}$ such that $P(i,j)$ represents the preference score of assigning job $i$ to staff member $j$.
+\label{sub:preferences}
+\noindent It is not uncommon in practice that district nurses do their planning using preferences.
+These might arise from multiple reasons, such as a preserving continuity of care, gender preference, patients and nurses speaking the same language, previous conflicts between a patient and a staff member, \model{etc}.
+We model this with a preference function, $P: N\times J \rightarrow \mathbb{R}$, such that $P(i,j)$ represents the preference score of assigning job $i$ to staff member $j$.
 The function $P$ should assign a higher (positive) score to preferred patient-nurse assignments and a lower (negative) score to assignments that are better avoided.
-We give an indication on how to choose the specific values of the function in Section~\ref{seq:adequate_weights}.
+We give an indication on how to choose the specific values of the function in Section~\ref{subsub:adequateweights}.
 
+%----------------------------------------------------------------------------------%
 \subsection{Travel times}
-To describe travel times, for each nurse we define $N$ functions $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, where $k^i_0$ refers to the start and end location of the nurse $i$.
+\label{sub:traveltimes}
+\noindent To describe travel times, for each nurse we define $N$ functions $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, where $k^i_0$ refers to the start and end location of the nurse $i$.
 Each nurse might start from a different depot (typically, their own home), a fact that is accounted for in these functions.
 Note that with this definition we allow to have different travel times between jobs for different nurses, and this can be used to model, for example, different means of transport.
-In our observations, however, all nurses used their own vehicles, so we assume that $T_i$ returns driving times for all nurses and, in particular that:
+In our observations, however, all nurses used their own vehicles, so we assume that $T_i$ returns driving times for all nurses and, in particular, that:
 \begin{equation}
-    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2),\ \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J
+    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2),\ \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J. \label{eqn:traveltimequality}
 \end{equation}
 
-Note that this function can be substituted by a more sophisticated one without any changes to the algorithms described here.
+\noindent Note that this function can be substituted by a more sophisticated one without any changes to the algorithms described here.
 
@@
 
-\subsection{Temporal constraints}\label{seq:sub temporal constraints}
-Each job $j$ has a service time $s_j$ and a time window that is defined by its earliest start time, $e_j$ and the latest start time $l_j$.
-
-It is a hard constraint that jobs cannot start before their earliest start time,
+%----------------------------------------------------------------------------------%
+\subsection{Temporal constraints}
+\label{sub:temporalconstraints}
+\noindent Each job $j$ has a service time $s_j$ and a time window $t_j$ that is defined by its earliest start time, $e_j$ and the latest start time $l_j$.
+It is a hard constraint that jobs cannot start before their earliest start time:
 \begin{align}
-    t_j \geq e_j, & \forall j \in J
+    t_j \geq e_j, & \forall j \in J, \label{eqn:jobstart}
 \end{align}
 
-However, they might start late. The tardiness with which a job $j$ starts, $z_j$ can be simply calculated by:
+\noindent however, they may start late. The tardiness with which a job $j$ starts, $z_j$, can be simply calculated using the following:
 \begin{equation}
-    z_j = \max(0, t_j - l_j)
+    z_j = \max(0, t_j - l_j). \label{eqn:tardiness}
 \end{equation}
 
-Note that tardiness here refers only to the difference between the job start and its original time window.
+\noindent Note that tardiness here refers only to the difference between the job start and its original time window.
 Jobs that have dependencies or require two nurses might be further restricted temporally.
-This might happen in practice, for example, in situations where a certain medication must be administered a few hours after the first dose.
-If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$, that depends on the start time of $j_1$, $[e^{(j_1)}_{j_2} + t_{j_1}, l^{(j_1)}_{j_2} + t_{j_1}]$, then, the same time window constraints hold:
+This might happen in practice, for example, in situations where a certain medication must be administered \model{a few hours (a specific time period/amount of time)} after the first dose.
+If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$ that depends on the start time of $j_1$, $[e^{(j_1)}_{j_2} + t_{j_1}, l^{(j_1)}_{j_2} + t_{j_1}]$.
+Then, the same time window constraints hold:
+\begin{subequations}
+	\begin{align}
+	t_{j_2} &\geq e^{(j_1)}_{j_2} + t_{j_1}, \label{eqn:timewindow1}\\[3pt] % & \forall j \in J
+	t_{j_2} &\leq l^{(j_1)}_{j_2} + t_{j_1}. \label{eqn:timewindow2} % & \forall j \in J
+	\end{align}
+\end{subequations}
+
+\noindent Similarly, $j_2$ will impose a time window on $j_1$, and in a valid instance of the problem it must hold that:
+\begin{subequations}
+	\begin{align}
+	e^{(j_1)}_{j_2} &= -e^{(j_2)}_{j_1}, \label{eqn:earliesttimedependecies}\\[3pt]
+	l^{(j_1)}_{j_2} &= -l^{(j_2)}_{j_1}. \label{eqn:latesttimedependencies}
+	\end{align}
+\end{subequations}
+
+\noindent Then, we can define the \emph{gap tardiness} of job $j_2$, $\tilde{z}_{j_2}$ as:
 \begin{equation}
-    t_{j_2} \geq e^{(j_1)}_{j_2} + t_{j_1}, % & \forall j \in J
-\end{equation}
-\begin{equation}
-    t_{j_2} \leq l^{(j_1)}_{j_2} + t_{j_1}, % & \forall j \in J
-\end{equation}
-Similarly, $j_2$ will impose a time window on $j_1$, and in a valid instance of the problem it must hold that:
-\begin{equation}
-    e^{(j_1)}_{j_2} = -e^{(j_2)}_{j_1}
-\end{equation}
-and
-\begin{equation}
-    l^{(j_1)}_{j_2} = -l^{(j_2)}_{j_1}
+    \tilde{z}_{j_2} = \max(0, t_{j_2} - (l^{(j_1)}_{j_2} + t_{j_1})). \label{eqn:gaptardiness}
 \end{equation}
 
-Then, we can define the \emph{gap tardiness} of job $j_2$, $\tilde{z}_{j_2}$ as:
+\noindent \model{Since some jobs have time windows (only \textit{some}?)} or might need to be synchronised, waiting times may occur.
+We denote by $w_{ij}$ the waiting time of nurse $i$ at job $j$.
+Note that $w_{ij} = 0$ if $j \notin r_i$; otherwise, $w_{ij}$ represents the difference between $t_j$ and the arrival time of nurse $i$ to job $j$.% , denoted by $a(i,j)$.
+
+The arrival time $a(i,j)$ of nurse $i$ at job $j \in r_i$ is calculated as follows:
 \begin{equation}
-    \tilde{z}_{j_2} = \max(0, t_{j_2} - (l^{(j_1)}_{j_2} + t_{j_1})]
+    a(i,j) = st(i) + \sum_{h=0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1}),
 \end{equation}
 
-Since some jobs have time windows or might need to be synchronised, waiting times might occur.
-We denote by $w_{ij}$ the waiting time of nurse $i$ at job $j$.
-Note that $w_{ij} = 0$ if $j \notin r_i$ and, otherwise, $w_{ij}$ represents the difference between $t_j$ and the arrival time of nurse $i$ to job $j$.% , denoted by $a(i,j)$.
+\noindent where $j'$ denotes the position of job $j$ in the route $r_i$, i.e.
+$j = k^i_{j'} \in r_i$.
+Note that there are no waiting or service times at the depot; that is, $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 $ for all $i \in N$.
 
-The arrival time $a(i,j)$ of the nurse $i$ at the job $j \in r_i$ is calculated as follows:
-
-\begin{equation}
-    a(i,j) = st(i) + \sum_{h \in 0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1})
-\end{equation}
-Where $j'$ denotes the position of job $j$ in the route $r_i$, \ie $j = k^i_{j'} \in r_i$.
-Note that there is not waiting or service times at the depot, \ie $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 \ \forall i \in \mathcal{N}$.
-
-In practice, waiting times might occur for a variety of reasons.
-In order to maintain simplicity in our model, we only consider three reasons for which a nurse $i$ might have waiting time $w_{ij} \geq 0$ at the job $j \in r_i$:
+In practice, waiting times may occur for a variety of reasons.
+In order to maintain simplicity in our model, we only consider three reasons for which a nurse $i$ might have waiting time $w_{ij} \geq 0$ at a job $j \in r_i$:
 \begin{itemize}
@@
     \begin{align}
-        w_{ij} \geq e_i - a(i,j) & \forall i \in N, \forall j \in J
+        w_{ij} \geq e_i - a(i,j) & \forall i \in N, \forall j \in J.
     \end{align}
-    \item Waiting for a second nurse $i'$ to arrive for a double service
+    \item Waiting for a second nurse $i'$ to arrive for a double service:
     \begin{align}
-        w_{i j} \geq t_j - a(i',j) & \forall j \in DS,\ such\ that\ j \in r_{i} \cap r_{i'}
+        w_{i j} \geq t_j - a(i',j) & \forall j \in DS,\ such\ that\ j \in r_{i} \cap r_{i'}.
     \end{align}
     \item Waiting for the minimum gap of a synchronised service. If the job $j$ depends on $j'$, then:
-    \begin{equation}\label{eq:waiting_for_synchronised}
-        w_{ij} \geq e^{(j')}_{j} - a(i,j)
+    \begin{equation}\label{eqn:waitingforsynchronised}
+        w_{ij} \geq e^{(j')}_{j} - a(i,j).
     \end{equation}
@@
 
-Further, we assume that waiting times are never larger than strictly necessary.
-To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based  in the order in which they appear in the solution.
-Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, and such that $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$ and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$.
+\noindent Furthermore, we assume that waiting times are never larger than strictly necessary.
+To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based on the order in which they appear in the solution.
+\correction{The equations below are for a single nurse/job only} Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, where $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$, and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$.
 Then, we enforce that:
 %  $i_1 < i_2$, \ie $j_1$ appears earlier in the vector $\mathcal{O}$ than $j_2$. Then, we enforce that:
-\begin{equation}\label{eq:earliest_possible_start_assumption}
-    w_{i_1 j_1} = \max \{ a(i,j_1), e_{j_1} \}
+\begin{equation}\label{eqn:earliestpossiblestartassumption1}
+    w_{i_1 j_1} = \max \{0, e_{j_1} - a(i_1,j_1)\}
+\end{equation}
+and
+\begin{equation}\label{eqn:earliestpossiblestartassumption2}
+    t_{j_1} = \max \{ a(i_1,j_1), e_{j_1} \}
 \end{equation}
 
-and
+\noindent Note that this implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
+To mitigate this effect, in Section~\ref{sec:dst} we describe strategies to search different orderings of nurses when solving the problem.
 
-\begin{equation}\label{eq:earliest_possible_start_assumption2}
-    t_{j_1} = \max \{ a(i,j_1), e_{j_1} \}
-\end{equation}
-
-Note that this implies that the order in which nurse routes appear in a solution may have an impact in the solution quality.
-To mitigate this effect, in Section~\ref{seq:DSS} we describe strategies to search different orderings of nurses when solving the problem.
-
-\subsection{Objective function}
+%----------------------------------------------------------------------------------%
+\subsection{Objective Function}
+\label{sub:objectivefunc}
 % The objective function is based on the quality of a solution, which consists on a few measures of interest that are added together with different weights.
 Performance measures to be maximised have positive weights, while those to be minimised have negative weights.
-We consider a weighted objective function that includes waiting time, travel time, tardiness, maximum tardiness, overtime, workload balance and preferences.
+\noindent We consider a weighted objective function that includes waiting time, travel time, tardiness, maximum tardiness, overtime, workload balance and preferences.
 The measures to be maximised have positive weights, while those to be minimised have negative weights.
@@
 \begin{equation}
-     \bar{W} = \sum_{i \in \mathcal{N}} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}
+     \bar{W} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}.
 \end{equation}
 
-\paragraph{Total Travel time}\ \\
-
+\paragraph{Total travel time}\ \\
 If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
 \begin{equation}
-    TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1})
+    TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1}),
 \end{equation}
-The total travel time in $\mathcal{S}$, $\bar{T}$ is calculated as:
+then the total travel time in $\mathcal{S}$, $\bar{T}$, is calculated as:
 \begin{equation}
-     \bar{T} = \sum_{i \in \mathcal{N}}  TR(r_i)
+     \bar{T} = \sum_{i \in N}  TR(r_i).
 \end{equation}
 
-\paragraph{Total Tardiness}\ \\
-The total tardiness in a solution, is simply the sum of the tardiness across all jobs,
+\paragraph{Total tardiness}\ \\
+The total tardiness in a solution, $\bar{Z}$, is simply the sum of the tardiness across all jobs:
 \begin{equation}
-    \bar{Z} = \sum_{j \in \mathcal{J}} z_j
+    \bar{Z} = \sum_{j \in J} z_j.
 \end{equation}
 
-\paragraph{Maximum Tardiness}\ \\
+\paragraph{Maximum tardiness}\ \\
 The maximum tardiness in a solution, $Z_{max}$, is given by:
 \begin{equation}
-    Z_{max} = \max_{j \in \mathcal{J}} z_j
+    Z_{max} = \max_{j \in J} z_j
 \end{equation}
@@
 The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$.
-If we denote the arrival of a nurse $i$ to its end point by $arr(i)$,
+If we denote the arrival of a nurse $i$ to its end point by $arr(i)$:
 \begin{equation}
-	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k
+	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k,
 \end{equation}
-Then, the total overtime can be calculated as
+then the total overtime can be calculated as:
 \begin{equation}
-     \bar{O} = \sum_{i \in \mathcal{N}} \max \{0, arr(i) - en(i)\}
+     \bar{O} = \sum_{i \in N} \max \{0, arr(i) - en(i)\}.
 \end{equation}
 
-\paragraph{Workload Balance}\ \\
-
+\paragraph{Workload balance}\ \\
 In order to produce practical solutions, maintaining a balance in the workload is fundamental.
@@
 If $\bar{B}$ is minimised, we favour solutions in which the lowest spare time of any given nurse is as large as possible, hence removing imbalance when possible.
-
 \begin{equation}
-    \bar{B} = \min_{i \in \mathcal{N}} \left\{\max \{0, en(i) - arr(i)\} \right\}
+    \bar{B} = \min_{i \in N} \left\{\max \{0, en(i) - arr(i)\} \right\}.
 \end{equation}
 
-\paragraph{Total Preference Score}\ \\
-
+\paragraph{Total preference score}\ \\
 The total preference score in the solution, $\bar{P}$, is computed as:
 \begin{equation}
-    \bar{P} = \sum_{i \in \mathcal{N}} \sum_{j \in r_i \subset \mathcal{R}} P(i,j)
+    \bar{P} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{R}} P(i,j).
 \end{equation}
-
-Finally, the full quality of the solution, $q(\mathcal{S})$ is then obtained as:
-\begin{equation}\label{eq:solution_quality}
+Finally, the full quality of the solution, $q(\mathcal{S})$, is then obtained as:
+\begin{equation}
     q(\mathcal{S}) = \alpha_1 \bar{T} + % Travel time 
@@
          \alpha_6 \bar{P} + % Preference score 
-         \alpha_7 Z_{max} 
+         \alpha_7 Z_{max}. \label{eqn:solutionquality}
 \end{equation}
 
-Note that some of the terms considered in the quality function \eqref{eq:solution_quality} might be considered infeasible in some settings (\eg overtime and tardiness).
+\noindent Note that some of the terms considered in the quality function \eqref{eqn:solutionquality} might be considered infeasible in some settings (e.g.
+overtime and tardiness).
 For this reason, the objective function that is used during the optimisation, $Q(\mathcal{S})$, is defined as follows:
-
-\begin{equation}\label{eq:real_obj_function}
+\begin{equation}
   \max Q(\mathcal{S}) =
@@
    q(\ensuremath{\mathcal{S}}) & \text{if \ensuremath{I(\mathcal{S}) \leq 0}} \\
-   q(\ensuremath{\mathcal{S}}) + \ensuremath{M_1 I(\mathcal{S})} + \ensuremath{M_2} & \text{if \ensuremath{I(\mathcal{S}) > 0}} \\
+   q(\ensuremath{\mathcal{S}}) + \ensuremath{M_1 I(\mathcal{S})} + \ensuremath{M_2} & \text{if \ensuremath{I(\mathcal{S}) > 0}}, \\ \label{eqn:realobjfunction}
   \end{cases}
@@
 
-where $I(\mathcal{S})$ represents the value of the infeasibility in the solution and hence solutions where $I(\mathcal{S}) > 0$ are considered infeasible.
-The measures that are included in the calculation of $I(\mathcal{S})$ might vary depending on the team being modelled.
-Typically, they include the total gap tardiness, $\tilde{Z}$:
+\noindent where $I(\mathcal{S})$ represents the value of the infeasibility in the solution; hence solutions where $I(\mathcal{S}) > 0$ are considered infeasible.
+The measures that are included in the calculation of $I(\mathcal{S})$ may vary depending on the \model{team (instance?)} being modelled.
+Typically, \model{they (?be specific)} include the total gap tardiness, $\tilde{Z}$:
 \begin{equation}
-    \tilde{Z} = \sum_{j \in \mathcal{J}} \tilde{z}_j
+    \tilde{Z} = \sum_{j \in J} \tilde{z}_j,
 \end{equation}
- and often the overtime in the solution, $\bar{O}$ and the tardiness in the solution, $\bar{Z}$.
+ and often the overtime in the solution, $\bar{O}$, and the tardiness in the solution, $\bar{Z}$.
 For example, if no overtime or gap tardiness is allowed, $I(\mathcal{S})$ might be defined as $I(\mathcal{S}) = \bar{O} + \tilde{Z}$.
 
-The terms $M_1$ and $M_2$ represent an arbitrary large value, which is used to guide the algorithm towards feasible solutions.
-We include $M_1$ to drive the search towards solutions with smaller infeasibility values and $M_2$ to prevent accepting infeasible solutions when the value of $I(\mathcal{S})$ is very small.
+The terms $M_1$ and $M_2$ represent an arbitrary large value which is used to guide the algorithm towards feasible solutions.
+We include $M_1$ to drive the search towards solutions with smaller infeasibility values, and $M_2$ to prevent accepting infeasible solutions when the value of $I(\mathcal{S})$ is very small.
 In our experiments we have used $M_1 = M_2 = -10^3$.
 
-\subsubsection{Setting adequate weights}\label{seq:adequate_weights}
-
-Having different measures on the objective function, some of which might represent quantities of different nature (\eg waiting time and preference of assignment), raises the question on how the weights should be chosen.
+%----------------------------------------------------------------------------------%
+\subsubsection{Setting adequate weights}
+\label{subsub:adequateweights}
+\noindent Having different measures on the objective function, some of which might represent quantities of different nature (e.g.
+waiting time and preference of assignment), raises the question on how the weights should be chosen.
 While the way to do this is better assessed by practitioners, we provide a simple guidance on how to calibrate these weights using time as a reference.
-As a base measure, we use the total time used to provide care to patients, \ie the sum of travel time and waiting time (service time is not included, as it does not depend on the scheduling).
+As a base measure, we use the total time used to provide care to patients, i.e.
+the sum of travel time and waiting time (service time is not included as it does not depend on the scheduling).
 As such, their weights are set to one, $\alpha_1 = \alpha_2 = 1$.
 Now, the remaining weights are set according to their relative importance to this base measure.
-This means, answering the question: How many minutes of total time would we be willing to sacrifice to improve this other measure by one minute?
+This means answering the question: How many minutes of total time would we be willing to sacrifice to improve this other measure by one minute?
 The answer to this question indicates the weight that should be assigned to that measure.
@@
 
-The weights we propose for the Decision Support Tool are listed in Table~\ref{tab:adequate_weights}.
+The weights we propose for the Decision Support Tool are listed in Table~\ref{table:adequateweights}.
 
@@
     \end{tabular}
-    \label{tab:adequate_weights}
+    \label{table:adequateweights}
 \end{table}
 
-Note that the weight for workload balance is smaller than one, meaning that the workload will be balanced by one minute if that costs less than one minute to the overall schedule.
+\noindent Note that the weight for workload balance is smaller than one, meaning that the workload will be balanced by one minute if that costs less than one minute to the overall schedule.
 This value might be adjusted depending on the number of nurses on the teams, as the impact on the total schedule might be easier to absorb if more nurses are present.
 
-\section{Decision Support Tool}\label{seq:DSS}
-
-The problem described in the previous section aims to capture all the real-world constraints we observed in practice, but also to include a high degree of flexibility in the modelling.
+%----------------------------------------------------------------------------------%
+\section{Decision Support Tool}
+\label{sec:dst}
+\noindent The problem described in the previous section aims to capture all the real-world constraints we observed in practice, but also to include a high degree of flexibility in the modelling.
 This occurs for example by the introduction of preferences (which can account for a number of features, as continuity of care, for example) or weights in the objective function.
@@
 The main component of the tool is a Greedy Randomised Adaptive Search Procedure (GRASP) optimisation algorithm, which aims to provide efficient solutions.
-As noted by other researchers \cite[\eg][]{BERTELS20062866, Rasmussen2012, Mankowska2014}, the problem presented in Section~\ref{seq:model description} and its similar versions are NP-Hard.
+As noted by other researchers \cite[e.g.][]{bertels2006, rasmussen2012, mankowska2014}, the problem presented in Section~\ref{sec:model} and its similar versions are NP-hard.
 
+%----------------------------------------------------------------------------------%
 \subsection{GRASP}
-
-The GRASP framework was originally proposed by \cite{FEO198967} and has been widely used and extended since then \citep{resende2016optimization}.
-In its basic form, it consists in generating greedy solutions with a controlled degree of randomness and then improving them with a local search.
-The general layout of our implementation of the GRASP algorithm is depicted in Figure~\ref{fig:grasp_diagram}.
+\label{sub:grasp}
+\noindent The GRASP framework was originally proposed by \cite{feo1989} and has been widely used and extended since \citep{resende2016}.
+In its basic form, GRASP consists in generating greedy solutions with a controlled degree of randomness and then improving them with a local search.
+The general layout of our implementation of the GRASP algorithm is depicted in Figure~\ref{fig:graspdiagram}.
 
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{im/scheme}
-    % \includegraphics{}
+    \includegraphics[width=0.7\textwidth]{images/scheme}
     \caption{Scheme of the proposed GRASP algorithm.}
-    \label{fig:grasp_diagram}
+    \label{fig:graspdiagram}
 \end{figure}
 
-Solutions are generated iteratively by the randomised constructive (described in Section~\ref{subseq:constructive}) and improved with local search (described in Section~\ref{subseq:local_search}).
+Solutions are generated iteratively by the randomised constructive (described in Section~\ref{sub:randomisedconstructive}) and improved with local search (described in Section~\ref{sub:localsearch}).
 Good solutions are kept in a solution pool if they are different enough from the ones currently in the pool.
-In order to speed the convergence of the algorithm, the solutions of the pool are used to perform path relinking (see Section~\ref{subseq:PR}).
+In order to speed the convergence of the algorithm, the solutions of the pool are used to perform path relinking (see Section~\ref{sub:pathrelinking}).
 After each iteration we check the stopping criterion.
@@
 
-\subsection{Randomised constructive}\label{subseq:constructive}
-
-The Randomised Constructive Algorithm (RCA) is used to create initial solutions for each iteration of the GRASP algorithm.
-It works by ranking the quality of all job-nurse assignments by the value of the objective function \eqref{eq:real_obj_function}, which might also be evaluated in partial solutions.
+%----------------------------------------------------------------------------------%
+\subsection{Randomised Constructive}
+\label{sub:randomisedconstructive}
+\noindent The Randomised Constructive Algorithm (RCA) is used to create initial solutions for each iteration of the GRASP algorithm.
+It works by ranking the quality of all job-nurse assignments by the value of the objective function \eqref{eqn:realobjfunction}, which might also be evaluated in partial solutions.
 These rankings are sorted by quality, and the highest quality ones are stored in a reduced candidate list (RCL).
@@
      \begin{equation}
-         C = B + \delta (B - W)
+         C = B + \delta (B - W).
      \end{equation}
@@
      \begin{equation}
-         C = B - \delta B
+         C = B - \delta B.
      \end{equation}
@@
 The parameter $\delta$ is randomly selected in each iteration, sampled from the interval $\delta \in [\texttt{grasp\_dl}, \texttt{grasp\_dl} + \texttt{grasp\_dr}]$.
-The values of \texttt{grasp\_dl} + \texttt{grasp\_dr}, together with the type of strategy to use, \texttt{rcl\_strategy}, are hyper-parameters of the algorithm and are optimised in Section~\ref{seq:parameter_tuning}.
+The values of \texttt{grasp\_dl} + \texttt{grasp\_dr}, together with the type of strategy to use, \texttt{rcl\_strategy}, are hyper-parameters of the algorithm and are optimised in Section~\ref{sub:parametertuning}.
 The RCA has two phases.
@@
 
-\subsection{Local search}\label{subseq:local_search}
-After each iteration of the constructive algorithm a local search procedure is applied to the resulting solution to improve its quality.
-The local search consists in a first improvement Variable Neighbourhood Search (VNS) framework \citep{Mladenovic1997}, a general metaheuristic algorithm that has proven to be successful in many application areas, including routing problems \cite{hansen2010variable}, due to its capacity to include a range of different local search movements.
+%----------------------------------------------------------------------------------%
+\subsection{Local Search}
+\label{sub:localsearch}
+\noindent After each iteration of the constructive algorithm a local search procedure is applied to the resulting solution to improve its quality.
+The local search consists in a first improvement Variable Neighbourhood Search (VNS) framework \citep{mladenovic1997}, a general metaheuristic algorithm that has proven to be successful in many application areas, including routing problems \citep{hansen2010}, due to its capacity to include a range of different local search movements.
 VNS takes as input an initial solution and an ordered neighbourhood structure.
@@
     \item NE03 Nurse order change: This neighbourhood focuses on exploring different orderings of the nurse list, $\mathcal{O}$. It searches over exchanges between pairs of nurses in the list. 
-    This can have an effect in the way waiting times are calculated (as detailed in Section~\ref{seq:sub temporal constraints}) and improve the overall quality.
+    This can have an effect in the way waiting times are calculated (as detailed in Section~\ref{sub:temporalconstraints}) and improve the overall quality.
     \item NE04 Remove and reinsert linked services: Two services that are dependent on each other (synchronised or double services) are removed simultaneously from the solution and reinserted in the best possible positions.
@@
 
-While some neighbourhoods proved key to the performance of the GRASP algorithm in our initial testing, the benefit of 2OPT and the nurse order change was not immediately evident.
-For this reason, in Section~\ref{seq:parameter_tuning} we analyse whether they are helpful to the overall performance of the algorithm or not.
+\noindent While some neighbourhoods proved key to the performance of the GRASP algorithm in our initial testing, the benefit of 2OPT and the nurse order change was not immediately evident.
+For this reason, in Section~\ref{sub:parametertuning} we analyse whether they are helpful to the overall performance of the algorithm or not.
 
-\subsection{Path Relinking}\label{subseq:PR}
-% \subsection{Description}
-The basic idea behind Path Relinking (PR) \citep{laguna1999grasp,resende2016optimization} is to use two existing solutions to generate a third one, which can potentially have higher quality than both of them.
+%----------------------------------------------------------------------------------%
+\subsection{Path Relinking}
+\label{sub:pathrelinking}
+\noindent The basic idea behind Path Relinking (PR) \citep{laguna1999,resende2016} is to use two existing solutions to generate a third one, which can potentially have higher quality than both of them.
 This is done by transforming one of them (called starting solution) step by step, by using local search movements, in the other solution (called guiding solution).
@@
 Note that, this intermediate solution is often not a local optimum (unlike the starting and the guiding solutions), so applying local search may lead to high quality solutions.
-The pseudo-code of the routine is described in Algorithm~\ref{alg::path_relinking}.
+The pseudo-code of the routine is described in Algorithm~\ref{alg:pathrelinking}.
 
-\input{pr_algorithm.tex}
+% \input{pr_algorithm.tex}
+\begin{algorithm}[h]
+\caption{Path Relinking}
+\label{alg:pathrelinking}
+\begin{algorithmic}[1]
+\State Let $\mathcal{S}_g = \{\mathcal{R}_g, \mathcal{O}_g\}$ be the guiding solution
+\State Let $\mathcal{S}_s = \{\mathcal{R}_s, \mathcal{O}_s\}$ be the starting solution
+\State Let $\mathcal{S}_w = \mathcal{S}_s$ be the working solution
+\State Let $\mathcal{S}_r = \emptyset$ be the relinked solution;  $Q(\mathcal{S}_r) = -\infty$
+\State \texttt{n\_movements} = n + m
+\While {$\mathcal{S}_w \neq \mathcal{S}_g$} \label{alg:prmainwhile}
+	\For{$h  = 1, \dots, \texttt{n\_movements}$}
+		\State $\mathcal{S}_{best} = \emptyset$; $Q(\mathcal{S}_{best}) = -\infty$
+		\For{$i  = 1, \dots, n$}
+			\State $\mathcal{S}^* = \mathcal{S}_w$
+			% Check nurse
+			% \STATE \COMMENT{Check nurse order movements}
+			\If{$\mathcal{O}^*[i] \neq \mathcal{O}_g[i]$}
+				\State Swap two elements of $\mathcal{O}^*$ such that $\mathcal{O}^*[i]$ equals $\mathcal{O}_g[i]$
+				\If{$Q(\mathcal{S}^*) \geq Q(\mathcal{S}_{best})$}
+					\State $\mathcal{S}_{best} = \mathcal{S^*}$
+				\EndIf
+			\EndIf
+		\EndFor % Finished with nurses
+		\For{$j  = 1, \dots, m$}
+			\State $\mathcal{S}^* = \mathcal{S}_w$
+			\State Remove job $j$ from $\mathcal{S^*}$
+			\State Re-insert job $j$ in $\mathcal{S^*}$ in the position(s) that match $\mathcal{S}_g$ \label{alg:prforinsertion}
+			% Update the best movement
+			\If{$Q(\mathcal{S}^*) \geq Q(\mathcal{S}_{best})$}
+				\State $\mathcal{S}_{best} = \mathcal{S^*}$
+			\EndIf
+		\EndFor % Finished with jobs
+		% Apply the best movement:
+		\State $\mathcal{S}_w = \mathcal{S}_{best}$
+		\If{$Q(\mathcal{S}_w) \geq Q(\mathcal{S}_r)$}
+			\State $\mathcal{S}_r = \mathcal{S}_w$
+		\EndIf
+	\EndFor
+\EndWhile
+\State $\mathcal{S}_r = \texttt{local\_search}(\mathcal{S}_r)$
+\Return $\mathcal{S}_r$
+\end{algorithmic}
+\end{algorithm}
 
-Note that double services are removed and reinserted simultaneously (see line~\ref{alg:pr:for_insertion}).
+Note that double services are removed and reinserted simultaneously (see line~\ref{alg:prforinsertion}).
 In the case that some jobs that are due to be inserted in a position later than the current length of the route, they are instead inserted in the last possible position.
-This might cause the solutions not to be equal after applying $n + m$ movements, so if this happens another pass is performed (line~\ref{alg:pr:main_while}).
+This might cause the solutions not to be equal after applying $n + m$ movements, so if this happens another pass is performed (line~\ref{alg:prmainwhile}).
 
@@
 The number of solutions in the pool is controlled by the parameter \texttt{pool\_size}.
-We determine a suitable value for this parameter in our experiments (see Section~\ref{seq:Computational Experiments}).
+We determine a suitable value for this parameter in our experiments (see Section~\ref{sec:computationalexperiments}).
 
@@
 Let schedule $\mathcal{S}_1$ be the pair $\mathcal{S}_1 = \{\mathcal{R}_1, \mathcal{O}_1\}$ and schedule $\mathcal{S}_2 = \{\mathcal{R}_2, \mathcal{O}_2\}$.
+Then:
 \begin{equation}
-    d(\mathcal{S}_1, \mathcal{S}_2) = d_o(\mathcal{O}_1, \mathcal{O}_2) + d_a(\mathcal{R}_1, \mathcal{R}_2) + d_r(\mathcal{R}_1, \mathcal{R}_2)
+    d(\mathcal{S}_1, \mathcal{S}_2) = d_o(\mathcal{O}_1, \mathcal{O}_2) + d_a(\mathcal{R}_1, \mathcal{R}_2) + d_r(\mathcal{R}_1, \mathcal{R}_2).
 \end{equation}
-
 This is a combination of three measures, the dissimilarity of the nurse order $d_o(\mathcal{O}_1, \mathcal{O}_2)$, of the nurse-job assignments, $d_a(\mathcal{R}_1, \mathcal{R}_2)$, and the differences in the routes themselves, $d_r(\mathcal{R}_1, \mathcal{R}_2)$.
 
-The distance between the solution orderings is simply calculated as the Hamming distance between the ordering vectors, \ie the number of different digits in them:
+The distance between the solution orderings is simply calculated as the Hamming distance between the ordering vectors, i.e. the number of different digits in them:
 \begin{equation}
-    d_o(\mathcal{O}_1, \mathcal{O}_2) = |\{i\in\{1,\dots,N\} : \mathcal{O}_1(i) \neq  \mathcal{O}_2(i)\}|
+    d_o(\mathcal{O}_1, \mathcal{O}_2) = |\{i\in\{1,\dots,n\} : \mathcal{O}_1(i) \neq  \mathcal{O}_2(i)\}|,
 \end{equation}
@@
 r^{(1)}_1, r^{(1)}_2, \dots, r^{(1)}_n\}$ and $\mathcal{R}_2 = \{r^{(2)}_1, r^{(2)}_2, \dots, r^{(2)}_n\}$ the routes in schedules $\mathcal{S}_1$ and $\mathcal{S}_2$ respectively, then:
-
 \begin{equation}
-    d_a(\mathcal{R}_1, \mathcal{R}_2) = |\{j\in\{1,\dots,J\} : \nexists i \in N such\ that\ j \in r^{(1)}_i \cap r^{(2)}_i\}|
+    d_a(\mathcal{R}_1, \mathcal{R}_2) = |\{j\in\{1,\dots,m\} : \nexists i \in N \text{ such that } j \in r^{(1)}_i \cap r^{(2)}_i\}|.
 \end{equation}
-
 Finally, the dissimilarities in the routes are captured by looking at the presence of different arcs in the solution.
@@
 \begin{equation}
-    d_r(\mathcal{R}_1, \mathcal{R}_2) = |A(\mathcal{R}_1)| - |A(\mathcal{R}_1) \cap A(\mathcal{R}_2)|
+    d_r(\mathcal{R}_1, \mathcal{R}_2) = |A(\mathcal{R}_1)| - |A(\mathcal{R}_1) \cap A(\mathcal{R}_2)|.
 \end{equation}
@@
 
+%----------------------------------------------------------------------------------%
 \subsubsection{Type of Path Relinking}
-Path relinking is performed at the end of the algorithm between all pairs of solutions in the pool.
+\label{subsub:typepathrelinking}
+\noindent Path relinking is performed at the end of the algorithm between all pairs of solutions in the pool.
 However, it is also possible to perform path relinking at intermediate steps of the algorithm too, each time a new solution is generated.
@@
 
-Finally, another important parameter of PR is the direction.
+\noindent Finally, another important parameter of PR is the direction.
 In forward path relinking, the quality of the starting solution is less or equal than the quality of the guiding solution.
@@
 
-Both, a suitable strategy and direction for PR are determined in Section~\ref{seq:parameter_tuning}.
-% \end{itemize}
+Both a suitable strategy and direction for PR are determined in Section~\ref{sub:parametertuning}.
 
-\section{Computational experiments}\label{seq:Computational Experiments}
+%----------------------------------------------------------------------------------%
+\section{Computational experiments}
+\label{sec:computationalexperiments}
+
+%----------------------------------------------------------------------------------%
 \subsection{Testing data}
-For our experiments, we use three datasets.
-Two are from two similar problems in the literature, presented in \cite{Mankowska2014} and \cite{AitHaddadene2016}, and are used to evaluate the algorithm performance in Section~\ref{seq:algorithm_performance}.
+\label{sub:testing}
+\noindent For our experiments, we use three datasets.
+Two are from two similar problems in the literature, presented in \cite{mankowska2014} and \cite{aithaddadene2016}, and are used to evaluate the algorithm performance in Section~\ref{sub:algorithmperformance}.
 The third dataset is composed of randomly generated instances based on road distances in the city of Southampton.
-This dataset is used for tuning the algorithm parameters in Section~\ref{seq:parameter_tuning} and to explore different weights of the objective function in Section~\ref{seq:exploring_obj_function}.
-Details on the instances and the process for generating them are detailed in \ref{appendix:RandomlyGeneratedData}.
+This dataset is used for tuning the algorithm parameters in Section~\ref{sub:parametertuning} and to explore different weights of the objective function in Section~\ref{sub:exploringobjfunction}.
+Details on the instances and the process for generating them are detailed in \ref{appendix:randomlygenerateddata}.
 
-\subsection{Parameter tuning}\label{seq:parameter_tuning}
-The GRASP algorithm is made up of several components, for each of which we have identified various potential configurations.
+%----------------------------------------------------------------------------------%
+\subsection{Parameter tuning}
+\label{sub:parametertuning}
+\noindent The GRASP algorithm is made up of several components, for each of which we have identified various potential configurations.
 %For example, we can choose whether to use or not some of the neighbourhoods described in Section~\ref{subseq:local_search}.
-In order to select the best possible configuration automatically and in a statistically sound manner, we used the software package \texttt{irace} \citep{Lopez-Ibanez2016} with a budget of 1000 runs.
-Table~\ref{tab:parameters} shows a list of the hyper-parameters and solving strategies tested.
+In order to select the best possible configuration automatically and in a statistically sound manner, we used the software package \texttt{irace} \citep{lopez2016} with a budget of 1000 runs.
+Table~\ref{table:parameters} shows a list of the hyper-parameters and solving strategies tested.
 In the first column we show the name of the hyper-parameter, in the second column a brief description, in the third column the type of variable we used for testing it, in the fourth column the values that were tested and in the fifth column the best value found by \texttt{irace}.
@@
     NurseChange             &   Change nurse order between GRASP iterations & B & Yes, No                        &  Yes  \\ % no_change_grasp
-    \texttt{pr\_strategy}   &  Path relinking strategy, as listed in Section~\ref{subseq:PR}& C & 1, 2, 3   &   1 \\ %  
+    \texttt{pr\_strategy}   &  Path relinking strategy, as listed in Section~\ref{sub:pathrelinking}& C & 1, 2, 3   &   1 \\ %  
     \texttt{pr\_direction}  &   Path relinking direction: forward (1), backward (2), forward and backward (3), chosen randomly at each iteration (4).& C & 1, 2, 3, 4       &   1 \\ % performPathRelinking
@@
     \end{tabular}
-    \label{tab:parameters}
+    \label{table:parameters}
 \end{table}
@@
 
-\subsection{Variations in the literature}\label{seq:literature data}
-While the problems reported by \cite{Mankowska2014} and \cite{AitHaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to considered.
-
-\cite{Mankowska2014} describes the following objective function:
+%----------------------------------------------------------------------------------%
+\subsection{Variations in the Literature}
+\label{sub:literaturedata}
+\noindent While the problems reported by \cite{mankowska2014} and \cite{aithaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to considered.
+\cite{mankowska2014} describes the following objective function:
 \begin{equation}
-    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max}
+    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max},
 \end{equation}
 where $Dist$ denotes the distance travelled by all nurses, $Travel$ denotes the total tardiness and $Travel^{max}$ denotes the maximum tardiness.
-For \cite{Mankowska2014} tardiness refers to the time between the end of its time window and the actual start time of a job.
+For \cite{mankowska2014} tardiness refers to the time between the end of its time window and the actual start time of a job.
 Tardiness is not allowed for interdependent jobs, where the dependency gap is a hard constraint.
@@
 
-In \cite{AitHaddadene2016} tardiness is not allowed, but preferences are taken into account, so their objective function is as follows:
+In \cite{aithaddadene2016} tardiness is not allowed, but preferences are taken into account, so their objective function is as follows:
 \begin{equation}
-    \min Dist + Pref
+    \min Dist + Pref,
 \end{equation}
-Where $Dist$ represents the cost of travelling for all nurses and $Pref$ the preference score.
-In Table~\ref{tab:adapted_obj_values} we specify the weights used in the objective function \eqref{eq:solution_quality} in order to represent the objective functions of these algorithms.
+where $Dist$ represents the cost of travelling for all nurses and $Pref$ the preference score.
+In Table~\ref{table:adaptedobjvalues} we specify the weights used in the objective function \eqref{eqn:solutionquality} in order to represent the objective functions of these algorithms.
 
@@
     \centering
-    \caption{Weights used in the objective function to match those used in the methods from  \cite{AitHaddadene2016} (AH) and \cite{Mankowska2014} (MK).
-A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, \ie they contribute to $I(S)$.}
+    \caption{Weights used in the objective function to match those used in the methods from  \cite{aithaddadene2016} (AH) and \cite{mankowska2014} (MK).
+A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, i.e.
+they contribute to $I(S)$.}
     \begin{tabular}{ccc}
@@
     \end{tabular}
-    \label{tab:adapted_obj_values}
+    \label{table:adaptedobjvalues}
 \end{table}
 
-Note that we multiply time measures by $60$, since we consider our objective function in minutes, so it is easier to understand by final users of the tool.
-Furthermore, since both articles present minimisation problems, we switch the sign of the solutions we report in Section~\ref{seq:algorithm_performance} for ease of comparison.
+\noindent Note that we multiply time measures by $60$, since we consider our objective function in minutes, so it is easier to understand by final users of the tool.
+Furthermore, since both articles present minimisation problems, we switch the sign of the solutions we report in Section~\ref{sub:algorithmperformance} for ease of comparison.
 
-Another difference between \cite{AitHaddadene2016} and our model is that for time dependent services, we define a precedence between them, while they define a gap.
-That means that our constraints precisely determine which job should be executed first, while \cite{AitHaddadene2016} leaves this decision to the algorithm, specifying only the minimum and maximum amount of time that might separate their start times.
+Another difference between \cite{aithaddadene2016} and our model is that for time dependent services, we define a precedence between them, while they define a gap.
+That means that our constraints precisely determine which job should be executed first, while \cite{aithaddadene2016} leaves this decision to the algorithm, specifying only the minimum and maximum amount of time that might separate their start times.
 This was accounted for in our algorithm by considering the dependency only at runtime and choosing the most convenient gap (positive or negative) in each objective function evaluation, depending on what the arrival times were.
 
-\subsection{Algorithm performance}\label{seq:algorithm_performance}
-
-While the problem description of our DST is slightly different from the examples seen in Section~\ref{seq:literature data}, we test the performance of GRASP algorithm, with the required modifications, in the instances provided by \cite{AitHaddadene2016} and \cite{Mankowska2014}.
+%----------------------------------------------------------------------------------%
+\subsection{Algorithm Performance}
+\label{sub:algorithmperformance}
+\noindent While the problem description of our DST is slightly different from the examples seen in Section~\ref{sub:literaturedata}, we test the performance of GRASP algorithm, with the required modifications, in the instances provided by \cite{aithaddadene2016} and \cite{mankowska2014}.
 
 For all the instances, we perform 5 runs and report the best and the average result.
-The full computational results and solving times are listed in the Appendix tables \ref{tab:mk_results}, \ref{tab:aith_results} and \ref{tab:longer_experiments}.
+The full computational results and solving times are listed in the Appendix tables \ref{table:mkresults}, \ref{table:aithresults} and \ref{table:longerexperiments}.
 
-In Table~\ref{tab:group_table_aith_combined} we present a summary of the results found by our DST for the three instance groups presented in \cite{AitHaddadene2016}.
+In Table~\ref{table:grouptableaithcombined} we present a summary of the results found by our DST for the three instance groups presented in \cite{AitHaddadene2016}.
 The results are the average of the best and average results found for the instances of the three instance groups \emph{G1}, \emph{G2} and \emph{G3}.
 We have allowed a time of 5 seconds for instances in the group \emph{G1}, 5 minutes for those in \emph{G2} and three hours for the instances in \emph{G3}.
-See \ref{appendix:allowing_extra_time} for a note on how solution quality varies when increasing solving time.
+See \ref{appendix:allowingextratime} for a note on how solution quality varies when increasing solving time.
 
@@
 
-These results suggest that our DST can find better objective values within reasonable planning times.
+\noindent These results suggest that our DST can find better objective values within reasonable planning times.
 Further, we are able to replicate the 23 optimal solutions provided in the paper.
@@
 
-Similarly, in Table~\ref{tab:group_table_mk_avg_best} we present the comparison of our DST against the performance of the AVNS algorithm presented in \cite{Mankowska2014}.
+Similarly, in Table~\ref{table:grouptablemkavgbest} we present the comparison of our DST against the performance of the AVNS algorithm presented in \cite{mankowska2014}.
 
@@
 	\centering
-	\caption{Performance of the GRASP algorithm (average of best and average results) compared to the AVNS algorithm presented in \cite{Mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}}
+	\caption{Performance of the GRASP algorithm (average of best and average results) compared to the AVNS algorithm presented in \cite{mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}}
 	\begin{tabular}{c|c|ccc}
@@
 	\end{tabular}
-	\label{tab:group_table_mk_avg_best}
+	\label{table:grouptablemkavgbest}
 \end{table}
@@
 GRASP clearly outperforms AVNS in the majority of instances, especially on those of larger sizes.
-Further, our algorithm is able to replicate most of the proven optimal solutions (9 of the 10 instances for which \cite{Mankowska2014} provides an optimal solution).
+Further, our algorithm is able to replicate most of the proven optimal solutions (9 of the 10 instances for which \cite{mankowska2014} provides an optimal solution).
 In the instance group A, where the AVNS finds all optimal solutions for the problem, GRASP fails to replicate one of these solutions, hence increasing the average.
-This is one of the cases where the assumption that a start time of a job should be as early as possible for the first nurse arriving \eqref{eq:earliest_possible_start_assumption2} proves to be too restrictive and removes the optimal solution from the search space of the metaheuristic.
+This is one of the cases where the assumption that a start time of a job should be as early as possible for the first nurse arriving \eqref{eqn:earliestpossiblestartassumption2} proves to be too restrictive and removes the optimal solution from the search space of the metaheuristic.
 Nevertheless, the loss in quality is under $5\%$ and the route itself found by the two algorithms is the same (but not the scheduled start times).
@@
 
-\subsection{Trade-offs in solution quality}\label{seq:exploring_obj_function}
-While the previous section sheds some light on the algorithmic performance of the GRASP algorithm and, in general, of our ability to solve efficiently HHCRSP problems of size similar to those faced by district nursing teams in the UK, it does not provide any insight on the actual applicability of these solutions.
-In this section we focus on the instances from \cite{AitHaddadene2016} that are solved to optimality (with up to 45 patients) and analyse how these solutions would look like from the perspective of a team of nurses undertaking the job.
+%----------------------------------------------------------------------------------%
+\subsection{Trade-offs in Solution Quality}
+\label{sub:exploringobjfunction}
+\noindent While the previous section sheds some light on the algorithmic performance of the GRASP algorithm and, in general, of our ability to solve efficiently HHCRSP problems of size similar to those faced by district nursing teams in the UK, it does not provide any insight on the actual applicability of these solutions.
+In this section we focus on the instances from \cite{aithaddadene2016} that are solved to optimality (with up to 45 patients) and analyse how these solutions would look like from the perspective of a team of nurses undertaking the job.
 This means asking questions like: Is the workload fairly distributed?
@@
 In order to answer these questions, we examine closely the values of a few key indicators of the quality of a solution: \emph{Total time}, \emph{Travel time}, \emph{Waiting time}, \emph{Preferences} and \emph{Longest Day}, which is the time it takes to perform the longest shift in the solution.
-We compare the indicators found by using the original objective function (listed in Table~\ref{tab:adapted_obj_values}) with those found by using our proposed objective function (the values listed in Table~\ref{tab:adequate_weights}, but with preferences being negative and time quantities scaled by $-0.3\times60$, in order to scale them to the data).
-The result of this comparison is shown in Table~\ref{tab:comparison_of_qualities}.
+We compare the indicators found by using the original objective function (listed in Table~\ref{table:adaptedobjvalues}) with those found by using our proposed objective function (the values listed in Table~\ref{table:adequateweights}, but with preferences being negative and time quantities scaled by $-0.3\times60$, in order to scale them to the data).
+The result of this comparison is shown in Table~\ref{table:comparisonofqualities}.
 
-\input{"quality_comparison_table.tex"}
+\input{quality_comparison_table.tex}
 
@@
 
-\section{Conclusion}\label{seq:Conclusion}
-
-In this article we propose a Decision Support Tool for the Home Health Care Routing and Scheduling Problem that is daily faced by district nursing teams.
+%----------------------------------------------------------------------------------%
+\section{Conclusion}
+\label{sec:conclusion}
+\noindent In this article we propose a Decision Support Tool for the Home Health Care Routing and Scheduling Problem that is daily faced by district nursing teams.
 We propose a model that includes the most common constraints observed in practice (skills, time windows, synchronisation, double services) and objectives (travel time, waiting time, assignment preferences, workload balance, tardiness, overtime).
@@
 
+%----------------------------------------------------------------------------------%
 \section{Funding}
-
-This article presents independent research partially supported by the National Institute for Health Research (NIHR) Collaboration for Leadership in Applied Health Research and Care (CLAHRC), NIHR Applied Research Collaboration (ARC) and Solent NHS Trust.
+\label{sec:funding}
+\noindent This article presents independent research partially supported by the National Institute for Health Research (NIHR) Collaboration for Leadership in Applied Health Research and Care (CLAHRC), NIHR Applied Research Collaboration (ARC) and Solent NHS Trust.
 The views expressed in this publication are those of the author(s) and not necessarily those of the National Health Service, the NIHR or the Department of Health and Social Care.
@@
 
-\input{bib_section.tex}
+\bibliographystyle{model5-names}
+\bibliography{includes/bibliography}
+
+%\input{bib_section.tex}
 
@@
 
-\section{Result tables}\label{a:Result_Tables}
+%----------------------------------------------------------------------------------%
+\section{Result Tables}
+\label{appendix:resulttables}
 
-Table~\ref{tab:mk_results} shows the results for the instances presented in \cite{Mankowska2014}:
+\noindent Table~\ref{table:mkresults} shows the results for the instances presented in \cite{mankowska2014}:
 \input{"mk_table.tex"}
 
-Table~\ref{tab:aith_results} shows the results for the instances presented in \cite{AitHaddadene2016}.
+\noindent Table~\ref{table:aithresults} shows the results for the instances presented in \cite{aithaddadene2016}.
 We have allowed a time of 5 seconds for the instances with $18$ patients, 5 minutes for the instances with $45$ patients and one hour for the instances with $73$ patients.
@@
 
-\section{Solving time and solution quality}\label{appendix:allowing_extra_time}
-In order to understand better how GRASP performs over time, we performed an extra set of experiments, where GRASP was allowed one, two and three hours on the largest instances from \cite{AitHaddadene2016}.
+%----------------------------------------------------------------------------------%
+\section{Solving Time and Solution Quality}
+\label{appendix:allowingextratime}
+\noindent In order to understand better how GRASP performs over time, we performed an extra set of experiments, where GRASP was allowed one, two and three hours on the largest instances from \cite{aithaddadene2016}.
 We compared these results to their ILS algorithm, which obtains the best results on large instances.
-They are shown in Table~\ref{tab:longer_experiments}.
-{\footnotesize
+They are shown in Table~\ref{table:longerexperiments}.
+{\scriptsize
 \begin{longtable}{c|cc|cc|cc|cc}
-\caption{Performance of the GRASP algorithm on the larger instances with more time compared with that of the ILS algorithm from \cite{AitHaddadene2016} (AH ILS)}
-\label{tab:longer_experiments}\\
+\caption{Performance of the GRASP algorithm on the larger instances with more time compared with that of the ILS algorithm from \cite{aithaddadene2016} (AH ILS)}
+\label{table:longerexperiments}\\
 \hline
@@
 
-These results suggest that allowing more time improves the performance of GRASP, to a quality similar to that of the ILS algorithm.
+\noindent These results suggest that allowing more time improves the performance of GRASP, to a quality similar to that of the ILS algorithm.
 Note that the current practice in district nursing teams is to plan work one day in advance, so times of up to three hours for running the algorithm are feasible in a practical setting.
-The average results also improve with more time, and are often better than the averages of ILS, suggesting GRASP is more stable,
+The average results also improve with more time, and are often better than the averages of ILS, suggesting GRASP is more stable.
 

### 2021-10-25 tex/ARCBoard/ArcBoard.tex
paragraph for ARC Board.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Improving Community Health Care Logistics Using OR -- ARC Board}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\noindent Throughout this project we worked alongside Abicare, a care home company based in various regions in the south of England and Wales, in order to improve upon the routing and scheduling of nurses to patients prefering domiciliary care. Business-as-usual planning at Abicare is manual, requiring staff in each area to create schedules for nurses by hand. The schedule must ensure that patients' needs are met, whilst also minimising the time and distance travelled by the nurses to each location; additional travel incurs extra costs. Furthermore, it is important to consider synchronised jobs, where the care of a patient requires two nurses, and the workload balance of each nurse. We therefore developed a Decision Support Tool which utilises a GRASP-VNS algorithm to minimise waiting times, travel times, mileage and tardiness for each of the nurses. This DST will be used in place of manual planning at Abicare, which will allow for the production of high-quality schedules and routes. In preliminary tests, which assessed instances in Basingstoke, Monmouth, and Aldershot, the DST was able to reduce the total time spent travelling by nurses in a single day up to two hours and 34 minutes, the total waiting time by up to seven hours and 30 minutes, and the total distance travelled by up to 154km. These encouraging results show that the DST will greatly benefit Abicare and the domiciliary care industry, allowing staff to focus on providing quality care to patients.

\end{document}

### 2021-10-25 tex/Plan/Plan.tex
document for plan for remainder of ARC Board.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}

\begin{document}
\title{Updated Project Plan}
%\author{Asyl L. Hawa}
\date{\today}
\maketitle

\section*{Objectives}
\begin{enumerate}[leftmargin=*]
	\item \textbf{Real Evaluation of the DST}\\
	This will be done over a period of time in two ways: (a) comparing the automated plans generated by the DST with plans created manually by staff; and (b) obtaining and assessing reviews from nurses regaring the new automated plans. This will provide us with a clearer understanding of the impact of the DST in practice.
	\item \textbf{Exact Methods for the RSP}\\
	An exact method would provide a guarantee that the problem is solved to optimality, however, it may not be suitable for daily planning due to the time and software that may be required. An exact method would be useful for Health Services for considering whether investments in planning are helpful or necessary, and will also be useful as a basis for other types of modelling.
	\item \textbf{Stochastic Demand}\\
	We will identify particular situations that require modifications to existing plans, e.g. changes in the number of nurses or new patients requiring treatment, and update the model to take these unplanned changes into consideration.
\end{enumerate}

%\begin{itemize}
%	\item Compare results from original plan with new DST plans at Abicare
%	\item start looking at exact methods
%	\item look at stochastic method
%	\item publish code and paper
%\end{itemize}

\section*{Outputs}
\begin{enumerate}[leftmargin=*]
	\item \textbf{Decision Support Tool}\\
	We aim to release the DST for the planning of domiciliary care to make the software readily available to health care teams to assist with creating routes and schedules for staff. This will also allow us to obtain feedback from the users and a comparison with previous schedules.
	\item \textbf{Publications}\\
	We endeavour to submit two articles for publication: a methodological paper covering findings regarding exact methods and adaptations to stochastic demand, which will be submitted to a mathematical journal, and a scientific paper for a healthcare journal that will present our work on the efficiency of our DST in a real setting in comparison to business-as-usual planning. 
\end{enumerate}

%\begin{itemize}
%	\item Release DST
%	\item Second paper with exact method and stochastic demand
%	\item publish paper comparing abicare to a healthcare journal
%	\item Resubmit paper to computers and OR
%\end{itemize}

\section*{Timeline}
Below is a guide for the remainder of the project, and outlines how the above targets will be achieved:

\begin{enumerate}[leftmargin=*]
	\item \textbf{18th October 2021 -- 31st October 2021}\\
	Finalise initial publication and prepare to submit to journal.
	\item \textbf{November 2021}\\
	Release the DST. Prepare a questionnaire for nurses and planners in industry, which will be used as a basis for comparison after the DST has been used for a period of time. Meet with Abicare to discuss the frequency in which the DST should be used, and how results should be collected for future review. Complete formal documentation for the DST.
	\item \textbf{December 2021 -- January 2022}\\
	Focus on adapting the DST for stochastic demand. Identify the situations that should be considered, and attempt to model the changes.
	\item \textbf{January 2022 -- February 2022}\\
	Turn attention to exact methods. Study the literature to determine methods that will/not be suitable for our DST, and develop a model that can be adapted for the problem.
	\item \textbf{March 2022}\\
	Obtain results from Abicare. Send out final questionnaire to nurses and planners and compare with initial questionnaire. Compare and analyse outputs.
	\item \textbf{April 2022 -- May 2022}\\
	Gather data and test stochastic demand and exact method changes to the model.
	\item \textbf{June 2022}\\
	Potentially release new version of DST including stochastic demand and exact method. Prepare articles for submission to mathematical and healthcare journals. Conclude project.
\end{enumerate}

\end{document}

### 2021-10-25 tex/Paper/main.tex
tidied up paper, made minor changes.

@@
 While the problem might be addressed by drawing from the extensive vehicle routing problem (VRP) literature (see, for example, \cite{drexl2012} for a comprehensive review of VRPs and synchronisation constraints), HHCRSP has developed its own body of literature.
-In particular, there are two recent review articles, \cite{cisse2017} and \cite{fikar2017}.
+In particular, there are recent review articles: \cite{cisse2017}, \cite{fikar2017}, \citet{dimascolo2021}, and \citet{grieco2020}.
 We refer the reader to them for a broader understanding of what is meant by HHCRSP and different solving approaches.
@@
 For example, \cite{rasmussen2012} tackles the HHCRSP problem in Denmark.
-They propose a branch-and-price scheme where generalised synchronisation constraints are relaxed and dealt with in the branching.
-% \correction{Only Rasmussen?}
+They propose a branch-and-price scheme where generalised synchronisation constraints are relaxed and dealt with in the branching.% \correction{Only Rasmussen?}
  % \correction{Cite relevant papers here}. The ILP formulations work well for small problems, but are often not practical to solve for problems with a significant number of patients. %around 50
@@
 Another is \cite{bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time.
+In addition, \citet{nikzad2021} considers a four-phase matheuristic that calls upon a variety of mixed integer models.
 
 Another approach, which is the one we use in this article, is to propose a metaheuristic algorithm independent of a mathematical solver.
+Various metaheuristic approaches have been widely studied in the literature: \citet{manavizadeh2020} considers a simulated annealing (SA) algorithm, whilst \citet{nasir2020} investigates the effectiveness of a hybrid genetic algorithm (GA).
+Furthermore, a memetic algorithm is proposed by \citet{shahnejat2021}, and four hybrid metaheuristics including combinations of GA and variable neighbourhood search (VNS) and SA with local search (LS) are developed by \citet{liu2021}.
 \cite{mankowska2014} and \cite{aithaddadene2016} propose an Adaptive Large Neighbourhood search and a hybrid GRASP combined with Iterated Local Search respectively.
@@
 %----------------------------------------------------------------------------------%
-\subsection{Skill mix}
+\subsection{Skill Mix}
 \label{sub:skillmix}
@@
 %----------------------------------------------------------------------------------%
-\subsection{Travel times}
+\subsection{Travel Times}
 \label{sub:traveltimes}
@@
 %----------------------------------------------------------------------------------%
-\subsection{Temporal constraints}
+\subsection{Temporal Constraints}
 \label{sub:temporalconstraints}
@@
 
-\paragraph{Total waiting time}\ \\
-The total waiting time in a solution $\bar{W}$, calculated as:
+\subsubsection*{1. Total waiting time}
+\noindent The total waiting time in a solution $\bar{W}$, calculated as:
 \begin{equation}
@@
 
-\paragraph{Total travel time}\ \\
-If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
+\subsubsection*{2. Total travel time}
+\noindent If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
 \begin{equation}
@@
 
-\paragraph{Total tardiness}\ \\
-The total tardiness in a solution, $\bar{Z}$, is simply the sum of the tardiness across all jobs:
+\subsubsection*{3. Total tardiness}
+\noindent The total tardiness in a solution, $\bar{Z}$, is simply the sum of the tardiness across all jobs:
 \begin{equation}
@@
 
-\paragraph{Maximum tardiness}\ \\
-The maximum tardiness in a solution, $Z_{max}$, is given by:
+\subsubsection*{4. Maximum tardiness}
+\noindent The maximum tardiness in a solution, $Z_{max}$, is given by:
 \begin{equation}
-    Z_{max} = \max_{j \in J} z_j
+    Z_{max} = \max_{j \in J} z_j.
 \end{equation}
 
-\paragraph{Total Overtime}\ \\
-The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$.
+\subsubsection*{5. Total overtime}
+\noindent The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$.
 If we denote the arrival of a nurse $i$ to its end point by $arr(i)$:
@@
 
-\paragraph{Workload balance}\ \\
-In order to produce practical solutions, maintaining a balance in the workload is fundamental.
+\subsubsection*{6. Workload balance}
+\noindent In order to produce practical solutions, maintaining a balance in the workload is fundamental.
 To achieve this, we quantify the minimum spare time on any nurse route in the schedule, $\bar{B}$.
@@
 
-\paragraph{Total preference score}\ \\
-The total preference score in the solution, $\bar{P}$, is computed as:
+\subsubsection*{7. Total preference score}
+\noindent The total preference score in the solution, $\bar{P}$, is computed as:
 \begin{equation}
@@
     \begin{tabular}{ccc}
-    \hline
+    \toprule
     \textbf{Weight}  & \textbf{Value} & \textbf{Description} \\
-    \hline
+    \midrule
     $\alpha_1$      &   -1   &   Travel time \\
@@
     $\alpha_7$      &   0   &   Maximum tardiness\\    
-    \hline
+    \bottomrule
     \end{tabular}
@@
 \begin{itemize}
-    \item PR01 Perform path relinking between every generated solution and another solution selected randomly from the pool.
-    \item PR02 Perform path relinking between every new solution and all solutions in the pool.
-    \item PR03 Do not perform path relinking during the algorithm.
+    \item PR01: Perform path relinking between every generated solution and another solution selected randomly from the pool.
+    \item PR02: Perform path relinking between every new solution and all solutions in the pool.
+    \item PR03: Do not perform path relinking during the algorithm.
 \end{itemize}
@@
 %----------------------------------------------------------------------------------%
-\subsection{Testing data}
+\subsection{Testing Data}
 \label{sub:testing}
@@
 %----------------------------------------------------------------------------------%
-\subsection{Parameter tuning}
+\subsection{Parameter Tuning}
 \label{sub:parametertuning}
@@
     \begin{tabular}{cp{5.5cm}ccc}
-    \hline
+    \toprule
     \textbf{Parameter}& \textbf{Explanation}  & \textbf{Type} & \textbf{Tested} & \textbf{Best} \\
-    \hline
+    \midrule
     NE03 & Use a local search neighbourhood to change nurse order & B & Yes, No & Yes \\ % no_change_ls
@@
     NurseChange & Change nurse order between GRASP iterations & B & Yes, No & Yes \\ % no_change_grasp
-    \texttt{pr\_strategy}   &  Path relinking strategy, as listed in Section~\ref{sub:pathrelinking}& C & 1, 2, 3   &   1 \\ %  
-    \texttt{pr\_direction}  &   Path relinking direction: forward (1), backward (2), forward and backward (3), chosen randomly at each iteration (4).& C & 1, 2, 3, 4       &   1 \\ % performPathRelinking
+    \texttt{pr\_strategy} & Path relinking strategy, as listed in Section~\ref{sub:pathrelinking} & C & 1, 2, 3 & 1 \\ %  
+    \texttt{pr\_direction} & Path relinking direction: forward (1), backward (2), forward and backward (3), chosen randomly at each iteration (4) & C & 1, 2, 3, 4 & 1 \\ % performPathRelinking
     \texttt{pool\_size} & Number of solutions in pool & I & [1,100] & 19 \\ % sols_in_pool
     \texttt{grasp\_dl} & Minimum value of $\delta$ & R & [0,0.5] & 0.25 \\ % grasp_dl
-    \texttt{grasp\_dr}              & Range of $\delta$ &   R & [0,1]                                               & 0.81\\ % grasp_dr
+    \texttt{grasp\_dr} & Range of $\delta$ & R & [0,1] & 0.81 \\ % grasp_dr
     \texttt{rcl\_strategy} & Strategy for generating the reduced candidate list & C & 1,2 & 2 \\ % rcl_strategy    
-    \hline
+    \bottomrule
     \end{tabular}
@@
 \label{sub:literaturedata}
-\noindent While the problems reported by \cite{mankowska2014} and \cite{aithaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to considered.
+\noindent While the problems reported by \cite{mankowska2014} and \cite{aithaddadene2016} are very similar to the one we describe, there are slight variations in how the problem is modelled that need to be considered.
 \cite{mankowska2014} describes the following objective function:
@@
 Tardiness is not allowed for interdependent jobs, where the dependency gap is a hard constraint.
-Further, the synchronisation is simplified for jobs with two skills, where the skill with the lower index must always be performed earlier than the skill with the largest index.
+Furthermore, the synchronisation is simplified for jobs with two skills, where the skill with the lower index must always be performed earlier than the skill with the higher index.
 Tardiness does not include overtime for the workers either, which is not mentioned in the model, so we have not set a limit for working hours of staff when solving these instances.
@@
     \centering
-    \caption{Weights used in the objective function to match those used in the methods from  \cite{aithaddadene2016} (AH) and \cite{mankowska2014} (MK).
+    \caption{Weights used in the objective function to match those used in the methods from  \citet{aithaddadene2016} (AH) and \citet{mankowska2014} (MK).
 A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, i.e.
@@
     \begin{tabular}{ccc}
-    \hline
+    \toprule
     \textbf{Weight}  & \textbf{Values for AH} & \textbf{Values for MK} \\
-    \hline
+    \midrule
     $\alpha_1$      &   $-0.3\times60$          &  $-\frac{1}{3}\times60$   \\
@@
     $\alpha_7$      &   -                       &  $-\frac{1}{3}\times60$             \\    
-    \hline
+    \bottomrule
     \end{tabular}
@@
 \label{sub:algorithmperformance}
-\noindent While the problem description of our DST is slightly different from the examples seen in Section~\ref{sub:literaturedata}, we test the performance of GRASP algorithm, with the required modifications, in the instances provided by \cite{aithaddadene2016} and \cite{mankowska2014}.
+\noindent While the problem description of our DST is slightly different from the examples seen in Section~\ref{sub:literaturedata}, we test the performance of the GRASP algorithm, with the required modifications, in the instances provided by \cite{aithaddadene2016} and \cite{mankowska2014}.
+For all the instances, we perform 5 runs and report the best and the average result.
+The full computational results and solving times are listed in the Appendix tables \ref{table:mkresults}, \ref{table:aithresults}, and \ref{table:longerexperiments}.
 
-For all the instances, we perform 5 runs and report the best and the average result.
-The full computational results and solving times are listed in the Appendix tables \ref{table:mkresults}, \ref{table:aithresults} and \ref{table:longerexperiments}.
-
-In Table~\ref{table:grouptableaithcombined} we present a summary of the results found by our DST for the three instance groups presented in \cite{AitHaddadene2016}.
+In Table~\ref{table:grouptableaithcombined} we present a summary of the results found by our DST for the three instance groups presented in \cite{aithaddadene2016}.
 The results are the average of the best and average results found for the instances of the three instance groups \emph{G1}, \emph{G2} and \emph{G3}.
@@
 \noindent These results suggest that our DST can find better objective values within reasonable planning times.
-Further, we are able to replicate the 23 optimal solutions provided in the paper.
-The average results show that GRASP is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.
+In addition, we are able to replicate the 23 optimal solutions provided in the paper.
+The average results show that \comp{GRASP} is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.
 
-Similarly, in Table~\ref{table:grouptablemkavgbest} we present the comparison of our DST against the performance of the AVNS algorithm presented in \cite{mankowska2014}.
+Similarly, in Table~\ref{table:grouptablemkavgbest} we present the comparison of our DST against the performance of the AVNS algorithm presented in \citet{mankowska2014}.
 
@@
 	\centering
-	\caption{Performance of the GRASP algorithm (average of best and average results) compared to the AVNS algorithm presented in \cite{mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}}
+	\caption{Performance of the \comp{GRASP algorithm} (average of best and average results) compared to the AVNS algorithm presented in \citet{mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}.
+Figures in bold indicate the best results for each instance group.
+\comp{mk only 2dp, why?}}
 	\begin{tabular}{c|c|ccc}
-		\hline
+		\toprule
 		        &  \emph{mk\_avns}    &   \multicolumn{3}{c}{GRASP} \\
-		Group   &   Result    &   Best        &   Avg         &   Time limit\\ \hline
+		Group   &   Result    &   Best        &   Avg         &   Time limit\\
+		\midrule
 		\emph{A}    &   \textbf{225.6}  &   226.41  &   226.41      &   5 s\\
@@
 		\emph{C}    &   713.6   &   \textbf{628.15} &   637.59      &   1 min\\
-		\emph{D}    &   930.3   &   \textbf{798.8}  &   814.53      &   5 min\\
-		\emph{E}    &   1064.7  &   \textbf{880.06} &   901         &   15 min\\
-		\emph{F}    &   1588    &   \textbf{1405.68}&   1491.09     &   30 min\\
-		\emph{G}    &   2161.2  &   \textbf{2042.1} &   2341.89     &   3 h\\
-
-		\hline
+		\emph{D}    &   930.3   &   \textbf{798.80}  &   814.53      &   5 min\\
+		\emph{E}    &   1064.7  &   \textbf{880.06} &   901.00       &   15 min\\
+		\emph{F}    &   1588.0    &   \textbf{1405.68}&   1491.09     &   30 min\\
+		\emph{G}    &   2161.2  &   \textbf{2042.10} &   2341.89     &   3 h\\
+		\bottomrule
 	\end{tabular}
@@
 
-GRASP clearly outperforms AVNS in the majority of instances, especially on those of larger sizes.
+\comp{GRASP} clearly outperforms AVNS in the majority of instances, especially on those of larger sizes.
 Further, our algorithm is able to replicate most of the proven optimal solutions (9 of the 10 instances for which \cite{mankowska2014} provides an optimal solution).
@@
 Nevertheless, the loss in quality is under $5\%$ and the route itself found by the two algorithms is the same (but not the scheduled start times).
-While this effect might still occur in larger instances, it is not evident from our results, suggesting that the efficiency gains seem to outweigh the potential decrease in quality.
+While this effect may still occur in larger instances, it is not evident from our results, suggesting that the efficiency gains seem to outweigh the potential decrease in quality.
 
@@
 \noindent While the previous section sheds some light on the algorithmic performance of the GRASP algorithm and, in general, of our ability to solve efficiently HHCRSP problems of size similar to those faced by district nursing teams in the UK, it does not provide any insight on the actual applicability of these solutions.
-In this section we focus on the instances from \cite{aithaddadene2016} that are solved to optimality (with up to 45 patients) and analyse how these solutions would look like from the perspective of a team of nurses undertaking the job.
-This means asking questions like: Is the workload fairly distributed?
+In this section, we focus on the instances from \cite{aithaddadene2016} that are solved to optimality (with up to 45 patients) and analyse how these solutions would appear from the perspective of a team of nurses undertaking the task.
+This involves asking questions such as: Is the workload fairly distributed?
 How long does it take to complete the shift?
@@
 \label{sec:conclusion}
-\noindent In this article we propose a Decision Support Tool for the Home Health Care Routing and Scheduling Problem that is daily faced by district nursing teams.
+\noindent In this article we propose a Decision Support Tool for the Home Health Care Routing and Scheduling Problem that is faced daily by district nursing teams.
 We propose a model that includes the most common constraints observed in practice (skills, time windows, synchronisation, double services) and objectives (travel time, waiting time, assignment preferences, workload balance, tardiness, overtime).
 This comprehensive approach provides flexibility in the modelling of the working environment in order to be applicable in general real settings.
-This allows for modelling a variety of situations, for example, teams with large workloads might find that overtime and tardiness are acceptable but should be penalised, while other teams with less workload might find those solutions infeasible and prefer to weight more patient preferences instead.
-At the core of the DST is a GRASP algorithm with Path Relinking that is capable to solve the underlying NP-Hard problem successfully for the usual instance sizes observed in practice.
-We show that the DST is competitive and outperforms other methods in the state-of-the-art literature, and it is able to identify most of the optimal solutions known for the problem.
+This allows for modelling a variety of situations, for example, teams with large workloads might find that overtime and tardiness are acceptable but should be penalised, whilst other teams with less workload might find those solutions infeasible and prefer to weight more patient preferences instead.
+
+At the core of the DST is a GRASP algorithm with Path Relinking that is capable of solving the underlying NP-hard problem successfully for the usual instance sizes observed in practice.
+We show that the DST is competitive and outperforms other methods in the state-of-the-art literature, and is able to identify most of the optimal solutions known for the problem.
 Furthermore, we show that current literature ignores route aspects that are important to decision makers, such as waiting time.

### 2021-10-25 tex/Paper/combined_table_aith_results.tex
files for paper.
updated citations and format.

@@
 \centering
-    \caption{Performance of the best and average results found by GRASP DST compared with the three algorithms presented in \cite{AitHaddadene2016} (ILS, \emph{ah\_ils}, basic GRASP,  \emph{ah\_grasp}, and GRASPxILS, \emph{ah\_gi}) for the three instance groups \emph{G1}, \emph{G2} and \emph{G3}.}
+    \caption{Performance of the best and average results found by GRASP DST compared with the three algorithms presented in \citet{aithaddadene2016} (ILS, \emph{ah\_ils}, basic GRASP,  \emph{ah\_grasp}, and GRASPxILS, \emph{ah\_gi}) for the three instance groups \emph{G1}, \emph{G2} and \emph{G3}.
+Figures in bold indicate the best results for each instance group.
+\comp{Should rename \textit{GRASP} so that it is clear that it is the algorithm from this paper.}}
     \begin{tabular}{cccccc}
-    \label{tab:group_table_aith_combined}\\\hline
-
-Result  &   Group   &   \emph{ah\_ils}  &   \emph{ah\_grasp}    &   \emph{ah\_gi}   &   \emph{GRASP}    \\  \hline
+    \toprule
+Result  &   Group   &   \emph{ah\_ils}  &   \emph{ah\_grasp}    &   \emph{ah\_gi}   &   \emph{GRASP}    \\  \midrule
 \multicolumn{6}{l}{\emph{Best}} \\                                              
@@
     &   \emph{G3}   &   -251.35 &   -231.54 &   -245.42 &   \textbf{-251.55}    \\  
-\multicolumn{6}{l}{\emph{Avg.}} \\                                              
+\multicolumn{6}{l}{\emph{Average}} \\                                              
     &   \emph{G1}   &   42.27   &   38.39   &   38.05   &   \textbf{38.04}  \\  
@@
     &   \emph{G3}   &   -240.82 &   -226.79 &   -239.44 &   \textbf{-247.01}    \\  
-    \hline
+    \bottomrule
     \end{tabular}
+\label{table:grouptableaithcombined}
 \end{table}

### 2021-10-25 tex/Plan/Plan.tex
updated project plan - weekly.

@@
 
+\begin{enumerate}[leftmargin=*]
+	\item \textbf{!8th October 2021 -- 31st October 2021}\\
+	Finish first paper
+	\item \textbf{1st November 2021 -- 7th November 2021}\\
+	Start on research for stochastic/exact method
+	\item \textbf{8th November 2021 -- 14th November 2021}\\
+	Continue research for stochastic/exact method
+	\item \textbf{15th November 2021 -- 21st November 2021}\\
+	Mee with Abicare for analysis of solutions, prep DST for release, finish documentation
+	\item \textbf{22nd November 2021 -- 28th November 2021}
+	Release DST and documentation, give DST to Abicare, send questionnaire
+	\item \textbf{29th November 2021 -- 5th December 2021}
+	Research for stochastic/exact method
+	\item \textbf{6th December 2021 -- 12th December 2021}
+	Research for stochastic/exact method
+	\item \textbf{13th December 2021 -- 19th December 2021}
+	Research for stochastic/exact method
+	\item \textbf{20th December 2021 -- 26th December 2021}
+	Start creating model for stochastic/exact method
+	\item \textbf{3rd January 2022 -- 9th December 2022}
+	Check and begin to implement model for stochastic/exact method
+	\item \textbf{10th January 2022 -- 16th January 2022}
+	Implement model into DST, debug
+	\item \textbf{17th January 2022 -- 23rd January 2022}
+	Start gathering data/creating instances to test model in DST
+	\item \textbf{24th January 2022 -- 30th January 2022}
+	Start testing model in DST
+	\item \textbf{31st January 2022 -- 6th February 2022}
+	Analyse data, begin paper
+	\item \textbf{7th February 2022 -- 13th February 2022}
+	Papers
+	\item \textbf{14th February 2022 -- 20th February 2022}
+	Papers
+	\item \textbf{21st February 2022 -- 27th February 2022}
+	Papers
+	\item \textbf{28th February 2022 -- 6th March 2022}
+	Papers
+	\item \textbf{7th March 2022 -- 13th March 2022}
+	Gather results from Abicare, compare, send another questionnaire
+	\item \textbf{14th March 2022 -- 20th March 2022}
+	Papers
+	\item \textbf{21st March 2022 -- 27th March 2022}
+	Papers
+	\item \textbf{28th March 2022 -- 3rd April 2022}
+	Papers
+	\item \textbf{4th April 2022 -- 10th April 2022}
+	Papers
+	\item \textbf{11th April 2022 -- 17th April 2022}
+	Papers
+	\item \textbf{18th April 2022 -- 24th April 2022}
+	Papers
+	\item \textbf{25th April 2022 -- 30th April 2022}
+	Conclude project.
+
+\end{enumerate}
+
 \end{document}

### 2021-12-18 tex/Documentation/Documentation.tex
updated documentation.

--- 2021-06-28 tex/Documentation/Documentation.tex
@@
 \input{includes/preamble.tex}
+\usepackage{parskip}
 
 \begin{document}
-\title{Documentation: Abicare DST}
+\title{\vspace{-25mm}Quick Start Guide: Carrot DST\vspace{-15mm}}
 %\author{Asyl L. Hawa}
@@
 
-\section{Disruption Management for the RT-HCSRP}
+This is a quick start guide for the Carrot (CARe ROuting Tool) DST. It will allow the user to run the program for a given input file using default settings.
 
-\subsection{Differences}
-\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
-\end{itemize}
+\section{Quick Start Guide}
+\begin{enumerate}[leftmargin=*]
+	\item Double-click on the \texttt{runprogram.bat} file to start the program. Two command prompts will appear:
+	\begin{enumerate}
+		\item The first window is for OSRM (Open Source Routing Machine). The last line in the window should be `[info] running and waiting for requests'. This line may take a few moments to appear.
+		\item The second window will show some outputs from the program, and will initially appear blank.
+	\end{enumerate}
+	\item A third window will then appear titled `User Input Parameters'. It may take some time for this window to pop up due to the packaging of the program. An image of this window is provided below.
+	\item In this window, simply select the desired area name in the corresponding field.
+If preferred, the time limit can also be altered -- currently it is set to 60 seconds.
+For `Select input file', use the `Browse' button to select the desired Excel input file containing data regarding clients and carers.
+The input file must be in the same format as the original Abicare file, of which a copy has been provided in the folder for reference.
+Finally, select a date for which a schedule and route is required using the `Calendar' button.
+All other fields may remain as they are.
+	\item Double-click `OK' on the main `User Input Parameters' window, which will start the algorithm.
+Another window will appear outlining the progress of the program.
+Some outputs from the program will also be printed to the second command prompt window.
+	\item When the program has terminated, all windows will close except the OSRM command prompt. This window can be closed manually on completion of the program.
+	\item All outputs from the program will be stored in a new folder titled `outputs` in the main CarrotDST folder.
+\end{enumerate}
 
-\subsection{Model and Solution Approach}
-\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item One
-	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
-		\item 
-	\end{enumerate}
-	\item Two
-	\begin{enumerate}[label=(\roman*),itemsep=-0.1em]
-		\item
-	\end{enumerate}
-	\item 
-\end{itemize}
-
-\subsection{Modifications}
-\begin{itemize}[label=\textcolor{myPurple}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
-\end{itemize}
-
-\section{Disruption Management of the VRP with Breakdown}
-
-\subsection{Differences}
-\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
-\end{itemize}
-
-\subsection{Model and Solution Approach}
-\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
-\end{itemize}
-
-\subsection{Modifications}
-\begin{itemize}[label=\textcolor{myYellow}{\textbullet},leftmargin=*,itemsep=-0.1em]
-	\item 
-\end{itemize}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.72\linewidth]{figures/userinput}%
+	\caption{User Input Parameters window.}
+\end{figure}
 

### 2021-12-18 tex/Paper/main.tex
updated files.

@@
 
+\author[1]{Asyl L. Hawa}
+\ead{A.Hawa@soton.ac.uk}
+
 \begin{abstract}
 The Home Health Care Routing and Scheduling Problem (HHCRSP) consists in assigning routes and schedules to nurses that deliver health care services to patients at their homes.
-A varied range of hard and soft constraints such as time windows, synchronisation of nurses, skill mix or patient and nurse preferences make this problem particularly difficult to solve manually.
+A varied range of hard and soft constraints such as time windows, synchronisation of nurses, skill mix, and patient and nurse preferences make this problem particularly difficult to solve manually.
 However, this is a challenge faced daily by many community nursing teams.
@@
 We compare the performance of the GRASP algorithm with state-of-the-art approaches for similar problems and show that it offers competitive results, while providing a wider range of possibilities when modelling real problems.
-For instance, the computational results show that considering waiting times in the objective function can reduce total operation times up to 19\% in comparison with alternative modelling approaches that minimise travel time only.
+For instance, the computational results show that considering waiting times in the objective function can reduce total operation times up to 24\% in comparison with alternative modelling approaches that minimise travel time only.
 
@@
 
-A key role in the health service to deliver this care are District Nurses (DN), specialised nursing teams that travel to patient's homes and perform a variety of tasks, ranging from insulin injections to dressing wounds.
+A key role in the health service to deliver this care are District Nurses (DN), specialised nursing teams that travel to patients' homes and perform a variety of tasks, ranging from insulin injections to dressing wounds.
 Nevertheless, DN numbers have fallen by 44\% \citep{nhs2018} between 2010 and 2018 in the NHS.
 On the other hand, while there is no official data on demand, the workforce perception is that demand has risen significantly, putting the service under pressure \citep{robertson2017}.
-There is, therefore, a strong need for efficiency within the community nursing service, to better utilise the available resources.
+There is, therefore, a strong need for efficiency within the community nursing service to better utilise the available resources.
 
 One of the first tasks that the DN team lead faces on a working day is to distribute the work for the team.
-In the DN teams we visited in the city of Southampton, UK, this entailed assigning a number of home health care visits (typically over a hundred) to the nurses available on the team, attending to hard requirements (skill matching, time windows, job dependencies) and soft ones (minimising travel and waiting times, maintaining workload balance, matching carer-patient preferences).
+In the DN teams we visited in the city of Southampton, UK, this entailed assigning a number of home health care visits (typically over a hundred) to the nurses available on the team, attending to hard requirements (skill matching, time windows, job dependencies) and soft ones (minimising travel and waiting times, maintaining workload balance, matching nurse-patient preferences).
 Despite the abundant literature related to similar problems, the planning is currently performed manually in a lengthy process for the planner that potentially leads to suboptimal schedules.
 Manual planning has three further weaknesses.
-First there are inconsistencies between the different plans of a nurse, and between plans of different nurses.
-Although a single DN may become highly skilled in planning overtime that knowledge is lost when they leave the organisation.
-Second, a poor planning results in turn in a poor utilisation of the workforce, which might include excessive travel and waiting times, unbalanced workloads.
-Third, it can also affect the quality of care, leading to missed time windows or lack of preference matching for patients.
+First, there are inconsistencies between the different plans of a nurse, and between plans of different nurses.
+Although a single DN may become highly skilled in planning over time, that knowledge is lost when they leave the organisation.
+Second, poor planning results in poor utilisation of the workforce, which may include excessive travel and waiting times, and unbalanced workloads.
+Third, maunal planning can affect the quality of care, leading to missed time windows or lack of preference matching for patients.
 
-In an effort to automate and improve this process, we model the planning of home health care visits as a rich vehicle routing problem, with constraints and objectives drawn from our interaction with community nursing teams.
-This problem is often known in the literature as Home Health Care Routing and Scheduling Problem (HHCRSP).
+In an effort to automate and improve this process, we model the planning of home health care visits as a rich vehicle routing problem with constraints and objectives drawn from our interaction with community nursing teams.
+This problem is often known in the literature as the Home Health Care Routing and Scheduling Problem (HHCRSP).
 In our case, the schedule takes place over a single day, as it is the preferred approach in practice.
-This is motivated by the last-minute changes due to staff availability, patient next-day referrals and the nature of continuous review of care, which often prompts to scheduling the next visit right after reviewing their condition on the previous visit.
+This is motivated by the last-minute changes due to staff availability, patient next-day referrals and the nature of continuous review of care, which often prompts to scheduling the next visit right after reviewing the patient's condition on the previous visit.
 A formal description of the problem is provided in Section~\ref{sec:model}.
@@
 We propose a Decision Support Tool (DST) that can produce high quality daily routes and schedules for a DN team, according to a flexible objective function.
-The tool requires as input a patient requirement list and the available staff and then runs an optimisation algorithm based on a weighted objective function.
+The tool requires as input a patient requirement list and the available staff, and then runs an optimisation algorithm based on a weighted objective function.
 The optimisation is based on a Greedy Randomised Adaptive Search Procedure (GRASP) with Path Relinking, which is detailed in Section~\ref{sec:dst}.
@@
 It tackles the most common constraints found in literature (e.g.
-time windows or skills) but also others often found in practice (e.g.
-multiple depots or workload balance).
+time windows and skills) but also others often found in practice (e.g.
+multiple depots and workload balance).
 We provide extensive computational results (Section~\ref{sec:computationalexperiments}) that show that our algorithm is competitive with state-of-the-art approaches and is flexible to allow real-world constraints and objectives; but also provide insight on the algorithm components that help achieve high quality routes and schedules.
@@
 %----------------------------------------------------------------------------------%
-\section{Literature review}
+\section{Literature Review}
 \label{sec:litreview}
@@
 Home Health Care and Home Care services across the world operate in different ways and, as such, the related scheduling problems have different objectives and constraints.
-For example, it might be possible to speed up the visits to some patients \citep{mosquera2019}, to allow nurses to share a minibus for part of their route, while walking the rest of it \citep{fikar2015}, for nurses to use public transport \citep{rest2016}, or to allow time preferences for patients \citep{decerle2018}.
+For example, it might be possible to speed up the visits to some patients \citep{mosquera2019}, to allow nurses to share a minibus for part of their route, whilst walking the rest of it \citep{fikar2015}, for nurses to use public transport \citep{rest2016}, or to allow time preferences for patients \citep{decerle2018}.
 % This variety makes difficult comparing approaches and translating the findings from one application to another.
@@
 These are usually based on a Mixed Integer Linear Program (MILP) formulation that is solved with the help of a commercial solver \citep{bredstrom2008,bachouch2011, mankowska2014, aithaddadene2016}.
-Typically, these formulations are three index formulations that involve two types of variables: $x_{ijk}$ binary variables that indicate if carer $k$ traverses the arc $i$ and $j$; and $t_{ik}$ real variables that determine at what time carer $k$ starts performing job $i$.
+Typically, these formulations are three index formulations that involve two types of variables: $x_{ijk}$ binary variables that indicate if nurse $k$ traverses the arc $i$ and $j$; and $t_{ik}$ real variables that determine at what time nurse $k$ starts performing job $i$.
 These formulations are limited in practice and can only solve to optimality small problems (e.g.
@@
 
-Often, in order to tackle larger instances, the MILPs are solved in a restricted manner and, while they do not guarantee optimality, they might provide feasible solutions suitable for day to day planning.
-This is often referred to as a matheuristic algorithm and can be done by reducing the size of the original problem by fixing some aspects and solving it again.
+Often, in order to tackle larger instances, the MILPs are solved in a restricted manner and, while they do not guarantee optimality, they might provide feasible solutions suitable for day-to-day planning.
+This is often referred to as a matheuristic algorithm, and can be done by reducing the size of the original problem by fixing some aspects and solving it again.
 One example of this technique is given in \cite{rasmussen2012} that also proposes clustering strategies to reduce the size of the problem, which allows them to solve problems with up to $150$ patients, although no longer proving optimality.
-Another is \cite{bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and carers in order to generate a series of restricted MILPs that can be solved in a short time.
+Another is \cite{bredstrom2008}, where they propose a heuristic that iteratively allows or forbids assignments between patients and nurses in order to generate a series of restricted MILPs that can be solved in a short time.
 In addition, \citet{nikzad2021} considers a four-phase matheuristic that calls upon a variety of mixed integer models.
@@
 One of the earliest works is due to \cite{begur1997}, who developed a decision support system based on the Clarke and Wright algorithm \citep{clarke1964} for vehicle routing and then relied on integration with a Geographic Information System (GIS), an interactive user interface and a human planner to alter the routes as necessary.
-The Laps Care system \citep{eveborn2006, eveborn2009} is a Swedish software for the HHCRSP, based on a repeated matching algorithm has been shown to be successful in saving planning time, minimising travel times and even reducing the number of missed visits.
+The Laps Care system \citep{eveborn2006, eveborn2009} is a Swedish software for the HHCRSP based on a repeated matching algorithm, and has been shown to be successful in saving planning time, minimising travel times and even reducing the number of missed visits.
 The authors estimated annual savings in 2009 of up to \euro 120 million across teams in Norway, Finland and Sweden \citep{eveborn2009}.
@@
 %----------------------------------------------------------------------------------%
-\section{A model for health home care services}
+\section{A Model for Health Home Care Services}
 \label{sec:model}
@@
 Each nurse $i$ might have a different starting time, $st(i)$, and end time, $en(i)$, of their working day.
-During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$, a subset of which, $DS \subset J$ are double services, i.e.
-they require two nurses to be present to be performed.
+During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$, a subset of which, $DS \subset J$, are double services; that is, they require two nurses to be present simultaneously.
 Note that jobs refer to individual tasks that a nurse needs to perform and are not necessarily identified with unique patients or locations, meaning that the same patient might have more than one job on the list.
 
-A schedule (not necessarily feasible) $\mathcal{S}$, is made of a pair   $\mathcal{S} = \{\mathcal{R}, \mathcal{O}\}$, where $\mathcal{O}$ is a permutation of the vector $o = (1,\dots,n)$ representing the order of the nurses in the solution and $\mathcal{R}$ is a set of $n$ routes, $\mathcal{R} = \{r_1,\dots,r_n\}$, one per nurse.
-Each route $r_i$ is itself an ordered list of $m_i$ jobs, $r_i = \{k^i_0, k^i_1,\dots,k^i_{m_i}, k^i_0\}$, that are the jobs assigned to nurse $i \in N$ and starts and ends at the depot of that nurse denoted by, $k^i_0$.
-Of course, a route might be empty, meaning that that particular nurse is not assigned any jobs and $m_i = 0$ and the route only contains the depot, $r_i = \{k^i_0, k^i_0\}$.
+A schedule (not necessarily feasible) $\mathcal{S}$ is made of a pair   $\mathcal{S} = \{\mathcal{R}, \mathcal{O}\}$, where $\mathcal{O}$ is a permutation of the vector $o = (1,\dots,n)$ representing the order of the nurses in the solution, and $\mathcal{R}$ is a set of $n$ routes, $\mathcal{R} = \{r_1,\dots,r_n\}$, one per nurse.
+Each route $r_i$ is itself an ordered list of $m_i$ jobs, $r_i = \{k^i_0, k^i_1,\dots,k^i_{m_i}, k^i_0\}$, that are the jobs assigned to nurse $i \in N$ and start and end at the depot of that nurse, denoted by $k^i_0$.
+Of course, a nurse's route may be empty, meaning that the nurse has not been assigned any jobs; thus $m_i = 0$ and the route only contains the depot, $r_i = \{k^i_0, k^i_0\}$.
 After a schedule is generated, for each job $j \in J$ we denote by $t_j$ its starting time.
-Note that this refers to the time at which the job starts, rather than the time at which the nurse (or nurses) performing it arrive to the job location, which might be different.
+Note that this refers to the time at which the job starts, rather than the time at which the assigned nurse (or nurses) arrive to the job location, which may differ.
 
-In the rest of the section, we formally define the constraints and objectives that commonly apply in practice and relate to the skill mix of the workforce, the preferences of some assignments, travel times and synchronisation, visits with two staff members and quality of the schedule.
+In the rest of this section we formally define the constraints and objectives that commonly apply in practice and relate to the skill mix of the workforce, the preferences of some assignments, travel times and synchronisation, visits with two nurses, and the quality of the schedule.
 
@@
 To model this, we define a function $S_1: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise.
-Naturally, for double service jobs, no single nurse is skilled to perform them $S_1(i,j) = 0, \forall i \in N, \forall j \in DS$.
-To account for this, we define a second function $S_2: N\times N\times J \rightarrow \{0,1\}$, that takes value $S_2(i_1,i_2,j) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform job $j$ and $0$ otherwise.
+Naturally, for double service jobs no single nurse is skilled to perform them, $S_1(i,j) = 0$ $\forall i \in N, \forall j \in DS$.
+To account for this, we define a second function, $S_2: N\times N\times J \rightarrow \{0,1\}$, that takes value $S_2(i_1,i_2,j) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform job $j$ and $0$ otherwise.
 
@@
 In particular, it could happen that only one staff member is qualified for all services and the other for none.
-An example of this would be visits that require a second person to provide support and ensure safety, for example with patients that have a record of behaving inappropriately towards staff.
-    \item \textbf{Strictly shared}.
+An example of this in practice would be visits that require a second person to provide support and ensure safety, for example with patients that have a record of behaving inappropriately towards staff.
+    \item \textbf{Strictly Shared}.
 In this case each member of the service does not need to be qualified for all services but should be attending the visit to perform at least one service.
@@
 The data is introduced as a list of skills and, if some skills need to be present in more than one staff member, they can simply appear duplicated on the list.
-We have, however, also implemented the Strictly Shared approach, in order to solve the instances from the literature.
+We have, however, also implemented the \emph{Strictly Shared} approach in order to solve the instances from the literature.
 
 It is not uncommon in the literature to model the skills of the workforce as hierarchical, i.e.
-each nurse is assigned a skill level and it is assumed that they possess all the skills that lower levels have, and some more.
+each nurse is assigned a skill level and it is assumed that they possess all the skills that lower levels have and more.
 The model might then introduce a penalty or limit for high skilled nurses performing too many low skilled jobs, known as downgrading \cite[see, for example,][]{fikar2015}.
-In the DN teams we observed, skills were not hierarchical, neither were any such limits in place.
-Nevertheless, nurses pointed out that high skilled nurses were typically assigned only the high skilled jobs, as they might need to perform other jobs (e.g.
+In the DN teams we observed skills were not hierarchical, neither were any such limits in place.
+Nevertheless, nurses pointed out that high skilled nurses were typically assigned only the high skilled jobs as they might need to perform other jobs (e.g.
 attending meetings or other administrative tasks).
@@
 \label{sub:preferences}
-\noindent It is not uncommon in practice that district nurses do their planning using preferences.
-These might arise from multiple reasons, such as a preserving continuity of care, gender preference, patients and nurses speaking the same language, previous conflicts between a patient and a staff member, \model{etc}.
-We model this with a preference function, $P: N\times J \rightarrow \mathbb{R}$, such that $P(i,j)$ represents the preference score of assigning job $i$ to staff member $j$.
+\noindent It is not uncommon in practice for district nurses to do their planning using preferences.
+These might arise from multiple reasons, such as a preserving continuity of care, gender preference, patients and nurses speaking the same language, or previous conflicts between a patient and a nurse.
+We model this with a preference function, $P: N\times J \rightarrow \mathbb{R}$, such that $P(i,j)$ represents the preference score of assigning job $i$ to nurse $j$.
 The function $P$ should assign a higher (positive) score to preferred patient-nurse assignments and a lower (negative) score to assignments that are better avoided.
@@
 \label{sub:traveltimes}
-\noindent To describe travel times, for each nurse we define $N$ functions $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, where $k^i_0$ refers to the start and end location of the nurse $i$.
-Each nurse might start from a different depot (typically, their own home), a fact that is accounted for in these functions.
-Note that with this definition we allow to have different travel times between jobs for different nurses, and this can be used to model, for example, different means of transport.
-In our observations, however, all nurses used their own vehicles, so we assume that $T_i$ returns driving times for all nurses and, in particular, that:
+\noindent To describe travel times, for each nurse we define $N$ functions, $T_i: J\cup\{k^i_0\} \times J\cup\{k^i_0\} \rightarrow \mathbb{R}$, $i \in N$, where $T_i(j_1, j_2)$ indicates the time it takes nurse $i$ to travel from job $j_1$ to job $j_2$, and $k^i_0$ refers to the start and end location of the nurse $i$.
+Each nurse might start from a different depot (typically their own home), a fact that is accounted for in these functions.
+Note that with this definition we accommodate for different travel times between jobs for different nurses which can be used to model, for example, different means of transport.
+In our observations, however, all nurses used their own vehicles; thus we assume $T_i$ returns driving times for all nurses and, in particular, that:
 \begin{equation}
-    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2),\ \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J. \label{eqn:traveltimequality}
+    T_{i_1}(j_1, j_2) = T_{i_2}(j_1, j_2)\quad  \forall i_1, i_2 \in N,\ \forall j_1, j_2 \in J. \label{eq:traveltimeequality}
 \end{equation}
 
-\noindent Note that this function can be substituted by a more sophisticated one without any changes to the algorithms described here.
+\noindent Note that this function can be substituted by a more sophisticated one without any changes to the algorithms described in the remainder of the paper.
 
@@
 \label{sub:temporalconstraints}
-\noindent Each job $j$ has a service time $s_j$ and a time window $t_j$ that is defined by its earliest start time, $e_j$ and the latest start time $l_j$.
+\noindent Each job $j$ has a service time $s_j$ and a time window $t_j$ that is defined by its earliest start time, $e_j$, and the latest start time, $l_j$.
 It is a hard constraint that jobs cannot start before their earliest start time:
-\begin{align}
-    t_j \geq e_j, & \forall j \in J, \label{eqn:jobstart}
-\end{align}
+\begin{equation}
+    t_j \geq e_j \quad \forall j \in J, \label{eq:jobstart}
+\end{equation}
 
@@
 \begin{equation}
-    z_j = \max(0, t_j - l_j). \label{eqn:tardiness}
+    z_j = \max(0, t_j - l_j). \label{eq:tardiness}
 \end{equation}
@@
 \noindent Note that tardiness here refers only to the difference between the job start and its original time window.
-Jobs that have dependencies or require two nurses might be further restricted temporally.
-This might happen in practice, for example, in situations where a certain medication must be administered \model{a few hours (a specific time period/amount of time)} after the first dose.
+Jobs that have dependencies or require two nurses may be further restricted temporally.
+This can occur in practice, for example, in situations where a second dose of medication must be administered a few hours after the first dose.
 If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$ that depends on the start time of $j_1$, $[e^{(j_1)}_{j_2} + t_{j_1}, l^{(j_1)}_{j_2} + t_{j_1}]$.
@@
 	\begin{align}
-	t_{j_2} &\geq e^{(j_1)}_{j_2} + t_{j_1}, \label{eqn:timewindow1}\\[3pt] % & \forall j \in J
-	t_{j_2} &\leq l^{(j_1)}_{j_2} + t_{j_1}. \label{eqn:timewindow2} % & \forall j \in J
+	t_{j_2} &\geq e^{(j_1)}_{j_2} + t_{j_1}, \label{eq:timewindow1}\\[3pt] % & \forall j \in J
+	t_{j_2} &\leq l^{(j_1)}_{j_2} + t_{j_1}. \label{eq:timewindow2} % & \forall j \in J
 	\end{align}
@@
 	\begin{align}
-	e^{(j_1)}_{j_2} &= -e^{(j_2)}_{j_1}, \label{eqn:earliesttimedependecies}\\[3pt]
-	l^{(j_1)}_{j_2} &= -l^{(j_2)}_{j_1}. \label{eqn:latesttimedependencies}
+	e^{(j_1)}_{j_2} &= -e^{(j_2)}_{j_1}, \label{eq:earliesttimedependecies}\\[3pt]
+	l^{(j_1)}_{j_2} &= -l^{(j_2)}_{j_1}. \label{eq:latesttimedependencies}
 	\end{align}
@@
 \begin{equation}
-    \tilde{z}_{j_2} = \max(0, t_{j_2} - (l^{(j_1)}_{j_2} + t_{j_1})). \label{eqn:gaptardiness}
+    \tilde{z}_{j_2} = \max(0, t_{j_2} - (l^{(j_1)}_{j_2} + t_{j_1})). \label{eq:gaptardiness}
 \end{equation}
 
-\noindent \model{Since some jobs have time windows (only \textit{some}?)} or might need to be synchronised, waiting times may occur.
+\noindent Since some jobs have time windows or might need to be synchronised, waiting times may occur.
 We denote by $w_{ij}$ the waiting time of nurse $i$ at job $j$.
@@
 \begin{equation}
-    a(i,j) = st(i) + \sum_{h=0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1}),
+    a(i,j) = st(i) + \sum_{h=0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1}), \label{eq:arrivaltime}
 \end{equation}
@@
 Note that there are no waiting or service times at the depot; that is, $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 $ for all $i \in N$.
+Furthermore, let $en_{\text{act}}(i)$ be the actual end time of a nurse $i \in N$, that is:
+\begin{equation}
+	en_{\text{act}}(i) = a(i,m_i) + w_{ik_{m_i}^i} + s_{k_{m_i}^i}, \label{eq:actualendtime}
+\end{equation}
+where $m_i$ is the final job in nurse $i$'s route $r_i$.
 
 In practice, waiting times may occur for a variety of reasons.
-In order to maintain simplicity in our model, we only consider three reasons for which a nurse $i$ might have waiting time $w_{ij} \geq 0$ at a job $j \in r_i$:
+In order to maintain simplicity in our model, we only consider four reasons for which a nurse $i$ might have waiting time $w_{ij} \geq 0$ at a job $j \in r_i$:
 \begin{itemize}
     \item Respecting a time window: 
-    \begin{align}
-        w_{ij} \geq e_i - a(i,j) & \forall i \in N, \forall j \in J.
-    \end{align}
+    \begin{equation}
+        w_{ij} \geq e_j - a(i,j) \quad \forall i \in N, \forall j \in J. \label{eq:waiting1TW}
+    \end{equation}
     \item Waiting for a second nurse $i'$ to arrive for a double service:
-    \begin{align}
-        w_{i j} \geq t_j - a(i',j) & \forall j \in DS,\ such\ that\ j \in r_{i} \cap r_{i'}.
-    \end{align}
+    \begin{equation}
+        w_{i j} \geq t_j - a(i',j) \quad \forall j \in DS \text{ such that } j \in r_{i} \cap r_{i'}. \label{eq:waiting2DS}
+    \end{equation}
     \item Waiting for the minimum gap of a synchronised service. If the job $j$ depends on $j'$, then:
-    \begin{equation}\label{eqn:waitingforsynchronised}
+    \begin{equation}\label{eq:waiting3dependent}
         w_{ij} \geq e^{(j')}_{j} - a(i,j).
     \end{equation}
+	\item Waiting time incurred due to breaks between shifts. This is discussed further in Section~\ref{sub:shiftwork}.
 \end{itemize}
@@
 To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based on the order in which they appear in the solution.
-\correction{The equations below are for a single nurse/job only} Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, where $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$, and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$.
-Then, we enforce that:
+Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, where $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$, and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$.
+Then, we enforce the following: \correction{Does this still hold?}
 %  $i_1 < i_2$, \ie $j_1$ appears earlier in the vector $\mathcal{O}$ than $j_2$. Then, we enforce that:
-\begin{equation}\label{eqn:earliestpossiblestartassumption1}
-    w_{i_1 j_1} = \max \{0, e_{j_1} - a(i_1,j_1)\}
-\end{equation}
-and
-\begin{equation}\label{eqn:earliestpossiblestartassumption2}
-    t_{j_1} = \max \{ a(i_1,j_1), e_{j_1} \}
-\end{equation}
-
+\begin{align}
+	w_{i_1 j_1} &= \max \{0, e_{j_1} - a(i_1,j_1)\}, \\ 
+		t_{j_1} &= \max \{ a(i_1,j_1), e_{j_1} \}.\label{eq:earliestpossiblestartassumption}
+\end{align}
+%\begin{equation}\label{eq:earliestpossiblestartassumption1}
+%    w_{i_1 j_1} = \max \{0, e_{j_1} - a(i_1,j_1)\},
+%\end{equation}
+%\begin{equation}\label{eq:earliestpossiblestartassumption2}
+%    t_{j_1} = \max \{ a(i_1,j_1), e_{j_1} \}.
+%\end{equation}
 \noindent Note that this implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
 To mitigate this effect, in Section~\ref{sec:dst} we describe strategies to search different orderings of nurses when solving the problem.
+
+%----------------------------------------------------------------------------------%
+\subsection{Shift Work}
+\label{sub:shiftwork}
+\noindent There may be cases where some nurses have multiple shifts in a single day, e.g.
+a morning shift and an evening shift; these nurses will not be available at all times during the day.
+Let $\Pi_i = \{\pi_1^i,\dotsc, \pi_{g_i}^i\}$ be the set of shifts for a nurse $i \in N$, where $g_i$ is the total number of shifts for nurse $i$ during the day.
+For each shift $\pi_f^i \in \Pi_i$, we denote the shift start time by $u(\pi_f^i)$ and the shift end time by $v(\pi_f^i)$.
+Therefore, the total contracted working time, $ct(i)$, for a nurse $i$ is calculated as:
+\begin{equation}
+	ct(i) = \displaystyle\sum_{f=1}^{g_i} v(\pi_f^i) - u(\pi_f^i). \label{eq:contractedworkingtime}
+\end{equation}
+\noindent Clearly, if $g_i = 1$ for a nurse $i \in N$, then the nurse only has one shift; thus $ct(i) = en(i)-st(i)$.
+Given these shifts, we must ensure that jobs are assigned to nurses during their actual working time.
+This also involves considering two cases that may occur:
+\begin{enumerate}[label={(\arabic*)}]
+	%\item A nurse $i$ is assigned a job $j$ with time window within the nurse's valid shift time, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period.
+In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the beginning of the nurse's next shift, and the start time for $j$ is moved forward to the start of the next shift.
+The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
+	%\item A nurse $i$ is assigned a job $j$ with time window within nurse $i$'s break period.
+In this case, the waiting time incurred is the entire break period, and the arrival time at $j$ is moved forward to the start of the next shift.
+Tardiness as calculated as above.
+	\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period.
+In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the end of the shift, and the start time for $j$ is moved forward to the start of $i$'s next shift.
+The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
+	\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time.
+In this case, no waiting time is incurred and the new start time for job $j$ and tardiness are as above.
+\end{enumerate}
+\begin{figure}[h!]
+	\centering	
+	\begin{subfigure}[h]{\textwidth}
+		\centering
+		\includestandalone[width=0.7\textwidth]{figures/shiftwork1}
+		\vspace{-2mm}\caption{}
+		\label{fig:itemsinfeas}
+		\vspace{2mm}
+	\end{subfigure}
+	\begin{subfigure}[h]{\textwidth}
+		\centering
+		\includestandalone[width=0.7\textwidth]{figures/shiftwork2}
+		\vspace{-2mm}\caption{}
+		\label{fig:itemsfeas}
+	\end{subfigure}
+	\caption{Examples of the alteration of a job $j$ depending on a nurse $i$'s shift times according to the scenarios above, when (a) the duration of $j$ extends past the end of $i$'s shift; and (b) the start time of $j$ occurs during a period of unavailable time.}
+	\label{fig:itemsaligned}
+\end{figure}
+%\correction{What happens if the next "available" slot is not enough for the job.}
+\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
+For simplicity, we will consider all nurses' break periods as waiting time, so that the arrival time calculations $a(i,j)$ for nurses $i \in N$ at jobs $j \in J$ (see \eqref{eq:arrivaltime}) remain valid.
+Then, the waiting times will be corrected by removing all relevant breaks, as shown in \eqref{eq:waitingtimenurse}.
+
+%There may be cases where some nurses have multiple shifts in a single day, e.g.
+a morning shift and an evening shift; these nurses will not be available at all times during the day.
+Therefore, we must ensure that jobs are assigned to nurses during their actual working times.
+This also involves considering the case where although a nurse may be working and available at the time that a job starts, the duration of the job may extend into the nurse's unavailable time, i.e.
+the job ends after the nurse's shift ends.
+In these instances, the job must be moved forward in the nurse's schedule such that the job starts at the beginning of the nurse's next available shift, and tardiness will be incurred.
+For convenience, we denote by $ct(i)$ the actual contracted working time of nurse $i \in N$, where $ct(i) = en(i) - st(i)$ if nurse $i$'s day comprises a single shift starting at $en(i)$ and ending at $st(i)$, and $ct(i) < en(i) - st(i)$ if nurse $i$'s day consists of one or more shifts.
 
@@
 
-\subsubsection*{1. Total waiting time}
-\noindent The total waiting time in a solution $\bar{W}$, calculated as:
-\begin{equation}
-     \bar{W} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}.
-\end{equation}
-
-\subsubsection*{2. Total travel time}
+\subsubsection*{1. Total travel time}
 \noindent If we denote by $TR(r_i)$ travel time associated with a route $r_i$, such that:
 \begin{equation}
-    TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1}),
+	TR(r_i) =  \sum_{h = 0}^{m_i - 1} T_i(k^i_h, k^i_{h + 1}), \label{eq:traveltimenurse}
 \end{equation}
@@
 \begin{equation}
-     \bar{T} = \sum_{i \in N}  TR(r_i).
+	\bar{T} = \sum_{i \in N}  TR(r_i). \label{eq:totaltraveltime}
+\end{equation}
+
+\subsubsection*{2. Total waiting time}
+\noindent Firstly, let $W(r_i)$ be the total waiting time for each nurse's route $r_i$:
+\begin{equation}
+	W(r_i) = \displaystyle\sum_{h=0}^{m_i} w_{i k_{h}^i} - \displaystyle\sum_{f=1}^{g'_i - 1} u(\pi_{f+1}^i) - v(\pi_f^i), \label{eq:waitingtimenurse}
+\end{equation}
+\noindent where $g'_i$ is the number of shifts for nurse $i$ that start before $en_{\text{act}}(i)$. Then, the total waiting time in a solution, $\bar{W}$, is calculated as:
+\begin{equation}
+%     \bar{W} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{S}} w_{ij}.
+     \bar{W} = \sum_{i \in N} W(r_i). \label{eq:totalwaitingtime}
 \end{equation}
@@
 \begin{equation}
-    \bar{Z} = \sum_{j \in J} z_j.
+    \bar{Z} = \sum_{j \in J} z_j. \label{eq:totaltardiness}
 \end{equation}
 
-\subsubsection*{4. Maximum tardiness}
-\noindent The maximum tardiness in a solution, $Z_{max}$, is given by:
+\subsubsection*{4. Total overtime}
+\noindent The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse $i \in N$ represents the time by which their actual finish time, $en_{\text{act}}(i)$, exceeds their end of shift time, $en(i)$.
 \begin{equation}
-    Z_{max} = \max_{j \in J} z_j.
+     \bar{O} = \sum_{i \in N} \max \{0, en_{\text{act}}(i) - en(i)\}. \label{eq:totalovertime}
 \end{equation}
 
-\subsubsection*{5. Total overtime}
-\noindent The total overtime in the solution, $\bar{O}$, is calculated as the sum of the overtime of all nurses, where the overtime for each nurse represents the time by which their finish time exceeds their end of shift time, $en(i)$.
-If we denote the arrival of a nurse $i$ to its end point by $arr(i)$:
+%If we denote the arrival of a nurse $i$ to its end point by $arr(i)$:
+%\begin{equation}
+%	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k,
+%\end{equation}
+
+%\subsubsection*{\alert{5. Workload balance OLD}}
+%\noindent \alert{In order to produce practical solutions, maintaining a balance in the workload is fundamental.
+To achieve this, we quantify the minimum spare time on any nurse route in the schedule, $\bar{B}$.
+If $\bar{B}$ is minimised, we favour solutions in which the lowest spare time of any given nurse is as large as possible, hence removing imbalance when possible.}
+%\begin{equation}
+%    \alert{\bar{B} = \min_{i \in N} \left\{\max \{0, en(i) - arr(i)\} \right\}.}
+%\end{equation}
+
+\subsubsection*{5. Workload balance}
+\noindent In order to produce practical solutions, maintaining a balance in the workload is fundamental.
+There are a few papers in the literature that discuss workload balance calculations for the HHCRSP.
+\citet{bowers2015} implement a constraint limiting each nurse to attend to no more than four patients, where all service times are equal, whilst \citet{bredstrom2008} state a goal of an average of five hours workload per nurse.
+Overtimes are considered by \citet{lanzarone2014}, who aim to minimise the nurses' maximum overtimes in a lexicographical manner, and \citet{mutingi2014} calculate the percentage deviation of each nurse's workload from the average workload, which includes travel times.
+One aspect which is overlooked in all of these models is the length of the nurses's shifts.
+In practice, nurses may have varying daily shift lengths depending on their contracts, which must be taken into account when attempting to balance the workload.
+Therefore, we designed a workload balance objective, \eqref{eq:workloadbalance}, which considers the total amount of workload with respect to each nurse's scheduled shift length.
+
+To achieve this, we calculate the maximum difference of the proportion of nurse workloads to the average workload, $\bar{B}$.
+The aim is for $\bar{B}$ to equal zero; thus we favour solutions in which $\bar{B}$ is as small as possible, hence removing imbalance where possible.
+
+Let $wl_i$ be the workload proportion for each nurse $i \in N$, where $wl_i$ is calculated as follows:
 \begin{equation}
-	arr(i) = st(i) + TR(r_i) + \sum_{k \in r_i} w_k + s_k,
+	wl_i = \frac{\displaystyle\sum_{h=0}^{m_i} s_{k_h^i}}{ct(i)}, \label{eq:workloadnurse}
 \end{equation}
-then the total overtime can be calculated as:
+that is, $wl_i$ is the proportion of actual working time to the total contracted working time for each nurse. From this, we calculate the average workload, $\bar{wl}$:
 \begin{equation}
-     \bar{O} = \sum_{i \in N} \max \{0, arr(i) - en(i)\}.
+	\bar{wl} = \frac{\displaystyle\sum_{i \in N} wl_i}{n}. \label{eq:averageworkload}
+\end{equation}
+Then, $\bar{B}$ can be defined as:
+\begin{equation}
+	\bar{B} = \max_{i \in N} \{wl_i - \bar{wl}\}. \label{eq:workloadbalance}
+\end{equation}
+\noindent Note that when calculating $\bar{B}$, we only consider the service time performed by each nurse.
+To account for the travel time and waiting time of each nurse, we developed two further objective function parameters: maximum travel time and maximum waiting time.
+
+\subsubsection*{6. Maximum travel time}
+\noindent The maximum travel time of all nurses' routes, $TR_{\text{max}}$, is calculated as:
+\begin{equation}
+	TR_{\text{max}} = \max_{i \in N} \{TR(r_i)\}, \label{eq:maxtraveltime}
+\end{equation}
+where $TR(r_i)$ is the travel time associated with a route $r_i$ as described in~\eqref{eq:traveltimenurse}.
+
+\subsubsection*{7. Maximum waiting time}
+\noindent The maximum waiting time of the nurses's routes, $W_{\text{max}}$, is determined as follows: 
+\begin{equation}
+	W_{\text{max}} = \max_{i \in N} \{W(r_i)\}. \label{eq:maxwaitingtime}
 \end{equation}
 
-\subsubsection*{6. Workload balance}
-\noindent In order to produce practical solutions, maintaining a balance in the workload is fundamental.
-To achieve this, we quantify the minimum spare time on any nurse route in the schedule, $\bar{B}$.
-If $\bar{B}$ is minimised, we favour solutions in which the lowest spare time of any given nurse is as large as possible, hence removing imbalance when possible.
+\subsubsection*{8. Total preference score}
+\noindent The total preference score in the solution, $\bar{P}$, is computed as:
 \begin{equation}
-    \bar{B} = \min_{i \in N} \left\{\max \{0, en(i) - arr(i)\} \right\}.
+    \bar{P} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{R}} P(i,j). \label{eq:totalpreference}
 \end{equation}
 
-\subsubsection*{7. Total preference score}
-\noindent The total preference score in the solution, $\bar{P}$, is computed as:
+\subsubsection*{9. Maximum tardiness}
+\noindent The maximum tardiness in a solution, $Z_{\text{max}}$, is given by:
 \begin{equation}
-    \bar{P} = \sum_{i \in N} \sum_{j \in r_i \subset \mathcal{R}} P(i,j).
-\end{equation}
-Finally, the full quality of the solution, $q(\mathcal{S})$, is then obtained as:
-\begin{equation}
-    q(\mathcal{S}) = \alpha_1 \bar{T} + % Travel time 
-         \alpha_2 \bar{W} + % Waiting time 
-         \alpha_3 \bar{Z} + % Tardiness 
-         \alpha_4 \bar{O} + % Overtime 
-         \alpha_5 \bar{B} + % Workload balance 
-         \alpha_6 \bar{P} + % Preference score 
-         \alpha_7 Z_{max}. \label{eqn:solutionquality}
+	Z_{\text{max}} = \max_{j \in J} z_j. \label{eq:maxtardiness}
 \end{equation}
 
-\noindent Note that some of the terms considered in the quality function \eqref{eqn:solutionquality} might be considered infeasible in some settings (e.g.
+Finally, the full quality of the solution, $q(\mathcal{S})$, is then obtained as:
+%\begin{equation}
+%    \model{q(\mathcal{S}) = \alpha_1 \bar{T} + % Travel time 
+%         \alpha_2 \bar{W} + % Waiting time 
+%         \alpha_3 \bar{Z} + % Tardiness 
+%         \alpha_4 \bar{O} + % Overtime 
+%         \alpha_5 \bar{B} + % Workload balance 
+%         \alpha_6 \bar{P} + % Preference score 
+%         \alpha_7 Z_{max}.} \label{eqn:solutionqualityold}
+%\end{equation}
+\begin{align}
+	q(\mathcal{S}) &= \alpha_1 \bar{T} % Travel time 
+	+ \alpha_2 \bar{W} % Waiting time 
+	+ \alpha_3 \bar{Z} % Tardiness 
+	+ \alpha_4 \bar{O} % Overtime 
+	+ \alpha_5 \bar{B} \nonumber \\% Workload balance 
+	&+ \alpha_6 TR_{\text{max}} %Maximum travel time
+	+ \alpha_7 W_{\text{max}} %Maximum waiting time
+	+ \alpha_8 \bar{P} % Preference score 
+	+ \alpha_9 Z_{\text{max}}. \label{eq:solutionquality}
+\end{align}
+
+\noindent Note that some of the terms considered in the quality function \eqref{eq:solutionquality} might be considered infeasible in some settings (e.g.
 overtime and tardiness).
@@
    q(\ensuremath{\mathcal{S}}) & \text{if \ensuremath{I(\mathcal{S}) \leq 0}} \\
-   q(\ensuremath{\mathcal{S}}) + \ensuremath{M_1 I(\mathcal{S})} + \ensuremath{M_2} & \text{if \ensuremath{I(\mathcal{S}) > 0}}, \\ \label{eqn:realobjfunction}
+   q(\ensuremath{\mathcal{S}}) + \ensuremath{M_1 I(\mathcal{S})} + \ensuremath{M_2} & \text{if \ensuremath{I(\mathcal{S}) > 0}}, \label{eq:realobjfunction}
   \end{cases}
@@
 \noindent where $I(\mathcal{S})$ represents the value of the infeasibility in the solution; hence solutions where $I(\mathcal{S}) > 0$ are considered infeasible.
-The measures that are included in the calculation of $I(\mathcal{S})$ may vary depending on the \model{team (instance?)} being modelled.
-Typically, \model{they (?be specific)} include the total gap tardiness, $\tilde{Z}$:
+The measures that are included in the calculation of $I(\mathcal{S})$ may vary depending on the specific problem setting being modelled.
+Typically, they include the total gap tardiness, $\tilde{Z}$:
 \begin{equation}
-    \tilde{Z} = \sum_{j \in J} \tilde{z}_j,
+    \tilde{Z} = \sum_{j \in J} \tilde{z}_j, \label{eq:totalgaptardiness}
 \end{equation}
@@
     \centering
-    \caption{Weights proposed for the objective function of the DST}
-    \begin{tabular}{ccc}
+    \caption{Weights proposed for the objective function of the DST.}
+    \begin{tabular}{ccl}
     \toprule
@@
     \midrule
-    $\alpha_1$      &   -1   &   Travel time \\
-    $\alpha_2$      &   -1   &   Waiting time \\
-    $\alpha_3$      &   -15   &   Tardiness \\
-    $\alpha_4$      &   -15   &   Overtime \\   
-    $\alpha_5$      &   $\frac{1}{2}$   &   Workload balance \\       
-    $\alpha_6$      &   1   &   Preference score \\    
-    $\alpha_7$      &   0   &   Maximum tardiness\\    
+    $\alpha_1$      &   -1   &   Total travel time \\
+    $\alpha_2$      &   -1   &   Total waiting time \\
+    $\alpha_3$      &   -15   &   Total tardiness \\
+    $\alpha_4$      &   -15   &   Total overtime \\   
+    $\alpha_5$      &   -2000   &   Workload balance\\       
+    $\alpha_6$      &   -5   &   Maximum travel time\\       
+    $\alpha_7$      &   -5   &   Maximum waiting time\\       
+    $\alpha_8$      &   1   &   Preference score \\    
+    $\alpha_9$      &   0   &   Maximum tardiness\\    
     \bottomrule
@@
 
-\noindent Note that the weight for workload balance is smaller than one, meaning that the workload will be balanced by one minute if that costs less than one minute to the overall schedule.
+%\noindent Note that the weight for workload balance is smaller than one, meaning that the workload will be balanced by one minute if that costs less than one minute to the overall schedule.
 This value might be adjusted depending on the number of nurses on the teams, as the impact on the total schedule might be easier to absorb if more nurses are present.
@@
 
-Solutions are generated iteratively by the randomised constructive (described in Section~\ref{sub:randomisedconstructive}) and improved with local search (described in Section~\ref{sub:localsearch}).
+\noindent Solutions are generated iteratively by the randomised constructive (described in Section~\ref{sub:randomisedconstructive}) and improved with local search (described in Section~\ref{sub:localsearch}).
 Good solutions are kept in a solution pool if they are different enough from the ones currently in the pool.
@@
 After each iteration we check the stopping criterion.
-In our implementation this was a time limit.
+In our implementation this is a time limit.
 The order of the nurse list is randomly generated for each iteration, as this might have an impact on waiting times.
@@
 \noindent The Randomised Constructive Algorithm (RCA) is used to create initial solutions for each iteration of the GRASP algorithm.
-It works by ranking the quality of all job-nurse assignments by the value of the objective function \eqref{eqn:realobjfunction}, which might also be evaluated in partial solutions.
+It works by ranking the quality of all job-nurse assignments by the value of the objective function \eqref{eq:realobjfunction}, which might also be evaluated in partial solutions.
 These rankings are sorted by quality, and the highest quality ones are stored in a reduced candidate list (RCL).
@@
 There are a number of ways in which an RCL can be created, relating to which is the `cut-off' value of the objective value that decides if an assignment enters the RCL or not.
-In our implementation we have tested two strategies, both depending on the hyper-parameter $\delta$.
-Let $B$ ($W$) be the best (worst) ranking found in a given iteration of the RCA, the strategies to find the cut-off value, $C$, are as follows:
-\begin{itemize}
-     \item[1)] Strategy 1:
+In our implementation we have tested two strategies, both depending on a hyper-parameter $\delta$.
+Let $B$ ($W$) be the best (worst) ranking found in a given iteration of the RCA.
+The strategies to find the cut-off value, $C$, are as follows:
+\begin{enumerate}[label={(\arabic*)}]
+     \item Strategy 1:
      \begin{equation}
-         C = B + \delta (B - W).
+         C = B + \delta (B - W). \label{eq:rcastrategy1}
      \end{equation}
-
-     \item[2)] Strategy 2:
+     \item Strategy 2:
      \begin{equation}
-         C = B - \delta B.
+         C = B - \delta B. \label{eq:rcastrategy2}
      \end{equation}
- \end{itemize} 
+ \end{enumerate} 
 The parameter $\delta$ is randomly selected in each iteration, sampled from the interval $\delta \in [\texttt{grasp\_dl}, \texttt{grasp\_dl} + \texttt{grasp\_dr}]$.
@@
 In the first part, each nurse is assigned a single job using this procedure.
-This ensures that all nurses will have an initial route, regardless of how far their starting location is, and helps constructing more balanced solutions.
-In the second part, jobs are ranked and assigned to the best position in any of the nurse routes, until no more jobs are left to assign.
+This ensures that all nurses will have an initial route, regardless of how far their starting location is, and helps to construct more balanced solutions.
+In the second part, jobs are ranked and assigned to the best position in any of the nurse routes until no more jobs are left to assign.
 
@@
 % \correction{Review names and add descriptions}
-\begin{itemize}
-    \item NE01 Moving a job: This neighbourhood involves removing one service from the route of one nurse and inserting it in the best possible position of another (or the same) nurse.
-    \item NE02 Exchange two services: Two jobs from different nurse routes are removed and reinserted in the corresponding position of the other route.
+\begin{enumerate}[label={(\arabic*)}]
+    \item \textit{NE01: Move a service}. This neighbourhood involves removing one job from the route of one nurse and inserting it in the best possible position of another (or the same) nurse.
+    \item \textit{NE02: Exchange two services}. Two jobs from different nurse routes are removed and reinserted in the corresponding position of the other route.
     % \correction{Check code for real ordering and numbers here}
-    \item NE03 Nurse order change: This neighbourhood focuses on exploring different orderings of the nurse list, $\mathcal{O}$. It searches over exchanges between pairs of nurses in the list. 
+    \item \textit{NE03: Nurse order change}.
+This neighbourhood focuses on exploring different orderings of the nurse list, $\mathcal{O}$.
+It searches over exchanges between pairs of nurses in the list.
     This can have an effect in the way waiting times are calculated (as detailed in Section~\ref{sub:temporalconstraints}) and improve the overall quality.
-    \item NE04 Remove and reinsert linked services: Two services that are dependent on each other (synchronised or double services) are removed simultaneously from the solution and reinserted in the best possible positions.
-    \item NE05 2OPT: This is the classical vehicle routing neighbourhood and operates in a single route. Subsets of each route are removed and reinserted in a reversed order.
-\end{itemize}
+    \item \textit{NE04: Remove and reinsert linked services}.
+Two services that are dependent on each other (synchronised or double services) are removed simultaneously from the solution and reinserted in the best possible positions.
+    \item \textit{NE05: 2OPT}. This is the classical vehicle routing neighbourhood and operates in a single route. Subsets of each route are removed and reinserted in a reversed order.
+\end{enumerate}
 
@@
 The intermediate solution of highest quality is then saved, and local search is applied on it.
-Note that, this intermediate solution is often not a local optimum (unlike the starting and the guiding solutions), so applying local search may lead to high quality solutions.
+Note that, this intermediate solution is often not a local optimum (unlike the starting and the guiding solutions), so applying local search may lead to higher quality solutions.
 The pseudo-code of the routine is described in Algorithm~\ref{alg:pathrelinking}.
@@
 
-Note that double services are removed and reinserted simultaneously (see line~\ref{alg:prforinsertion}).
+\noindent Note that double services are removed and reinserted simultaneously (see line~\ref{alg:prforinsertion}).
 In the case that some jobs that are due to be inserted in a position later than the current length of the route, they are instead inserted in the last possible position.
@@
 \begin{equation}
-    d(\mathcal{S}_1, \mathcal{S}_2) = d_o(\mathcal{O}_1, \mathcal{O}_2) + d_a(\mathcal{R}_1, \mathcal{R}_2) + d_r(\mathcal{R}_1, \mathcal{R}_2).
+    d(\mathcal{S}_1, \mathcal{S}_2) = d_o(\mathcal{O}_1, \mathcal{O}_2) + d_a(\mathcal{R}_1, \mathcal{R}_2) + d_r(\mathcal{R}_1, \mathcal{R}_2). \label{eq:dissimilarityfunction}
 \end{equation}
@@
 \begin{equation}
-    d_o(\mathcal{O}_1, \mathcal{O}_2) = |\{i\in\{1,\dots,n\} : \mathcal{O}_1(i) \neq  \mathcal{O}_2(i)\}|,
+    d_o(\mathcal{O}_1, \mathcal{O}_2) = |\{i\in\{1,\dots,n\} : \mathcal{O}_1(i) \neq  \mathcal{O}_2(i)\}|, \label{eq:dissimilaritynurseorder}
 \end{equation}
@@
 \begin{equation}
-    d_a(\mathcal{R}_1, \mathcal{R}_2) = |\{j\in\{1,\dots,m\} : \nexists i \in N \text{ such that } j \in r^{(1)}_i \cap r^{(2)}_i\}|.
+    d_a(\mathcal{R}_1, \mathcal{R}_2) = |\{j\in\{1,\dots,m\} : \nexists i \in N \text{ such that } j \in r^{(1)}_i \cap r^{(2)}_i\}|. \label{eq:dissimilarityassignments}
 \end{equation}
@@
 \begin{equation}
-    d_r(\mathcal{R}_1, \mathcal{R}_2) = |A(\mathcal{R}_1)| - |A(\mathcal{R}_1) \cap A(\mathcal{R}_2)|.
+    d_r(\mathcal{R}_1, \mathcal{R}_2) = |A(\mathcal{R}_1)| - |A(\mathcal{R}_1) \cap A(\mathcal{R}_2)|. \label{eq:dissimilarityroutes}
 \end{equation}
@@
 Since this is potentially very time consuming, we have we have investigated the following three strategies for relinking intermediate solutions:
-\begin{itemize}
-    \item PR01: Perform path relinking between every generated solution and another solution selected randomly from the pool.
-    \item PR02: Perform path relinking between every new solution and all solutions in the pool.
-    \item PR03: Do not perform path relinking during the algorithm.
-\end{itemize}
+\begin{enumerate}[label={(\arabic*)}]
+    \item \textit{PR01}: Perform path relinking between every generated solution and another solution selected randomly from the pool.
+    \item \textit{PR02}: Perform path relinking between every new solution and all solutions in the pool.
+    \item \textit{PR03}: Do not perform path relinking during the algorithm.
+\end{enumerate}
 
@@
     \texttt{grasp\_dr} & Range of $\delta$ & R & [0,1] & 0.81 \\ % grasp_dr
-    \texttt{rcl\_strategy} & Strategy for generating the reduced candidate list & C & 1,2 & 2 \\ % rcl_strategy    
+    \texttt{rcl\_strategy} & Strategy for generating the reduced candidate list & C & 1, 2 & 2 \\ % rcl_strategy    
     \bottomrule
@@
 
-For the remainder of the section we use the parameters found by \texttt{irace} in all our experiments.
+\noindent For the remainder of the section we use the parameters found by \texttt{irace} in all our experiments.
 
@@
 \begin{equation}
-    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max},
+    \min \lambda_1 Dist + \lambda_2 Travel + \lambda_3 Travel^{max}, \label{eq:mankowskaobj}
 \end{equation}
@@
 \begin{equation}
-    \min Dist + Pref,
+    \min Dist + Pref, \label{eq:aithobj}
 \end{equation}
 where $Dist$ represents the cost of travelling for all nurses and $Pref$ the preference score.
-In Table~\ref{table:adaptedobjvalues} we specify the weights used in the objective function \eqref{eqn:solutionquality} in order to represent the objective functions of these algorithms.
+In Table~\ref{table:adaptedobjvalues} we specify the weights used in the objective function \eqref{eq:solutionquality} in order to represent the objective functions of these algorithms.
 
@@
     \caption{Weights used in the objective function to match those used in the methods from  \citet{aithaddadene2016} (AH) and \citet{mankowska2014} (MK).
-A weight of zero indicates that the measure is not considered, while a dash indicates that those measures that were considered infeasible, i.e.
+A weight of zero indicates that the measure is not considered, while a dash indicates those measures that were considered infeasible, i.e.
 they contribute to $I(S)$.}
@@
     $\alpha_5$      &   $0$                     &  $0$ \\       
-    $\alpha_6$      &   $-1$                    &  $0$                      \\    
-    $\alpha_7$      &   -                       &  $-\frac{1}{3}\times60$             \\    
+    $\alpha_6$      &   0                       &  0 \\    
+    $\alpha_7$      &   0                       &  0 \\    
+    $\alpha_8$      &   $-1$                    &  $0$ \\    
+    $\alpha_9$      &   -                       &  $-\frac{1}{3}\times60$ \\    
     \bottomrule
@@
 In addition, we are able to replicate the 23 optimal solutions provided in the paper.
-The average results show that \comp{GRASP} is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.
+The average results show that GRASP is also consistent across runs, finding the optimal solutions for all runs in group \emph{G1} and in most of the runs in group \emph{G2}.
 
 Similarly, in Table~\ref{table:grouptablemkavgbest} we present the comparison of our DST against the performance of the AVNS algorithm presented in \citet{mankowska2014}.
+GRASP clearly outperforms AVNS in the majority of instances, especially on those of larger sizes.
+Further, our algorithm is able to replicate most of the proven optimal solutions (9 of the 10 instances for which \cite{mankowska2014} provides an optimal solution).
+In the instance group A, where the AVNS finds all optimal solutions for the problem, GRASP fails to replicate one of these solutions, hence increasing the average.
+This is one of the cases where the assumption that a start time of a job should be as early as possible for the first nurse arriving \eqref{eq:earliestpossiblestartassumption} proves to be too restrictive and removes the optimal solution from the search space of the metaheuristic.
+Nevertheless, the loss in quality is under $5\%$ and the route itself found by the two algorithms is the same (but not the scheduled start times).
+While this effect may still occur in larger instances, it is not evident from our results, suggesting that the efficiency gains seem to outweigh the potential decrease in quality.
 
-\begin{table}
+\begin{table}[h]
 	\centering
-	\caption{Performance of the \comp{GRASP algorithm} (average of best and average results) compared to the AVNS algorithm presented in \citet{mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}.
-Figures in bold indicate the best results for each instance group.
-\comp{mk only 2dp, why?}}
-	\begin{tabular}{c|c|ccc}
+	\caption{Performance of the GRASP algorithm (average of best and average results) compared to the AVNS algorithm presented in \citet{mankowska2014} (\emph{mk\_avns}) for the instance groups \emph{A} -- \emph{G}.
+Figures in bold indicate the best results for each instance group.}
+	\begin{tabular}{crcrrr}
 		\toprule
-		        &  \emph{mk\_avns}    &   \multicolumn{3}{c}{GRASP} \\
-		Group   &   Result    &   Best        &   Avg         &   Time limit\\
+		        &      & \phantom{a} &   \multicolumn{3}{c}{GRASP} \\
+		\cmidrule{4-6}
+		Group   &   \emph{mk\_avns}    & &   \multicolumn{1}{c}{Best}        &   \multicolumn{1}{c}{Average}         &   \multicolumn{1}{c}{Time Limit}\\
 		\midrule
-		\emph{A}    &   \textbf{225.6}  &   226.41  &   226.41      &   5 s\\
-		\emph{B}    &   475.1   &   \textbf{408.25} &   408.31      &   30 s\\
-		\emph{C}    &   713.6   &   \textbf{628.15} &   637.59      &   1 min\\
-		\emph{D}    &   930.3   &   \textbf{798.80}  &   814.53      &   5 min\\
-		\emph{E}    &   1064.7  &   \textbf{880.06} &   901.00       &   15 min\\
-		\emph{F}    &   1588.0    &   \textbf{1405.68}&   1491.09     &   30 min\\
-		\emph{G}    &   2161.2  &   \textbf{2042.10} &   2341.89     &   3 h\\
+		\emph{A}    &   \textbf{225.6}  &&  226.41  &   226.41      &   5 seconds\\
+		\emph{B}    &   475.1   &&   \textbf{408.25} &   408.31      &   30 seconds\\
+		\emph{C}    &   713.6   &&   \textbf{628.15} &   637.59      &   60 seconds\\
+		\emph{D}    &   930.3   &&   \textbf{798.80}  &   814.53      &   5 minutes\\
+		\emph{E}    &   1064.7  &&   \textbf{880.06} &   901.00       &   15 minutes\\
+		\emph{F}    &   1588.0    &&   \textbf{1405.68}&   1491.09     &   30 minutes\\
+		\emph{G}    &   2161.2  &&   \textbf{2042.10} &   2341.89     &   3 hours\\
 		\bottomrule
@@
 \end{table}
-
-\comp{GRASP} clearly outperforms AVNS in the majority of instances, especially on those of larger sizes.
-Further, our algorithm is able to replicate most of the proven optimal solutions (9 of the 10 instances for which \cite{mankowska2014} provides an optimal solution).
-In the instance group A, where the AVNS finds all optimal solutions for the problem, GRASP fails to replicate one of these solutions, hence increasing the average.
-This is one of the cases where the assumption that a start time of a job should be as early as possible for the first nurse arriving \eqref{eqn:earliestpossiblestartassumption2} proves to be too restrictive and removes the optimal solution from the search space of the metaheuristic.
-Nevertheless, the loss in quality is under $5\%$ and the route itself found by the two algorithms is the same (but not the scheduled start times).
-While this effect may still occur in larger instances, it is not evident from our results, suggesting that the efficiency gains seem to outweigh the potential decrease in quality.
 
@@
 
-In order to answer these questions, we examine closely the values of a few key indicators of the quality of a solution: \emph{Total time}, \emph{Travel time}, \emph{Waiting time}, \emph{Preferences} and \emph{Longest Day}, which is the time it takes to perform the longest shift in the solution.
+In order to answer these questions, we examine closely the values of a few key indicators of the quality of a solution: \emph{Total time}, \emph{Travel time}, \emph{Waiting time}, \emph{Maximum travel time}, \emph{Maximum waiting time}, \emph{Maximum workload balance difference}, and \emph{Preferences}.
 We compare the indicators found by using the original objective function (listed in Table~\ref{table:adaptedobjvalues}) with those found by using our proposed objective function (the values listed in Table~\ref{table:adequateweights}, but with preferences being negative and time quantities scaled by $-0.3\times60$, in order to scale them to the data).
@@
 
-\input{quality_comparison_table.tex}
+\input{qualitycomparisontable.tex}
 
-The comparison clearly indicates that there is a significant amount of waiting time that can be reduced by rearranging the routes, and it is often not considered in the literature.
-We show that these waiting times can be reduced, sometimes entirely, and that these reductions have a non-negligible impact in the total time of operation, generating on average a $10.1\%$ reduction, which can be up to $19.7\%$ of the total time in some instances.
-Further, these gains in time can be spread out evenly, and it is often possible to achieve a reduction of the workload for those staff members that are already doing the most, reducing up to $11.9\%$ their shift time.
-Nonetheless, these gains come at the expense of a lower matching of preferences (on average, up by $22.9\%$) and travel time (on average, up by $12.7\%$).
-For this example, we have considered equal weights for travel and waiting times, but it would be up to individual teams to determine appropriate costs for these, considering mileage costs and environmental factors when weighing travel time.
+\noindent The comparison clearly indicates that there is a significant amount of waiting time that can be reduced by rearranging the routes, which is often not considered in the literature.
+We show that these waiting times can be reduced over all 23 instances by over 30\%.
+In 13 of the instances, waiting times are reduced by at least 90\%; the waiting time is eliminated entirely in three of these instances.
+Subsequently, the maximum waiting time has also been reduced across all instances by 82.33\% on average.
+These reductions have a non-negligible impact on the total time of operation, generating on average a $11.59\%$ reduction, which can be up to $24.46\%$ of the total time in some instances.
+
+In 10 of the 23 cases, the travel time has also been lowered by up to $16.25\%$, resulting in a decrease in maximum travel time in 16 instances by up to $29.13\%$.
+Furthermore, we have been able to achieve a better workload balance for nurses in the schedules, improving the factor in all instances by an average of $43.15\%$ \correction{explain}.
+Nonetheless, these gains come at the expense of a lower matching of preference (up $123.93\%$ on average).
+For these experiments, we have considered equal weights for total travel and total waiting time, as well as equal weights for maximum travel time and maximum waiting time, however it would be up to individual teams to determine appropriate costs for these, taking into account mileage costs and environmental factor when weighing travel time, and considering nurse satisfaction when weighing waiting time.
+
+\begin{figure}[h!]
+	\centering	
+	\begin{subfigure}[h]{\textwidth}
+		\centering
+		\includestandalone[width=\textwidth]{figures/aithgantt18-4-m-2c}
+		\caption{}
+		\label{fig:aith}
+		\vspace{2mm}
+	\end{subfigure}
+	\begin{subfigure}[h]{\textwidth}
+		\centering
+		\includestandalone[width=\textwidth]{figures/papergantt18-4-m-2c}
+		\caption{}
+		\label{fig:dst}
+	\end{subfigure}
+	\caption{A comparison of schedules for instance 18-4-m-2c using (a) the \textit{ah\_gi} algorithm of \cite{aithaddadene2016} and (b) our DST using GRASP and VNS, where in (b) all objectives have been reduced.
+The dotted segments indicate travel times, whilst the red lined segments indicate waiting periods.}
+	\label{fig:18-4-l-2agantt}
+\end{figure}
+
+%The total travel time has also been reduced in 12 of the 23 cases by up to 36.41\%, whilst the maximum travel time has been reduced in 14 cases by up to 32.35\%.
+Nonetheless, these gains come at the expense of a lower matching of preferences (on average, up by $132.6\%$) and maximum workload balance difference (up by 14.56\% on average).
+For this example, we have considered equal weights for travel time, waiting time, and workload balance difference, but it would be up to individual teams to determine appropriate costs for these, considering mileage costs and environmental factors when weighing travel time.
+
+%Further, these gains in time can be spread out evenly, \comp{and it is often possible to achieve a reduction of the workload for those staff members that are already doing the most, reducing up to $11.9\%$ their shift time - longest day no longer exists, need to discuss reduction in travel and waiting times as well as overall time}.
 
@@
 \noindent In this article we propose a Decision Support Tool for the Home Health Care Routing and Scheduling Problem that is faced daily by district nursing teams.
-We propose a model that includes the most common constraints observed in practice (skills, time windows, synchronisation, double services) and objectives (travel time, waiting time, assignment preferences, workload balance, tardiness, overtime).
+We propose a model that includes the most common constraints observed in practice (skills, time windows, synchronisation, double services, shifts) and objectives (travel time, waiting time, assignment preferences, workload balance, tardiness, overtime).
 This comprehensive approach provides flexibility in the modelling of the working environment in order to be applicable in general real settings.
@@
 Furthermore, we show that current literature ignores route aspects that are important to decision makers, such as waiting time.
-By including waiting times, we can reduce total operation times by up to 19\% when compared to optimal solutions in the literature that minimise travel times only.
+By including waiting times, we can reduce total operation times by up to 24\% when compared to optimal solutions in the literature that minimise travel times and preferences only.
 

### 2021-12-18 tex/Paper/figures/aithgantt18-4-l-2a.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (13.9, 7);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,6) rectangle (0.22,7);
\draw [thick] (0.22,6) -- (0.22,7); % start of job 10
\draw [thick] (1.98,6) -- (1.98,7); % end of job 10
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.98,6) rectangle (2.14,7);
\draw [thick] (2.14,6) -- (2.14,7); % start of job 14
\draw [thick] (3.4,6) -- (3.4,7); % end of job 14
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.4,6) rectangle (3.68,7);
\draw [thick] (3.68,6) -- (3.68,7); % start of job 3
\draw [thick] (5.72,6) -- (5.72,7); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.72,6) rectangle (5.86,7);
\draw [thick] (5.86,6) -- (5.86,7); % start of job 11
\draw [thick] (6.8,6) -- (6.8,7); % end of job 11
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.8,6) rectangle (7.32,7);
\draw [thick] (7.32,6) -- (7.32,7); % start of job 17
\draw [thick] (9.28,6) -- (9.28,7); % end of job 17
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.28,6) rectangle (9.52,7);
\draw [thick] (9.52,6) -- (9.52,7); % start of job 2
\draw [thick] (10.16,6) -- (10.16,7); % end of job 2
\draw [thick, pattern = north west lines, pattern color=black!30!white] (10.16,6) rectangle (10.22,7);
\draw [thick] (10.22,6) -- (10.22,7); % start of job 0 (DS)
\draw [thick] (13,6) -- (13,7); % end of job 0 (DS)
\draw [thick, pattern = north west lines, pattern color=black!30!white] (13,6) rectangle (13.42,7);
\draw [thick] (13.42,6) -- (13.42,7); % end of shift
\draw [thick, fill=black!20!white] (13.42,6) rectangle (13.9,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
\node at (1.1,6.6) {\scriptsize{job}}; 
\node at (1.1,6.35) {\scriptsize{10}};
\node at (2.77,6.6) {\scriptsize{job}};
\node at (2.77,6.35) {\scriptsize{14}};
\node at (4.7,6.6) {\scriptsize{job}};
\node at (4.7,6.35) {\scriptsize{3}};
\node at (6.33,6.6) {\scriptsize{job}};
\node at (6.33,6.35) {\scriptsize{11}};
\node at (8.3,6.6) {\scriptsize{job}};
\node at (8.3,6.35) {\scriptsize{17}};
\node at (9.84,6.6) {\scriptsize{job}};
\node at (9.84,6.35) {\scriptsize{2}};
\node at (11.61,6.6) {\scriptsize{job}};
\node at (11.61,6.35) {\scriptsize{0 (DS)}};

% Nurse 1
\draw [thick] (0,4) rectangle (13.9, 5);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,4) rectangle (0.26,5);
\draw [thick] (0.26,4) -- (0.26,5); % start of job 16
\draw [thick] (1.64,4) -- (1.64,5); % end of job 16
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.64,4) rectangle (1.96,5);
\draw [thick] (1.96,4) -- (1.96,5); % start of job 15
\draw [thick] (3,4) -- (3,5); % end of job 15
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3,4) rectangle (3.38,5);
\draw [thick] (3.38,4) -- (3.38,5); % start of job 1
\draw [thick] (4.12,4) -- (4.12,5); % end of job 1
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.12,4) rectangle (4.52,5);
\draw [thick] (4.52,4) -- (4.52,5); % start of job 7
\draw [thick] (6.66,4) -- (6.66,5); % end of job 7
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.66,4) rectangle (6.72,5);
\draw [thick] (6.72,4) -- (6.72,5); % start of job 12
\draw [thick] (9.08,4) -- (9.08,5); % end of job 12
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.08,4) rectangle (9.12,5);
\draw [thick] (9.12,4) -- (9.12,5); % ARRIVAL at job 18 (waiting time!)
\draw [thick, pattern = north west lines, pattern color=tRed] (9.12,4) rectangle (9.32,5);
\draw [thick] (9.32,4) -- (9.32,5); % start of job 18
\draw [thick] (9.96,4) -- (9.96,5); % end of job 18
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.96,4) rectangle (10.02,5);
\draw [thick] (10.02,4) -- (10.02,5); % ARRIVAL at job 0 (DS) (waiting time!)
\draw [thick, pattern = north west lines, pattern color=tRed] (10.02,4) rectangle (10.22,5);
\draw [thick] (10.22,4) -- (10.22,5); % start of job 0 (DS)
\draw [thick] (13,4) -- (13,5); % end of job 0 (DS)
\draw [thick, pattern = north west lines, pattern color=black!30!white] (13,4) rectangle (13.42,5);
\draw [thick] (13.42,4) -- (13.42,5); % end of shift
\draw [thick, fill=black!20!white] (13.42,4) rectangle (13.9,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
\node at (0.95,4.6) {\scriptsize{job}}; 
\node at (0.95,4.35) {\scriptsize{16}};
\node at (2.48,4.6) {\scriptsize{job}}; 
\node at (2.48,4.35) {\scriptsize{15}};
\node at (3.75,4.6) {\scriptsize{job}}; 
\node at (3.75,4.35) {\scriptsize{1}};
\node at (5.59,4.6) {\scriptsize{job}}; 
\node at (5.59,4.35) {\scriptsize{7}};
\node at (7.9,4.6) {\scriptsize{job}}; 
\node at (7.9,4.35) {\scriptsize{12}};
\node at (9.64,4.6) {\scriptsize{job}}; 
\node at (9.64,4.35) {\scriptsize{18}};
\node at (11.61,4.6) {\scriptsize{job}}; 
\node at (11.61,4.35) {\scriptsize{0 (DS)}};

% Nurse 2
\draw [thick] (0,2) rectangle (13.9, 3);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,2) rectangle (0.26,3);
\draw [thick] (0.26,2) -- (0.26,3); % start of job 8
\draw [thick] (1,2) -- (1,3); % end of job 8
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1,2) rectangle (1.18,3);
\draw [thick] (1.18,2) -- (1.18,3); % start of job 13
\draw [thick] (3.8,2) -- (3.8,3); % end of job 13
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.8,2) rectangle (3.92,3);
\draw [thick] (3.92,2) -- (3.92,3); % start of job 4
\draw [thick] (5.36,2) -- (5.36,3); % end of job 4
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.36,2) rectangle (5.66,3);
\draw [thick] (5.66,2) -- (5.66,3); % start of job 9
\draw [thick] (6.16,2) -- (6.16,3); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.16,2) rectangle (6.34,3);
\draw [thick] (6.34,2) -- (6.34,3); % end of shift
\draw [thick,fill=black!20!white] (6.34,2) rectangle (13.9,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
\node at (0.63,2.6) {\scriptsize{job}}; 
\node at (0.63,2.35) {\scriptsize{8}};
\node at (2.49,2.6) {\scriptsize{job}}; 
\node at (2.49,2.35) {\scriptsize{13}};
\node at (4.64,2.6) {\scriptsize{job}}; 
\node at (4.64,2.35) {\scriptsize{4}};
\node at (5.91,2.6) {\scriptsize{job}}; 
\node at (5.91,2.35) {\scriptsize{9}};

% Nurse 3
\draw [thick] (0,0) rectangle (13.9, 1);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,0) rectangle (0.36,1);
\draw [thick] (0.36,0) -- (0.36,1); % start of job 6
\draw [thick] (1.34,0) -- (1.34,1); % end of job 6
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.34,0) rectangle (1.58,1);
\draw [thick] (1.58,0) -- (1.58,1); % start of job 5
\draw [thick] (3.08,0) -- (3.08,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.08,0) rectangle (3.22,1);
\draw [thick] (3.22,0) -- (3.22,1); % end of shift
\draw [thick, fill=black!20!white] (3.22,0) rectangle (13.9,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
\node at (0.85,0.6) {\scriptsize{job}}; 
\node at (0.85,0.35) {\scriptsize{6}};
\node at (2.33,0.6) {\scriptsize{job}}; 
\node at (2.33,0.35) {\scriptsize{5}};

\end{tikzpicture}

\end{document}

%\draw [thick] (0,6) rectangle (6.95, 7);
%\draw [thick] (0.11,6) -- (0.11,7); % start of job 10
%\draw [thick] (0.99,6) -- (0.99,7); % end of job 10
%\draw [thick] (1.07,6) -- (1.07,7); % start of job 14
%\draw [thick] (1.7,6) -- (1.7,7); % end of job 14
%\draw [thick] (1.84,6) -- (1.84,7); % start of job 3
%\draw [thick] (2.86,6) -- (2.86,7); % end of job 3
%\draw [thick] (2.93,6) -- (2.93,7); % start of job 11
%\draw [thick] (3.4,6) -- (3.4,7); % end of job 11
%\draw [thick] (3.66,6) -- (3.66,7); % start of job 17
%\draw [thick] (4.64,6) -- (4.64,7); % end of job 17
%\draw [thick] (4.76,6) -- (4.76,7); % start of job 2
%\draw [thick] (5.08,6) -- (5.08,7); % end of job 2
%\draw [thick] (5.11,6) -- (5.11,7); % start of job 0 (DS)
%\draw [thick] (6.5,6) -- (6.5,7); % end of job 0 (DS)
%\draw [thick] (6.71,6) -- (6.71,7); % end of shift

### 2021-12-18 tex/Paper/figures/aithgantt18-4-l-2c.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (14.04, 7);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,6) rectangle (0.24,7);
\draw [thick] (0.24,6) -- (0.24,7); % start of job 12
\draw [thick] (2.10,6) -- (2.10,7); % end of job 12
\draw [thick, pattern = north west lines, pattern color=black!30!white] (2.10,6) rectangle (2.28,7);
\draw [thick] (2.28,6) -- (2.28,7); % start of job 16
\draw [thick] (3.76,6) -- (3.76,7); % end of job 16
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.76,6) rectangle (3.94,7);
\draw [thick] (3.94,6) -- (3.94,7); % start of job 14
\draw [thick] (5.50,6) -- (5.50,7); % end of job 14
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.50,6) rectangle (6.06,7);
\draw [thick] (6.06,6) -- (6.06,7); % ARRIVE at job 3
\draw [thick, pattern = north east lines, pattern color=tRed] (6.06,6) rectangle (6.96,7);
\draw [thick] (6.96,6) -- (6.96,7); % start of job 3
\draw [thick] (8.12,6) -- (8.12,7); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.12,6) rectangle (8.48,7);
\draw [thick] (8.48,6) -- (8.48,7); % start of job 17
\draw [thick] (10.72,6) -- (10.72,7); % end of job 17
\draw [thick, pattern = north west lines, pattern color=black!30!white] (10.72,6) rectangle (10.98,7);
\draw [thick] (10.98,6) -- (10.98,7); % start of job 1
\draw [thick] (12.94,6) -- (12.94,7); % end of job 1
\draw [thick, pattern = north west lines, pattern color=black!30!white] (12.94,6) rectangle (13.18,7);
\draw [thick] (13.18,6) -- (13.18,7); % end of shift
\draw [thick, fill=black!20!white] (13.18,6) rectangle (14.04,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
\node at (1.17,6.6) {\scriptsize{job}}; 
\node at (1.17,6.35) {\scriptsize{12}};
\node at (3.02,6.6) {\scriptsize{job}};
\node at (3.02,6.35) {\scriptsize{16}};
\node at (4.72,6.6) {\scriptsize{job}};
\node at (4.72,6.35) {\scriptsize{14}};
\node at (7.54,6.6) {\scriptsize{job}};
\node at (7.54,6.35) {\scriptsize{3(DS)}};
\node at (9.60,6.6) {\scriptsize{job}};
\node at (9.60,6.35) {\scriptsize{17}};
\node at (11.96,6.6) {\scriptsize{job}};
\node at (11.96,6.35) {\scriptsize{1}};

% Nurse 1
\draw [thick] (0,4) rectangle (14.04, 5);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,4) rectangle (0.34,5);
\draw [thick] (0.34,4) -- (0.34,5); % ARRIVE at job 9
\draw [thick, pattern = north east lines, pattern color=tRed] (0.34,4) rectangle (4.18,5);
\draw [thick] (4.18,4) -- (4.18,5); % start of job 9
\draw [thick] (5.48,4) -- (5.48,5); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.48,4) rectangle (6.10,5);
\draw [thick] (6.10,4) -- (6.10,5); % start of job 8
\draw [thick] (7.86,4) -- (7.86,5); % end of job 8
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.86,4) rectangle (8.27,5);
\draw [thick] (8.27,4) -- (8.27,5); % end of shift
\draw [thick, fill=black!20!white] (8.27,4) rectangle (14.04,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
\node at (4.83,4.6) {\scriptsize{job}}; 
\node at (4.83,4.35) {\scriptsize{9(DS)}};
\node at (6.98,4.6) {\scriptsize{job}}; 
\node at (6.98,4.35) {\scriptsize{8}};

% Nurse 2
\draw [thick] (0,2) rectangle (14.04, 3);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,2) rectangle (0.58,3);
\draw [thick] (0.58,2) -- (0.58,3); % start of job 4
\draw [thick] (2.44,2) -- (2.44,3); % end of job 4
\draw [thick, pattern = north west lines, pattern color=black!30!white] (2.44,2) rectangle (2.88,3);
\draw [thick] (2.88,2) -- (2.88,3); % start of job 15
\draw [thick] (4.10,2) -- (4.10,3); % end of job 15
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.10,2) rectangle (4.18,3);
\draw [thick] (4.18,2) -- (4.18,3); % start of job 9
\draw [thick] (5.48,2) -- (5.48,3); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.48,2) rectangle (5.86,3);
\draw [thick] (5.86,2) -- (5.86,3); % start of job 13
\draw [thick] (6.84,2) -- (6.84,3); % end of job 13
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.84,2) rectangle (6.96,3);
\draw [thick] (6.96,2) -- (6.96,3); % start of job 3
\draw [thick] (8.12,2) -- (8.12,3); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.12,2) rectangle (8.52,3);
\draw [thick] (8.52,2) -- (8.52,3); % end of shift
\draw [thick,fill=black!20!white] (8.52,2) rectangle (14.04,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
\node at (1.51,2.6) {\scriptsize{job}}; 
\node at (1.51,2.35) {\scriptsize{4}};
\node at (3.49,2.6) {\scriptsize{job}}; 
\node at (3.49,2.35) {\scriptsize{15}};
\node at (4.83,2.6) {\scriptsize{job}}; 
\node at (4.83,2.35) {\scriptsize{9(DS)}};
\node at (6.35,2.6) {\scriptsize{job}}; 
\node at (6.35,2.35) {\scriptsize{13}};
\node at (7.54,2.6) {\scriptsize{job}}; 
\node at (7.54,2.35) {\scriptsize{3(DS)}};

% Nurse 3
\draw [thick] (0,0) rectangle (14.04, 1);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,0) rectangle (0.18,1);
\draw [thick] (0.18,0) -- (0.18,1); % start of job 6
\draw [thick] (1.34,0) -- (1.34,1); % end of job 6
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.34,0) rectangle (1.50,1);
\draw [thick] (1.50,0) -- (1.50,1); % start of job 5
\draw [thick] (3.52,0) -- (3.52,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.52,0) rectangle (4.04,1);
\draw [thick] (4.04,0) -- (4.04,1); % start of job 5
\draw [thick] (5.20,0) -- (5.20,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.20,0) rectangle (5.26,1);
\draw [thick] (5.26,0) -- (5.26,1); % start of job 5
\draw [thick] (7.68,0) -- (7.68,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.68,0) rectangle (7.88,1);
\draw [thick] (7.88,0) -- (7.88,1); % start of job 5
\draw [thick] (9.26,0) -- (9.26,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.26,0) rectangle (9.60,1);
\draw [thick] (9.60,0) -- (9.60,1); % start of job 5
\draw [thick] (11.58,0) -- (11.58,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (11.58,0) rectangle (11.60,1);
\draw [thick] (11.60,0) -- (11.60,1); % start of job 5
\draw [thick] (13.48,0) -- (13.48,1); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (13.48,0) rectangle (13.78,1);
\draw [thick] (13.78,0) -- (13.78,1); % end of shift
\draw [thick, fill=black!20!white] (13.78,0) rectangle (14.04,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
\node at (0.76,0.6) {\scriptsize{job}}; 
\node at (0.76,0.35) {\scriptsize{6}};
\node at (2.51,0.6) {\scriptsize{job}}; 
\node at (2.51,0.35) {\scriptsize{5}};
\node at (4.62,0.6) {\scriptsize{job}}; 
\node at (4.62,0.35) {\scriptsize{0}};
\node at (6.47,0.6) {\scriptsize{job}}; 
\node at (6.47,0.35) {\scriptsize{7}};
\node at (8.57,0.6) {\scriptsize{job}}; 
\node at (8.57,0.35) {\scriptsize{11}};
\node at (10.59,0.6) {\scriptsize{job}}; 
\node at (10.59,0.35) {\scriptsize{2}};
\node at (12.54,0.6) {\scriptsize{job}}; 
\node at (12.54,0.35) {\scriptsize{10}};

\end{tikzpicture}

\end{document}

%\draw [thick] (0,6) rectangle (6.95, 7);
%\draw [thick] (0.11,6) -- (0.11,7); % start of job 10
%\draw [thick] (0.99,6) -- (0.99,7); % end of job 10
%\draw [thick] (1.07,6) -- (1.07,7); % start of job 14
%\draw [thick] (1.7,6) -- (1.7,7); % end of job 14
%\draw [thick] (1.84,6) -- (1.84,7); % start of job 3
%\draw [thick] (2.86,6) -- (2.86,7); % end of job 3
%\draw [thick] (2.93,6) -- (2.93,7); % start of job 11
%\draw [thick] (3.4,6) -- (3.4,7); % end of job 11
%\draw [thick] (3.66,6) -- (3.66,7); % start of job 17
%\draw [thick] (4.64,6) -- (4.64,7); % end of job 17
%\draw [thick] (4.76,6) -- (4.76,7); % start of job 2
%\draw [thick] (5.08,6) -- (5.08,7); % end of job 2
%\draw [thick] (5.11,6) -- (5.11,7); % start of job 0 (DS)
%\draw [thick] (6.5,6) -- (6.5,7); % end of job 0 (DS)
%\draw [thick] (6.71,6) -- (6.71,7); % end of shift

### 2021-12-18 tex/Paper/figures/aithgantt18-4-m-2c.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (14.6, 7);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,6) rectangle (0.30,7);
\draw [thick] (0.30,6) -- (0.30,7); % start of job 0
\draw [thick] (1.46,6) -- (1.46,7); % end of job 0
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (1.46,6) rectangle (1.92,7);
\draw [thick] (1.92,6) -- (1.92,7); % start of job 1
\draw [thick] (3.76,6) -- (3.76,7); % end of job 1
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (3.76,6) rectangle (4.20,7);
\draw [thick] (4.20,6) -- (4.20,7); % start of job 7
\draw [thick] (5.62,6) -- (5.62,7); % end of job 7
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (5.62,6) rectangle (6.10,7);
\draw [thick] (6.10,6) -- (6.10,7); % start of job 16
\draw [thick] (8.20,6) -- (8.20,7); % end of job 16
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (8.20,6) rectangle (8.34,7);
\draw [thick] (8.34,6) -- (8.34,7); % ARRIVE at job 13
\draw [thick, pattern = north west lines, pattern color=tRed] (8.34,6) rectangle (10.80,7);
\draw [thick] (10.80,6) -- (10.80,7); % start of job 13
\draw [thick] (11.72,6) -- (11.72,7); % end of job 13
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (11.72,6) rectangle (11.90,7);
\draw [thick] (11.90,6) -- (11.90,7); % end of shift
\draw [thick, fill=black!20!white] (11.90,6) rectangle (14.6,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
%\node at (0.88,6.6) {\scriptsize{job}}; 
%\node at (0.88,6.35) {\scriptsize{0}};
\node at (0.88,6.5) {\scriptsize{0}};
%\node at (2.84,6.6) {\scriptsize{job}};
%\node at (2.84,6.35) {\scriptsize{1}};
\node at (2.84,6.5) {\scriptsize{1}};
%\node at (4.91,6.6) {\scriptsize{job}};
%\node at (4.91,6.35) {\scriptsize{7}};
\node at (4.91,6.5) {\scriptsize{7}};
%\node at (7.15,6.6) {\scriptsize{job}};
%\node at (7.15,6.35) {\scriptsize{16}};
\node at (7.15,6.5) {\scriptsize{16}};
%\node at (11.26,6.6) {\scriptsize{job}};
%\node at (11.26,6.35) {\scriptsize{13}};
\node at (11.26,6.5) {\scriptsize{13}};

% Nurse 1
\draw [thick] (0,4) rectangle (14.6, 5);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,4) rectangle (0.30,5);
\draw [thick] (0.30,4) -- (0.30,5); % start of job 15
\draw [thick] (1.98,4) -- (1.98,5); % end of job 15
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (1.98,4) rectangle (2.42,5);
\draw [thick] (2.42,4) -- (2.42,5); % ARRIVE at job 3
\draw [thick, pattern = north west lines, pattern color=tRed] (2.42,4) rectangle (6.22,5);
\draw [thick] (6.22,4) -- (6.22,5); % start of job 3
\draw [thick] (8.26,4) -- (8.26,5); % end of job 3
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (8.26,4) rectangle (8.92,5);
\draw [thick] (8.92,4) -- (8.92,5); % ARRIVE at job 2
\draw [thick, pattern = north west lines, pattern color=tRed] (8.92,4) rectangle (10.18,5);
\draw [thick] (10.18,4) -- (10.18,5); % start of job 2
\draw [thick] (11.22,4) -- (11.22,5); % end of job 2
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (11.22,4) rectangle (11.58,5);
\draw [thick] (11.58,4) -- (11.58,5); % end of shift
\draw [thick, fill=black!20!white] (11.58,4) rectangle (14.6,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
%\node at (1.14,4.6) {\scriptsize{job}}; 
%\node at (1.14,4.35) {\scriptsize{15}};
\node at (1.14,4.5) {\scriptsize{15}};
%\node at (7.24,4.6) {\scriptsize{job}}; 
%\node at (7.24,4.35) {\scriptsize{3}};
\node at (7.24,4.5) {\scriptsize{3}};
%\node at (10.70,4.6) {\scriptsize{job}}; 
%\node at (10.70,4.35) {\scriptsize{2}};
\node at (10.70,4.5) {\scriptsize{2}};

% Nurse 2
\draw [thick] (0,2) rectangle (14.6, 3);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,2) rectangle (0.42,3);
\draw [thick] (0.42,2) -- (0.42,3); % start of job 14
\draw [thick] (1.44,2) -- (1.44,3); % end of job 14
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (1.44,2) rectangle (1.68,3);
\draw [thick] (1.68,2) -- (1.68,3); % start of job 6
\draw [thick] (2.32,2) -- (2.32,3); % end of job 6
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (2.32,2) rectangle (2.42,3);
\draw [thick] (2.42,2) -- (2.42,3); % ARRIVE at job 4
\draw [thick, pattern = north west lines, pattern color=tRed] (2.42,2) rectangle (7.12,3);
\draw [thick] (7.12,2) -- (7.12,3); % start of job 4
\draw [thick] (8.34,2) -- (8.34,3); % end of job 4
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (8.34,2) rectangle (8.44,3);
\draw [thick] (8.44,2) -- (8.44,3); % start of job 9
\draw [thick] (9.98,2) -- (9.98,3); % end of job 9
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (9.98,2) rectangle (10.38,3);
\draw [thick] (10.38,2) -- (10.38,3); % start of job 18
\draw [thick] (11.42,2) -- (11.42,3); % end of job 18
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (11.42,2) rectangle (11.78,3);
\draw [thick] (11.78,2) -- (11.78,3); % end of shift
\draw [thick,fill=black!20!white] (11.78,2) rectangle (14.6,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
%\node at (0.93,2.6) {\scriptsize{job}}; 
%\node at (0.93,2.35) {\scriptsize{14}};
\node at (0.93,2.5) {\scriptsize{14}};
%\node at (2.00,2.6) {\scriptsize{job}}; 
%\node at (2.00,2.35) {\scriptsize{6}};
\node at (2.00,2.5) {\scriptsize{6}};
%\node at (7.73,2.6) {\scriptsize{job}}; 
%\node at (7.73,2.35) {\scriptsize{4}};
\node at (7.73,2.5) {\scriptsize{4}};
%\node at (9.21,2.6) {\scriptsize{job}}; 
%\node at (9.21,2.35) {\scriptsize{9}};
\node at (9.21,2.5) {\scriptsize{9}};
%\node at (10.90,2.6) {\scriptsize{job}}; 
%\node at (10.90,2.35) {\scriptsize{18}};
\node at (10.90,2.5) {\scriptsize{18}};

% Nurse 3
\draw [thick] (0,0) rectangle (14.6, 1);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,0) rectangle (0.34,1);
\draw [thick] (0.34,0) -- (0.34,1); % start of job 11
\draw [thick] (0.64,0) -- (0.64,1); % end of job 11
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0.64,0) rectangle (1.22,1);
\draw [thick] (1.22,0) -- (1.22,1); % start of job 10
\draw [thick] (1.70,0) -- (1.70,1); % end of job 10
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (1.70,0) rectangle (2.20,1);
\draw [thick] (2.20,0) -- (2.20,1); % start of job 5
\draw [thick] (4.42,0) -- (4.42,1); % end of job 5
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (4.42,0) rectangle (4.56,1);
\draw [thick] (4.56,0) -- (4.56,1); % ARRIVE at job 8
\draw [thick, pattern = north west lines, pattern color=tRed] (4.56,0) rectangle (4.94,1);
\draw [thick] (4.94,0) -- (4.94,1); % start of job 8
\draw [thick] (6.20,0) -- (6.20,1); % end of job 8
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (6.20,0) rectangle (6.42,1);
\draw [thick] (6.42,0) -- (6.42,1); % start of job 19
\draw [thick] (8.46,0) -- (8.46,1); % end of job 19
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (8.46,0) rectangle (8.72,1);
\draw [thick] (8.72,0) -- (8.72,1); % start of job 17
\draw [thick] (10.30,0) -- (10.30,1); % end of job 17
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (10.30,0) rectangle (10.96,1);
\draw [thick] (10.96,0) -- (10.96,1); % start of job 12
\draw [thick] (12.48,0) -- (12.48,1); % end of job 12
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (12.48,0) rectangle (12.96,1);
\draw [thick] (12.96,0) -- (12.96,1); % end of shift
\draw [thick, fill=black!20!white] (12.96,0) rectangle (14.6,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
%\node at (0.49,0.6) {\scriptsize{job}}; 
%\node at (0.49,0.35) {\scriptsize{11}};
\node at (0.49,0.5) {\scriptsize{11}};
%\node at (1.46,0.6) {\scriptsize{job}}; 
%\node at (1.46,0.35) {\scriptsize{10}};
\node at (1.46,0.5) {\scriptsize{10}};
%\node at (3.31,0.6) {\scriptsize{job}}; 
%\node at (3.31,0.35) {\scriptsize{5}};
\node at (3.31,0.5) {\scriptsize{5}};
%\node at (5.57,0.6) {\scriptsize{job}}; 
%\node at (5.57,0.35) {\scriptsize{8}};
\node at (5.57,0.5) {\scriptsize{8}};
%\node at (7.44,0.6) {\scriptsize{job}}; 
%\node at (7.44,0.35) {\scriptsize{19}};
\node at (7.44,0.5) {\scriptsize{19}};
%\node at (9.51,0.6) {\scriptsize{job}}; 
%\node at (9.51,0.35) {\scriptsize{17}};
\node at (9.51,0.5) {\scriptsize{17}};
%\node at (11.72,0.6) {\scriptsize{job}}; 
%\node at (11.72,0.35) {\scriptsize{12}};
\node at (11.72,0.5) {\scriptsize{12}};

\end{tikzpicture}

\end{document}

%\draw [thick] (0,6) rectangle (6.95, 7);
%\draw [thick] (0.11,6) -- (0.11,7); % start of job 10
%\draw [thick] (0.99,6) -- (0.99,7); % end of job 10
%\draw [thick] (1.07,6) -- (1.07,7); % start of job 14
%\draw [thick] (1.7,6) -- (1.7,7); % end of job 14
%\draw [thick] (1.84,6) -- (1.84,7); % start of job 3
%\draw [thick] (2.86,6) -- (2.86,7); % end of job 3
%\draw [thick] (2.93,6) -- (2.93,7); % start of job 11
%\draw [thick] (3.4,6) -- (3.4,7); % end of job 11
%\draw [thick] (3.66,6) -- (3.66,7); % start of job 17
%\draw [thick] (4.64,6) -- (4.64,7); % end of job 17
%\draw [thick] (4.76,6) -- (4.76,7); % start of job 2
%\draw [thick] (5.08,6) -- (5.08,7); % end of job 2
%\draw [thick] (5.11,6) -- (5.11,7); % start of job 0 (DS)
%\draw [thick] (6.5,6) -- (6.5,7); % end of job 0 (DS)
%\draw [thick] (6.71,6) -- (6.71,7); % end of shift

### 2021-12-18 tex/Paper/figures/papergantt18-4-l-2a.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (13.9, 7);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,6) rectangle (0.18,7);
\draw [thick] (0.18,6) -- (0.18,7); % start of job 9
\draw [thick] (0.68,6) -- (0.68,7); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0.68,6) rectangle (0.78,7);
\draw [thick] (0.78,6) -- (0.78,7); % start of job 8
\draw [thick] (1.52,6) -- (1.52,7); % end of job 8
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.52,6) rectangle (1.7,7);
\draw [thick] (1.7,6) -- (1.7,7); % start of job 13
\draw [thick] (4.32,6) -- (4.32,7); % end of job 13
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.32,6) rectangle (4.4,7);
\draw [thick] (4.4,6) -- (4.4,7); % start of job 3
\draw [thick] (6.44,6) -- (6.44,7); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.44,6) rectangle (6.48,7);
\draw [thick] (6.48,6) -- (6.48,7); % start of job 4
\draw [thick] (7.92,6) -- (7.92,7); % end of job 4
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.92,6) rectangle (8.06,7);
\draw [thick] (8.06,6) -- (8.06,7); % start of job 11
\draw [thick] (9,6) -- (9,7); % end of job 11
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9,6) rectangle (9.38,7);
\draw [thick] (9.38,6) -- (9.38,7); % end of shift
\draw [thick, fill=black!20!white] (9.38,6) rectangle (13.9,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
\node at (0.43,6.6) {\scriptsize{job}}; 
\node at (0.43,6.35) {\scriptsize{9}};
\node at (1.15,6.6) {\scriptsize{job}};
\node at (1.15,6.35) {\scriptsize{8}};
\node at (3.01,6.6) {\scriptsize{job}};
\node at (3.01,6.35) {\scriptsize{13}};
\node at (5.42,6.6) {\scriptsize{job}};
\node at (5.42,6.35) {\scriptsize{3}};
\node at (7.2,6.6) {\scriptsize{job}};
\node at (7.2,6.35) {\scriptsize{4}};
\node at (8.53,6.6) {\scriptsize{job}};
\node at (8.53,6.35) {\scriptsize{11}};

% Nurse 1
\draw [thick] (0,4) rectangle (13.9, 5);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,4) rectangle (0.26,5);
\draw [thick] (0.26,4) -- (0.26,5); % start of job 16
\draw [thick] (1.64,4) -- (1.64,5); % end of job 16
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.64,4) rectangle (1.78,5);
\draw [thick] (1.78,4) -- (1.78,5); % start of job 1
\draw [thick] (2.52,4) -- (2.52,5); % end of job 1
\draw [thick, pattern = north west lines, pattern color=black!30!white] (2.52,4) rectangle (2.98,5);
\draw [thick] (2.98,4) -- (2.98,5); % ARRIVAL at job 18
\draw [thick, pattern = north west lines, pattern color=tRed] (2.98,4) rectangle (3.06,5);
\draw [thick] (3.06,4) -- (3.06,5); % start of job 18
\draw [thick] (3.7,4) -- (3.7,5); % end of job 18
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.7,4) rectangle (3.74,5);
\draw [thick] (3.74,4) -- (3.74,5); % start of job 12
\draw [thick] (6.10,4) -- (6.10,5); % end of job 12
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.10,4) rectangle (6.56,5);
\draw [thick] (6.56,4) -- (6.56,5); % start of job 6
\draw [thick] (7.54,4) -- (7.54,5); % end of job 6
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.54,4) rectangle (7.78,5);
\draw [thick] (7.78,4) -- (7.78,5); % start of job 5
\draw [thick] (9.28,4) -- (9.28,5); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.28,4) rectangle (9.42,5);
\draw [thick] (9.42,4) -- (9.42,5); % end of shift
\draw [thick, fill=black!20!white] (9.42,4) rectangle (13.9,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
\node at (0.95,4.6) {\scriptsize{job}}; 
\node at (0.95,4.35) {\scriptsize{16}};
\node at (2.15,4.6) {\scriptsize{job}}; 
\node at (2.15,4.35) {\scriptsize{1}};
\node at (3.38,4.6) {\scriptsize{job}}; 
\node at (3.38,4.35) {\scriptsize{18}};
\node at (4.92,4.6) {\scriptsize{job}}; 
\node at (4.92,4.35) {\scriptsize{12}};
\node at (7.05,4.6) {\scriptsize{job}}; 
\node at (7.05,4.35) {\scriptsize{6}};
\node at (8.53,4.6) {\scriptsize{job}}; 
\node at (8.53,4.35) {\scriptsize{5}};

% Nurse 2
\draw [thick] (0,2) rectangle (13.9, 3);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,2) rectangle (0.42,3);
\draw [thick] (0.42,2) -- (0.42,3); % start of job 0 (DS)
\draw [thick] (3.2,2) -- (3.2,3); % end of job 0 (DS)
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.2,2) rectangle (3.26,3);
\draw [thick] (3.26,2) -- (3.26,3); % start of job 2
\draw [thick] (3.9,2) -- (3.9,3); % end of job 2
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.9,2) rectangle (4.14,3);
\draw [thick] (4.14,2) -- (4.14,3); % start of job 17
\draw [thick] (6.1,2) -- (6.1,3); % end of job 17
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.1,2) rectangle (6.56,3);
\draw [thick] (6.56,2) -- (6.56,3); % start of job 10
\draw [thick] (8.32,2) -- (8.32,3); % end of job 10
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.32,2) rectangle (8.48,3);
\draw [thick] (8.48,2) -- (8.48,3); % start of job 14
\draw [thick] (9.74,2) -- (9.74,3); % end of job 14
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.74,2) rectangle (10.08,3);
\draw [thick] (10.08,2) -- (10.08,3); % end of shift
\draw [thick, fill=black!20!white] (10.08,2) rectangle (13.9,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
\node at (1.81,2.6) {\scriptsize{job}}; 
\node at (1.81,2.35) {\scriptsize{0 (DS)}};
\node at (3.58,2.6) {\scriptsize{job}}; 
\node at (3.58,2.35) {\scriptsize{2}};
\node at (5.12,2.6) {\scriptsize{job}}; 
\node at (5.12,2.35) {\scriptsize{17}};
\node at (7.44,2.6) {\scriptsize{job}}; 
\node at (7.44,2.35) {\scriptsize{10}};
\node at (9.11,2.6) {\scriptsize{job}}; 
\node at (9.11,2.35) {\scriptsize{14}};

% Nurse 3
\draw [thick] (0,0) rectangle (13.9, 1);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,0) rectangle (0.42,1);
\draw [thick] (0.42,0) -- (0.42,1); % start of job 0 (DS)
\draw [thick] (3.20,0) -- (3.20,1); % end of job  0 (DS)
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.2,0) rectangle (3.28,1);
\draw [thick] (3.28,0) -- (3.28,1); % start of job 7
\draw [thick] (5.42,0) -- (5.42,1); % end of job 7
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.42,0) rectangle (6.14,1);
\draw [thick] (6.14,0) -- (6.14,1); % start of job 15
\draw [thick] (7.18,0) -- (7.18,1); % end of job 15
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.18,0) rectangle (7.7,1);
\draw [thick] (7.70,0) -- (7.70,1); % end of shift
\draw [thick, fill=black!20!white] (7.7,0) rectangle (13.9,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
\node at (1.81,0.6) {\scriptsize{job}}; 
\node at (1.81,0.35) {\scriptsize{0 (DS)}};
\node at (4.35,0.6) {\scriptsize{job}}; 
\node at (4.35,0.35) {\scriptsize{7}};
\node at (6.66,0.6) {\scriptsize{job}}; 
\node at (6.66,0.35) {\scriptsize{15}};

\end{tikzpicture}

\end{document}

### 2021-12-18 tex/Paper/figures/papergantt18-4-l-2c.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (14.04, 7);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,6) rectangle (0.34,7);
\draw [thick] (0.34,6) -- (0.34,7); % start of job 9
\draw [thick] (1.64,6) -- (1.64,7); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.64,6) rectangle (2.12,7);
\draw [thick] (2.12,6) -- (2.12,7); % start of job 3
\draw [thick] (3.28,6) -- (3.28,7); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.28,6) rectangle (3.64,7);
\draw [thick] (3.64,6) -- (3.64,7); % start of job 17
\draw [thick] (5.88,6) -- (5.88,7); % end of job 17
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.88,6) rectangle (6.00,7);
\draw [thick] (6.00,6) -- (6.00,7); % start of job 8
\draw [thick] (7.76,6) -- (7.76,7); % end of job 8
\draw [thick, pattern = north west lines, pattern color=black!30!white] (7.76,6) rectangle (7.92,7);
\draw [thick] (7.92,6) -- (7.92,7); % start of job 1
\draw [thick] (9.88,6) -- (9.88,7); % end of job 1
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.88,6) rectangle (10.12,7);
\draw [thick] (10.12,6) -- (10.12,7); % end of shift
\draw [thick, fill=black!20!white] (10.12,6) rectangle (14.04,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
\node at (0.99,6.6) {\scriptsize{job}}; 
\node at (0.99,6.35) {\scriptsize{9(DS)}};
\node at (2.70,6.6) {\scriptsize{job}};
\node at (2.70,6.35) {\scriptsize{3(DS)}};
\node at (4.76,6.6) {\scriptsize{job}};
\node at (4.76,6.35) {\scriptsize{17}};
\node at (6.88,6.6) {\scriptsize{job}};
\node at (6.88,6.35) {\scriptsize{8}};
\node at (8.90,6.6) {\scriptsize{job}};
\node at (8.90,6.35) {\scriptsize{1}};

% Nurse 1
\draw [thick] (0,4) rectangle (14.04, 5);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,4) rectangle (0.12,5);
\draw [thick] (0.12,4) -- (0.12,5); % start of job 16
\draw [thick] (1.60,4) -- (1.60,5); % end of job 16
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.60,4) rectangle (1.78,5);
\draw [thick] (1.78,4) -- (1.78,5); % start of job 14
\draw [thick] (3.34,4) -- (3.34,5); % end of job 14
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.34,4) rectangle (3.64,5);
\draw [thick] (3.64,4) -- (3.64,5); % start of job 12
\draw [thick] (5.50,4) -- (5.50,5); % end of job 12
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.50,4) rectangle (5.74,5);
\draw [thick] (5.74,4) -- (5.74,5); % end of shift
\draw [thick, fill=black!20!white] (5.74,4) rectangle (14.04,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
\node at (0.86,4.6) {\scriptsize{job}}; 
\node at (0.86,4.35) {\scriptsize{16}};
\node at (2.56,4.6) {\scriptsize{job}}; 
\node at (2.56,4.35) {\scriptsize{14}};
\node at (4.57,4.6) {\scriptsize{job}}; 
\node at (4.57,4.35) {\scriptsize{12}};

% Nurse 2
\draw [thick] (0,2) rectangle (14.04, 3);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,2) rectangle (0.32,3);
\draw [thick] (0.32,2) -- (0.32,3); % start of job 5
\draw [thick] (2.34,2) -- (2.34,3); % end of job 5
\draw [thick, pattern = north west lines, pattern color=black!30!white] (2.34,2) rectangle (2.60,3);
\draw [thick] (2.60,2) -- (2.60,3); % start of job 4
\draw [thick] (4.46,2) -- (4.46,3); % end of job 4
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.46,2) rectangle (4.86,3);
\draw [thick] (4.86,2) -- (4.86,3); % start of job 6
\draw [thick] (6.02,2) -- (6.02,3); % end of job 6
\draw [thick, pattern = north west lines, pattern color=black!30!white] (6.02,2) rectangle (6.48,3);
\draw [thick] (6.48,2) -- (6.48,3); % start of job 2
\draw [thick] (8.45,2) -- (8.45,3); % end of job 2
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.45,2) rectangle (8.49,3);
\draw [thick] (8.49,2) -- (8.49,3); % start of job 10
\draw [thick] (10.36,2) -- (10.36,3); % end of job 10
\draw [thick, pattern = north west lines, pattern color=black!30!white] (10.36,2) rectangle (10.66,3);
\draw [thick] (10.66,2) -- (10.66,3); % end of shift
\draw [thick,fill=black!20!white] (10.66,2) rectangle (14.04,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
\node at (1.33,2.6) {\scriptsize{job}}; 
\node at (1.33,2.35) {\scriptsize{5}};
\node at (3.53,2.6) {\scriptsize{job}}; 
\node at (3.53,2.35) {\scriptsize{4}};
\node at (5.44,2.6) {\scriptsize{job}}; 
\node at (5.44,2.35) {\scriptsize{6}};
\node at (7.47,2.6) {\scriptsize{job}}; 
\node at (7.47,2.35) {\scriptsize{2}};
\node at (9.43,2.6) {\scriptsize{job}}; 
\node at (9.43,2.35) {\scriptsize{10}};

% Nurse 3
\draw [thick] (0,0) rectangle (14.04, 1);
\draw [thick, pattern = north west lines, pattern color=black!30!white] (0,0) rectangle (0.34,1);
\draw [thick] (0.34,0) -- (0.34,1); % start of job 9
\draw [thick] (1.64,0) -- (1.64,1); % end of job 9
\draw [thick, pattern = north west lines, pattern color=black!30!white] (1.64,0) rectangle (2.12,1);
\draw [thick] (2.12,0) -- (2.12,1); % start of job 3
\draw [thick] (3.28,0) -- (3.28,1); % end of job 3
\draw [thick, pattern = north west lines, pattern color=black!30!white] (3.28,0) rectangle (3.34,1);
\draw [thick] (3.34,0) -- (3.34,1); % start of job 11
\draw [thick] (4.72,0) -- (4.72,1); % end of job 11
\draw [thick, pattern = north west lines, pattern color=black!30!white] (4.72,0) rectangle (4.80,1);
\draw [thick] (4.80,0) -- (4.80,1); % start of job 13
\draw [thick] (5.78,0) -- (5.78,1); % end of job 13
\draw [thick, pattern = north west lines, pattern color=black!30!white] (5.78,0) rectangle (5.90,1);
\draw [thick] (5.90,0) -- (5.90,1); % start of job 7
\draw [thick] (8.32,0) -- (8.32,1); % end of job 7
\draw [thick, pattern = north west lines, pattern color=black!30!white] (8.32,0) rectangle (8.38,1);
\draw [thick] (8.38,0) -- (8.38,1); % start of job 0
\draw [thick] (9.54,0) -- (9.54,1); % end of job 0
\draw [thick, pattern = north west lines, pattern color=black!30!white] (9.54,0) rectangle (9.82,1);
\draw [thick] (9.82,0) -- (9.82,1); % start of job 15
\draw [thick] (11.04,0) -- (11.04,1); % end of job 15
\draw [thick, pattern = north west lines, pattern color=black!30!white] (11.04,0) rectangle (11.30,1);
\draw [thick] (11.30,0) -- (11.30,1); % end of shift
\draw [thick, fill=black!20!white] (11.30,0) rectangle (14.04,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
\node at (0.99,0.6) {\scriptsize{job}}; 
\node at (0.99,0.35) {\scriptsize{9(DS)}};
\node at (2.70,0.6) {\scriptsize{job}}; 
\node at (2.70,0.35) {\scriptsize{3(DS)}};
\node at (4.03,0.6) {\scriptsize{job}}; 
\node at (4.03,0.35) {\scriptsize{11}};
\node at (5.29,0.6) {\scriptsize{job}}; 
\node at (5.29,0.35) {\scriptsize{13}};
\node at (7.11,0.6) {\scriptsize{job}}; 
\node at (7.11,0.35) {\scriptsize{7}};
\node at (8.96,0.6) {\scriptsize{job}}; 
\node at (8.96,0.35) {\scriptsize{0}};
\node at (10.43,0.6) {\scriptsize{job}}; 
\node at (10.43,0.35) {\scriptsize{15}};

\end{tikzpicture}

\end{document}

### 2021-12-18 tex/Paper/figures/papergantt18-4-m-2c.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-1) grid (15,8);

% Nurse 0
\draw [thick] (0,6) rectangle (14.6, 7);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,6) rectangle (0.24,7);
\draw [thick] (0.24,6) -- (0.24,7); % start of job 1
\draw [thick] (2.08,6) -- (2.08,7); % end of job 1
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (2.08,6) rectangle (2.38,7);
\draw [thick] (2.38,6) -- (2.38,7); % start of job 16
\draw [thick] (4.48,6) -- (4.48,7); % end of job 16
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (4.48,6) rectangle (4.86,7);
\draw [thick] (4.86,6) -- (4.86,7); % ARRIVE at job 3
\draw [thick, pattern = north west lines, pattern color=tRed] (4.86,6) rectangle (5.22,7);
\draw [thick] (5.22,6) -- (5.22,7); % start of job 3
\draw [thick] (7.26,6) -- (7.26,7); % end of job 3
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (7.26,6) rectangle (7.92,7);
\draw [thick] (7.92,6) -- (7.92,7); % ARRIVE at job 2
\draw [thick, pattern = north west lines, pattern color=tRed] (7.92,6) rectangle (8.28,7);
\draw [thick] (8.28,6) -- (8.28,7); % start of job 2
\draw [thick] (9.32,6) -- (9.32,7); % end of job 2
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (9.32,6) rectangle (9.72,7);
\draw [thick] (9.72,6) -- (9.72,7); % ARRIVE at job 13
\draw [thick, pattern = north west lines, pattern color=tRed] (9.72,6) rectangle (10.80,7);
\draw [thick] (10.80,6) -- (10.80,7); % start of job 13
\draw [thick] (11.72,6) -- (11.72,7); % end of shift
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (11.72,6) rectangle (11.90,7);
\draw [thick] (11.90,6) -- (11.90,7); % end of shift
\draw [thick, fill=black!20!white] (11.90,6) rectangle (14.6,7);

\node [right] at (-0.1,7.2) {\scriptsize{Nurse 0}};
%\node at (1.16,6.6) {\scriptsize{job}}; 
%\node at (1.16,6.35) {\scriptsize{1}};
\node at (1.16,6.5) {\scriptsize{1}};
%\node at (3.34,6.6) {\scriptsize{job}};
%\node at (3.34,6.35) {\scriptsize{16}};
\node at (3.34,6.5) {\scriptsize{16}};
%\node at (6.24,6.6) {\scriptsize{job}};
%\node at (6.24,6.35) {\scriptsize{3}};
\node at (6.24,6.5) {\scriptsize{3}};
%\node at (8.80,6.6) {\scriptsize{job}};
%\node at (8.80,6.35) {\scriptsize{2}};
\node at (8.80,6.5) {\scriptsize{2}};
%\node at (11.26,6.6) {\scriptsize{job}};
%\node at (11.26,6.35) {\scriptsize{13}};
\node at (11.26,6.5) {\scriptsize{13}};

% Nurse 1
\draw [thick] (0,4) rectangle (14.6, 5);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,4) rectangle (0.30,5);
\draw [thick] (0.30,4) -- (0.30,5); % start of job 0
\draw [thick] (1.46,4) -- (1.46,5); % end of job 0
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (1.86,4) rectangle (1.46,5);
\draw [thick] (1.86,4) -- (1.86,5); % start of job 15
\draw [thick] (3.54,4) -- (3.54,5); % end of job 15
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (4.14,4) rectangle (3.54,5);
\draw [thick] (4.14,4) -- (4.14,5); % start of job 7
\draw [thick] (5.56,4) -- (5.56,5); % end of job 7
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (6.04,4) rectangle (5.56,5);
\draw [thick] (6.04,4) -- (6.04,5); % end of shift
\draw [thick, fill=black!20!white] (6.04,4) rectangle (14.6,5);

\node [right] at (-0.1,5.2) {\scriptsize{Nurse 1}};
%\node at (0.88,4.6) {\scriptsize{job}}; 
%\node at (0.88,4.35) {\scriptsize{9}};
\node at (0.88,4.5) {\scriptsize{9}};
%\node at (2.70,4.6) {\scriptsize{job}}; 
%\node at (2.70,4.35) {\scriptsize{15}};
\node at (2.70,4.5) {\scriptsize{15}};
%\node at (4.85,4.6) {\scriptsize{job}}; 
%\node at (4.85,4.35) {\scriptsize{7}};
\node at (4.85,4.5) {\scriptsize{7}};

% Nurse 2
\draw [thick] (0,2) rectangle (14.6, 3);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,2) rectangle (0.26,3);
\draw [thick] (0.26,2) -- (0.26,3); % start of job 10
\draw [thick] (0.74,2) -- (0.74,3); % end of job 10
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0.74,2) rectangle (1.24,3);
\draw [thick] (1.24,2) -- (1.24,3); % start of job 5
\draw [thick] (3.46,2) -- (3.46,3); % end of job 5
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (3.46,2) rectangle (3.60,3);
\draw [thick] (3.60,2) -- (3.60,3); % ARRIVE at job 8
\draw [thick, pattern = north west lines, pattern color=tRed] (3.60,2) rectangle (4.94,3);
\draw [thick] (4.94,2) -- (4.94,3); % start of job 8
\draw [thick] (6.20,2) -- (6.20,3); % end of job 8
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (6.20,2) rectangle (6.34,3);
\draw [thick] (6.34,2) -- (6.34,3); % start of job 17
\draw [thick] (7.92,2) -- (7.92,3); % end of job 17
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (7.92,2) rectangle (8.48,3);
\draw [thick] (8.48,2) -- (8.48,3); % start of job 18
\draw [thick] (9.52,2) -- (9.52,3); % end of job 18
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (9.52,2) rectangle (9.70,3);
\draw [thick] (9.70,2) -- (9.70,3); % start of job 12
\draw [thick] (11.22,2) -- (11.22,3); % end of job 12
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (11.22,2) rectangle (11.70,3);
\draw [thick] (11.70,2) -- (11.70,3); % end of shift
\draw [thick,fill=black!20!white] (11.70,2) rectangle (14.6,3);

\node [right] at (-0.1,3.2) {\scriptsize{Nurse 2}};
%\node at (0.50,2.6) {\scriptsize{job}}; 
%\node at (0.50,2.35) {\scriptsize{10}};
\node at (0.50,2.5) {\scriptsize{10}};
%\node at (2.35,2.6) {\scriptsize{job}}; 
%\node at (2.35,2.35) {\scriptsize{5}};
\node at (2.35,2.5) {\scriptsize{5}};
%\node at (5.57,2.6) {\scriptsize{job}}; 
%\node at (5.57,2.35) {\scriptsize{8}};
\node at (5.57,2.5) {\scriptsize{8}};
%\node at (7.13,2.6) {\scriptsize{job}}; 
%\node at (7.13,2.35) {\scriptsize{17}};
\node at (7.13,2.5) {\scriptsize{17}};
%\node at (9.00,2.6) {\scriptsize{job}}; 
%\node at (9.00,2.35) {\scriptsize{18}};
\node at (9.00,2.5) {\scriptsize{18}};
%\node at (10.46,2.6) {\scriptsize{job}}; 
%\node at (10.46,2.35) {\scriptsize{12}};
\node at (10.46,2.5) {\scriptsize{12}};

% Nurse 3
\draw [thick] (0,0) rectangle (14.6, 1);
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0,0) rectangle (0.34,1);
\draw [thick] (0.34,0) -- (0.34,1); % start of job 11
\draw [thick] (0.64,0) -- (0.64,1); % end of job 11
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (0.64,0) rectangle (1.28,1);
\draw [thick] (1.28,0) -- (1.28,1); % start of job 4
\draw [thick] (2.30,0) -- (2.30,1); % end of job 4
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (2.30,0) rectangle (2.54,1);
\draw [thick] (2.54,0) -- (2.54,1); % start of job 6
\draw [thick] (3.18,0) -- (3.18,1); % end of job 6
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (3.18,0) rectangle (3.50,1);
\draw [thick] (3.50,0) -- (3.50,1); % ARRIVE at job 19
\draw [thick, pattern = north west lines, pattern color=tRed] (3.50,0) rectangle (5.02,1);
\draw [thick] (5.02,0) -- (5.02,1); % start of job 19
\draw [thick] (7.06,0) -- (7.06,1); % end of job 19
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (7.06,0) rectangle (7.36,1);
\draw [thick] (7.36,0) -- (7.36,1); % start of job 4
\draw [thick] (8.58,0) -- (8.58,1); % end of job 4
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (8.58,0) rectangle (8.68,1);
\draw [thick] (8.68,0) -- (8.68,1); % start of job 9
\draw [thick] (10.22,0) -- (10.22,1); % end of job 9
\draw [thick, pattern = crosshatch dots, pattern color=black!30!white] (10.22,0) rectangle (10.54,1);
\draw [thick] (10.54,0) -- (10.54,1); % end of shift
\draw [thick, fill=black!20!white] (10.54,0) rectangle (14.6,1);

\node [right] at (-0.1,1.2) {\scriptsize{Nurse 3}};
%\node at (0.49,0.6) {\scriptsize{job}}; 
%\node at (0.49,0.35) {\scriptsize{11}};
\node at (0.49,0.5) {\scriptsize{11}};
%\node at (1.79,0.6) {\scriptsize{job}}; 
%\node at (1.79,0.35) {\scriptsize{4}};
\node at (1.79,0.5) {\scriptsize{4}};
%\node at (2.86,0.6) {\scriptsize{job}}; 
%\node at (2.86,0.35) {\scriptsize{6}};
\node at (2.86,0.5) {\scriptsize{6}};
%\node at (6.04,0.6) {\scriptsize{job}}; 
%\node at (6.04,0.35) {\scriptsize{19}};
\node at (6.04,0.5) {\scriptsize{19}};
%\node at (7.97,0.6) {\scriptsize{job}}; 
%\node at (7.97,0.35) {\scriptsize{4}};
\node at (7.97,0.5) {\scriptsize{4}};
%\node at (9.45,0.6) {\scriptsize{job}}; 
%\node at (9.45,0.35) {\scriptsize{9}};
\node at (9.45,0.5) {\scriptsize{9}};

\end{tikzpicture}

\end{document}

%\draw [thick] (0,6) rectangle (6.95, 7);
%\draw [thick] (0.11,6) -- (0.11,7); % start of job 10
%\draw [thick] (0.99,6) -- (0.99,7); % end of job 10
%\draw [thick] (1.07,6) -- (1.07,7); % start of job 14
%\draw [thick] (1.7,6) -- (1.7,7); % end of job 14
%\draw [thick] (1.84,6) -- (1.84,7); % start of job 3
%\draw [thick] (2.86,6) -- (2.86,7); % end of job 3
%\draw [thick] (2.93,6) -- (2.93,7); % start of job 11
%\draw [thick] (3.4,6) -- (3.4,7); % end of job 11
%\draw [thick] (3.66,6) -- (3.66,7); % start of job 17
%\draw [thick] (4.64,6) -- (4.64,7); % end of job 17
%\draw [thick] (4.76,6) -- (4.76,7); % start of job 2
%\draw [thick] (5.08,6) -- (5.08,7); % end of job 2
%\draw [thick] (5.11,6) -- (5.11,7); % start of job 0 (DS)
%\draw [thick] (6.5,6) -- (6.5,7); % end of job 0 (DS)
%\draw [thick] (6.71,6) -- (6.71,7); % end of shift

### 2021-12-18 tex/Paper/figures/shiftwork1.tex
new tikz figures.

\documentclass[usenames,dvipsnames,tikz]{standalone}
\usetikzlibrary{patterns}
%\usetikzlibrary{shapes.geometric}
%\usepackage{xcolor}
\colorlet{tBlue}{RoyalBlue!35!Cerulean}
\colorlet{tRed}{Red}
%\definecolor{tGreen}{HTML}{569909}
%\definecolor{tOrange}{HTML}{FA7602}
\usepackage{tikz}
\usepackage{standalone}
\begin{document}

\begin{tikzpicture}
%\draw [help lines] (-1,-2) grid (15,7);

\draw [thick] (0,2) rectangle (8,4);
\draw [thick, fill=black!10!white] (2,2) rectangle (5,4);
%\draw [thick, pattern = north west lines, pattern color=black!40!white] (2,2) rectangle (5,4);
\draw [thick] (2,2) -- (2,4); % start of unavailable shift time
\draw [thick] (5,2) -- (5,4); % end of unavailable shift time

% Original job - duration 2.75, arrival 1.25, start TW 1, end TW 1.5, end 4
\draw [thick, tRed] (1,3) -- (4,3); % line for duration of original job, length 3, includes start of TW
\draw [thick, tRed] (1,2.85) -- (1,3.15); % vertical line for start of TW 
\draw [thick, tRed] (1.5,2.85) -- (1.5,3.15); % vertical line for end of TW
\draw [fill=white, thick] (1.25,3) circle [radius = 0.08]; % original arrival time at job
\draw [fill=white, thick] (4,3) circle [radius = 0.08]; % original end time of job

% New job - duration 2.75, arrival/start 5, end 7.75
\draw [thick, tBlue] (5,3) -- (7.75,3); % line for duration of new job, length 2.75,
\draw [fill=white, thick] (5,3) circle [radius = 0.08]; % new arrival time at job
\draw [fill=white, thick] (7.75,3) circle [radius = 0.08]; % new end time of job

% Waiting time
\draw [thick] (1.25, 1.9) -- (1.25, 1.65) -- (2, 1.65) -- (2,1.9);
\node at (1.625, 1.45) {\scriptsize{$w_{ij}$}};

% Tardiness
\draw [thick] (1.5, 1.35) -- (1.5, 1.1) -- (5, 1.1) -- (5,1.35);
\node at (3.25, 0.95) {\scriptsize{$z_j$}};

% Unavailable time
\draw [thick] (2,4.4) -- (2,4.65) -- (5,4.65) -- (5,4.4);
\node at (3.5, 4.8) {\scriptsize{break}};

% Labels
\node at (0,4.2) {\scriptsize{$u(\pi_1^i)$}};
\node at (2,4.2) {\scriptsize{$v(\pi_1^i)$}};
\node at (5,4.2) {\scriptsize{$u(\pi_2^i)$}};
\node at (8,4.2) {\scriptsize{$v(\pi_2^i)$}};
\node at (1.25, 3.77) {\scriptsize{original}};
\node at (1.25, 3.52) {\scriptsize{$a(i,j)$}};
\draw [thick, tRed, ->] (1.25, 3.4) -- (1.25, 3.15);
\node at (0.95, 2.65) {\scriptsize{$e_j$}};
\node at (1.55, 2.65) {\scriptsize{$l_j$}};
\node at (5.45, 3.77) {\scriptsize{new}};
\node at (5.45, 3.52) {\scriptsize{$a(i,j)$}};
\draw [thick, tBlue, ->] (5.35, 3.35) -- (5.1, 3.15);

%\draw [thick] (0,0) rectangle (4,1.55);
%\draw [thick, dashed] (0.5,0) -- (0.5, 1.55);
%\draw [thick, dashed] (3,0) -- (3, 1.55);
%\node at (2,0.75) {\LARGE{$i$}};
%\draw [ultra thick, tBlue] (1,3) -- (1.8,3);
%\draw [tRed] (1.8,3) -- (2.8,3);
%\draw [fill=white, thick] (1,3) circle [radius = 0.08];
%\node [above] at (1,3.05) {\large{$v_1$}};
%\node at (1.025,2.5) {$4$};

\end{tikzpicture}

\end{document}

### 2021-12-18 tex/Paper/workloadbalance.tex
new file.

\documentclass[a4paper]{article}

\input{includes/preamble.tex}
\usepackage{fullpage}

\begin{document}
\title{Workload Balance}
%\author{Asyl L. Hawa}
\date{}
\maketitle

\section*{Objectives}
\begin{enumerate}[leftmargin=*]
	\item Bowers et al. (2015) -- (page 198) A simple workload balancing requirement is incorporated with a constraint that no member of staff should be allocated more than 4 patients. This assumes that all pateints require the same staff-time, excluding travel. A more sophisticated algorithm could consider the complete workloads and the likely time commitments associated with each patient visit, including the time for direct patient contact and indirect administrative tasks.
	\item Bredstrom and Ronnqvist (2008) -- (page 21) The difference in workload is measured with the pairwise vehicle difference sum of service durations. (pages 25--26) -- Balancing constraint given (eqn 9), goal of having a mean of 5 hours workload (excluding travelling time) for each staff member. (page 27) Objective is to minimise the maximal difference in workload (excluding travel time, only considering the sum of durations/service time).
	\item Lanzarone and Matta (2014) -- (page 49) The problem is formalised as a minimisation of nurses' maximum overtimes in a lexicographic way: firstly, we minimise the highest value, then the highest value of the remaining ones, and so forth. Operatively, the goal is analytically pursued by minimising the maximum increase of a stochastic cost function, which depends on the time spent by nurses when providing visits in surplus to their capacity. (page 50) OF: minimisation of the maximum assigned workload, which corresponds to reducing nurses' overtime: OF: minimisation of the difference between the maximum and average assigned workload, which corresponds to balancing the workloads among the nurses. Constraints related to maxmimum acceptable workloads and the assignment of each patient to exactly one nurse of each type are imposed, and the obj of the assignment is to balance the nurses' workloads by minimising a weighted sum of the number of provided visits, of the assigned patients, and of the distances travelled. (page 51) Equation (1) shows the formulation for min the max overtimes of nurses in lexicographical order.
	\item Mutingi and Mbohwa (2014) -- (page 210) Workload is measured as a sum of travel times and duration of visits. (page 212) For each nurse $i$ the workload is $w_i$, the percentage deviation $d_i$ of $w_i$ from the average workload $\overline{w}$ is calculated as $d_i =100.\frac{w_i - \overline{w}}{\overline{w}}$. Function is supposed to be as minimal as possible if the schedule is to be satisfactory to the nurse, the patient, and the management/decision maker. Function is converted using fuzzy interval-valued linear membership functions. In practice, values of $a$ and $b$ for the membership function are derived from expert knowledge and intuition -- for instance, average workload $\overline{w}$ is set as standard or target workload s.t. the var of individual workloads from $\overline{w}$ should at best be within the range $[0,a]$. Here, $a$ is the agreeable deviation of working time (hours) from the target $\overline{w}$, considered unlikely to cause any dissatisfaction among nurses. To find $b$, need to find the min of the max acceptable deviation of working time from the target. (page 213) Fitness function has user defined weights to signify the relative important of workload, time windows, and distance travelled.
\end{enumerate}

\section{Workload Balance}

Currently, 
\begin{equation}
	\bar{B} = \min_{i \in N} \{\max\{0, en(i) - arr(i)\}\},
\end{equation}
which only considers overtime.
Instead:
\begin{equation}
	wl_i = \frac{\displaystyle\sum_{h=0}^{mi} s_{k_{h}^{i}}}{en(i) - st(i)}
\end{equation}
where $s_{k_{h}^{i}}$ is the service time for each job $k_{h}^{i}$ in nurse $i$'s route $r_i = \{k_{0}^{i}, k_{1}^{i}, k_{2}^{i}, \dotsc, k_{mi}^{i}, k_{0}^{i}\}$. Thus, $wl_i$ is the sum of all service times of jobs performed by nurse $i$ divided by the length of nurse $i$'s shift.

Then, let

\begin{equation}
	\bar{wl} = \frac{\displaystyle\sum_{i \in N} wl_i}{n},
\end{equation}
which is the average of all workload times.

Finally,
\begin{equation}
	\bar{B} = \max_{i \in N}\{wl_i - \bar{wl}\},
\end{equation}	
that is, we take the maximum of all differences of workloads to the average.

\end{document}

### 2021-12-18 tex/Plan/Plan.tex
updated files.

--- 2021-10-25 tex/Plan/Plan.tex
@@
 	\item \textbf{Real Evaluation of the DST}\\
-	This will be done over a period of time in two ways: (a) comparing the automated plans generated by the DST with plans created manually by staff; and (b) obtaining and assessing reviews from nurses regaring the new automated plans.
+	This will be done over a 3-month period in two ways: (a) comparing the automated plans generated by the DST with plans created manually by staff; and (b) obtaining and assessing reviews from nurses regarding the new automated plans.
 This will provide us with a clearer understanding of the impact of the DST in practice.
@@
 	An exact method would provide a guarantee that the problem is solved to optimality, however, it may not be suitable for daily planning due to the time and software that may be required.
-An exact method would be useful for Health Services for considering whether investments in planning are helpful or necessary, and will also be useful as a basis for other types of modelling.
-	\item \textbf{Stochastic Demand}\\
+An exact method would be useful for Health Services for considering whether investments in planning are helpful or necessary, and will also be useful as a basis for other types of modelling such as stochastic modelling (see Objective 3).
+	\item \textbf{Stochastic Model for Unexpected Changed in Plans}\\
 	We will identify particular situations that require modifications to existing plans, e.g.
@@
 \begin{enumerate}[leftmargin=*]
-	\item \textbf{Decision Support Tool}\\
-	We aim to release the DST for the planning of domiciliary care to make the software readily available to health care teams to assist with creating routes and schedules for staff.
+	\item Release a stable version of the DST along with documentation and examples for the planning of domiciliary care to make the software readily available to health care teams to assist with creating routes and schedules for staff.
 This will also allow us to obtain feedback from the users and a comparison with previous schedules.
-	\item \textbf{Publications}\\
-	We endeavour to submit two articles for publication: a methodological paper covering findings regarding exact methods and adaptations to stochastic demand, which will be submitted to a mathematical journal, and a scientific paper for a healthcare journal that will present our work on the efficiency of our DST in a real setting in comparison to business-as-usual planning.
+	\item A methodological paper covering findings regarding exact methods (Objective 2) and adaptations for stochastic demand (Objective 3), which will be submitted to a highly-ranked mathematical journal such as the European Journal of Operational Research (EJOR).
+	\item A scientific paper for a healthcare journal that will present our work on the efficiency of our DST in a real setting in comparison to business-as-usual planning (Objective 1). 
 \end{enumerate}
@@
 
+\textbf{18th October 2021 -- 7th November 2021}\\
+Finalise initial publication and prepare to submit to journal.
+
+\textbf{8th November 2021 -- 14th November 2021}\\
+Begin research on exact methods for the HHCRSP and stochastic methods for changes in schedules.
+
+\textbf{15th November 2021 -- 21st November 2021}\\
+Meet with Abicare to discuss the frequency in which the DST should be used and how results should be collected for future review.
+Prepare the DST for release, and prepare a questionnaire for nurses and planners at Abicare which will be used as a basis for comparison after the DST has been trialed.
+
+\textbf{22nd November 2021 -- 28th November 2021}\\
+Complete the documentation and release the DST. Send out the initial questionnaire to nurses and planners at Abicare.
+
+\textbf{29th November 2021 -- 19th December 2021}\\
+Continue research for exact and stochastic methods.
+
+\textbf{20th December 2021 -- 26th December 2021}\\
+Start creating models for exact and stochastic methods.
+
+\textbf{3rd January 2022 -- 16th January 2022}\\
+Implement models for exact and stochastic methods into DST.
+
+\textbf{17th January 2022 -- 23rd January 2022}\\
+Start gathering data/creating instances to test models in DST.
+
+\textbf{24th January 2022 -- 6th February 2022}\\
+Begin running experiments using the DST and analyse results.
+
+\textbf{31st January 2022 -- 6th February 2022}\\
+Fix any issues with DST (debugging) and continue analysing results.
+
+\textbf{7th February 2022 -- 17th April 2022}\\
+Work on publications for mathematical and healthcare journals.
+
+\textbf{7th March 2022 -- 13th March 2022}\\
+Gather schedules from Abicare using DST, and send another questionnaire to nurses and planners.
+
+\textbf{18th April 2022 -- 24th April 2022}\\
+Continue working on publications for mathematical and healthcare journals.
+
+\textbf{25th April 2022 -- 30th April 2022}\\
+Conclude project.
+
+\begin{comment}
 	\begin{enumerate}[leftmargin=*]
@@
 
+	\begin{enumerate}[leftmargin=*]
+	\item \textbf{18th October 2021 -- 7th November 2021}\\
+	Finalise initial publication and prepare to submit to journal.
+	\item \textbf{8th November 2021 -- 14th November 2021}\\
+	Begin research on exact methods for the HHCRSP and stochastic methods for changes in schedules.
+	\item \textbf{15th November 2021 -- 21st November 2021}\\
+	Meet with Abicare to discuss the frequency in which the DST should be used and how results should be collected for future review.
+Prepare the DST for release, and prepare a questionnaire for nurses and planners at Abicare which will be used as a basis for comparison after the DST has been trialed.
+	\item \textbf{22nd November 2021 -- 28th November 2021}\\
+	Complete the documentation for the DST, and release the DST. Send out the initial questionnaire to nurses and planners at Abicare.
+	\item \textbf{29th November 2021 -- 19th December 2021}\\
+	Continue research for exact and stochastic methods.
+	\item \textbf{20th December 2021 -- 26th December 2021}\\
+	Start creating models for exact and stochastic methods.
+	\item \textbf{3rd January 2022 -- 16th January 2022}\\
+	Implement models for exact and stochastic methods into DST.
+	\item \textbf{17th January 2022 -- 23rd January 2022}\\
+	Start gathering data/creating instances to test models in DST.
+	\item \textbf{24th January 2022 -- 6th February 2022}\\
+	Begin running experiments using the DST and analyse results.
+	\item \textbf{31st January 2022 -- 6th February 2022}\\
+	Fix any issues with DST (debugging) and continue analysing results.
+	\item \textbf{7th February 2022 -- 17th April 2022}\\
+	Work on publications for mathematical and healthcare journals.
+	\item \textbf{7th March 2022 -- 13th March 2022}\\
+	Gather schedules from Abicare using DST, and send another questionnaire to nurses and planners.
+	\item \textbf{18th April 2022 -- 24th April 2022}\\
+	Continue working on publications for mathematical and healthcare journals.
+	\item \textbf{25th April 2022 -- 30th April 2022}\\
+	Conclude project.
+	\end{enumerate}	
+\end{comment}
+
 \end{document}

### 2021-12-18 code/CarrotAll/constructive_wrapper.py

@@ -1,0 +1,215 @@
+#-----------------------#
+# CARROT - CARe ROuting Tool
+# constructive_wrapper.py
+# main file to run the program
+# 01/11/2020
+#-----------------------#
+
+import os
+import time
+import numpy as np
+import pandas as pd
+from datetime import datetime
+from tkinter import *
+
+# Own modules:
+import instance_handler_dfs as ihd
+import tools_and_scripts.get_data_dfs as gdd
+import tools_and_scripts.tkinter_userinput as tui
+import tools_and_scripts.week_counter as wc
+
+def main():
+    
+    # Get user input variables using tkinter entry box:
+    # area, tw_interval, wb_balance, quality_measure, max_time_seconds, create_html_website, create_python_plots, codepoint_directory, input_filename, date_selected = tui.get_user_input_variables()
+    area = 'Hampshire'
+    tw_interval = 15
+    wb_balance = 1
+    quality_measure = 'default' # default = paper
+    max_time_seconds = 60
+    create_html_website = False
+    create_python_plots = False
+    input_filename = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/CarrotAll/data/abicare/2021.07.15 - UOS - Clients - Carers - Rota - Anonymised.xlsx'
+    codepoint_directory = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/CarrotAll/data/codepo_gb'
+    date_selected = pd.Timestamp("2021-07-27")
+    instance_type = 'ait_h' # Can be either 'abicare' or 'ait_h'
+    # random_seed = 13027 # Should random seed be added to tkinter box?
+    # random_seed = -1 # Only if run_in_parallel = False. -1 is a true random seed, otherwise set to this value
+
+    # Convert date_selected from datetime_date to Timestamp
+    # planning_date = pd.Timestamp(date_selected)
+
+    # Create options vector
+    options_vector = ihd.default_options_vector() 
+
+    if instance_type == 'abicare':
+        random_seed = 13027 # Should random seed be added to tkinter box?
+
+        planning_date = date_selected
+        
+        # Obtain day of week index of planning_date for both 2-week and 8-week schedules
+        week_period2 = 2
+        week_period8 = 8
+        dayindex_2weeks = wc.calculate_cycle_day(planning_date, week_period2)
+        dayindex_8weeks = wc.calculate_cycle_day(planning_date, week_period8)
+
[…]

### 2021-12-18 code/CarrotAll/instance_handler_dfs.py

@@ -1,0 +1,1878 @@
+#-----------------------#
+# CARROT - CARe ROuting Tool
+# instance_handler_dfs.py
+# main file containing functions to build and solve the instance of the problem
+# 06/05/2021
+#-----------------------#
+
+import os
+import osrm
+import math
+import time
+import ctypes
+import folium
+import datetime
+import requests # For osrm API
+import subprocess
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from math import sqrt
+from datetime import timedelta
+from numpy.ctypeslib import ndpointer
+
+import literature_conversions as lc
+
+# Display all rows and columns of dataframes in command prompt:
+pd.set_option("display.max_rows", None, "display.max_columns", None)
+
+def create_solve_inst(client_df, nurseshift_df, nurseday_df, planning_date, options_vector, wb_balance, quality_measure, max_time_seconds, random_seed):
+    
+    # Create instance of the INSTANCE class
+    inst = INSTANCE()
+   
+    # Build/fill inst object
+    inst = convert_dfs_inst(inst, client_df, nurseshift_df, nurseday_df, planning_date)
+        
+    inst.lambda_1 = 1
+    inst.lambda_2 = 1
+    inst.lambda_3 = 1
+    inst.lambda_4 = 1
+    inst.lambda_5 = 1
+    inst.lambda_6 = 10
+    inst.quality_measure = quality_measure
+    inst.MAX_TIME_SECONDS = max_time_seconds
+    options_vector[55] = wb_balance # alpha_5 Workload balance (from user input)
+
+    # Create instances of JOB and NURSE classes for inst
+    inst.instantiate_job_nurse_objects(client_df, nurseday_df)
+
+    inst.algorithmOptions = options_vector
+    if quality_measure == 'default':
+        inst.algorithmOptions[0] = 6.0 # paper (default)
+    elif quality_measure == 'ait h':
+        inst.algorithmOptions[0] = 0.0 # ait_h
+    elif quality_measure == 'mk':
+        inst.algorithmOptions[0] = 1.0 # mankowska
+    elif quality_measure == 'wb':
+        inst.algorithmOptions[0] = 5.0 # workload_balance
+    else:
+        print('[ERROR]: (instance_handler.py) incorrect quality measure provided.')
[…]

### 2021-12-18 code/CarrotAll/literature_conversions.py

@@ -1,0 +1,279 @@
+#-----------------------#
+# CARROT - CARe ROuting Tool
+# literature_conversions.py
+# file containing functions to convert literature instances (ait_h, mankowska etc) to instance for the program.
+# 09/11/2021
+#-----------------------#
+
+import numpy as np
+import imp
+import copy
+
+def convert_aith_instance(filename, inst):
+
+    # Load Ait Hadadene file
+    ldi = imp.load_source('ait_h', filename)
+
+    inst.nJobs = ldi.NbClients
+    inst.nNurses = ldi.NbVehicules
+    inst.nSkills = ldi.NbServices
+
+    # inst.nurseWorkingTimes = np.zeros((inst.nNurses, 3), dtype=np.int)
+    inst.nurseWorkingTimes = np.zeros((inst.nNurses, 5), dtype=np.int)# [0] = start time, [1] = end time, [2] = total length of day, [3] = number of shifts (always 1) [4] = duration of actual working times (always the same as total length of day [2])
+    for i in range(inst.nNurses):
+        inst.nurseWorkingTimes[i][0] = ldi.WindowsD[i][0]
+        inst.nurseWorkingTimes[i][1] = ldi.WindowsD[i][1]
+        inst.nurseWorkingTimes[i][2] = ldi.WindowsD[i][1] - ldi.WindowsD[i][0]
+        inst.nurseWorkingTimes[i][3] = 1
+        inst.nurseWorkingTimes[i][4] = inst.nurseWorkingTimes[i][2]
+
+    inst.nurseSkills = np.ascontiguousarray(ldi.Offres, dtype=np.int)
+    inst.jobTimeInfo = np.zeros((inst.nJobs, 3), dtype=np.int)
+    inst.jobSkillsRequired = np.ascontiguousarray(ldi.Demandes, dtype=np.int)
+
+    # print(ldi.WindowsC)
+    # print(len(ldi.WindowsC))
+    # print(inst.nJobs)
+    for job in range(inst.nJobs):
+        inst.jobTimeInfo[job][0] = ldi.WindowsC[job][0]
+        inst.jobTimeInfo[job][1] = ldi.WindowsC[job][1]
+        for sk in range(ldi.NbServices):
+            if (inst.jobSkillsRequired[job][sk] > 0):
+                inst.jobTimeInfo[job][2] = ldi.DureeDeVisite[job][sk]
+
[…]

### 2021-12-18 code/CarrotAll/src/GRASP_VNS_ALL/constructive.c

@@ -1,0 +1,4367 @@
+/*--------------/
+GRASP_VNS
+constructive.c
+UoS
+10/07/2021
+/--------------*/
+
+//#define _CRTDBG_MAP_ALLOC
+#include <stdlib.h>
+//#include <crtdbg.h>
+
+
+//#include <stdlib.h> // To have min in linux
+#include "Python.h"
+#include "constructive.h"
+#include "grasp.h"
+
+// #ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+// #    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+// #else
+// #  define MODULE_API
+// #endif
+
+int main(int argc, char const* argv[]){
+
+    //_CrtSetDbgFlag ( _CRTDBG_ALLOC_MEM_DF | _CRTDBG_LEAK_CHECK_DF );
+    //_CrtSetReportMode( _CRT_WARN, _CRTDBG_MODE_DEBUG );
+
+    struct INSTANCE inst = GenerateInstance();
+    struct INSTANCE* ip = &inst;
+    ip->verbose = 4;
+    return MainWithOutput(ip, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL);
+}
+
+MODULE_API int python_entry(int nJobs_data, int nNurses_data, int nSkills_data, int verbose_data, float MAX_TIME_SECONDS, int twInterval_data, bool excludeNurseTravel_data,
+                            double* od_data, double* nurseTravelFromDepot_data, double* nurseTravelToDepot_data, int* unavailMatrix_data, int* nurseUnavail_data,
+                            int* nurseWorkingTimes_data, int* jobTimeInfo_data, int* jobRequirements_data, int* nurseSkills_data, int* solMatrixPointer,
+                            int* doubleService_data, int* dependsOn_data, int* mkMinD_data, int* mkMaxD_data, int* capabilityOfDoubleServices_data, double* prefScore_data,
+                            double* algorithmOptions_data, double* timeMatrixPointer, double* nurseWaitingTimePointer, double* nurseTravelTimePointer, double* violatedTWPointer,
+                            double* nurseWaitingMatrixPointer, double* nurseTravelMatrixPointer, double* totalsArrayPointer, int randomSeed){
+
+    // printf("Inside C function. Random seed = %d.\n", randomSeed);
+
+    double eps = 1e-6;
[…]

### 2021-12-18 code/CarrotAll/src/GRASP_VNS_ALL/constructive.h

@@ -1,0 +1,196 @@
+/*--------------/
+GRASP_VNS
+constructive.h
+UoS
+10/07/2021
+/--------------*/
+
+
+#include <stdio.h>
+#include <stdbool.h> // Required to use 'bool' variable type
+#include <stdlib.h>
+#include <time.h>
+
+#ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+#    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+#else
+#  define MODULE_API
+#endif
+
+
+#define bigM 9999999;
+
+struct INSTANCE {
+	int nJobs; // Number of jobs
+    int nJobsIncDS; //Number of tasks in total, nJobs+nDS
+	int nNurses; // Number of carers
+	int nSkills; //Number of skills
+	int verbose; // Additional information
+	int isFeasible; // Feasibility (deprecated)
+	double objTime; // Objective time
+	double objWaiting; // Objective waiting time, NEW 07/01/2021
+	double objTravel; // Objective travel time
+	double objService; // Objective service time
+	double objTardiness; // Objective tardiness
+	double objMaxTardiness; // Objective max tardiness
+	double objMKTardiness; // Objective total MK tardiness
+	double objMKAllowedTardiness; // Objective mk_allowed_tardiness
+	double objOvertime; // Objective overtime
+	double objMaxOvertime; // Objective max overtime
+	double objMinSpareTime;
+	double objMaxSpareTime;
+	double objShortestDay; // Objective shortest day, NEW 07/01/2021
+	double objLongestDay; // Objective longest day
+	double objAitHQuality;
+	double objMKQuality;
+	double objWBQuality;
+	double objPaperQuality;
+	double objQuality;
+    double objMaxTravelTime; // NB new 16/11/2021
+    double objMaxWaitingTime; // NB new 16/11/2021
+    double objMaxDiffWorkload; // NB new 16/11/2021
+	float totalPref; // Total preference score
+	int qualityMeasure; // Temp. Allow to choose (Mk, Ait h., etc.) Ideally we want a vector with weights! This should be set ONLY in "InstanceFromPython" (read_instance.c). The value is hardcoded at the moment.
+	float MAX_TIME_SECONDS; // Maximum time limit in seconds
+	int twInterval; // Time window interval
[…]

### 2021-12-18 code/CarrotAll/src/GRASP_VNS_ALL/grasp.h

@@ -1,0 +1,36 @@
+/*--------------/
+GRASP_VNS
+grasp.h
+UoS
+10/07/2021
+/--------------*/
+
+#include "math.h"
+
+
+int GRASP(struct INSTANCE * ip);
+void RandomisedConstructive(struct INSTANCE * ip, int randomness, double delta, int rclStrategy);
+void GRASPLS(struct INSTANCE * ip, int LS_ITERS);
+void GenerateRCL(double delta, int* rclSeeds, int* rclSize, double* rankValues, int rvSize, double bestRank, double worstRank, int strategy);
+void RCLPick(int ** bestIndices, int * rclSeeds, int rclSize, int * cNurse, int * cJob);
+void BestIndexPick(int ** bestIndices, int GRASP_param, int * cNurse, int * cJob);
+int PickInteger(int max_int);
+void UpdateTopValues(int GRASP_param, int ** bestIndices, double * bestValues, double newValue, int bestNurse, int bestJob);
+void CleanSolutionFromStruct(struct INSTANCE * ip);
+int IdentifyEarlyInsert(struct INSTANCE * ip, struct INSTANCE * guiding, int guidingNurse, int guidingNursePos);
+void BestWorstPoolQualities(double * poolBest, double * poolWorst, struct INSTANCE ** pool, double * poolQuality, int poolSize);
+double SolutionDissimilarity(struct INSTANCE * input1, struct INSTANCE * input2);
+int FindArcDestination(int sourceNurse, int sourceJob, struct INSTANCE * ip);
+void PrintRecalculatePoolContents(struct INSTANCE ** pool, double * poolQuality, int solutionsInPool);
+void CalculatePrintDissimilarityMatrix(struct INSTANCE ** pool, int solutionsInPool);
+double ForwardPR(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output);
+double ForwardBackwardPR(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output1);
+double BackwardPR(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output);
+double DirectedPR(struct INSTANCE * input1, struct INSTANCE * input2, double q1, double q2, struct INSTANCE * output, int direction);
+double PathRelinking(struct INSTANCE * ip, struct INSTANCE * guiding);
[…]

### 2021-12-18 code/CarrotAll/src/GRASP_VNS_ALL/read_instance.c

@@ -1,0 +1,1475 @@
+/*--------------/
+GRASP_VNS
+read_instance.c
+UoS
+10/07/2021
+/--------------*/
+
+
+#include "constructive.h"
+void OverwriteInstance(struct INSTANCE* ip, struct INSTANCE* oi) {
+	/*
+	Copies the contents of the solution of "oi" into "ip"
+	without creating any new allocations.
+	Only copies the solution, no algorithm options or data
+	*/
+
+	for (int i = 0; i < oi->nNurses; ++i){
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->solMatrix[i][j] = oi->solMatrix[i][j];
+        }
+    }
+
+	for (int i = 0; i < oi->nNurses; ++i){
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->timeMatrix[i][j] = oi->timeMatrix[i][j];
+        }
+    }
+
+	for (int i = 0; i < oi->nNurses; ++i){
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->allNurseRoutes[i][j] = oi->allNurseRoutes[i][j];
+        }
+    }
+
+    for (int i = 0; i < oi->nNurses; ++i){ // NB: NEW: 03/06/2021
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->nurseWaitingMatrix[i][j] = oi->nurseWaitingMatrix[i][j];
+        }
+    }
+
+    for (int i = 0; i < oi->nNurses; ++i){ // NB: NEW: 03/06/2021
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->nurseTravelMatrix[i][j] = oi->nurseTravelMatrix[i][j];
+        }
+    }
+
+	for (int i = 0; i < oi->nNurses; ++i) {
+		ip->nurseOrder[i] = oi->nurseOrder[i];
+		ip->nurseWaitingTime[i] = oi->nurseWaitingTime[i];
+		ip->nurseTravelTime[i] = oi->nurseTravelTime[i];
+	}
+
+	for (int i = 0; i < oi->nJobs; ++i) {
+		ip->nurseRoute[i] = oi->nurseRoute[i];
+		ip->violatedTW[i] = oi->violatedTW[i];
+		ip->violatedTWMK[i] = oi->violatedTWMK[i];
+	}
+
+} //END OF OverwriteInstance function.
+
[…]

### 2021-12-18 code/CarrotAll/tools_and_scripts/class_cpo_df.py

@@ -1,0 +1,114 @@
+#--------------------#
+# CARROT - CARe ROuting Tool
+# class_cpo_df.py
+# class CPO_DF(): CodepointOpen Dataframe class
+# This class opens all csv files in codepo_gb folder, and goes through all files to find postcodes for the given CC/DC codes.
+# These dataframes are put into a dictionary, which are then put into a list and merged together into a single dataframe, df.
+# As codepoint open only contains eastings and northings, geopandas is used to convert these to longitudes and latitudes, which are added as columns to df.
+# This class is used to obtain the longitudes and latitudes of any postcode in the given CC/DC areas, directly from codepoint open rather than using postcodelookup and fullcodepoint files.
+# Note: codepoint open does not contain the number of addresses per postcode (i.e. 'RP' column).
+# Used in convert_dict_inst.pu and compre_abicare_osrm.py.
+# 16/12/2020
+#--------------------#
+
+import os
+import glob
+import numpy as np
+import pandas as pd
+import geopandas
+
+class CPO_DF(): # Codepoint Open DataFrame
+    def __init__(self, foldername):
+    # def __init__(self, foldername=r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\codepo_gb'):
+    # def __init__(self, foldername=r'data\codepo_gb'):
+    # def __init__(self, foldername=r'..\data\codepo_gb'):
+        self.foldername = foldername # CP_Open_Folder = r'..\data\codepo_gb'
+        all_cpopen_csvs = glob.glob(os.path.join(self.foldername, 'data\\CSV', '*.csv'))
+
+        # Get the headers in a list:
+        with open(os.path.join(self.foldername, 'Doc', 'Code-Point_Open_Column_Headers.csv')) as f:
+            raw_headers = f.readline()
+            cp_open_headers = raw_headers.split(',') # cp_open_headers:  ['PC', 'PQ', 'EA', 'NO', 'CY', 'RH', 'LH', 'CC', 'DC', 'WC\n'], it's a list of strings.
+        
+        # We are interested in the areas 'Wiltshire', 'Hampshire', 'Berkshire/Buckinghamshire', and 'Monmouthshire', which we search with the codes below (CC county code, DC district code).
+        # Wiltshire: DC E06000054, Hampshire: CC E10000014, Berkshire/Buckinghamshire: DC E06000060, Monmouthshire: DC W06000019, W06000020 and W06000021, W06000022
+        utas = [['CC', 'E10000030'], 
[…]

### 2021-12-18 code/CarrotAll/tools_and_scripts/get_data_dfs.py

@@ -1,0 +1,571 @@
+#--------------------#
+# CARROT - CARe ROuting Tool
+# get_data_dfs.py
+# Retrieve information about carers and clients from abicare's new dataset format (client_carers_details.xlsx)
+# 25/06/2021
+#--------------------#
+
+import math
+import pickle
+import datetime
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+import openpyxl
+import tools_and_scripts.class_cpo_df as ccd
+import tools_and_scripts.week_counter as wc
+
+# filename = 'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\abicare\client_carer_details.xlsx' #File from Abicare
+
+def time_dif_to_minutes(endtime, starttime, round=True):
+    mins = (endtime - starttime).seconds/60
+    if round:
+        return np.round(mins, 0)
+    else:
+        return mins
+
+def get_info_create_dfs(area = 'None', tw_interval = 15, planning_date=None, dayindex_2weeks=None, dayindex_8weeks=None, print_statements=False, filename='None', foldername='None'):
+
+    print('Collecting data from input file.\n')
+
+    if planning_date == None or dayindex_2weeks == None or dayindex_8weeks == None:
+        print('[ERROR]: invalid timestamp for dayindex_2weeks and dayindex_8weeks.')
+        print('Terminating program.')
+        exit(-1)
+
+    planning_day = planning_date.dayofweek
+
+    # NOTE: NEED TO ADD NUMBER OF CARERS (DOUBLE SERVICE) JOBS 
+    cpo_inst = ccd.CPO_DF(foldername)
+
+    # data_filename = r'data\abicare\clientcarerdetails.xlsx'
+    # data_filename = r'data\abicare\client_carer_details.xlsx'
+    data_filename = filename
+
+    ######################## -------------------------------- CLIENT -------------------------------- ########################
+
+    clientdetails_sheetname = 'Client Contracts Data'
+
+    # First sheet: Clients Details
+    clientdf_xl = pd.read_excel(data_filename, sheet_name=clientdetails_sheetname, header=10)
+
+    
+
+    clientdf_xl.rename(columns={'CAREROOMSROOM': 'Client ID',
+                                    'CAREROOMSSENSOR': 'Area', 
+                                    'CAREROOMSContact1Postcode': 'Postcode', 
+                                    'HomeVisitsStartDate': 'Start Date', 
[…]

### 2021-12-18 code/CarrotAll/tools_and_scripts/tkinter_userinput.py

@@ -1,0 +1,338 @@
+#--------------------#
+# CARROT - CARe ROuting Tool
+# tkinter_userinput.py
+# Tkinter pop-up for users to be able to select parameter values
+# 21/06/2021
+#--------------------#
+
+import os
+from sys import exit
+from tkinter import *
+from tkinter import filedialog
+from tkcalendar import Calendar
+
+class UserInputBox:
+    def __init__(self, top):
+        self.areaName = 'Hampshire'
+        self.twInterval = 15
+        self.wbBalance = 1
+        self.qualityMeasure = 'default'
+        self.maxTimeSeconds = 60
+        self.createHtmlWebsite = True
+        self.createPythonPlots = True
+        self.codepointDir = ''
+        self.inputFilename = ''
+        self.dateSelected = 0
+        self.inputBox = Frame(top, width=510, height=370)
+        self.inputBox.pack()
+        self.t_areaName = StringVar()
+        self.t_twInterval = IntVar()
+        self.t_wbBalance = DoubleVar()
+        self.t_maxTimeSeconds = IntVar()
+        self.t_qualityMeasure = StringVar()
+        self.t_createHtmlWebsite = BooleanVar()
+        self.t_createPythonPlots = BooleanVar()
+        self.t_codepointDir = StringVar()
+        self.t_inputFilename = StringVar()
+        self.t_dateSelected = 0
+
+
+        title_label = Label(self.inputBox, text='User Input Parameters')
+        title_label.configure(font=('Helvetica', 12))
+        title_label.place(x=25, y=15)
+
+        area_label = Label(self.inputBox, text='Area name')
+        area_label.configure(font=('Helvetica', 10))
+        area_label.place(x=25, y=45)
+        area_text = Entry(self.inputBox, textvariable=self.t_areaName, bg="white")
+        area_text.place(x=180, y=47, width=100, height=17)
+
+        tw_label = Label(self.inputBox, text='Time window interval')
+        tw_label.configure(font=('Helvetica', 10))
+        tw_label.place(x=25, y=70)
+        tw_text = Entry(self.inputBox, textvariable=self.t_twInterval, bg="white")
+        tw_text.place(x=180, y=72, width=100, height=17)
+
+        wb_label = Label(self.inputBox, text='Workload balance')
+        wb_label.configure(font=('Helvetica', 10))
+        wb_label.place(x=25, y=95)
+        wb_text = Entry(self.inputBox, textvariable=self.t_wbBalance, bg="white")
[…]

### 2021-12-18 code/CarrotTkinter/constructive_wrapper.py

@@ -1,0 +1,100 @@
+#-----------------------#
+# CARROT - CARe ROuting Tool
+# constructive_wrapper.py
+# main file to run the program
+# 01/11/2020
+#-----------------------#
+
+import os
+import time
+import numpy as np
+import pandas as pd
+from datetime import datetime
+from tkinter import *
+
+# Own modules:
+import instance_handler_dfs as ihd
+import tools_and_scripts.get_data_dfs as gdd
+import tools_and_scripts.tkinter_userinput as tui
+import tools_and_scripts.week_counter as wc
+
+def main():
+    
+    # Get user input variables using tkinter entry box:
+    area, max_time_seconds, tw_interval, wb_coeff, maxtravel_coeff, maxwaiting_coeff, quality_measure, create_html_website, create_python_plots, codepoint_directory, input_filename, date_selected = tui.get_user_input_variables()
+    # area = 'Hampshire'
+    # tw_interval = 15
+    # wb_balance = 1
+    # quality_measure = 'default' # default = paper
+    # max_time_seconds = 60
+    # create_html_website = False
+    # create_python_plots = False
+    # input_filename = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/Carrot/data/abicare/2021.07.15 - UOS - Clients - Carers - Rota - Anonymised.xlsx'
+    # codepoint_directory = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/Carrot/data/codepo_gb'
+    # date_selected = pd.Timestamp("2021-07-27")
+
+    root = Tk()
+    # root.title('User Input Parameters')
+    # inputBox = Frame(root, width=510, height=420)
+    # inputBox.pack()
+
+    # ROOT = Tk()
+    LABEL = Label(root, text="Convert")
+    LABEL.pack()
+    LOOP_ACTIVE = True
+    while LOOP_ACTIVE:
+        root.update()
+        
+        LOOP_ACTIVE = False
+
+    # Convert date_selected from datetime_date to Timestamp
+    planning_date = pd.Timestamp(date_selected)
+    # planning_date = date_selected
+    
+    # Obtain day of week index of planning_date for both 2-week and 8-week schedules
+    week_period2 = 2
+    week_period8 = 8
+    dayindex_2weeks = wc.calculate_cycle_day(planning_date, week_period2)
+    dayindex_8weeks = wc.calculate_cycle_day(planning_date, week_period8)
+
[…]

### 2021-12-18 code/CarrotTkinter/literature_conversions.py

@@ -1,0 +1,170 @@
+#-----------------------#
+# CARROT - CARe ROuting Tool
+# literature_conversions.py
+# file containing functions to convert literature instances (ait_h, mankowska etc) to instance for the program.
+# 09/11/2021
+#-----------------------#
+
+import numpy as np
+import imp
+import copy
+
+def convert_aith_instance(filename, inst):
+
+    # Load Ait Hadadene file
+    ldi = imp.load_source('ait_h', filename)
+
+    inst.nJobs = ldi.NbClients
+    inst.nNurses = ldi.NbVehicules
+    inst.nSkills = ldi.NbServices
+
+    inst.nurseWorkingTimes = np.zeros((inst.nNurses, 3), dtype=np.int)
+    for i in range(inst.nNurses):
+        inst.nurseWorkingTimes[i][0] = ldi.WindowsD[i][0]
+        inst.nurseWorkingTimes[i][1] = ldi.WindowsD[i][1] - ldi.WindowsD[i][0]
+        inst.nurseWorkingTimes[i][2] = ldi.WindowsD[i][1]
+
+    inst.nurseSkills = np.ascontiguousarray(ldi.Offres, dtype=np.int)
+    inst.jobTimeInfo = np.zeros((inst.nJobs, 3), dtype=np.int)
+    inst.jobSkillsRequired = np.ascontiguousarray(ldi.Demandes, dtype=np.int)
+
+    # print(ldi.WindowsC)
+    # print(len(ldi.WindowsC))
+    # print(inst.nJobs)
+    for job in range(inst.nJobs):
+        inst.jobTimeInfo[job][0] = ldi.WindowsC[job][0]
+        inst.jobTimeInfo[job][1] = ldi.WindowsC[job][1]
+        for sk in range(ldi.NbServices):
+            if (inst.jobSkillsRequired[job][sk] > 0):
+                inst.jobTimeInfo[job][2] = ldi.DureeDeVisite[job][sk]
+
+    inst.doubleService = np.zeros(inst.nJobs, dtype=np.int)
+    inst.mk_mind = -1*np.ones(inst.nJobs)
+    inst.mk_maxd = -1*np.ones(inst.nJobs)
+
+    for job,gapVal in enumerate(ldi.gap):
+        if gapVal == 0:
+            inst.doubleService[job] = 1
+        elif gapVal > 0:
+            inst.mk_mind[job] = gapVal
+            inst.mk_maxd[job] = gapVal
+
+    # print(inst.mk_mind)
+    # To be filled later!
[…]

### 2021-12-18 code/CarrotTkinter/src/GRASP_VNS/constructive.c

@@ -1,0 +1,4125 @@
+/*--------------/
+GRASP_VNS
+constructive.c
+UoS
+10/07/2021
+/--------------*/
+
+
+#include <stdlib.h> // To have min in linux
+#include "Python.h"
+#include "constructive.h"
+#include "grasp.h"
+
+// #ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+// #    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+// #else
+// #  define MODULE_API
+// #endif
+
+int main(int argc, char const* argv[]){
+    struct INSTANCE inst = GenerateInstance();
+    struct INSTANCE* ip = &inst;
+    ip->verbose = 4;
+    return MainWithOutput(ip, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL, NULL);
+}
+
+MODULE_API int python_entry(int nJobs_data, int nNurses_data, int nSkills_data, int verbose_data, float MAX_TIME_SECONDS, int twInterval_data, bool excludeNurseTravel_data,
+                            double* od_data, double* nurseTravelFromDepot_data, double* nurseTravelToDepot_data, int* unavailMatrix_data, int* nurseUnavail_data,
+                            int* nurseWorkingTimes_data, int* jobTimeInfo_data, int* jobRequirements_data, int* nurseSkills_data, int* solMatrixPointer,
+                            int* doubleService_data, int* dependsOn_data, int* mkMinD_data, int* mkMaxD_data, int* capabilityOfDoubleServices_data, double* prefScore_data,
+                            double* algorithmOptions_data, double* timeMatrixPointer, double* nurseWaitingTimePointer, double* nurseTravelTimePointer, double* violatedTWPointer,
+                            double* nurseWaitingMatrixPointer, double* nurseTravelMatrixPointer, double* totalsArrayPointer, int randomSeed){
+
+    // printf("Inside C function. Random seed = %d.\n", randomSeed);
+
+    double eps = 1e-6;
+    int printInputData = (int)algorithmOptions_data[99] + eps;
+
+    int qualityMeasureModAPI = (int)(algorithmOptions_data[0] + eps); // printf("Quality measure	= %d\n\n", qualityMeasureModAPI);
[…]

### 2021-12-18 code/CarrotTkinter/src/GRASP_VNS/constructive.h

@@ -1,0 +1,194 @@
+/*--------------/
+GRASP_VNS
+constructive.h
+UoS
+10/07/2021
+/--------------*/
+
+
+#include <stdio.h>
+#include <stdbool.h> // Required to use 'bool' variable type
+#include <stdlib.h>
+#include <time.h>
+
+#ifdef _WIN32
+// #  ifdef MODULE_API_EXPORTS
+#    define MODULE_API __declspec(dllexport)
+// #  else
+// #    define MODULE_API __declspec(dllimport)
+// #  endif
+#else
+#  define MODULE_API
+#endif
+
+
+#define bigM 9999999;
+
+struct INSTANCE {
+	int nJobs; // Number of jobs
+    int nJobsIncDS; //Number of tasks in total, nJobs+nDS
+	int nNurses; // Number of carers
+	int nSkills; //Number of skills
+	int verbose; // Additional information
+	int isFeasible; // Feasibility (deprecated)
+	double objTime; // Objective time
+	double objWaiting; // Objective waiting time, NEW 07/01/2021
+	double objTravel; // Objective travel time
+	double objService; // Objective service time
+	double objTardiness; // Objective tardiness
+	double objMaxTardiness; // Objective max tardiness
+	double objMKTardiness; // Objective total MK tardiness
+	double objMKAllowedTardiness; // Objective mk_allowed_tardiness
+	double objOvertime; // Objective overtime
+	double objMaxOvertime; // Objective max overtime
+	double objMinSpareTime;
+	double objMaxSpareTime;
+	double objShortestDay; // Objective shortest day, NEW 07/01/2021
+	double objLongestDay; // Objective longest day
+	double objAitHQuality;
+	double objMKQuality;
+	double objWBQuality;
+	double objPaperQuality;
+	double objQuality;
+    double objMaxTravelTime; // NB new 16/11/2021
+    double objMaxWaitingTime; // NB new 16/11/2021
+    double objMaxDiffWorkload; // NB new 16/11/2021
+	float totalPref; // Total preference score
+	int qualityMeasure; // Temp. Allow to choose (Mk, Ait h., etc.) Ideally we want a vector with weights! This should be set ONLY in "InstanceFromPython" (read_instance.c). The value is hardcoded at the moment.
+	float MAX_TIME_SECONDS; // Maximum time limit in seconds
+	int twInterval; // Time window interval
[…]

### 2021-12-18 code/CarrotTkinter/src/GRASP_VNS/grasp.c

@@ -1,0 +1,2379 @@
+/*--------------/
+GRASP_VNS
+grasp.c
+UoS
+10/07/2021
+/--------------*/
+
+
+#include "float.h"
+#include "math.h"
+#include "constructive.h"
+#include "grasp.h"
+
+int GRASP(struct INSTANCE* ip) {
+
+	// PR01 Perform path relinking between every generated solution and a random solution from the pool.
+	// PR02 Perform path relinking between some new solutions (above a threshold) and all solutions in the pool.
+	// PR03 Perform path relinking at the end of the algorithm, between all pairs of solutions in the pool.
+	// PR04 Do not perform path relinking at all.
+	
+	clock_t start = clock();
+	clock_t end = clock();
+	double eps = 1e-6;
+	double elapsedTime = 0.0;
+	double average_gen_quality = 0.0;
+	long double sum_of_qualities = 0.0;
+	int given_quality = ip->qualityMeasure;
+	int solutions_in_pool = (int)(ip->algorithmOptions[8] + eps); // Allocate space for a pool of solutions (aO[8] = 10)
+	int rcl_strategy = (int)(ip->algorithmOptions[10] + eps); //Type of strategy for cut-off value in RCL (used in RandomisedConstructive). (aO[10] = 1 (strategy 1 C=B-delta*(B-W)))
+
+	double* pool_quality = malloc(solutions_in_pool * sizeof(double));
+	struct INSTANCE** pool = (struct INSTANCE**)malloc(solutions_in_pool * sizeof(struct INSTANCE*));
+	for (int i = 0; i < solutions_in_pool; i++) {
+		pool[i] = NULL;
+		pool_quality[i] = -DBL_MAX;
+	}
+
+	int PR_STRATEGY = (int)(ip->algorithmOptions[9] + eps); // 1, 2, 3 or 4, as above (aO[9] = 1, so PR with one random solution in pool)
+	int PR_DIRECTION_PARAMETER = (int)(ip->algorithmOptions[11] + eps); // 0 forward, 1 backward, 2 forward & backward, 3 randomly chosen (including f&b) (aO[11] = 0, so forward)
+	// int performPathRelinking = (int) ip->algorithmOptions[7] + eps;
+
+	if (ip->verbose > 5) {
+		printf("Started GRASP procedure...\n");
+	}
+
+	int doLS = 1;
+	int doLSAfter = 0;
+	int randomGraspParam = 0;
[…]

### 2021-12-18 code/CarrotTkinter/src/GRASP_VNS/read_instance.c

@@ -1,0 +1,1303 @@
+/*--------------/
+GRASP_VNS
+read_instance.c
+UoS
+10/07/2021
+/--------------*/
+
+
+#include "constructive.h"
+void OverwriteInstance(struct INSTANCE* ip, struct INSTANCE* oi) {
+	/*
+	Copies the contents of the solution of "oi" into "ip"
+	without creating any new allocations.
+	Only copies the solution, no algorithm options or data
+	*/
+
+	for (int i = 0; i < oi->nNurses; ++i){
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->solMatrix[i][j] = oi->solMatrix[i][j];
+        }
+    }
+
+	for (int i = 0; i < oi->nNurses; ++i){
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->timeMatrix[i][j] = oi->timeMatrix[i][j];
+        }
+    }
+
+	for (int i = 0; i < oi->nNurses; ++i){
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->allNurseRoutes[i][j] = oi->allNurseRoutes[i][j];
+        }
+    }
+
+    for (int i = 0; i < oi->nNurses; ++i){ // NB: NEW: 03/06/2021
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->nurseWaitingMatrix[i][j] = oi->nurseWaitingMatrix[i][j];
+        }
+    }
+
+    for (int i = 0; i < oi->nNurses; ++i){ // NB: NEW: 03/06/2021
+        for(int j = 0; j < oi->nJobs; ++j){
+            ip->nurseTravelMatrix[i][j] = oi->nurseTravelMatrix[i][j];
+        }
+    }
+
+	for (int i = 0; i < oi->nNurses; ++i) {
+		ip->nurseOrder[i] = oi->nurseOrder[i];
+		ip->nurseWaitingTime[i] = oi->nurseWaitingTime[i];
+		ip->nurseTravelTime[i] = oi->nurseTravelTime[i];
+	}
+
+	for (int i = 0; i < oi->nJobs; ++i) {
+		ip->nurseRoute[i] = oi->nurseRoute[i];
+		ip->violatedTW[i] = oi->violatedTW[i];
+		ip->violatedTWMK[i] = oi->violatedTWMK[i];
+	}
+
+} //END OF OverwriteInstance function.
+
[…]

### 2021-12-18 code/CarrotTkinter/tools_and_scripts/get_data_dfs.py

@@ -1,0 +1,543 @@
+#--------------------#
+# CARROT - CARe ROuting Tool
+# get_data_dfs.py
+# Retrieve information about carers and clients from abicare's new dataset format (client_carers_details.xlsx)
+# 25/06/2021
+#--------------------#
+
+import math
+import pickle
+import datetime
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+import openpyxl
+import tools_and_scripts.class_cpo_df as ccd
+import tools_and_scripts.week_counter as wc
+
+# filename = 'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\abicare\client_carer_details.xlsx' #File from Abicare
+
+def time_dif_to_minutes(endtime, starttime, round=True):
+    mins = (endtime - starttime).seconds/60
+    if round:
+        return np.round(mins, 0)
+    else:
+        return mins
+
+def get_info_create_dfs(area = 'None', tw_interval = 15, planning_date=None, dayindex_2weeks=None, dayindex_8weeks=None, print_statements=False, filename='None', foldername='None'):
+
+    print('Collecting data from input file.\n')
+
+    if planning_date == None or dayindex_2weeks == None or dayindex_8weeks == None:
+        print('[ERROR]: invalid timestamp for dayindex_2weeks and dayindex_8weeks.')
+        print('Terminating program.')
+        exit(-1)
+
+    planning_day = planning_date.dayofweek
+
+    # NOTE: NEED TO ADD NUMBER OF CARERS (DOUBLE SERVICE) JOBS 
+    cpo_inst = ccd.CPO_DF(foldername)
+
+    # data_filename = r'data\abicare\clientcarerdetails.xlsx'
+    # data_filename = r'data\abicare\client_carer_details.xlsx'
+    data_filename = filename
+
+    ######################## -------------------------------- CLIENT -------------------------------- ########################
+
+    clientdetails_sheetname = 'Client Contracts Data'
+
+    # First sheet: Clients Details
+    clientdf_xl = pd.read_excel(data_filename, sheet_name=clientdetails_sheetname, header=10)
+
+    
+
+    clientdf_xl.rename(columns={'CAREROOMSROOM': 'Client ID',
+                                    'CAREROOMSSENSOR': 'Area', 
+                                    'CAREROOMSContact1Postcode': 'Postcode', 
+                                    'HomeVisitsStartDate': 'Start Date', 
[…]

### 2021-12-18 code/CarrotTkinter/tools_and_scripts/retrieve_info_dfs.py

@@ -1,0 +1,362 @@
+#--------------------#
+# retrieve_info_dfs.py
+# 21/04/2021
+# NOT USED
+#--------------------#
+
+import math
+import pickle
+import datetime
+import openpyxl
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+
+# Own modules:
+import tools_and_scripts.class_cpo_df as ccd
+
+def time_dif_to_minutes(endtime, starttime, round=True):
+    mins = (endtime - starttime).seconds/60
+    if round:
+        return np.round(mins, 0)
+    else:
+        return mins
+
+def carer_works_this_slot(slot): # Function finds out whether there is 'Unavailable' in the given cell 'slot' in Carer Availability
+    works = True
+    if type(slot) == str: # If there is the string 'Unavailable' in the given 'slot', then the carer does not work that time
+        works = False
+    else: # Else, the 'slot' is empty, and so the carer is available to work that time.
+        works = np.isnan(slot)
+    return works
+
+# filename = 'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\abicare\clientcarerdetails.xlsx' #File from Abicare
+
+def retrieve_dfs(area = 'None', tw_interval = 15, print_statements=True, filename='None', foldername='None'):
+
+    # cpo_inst = ccd.CPO_DF(r'C:\Users\ah4c20\Asyl\PostDoc\SOCIALCARE\code\screpo\data\codepo_gb')
+    cpo_inst = ccd.CPO_DF(foldername)
+
+    # data_filename = r'data\abicare\clientcarerdetails.xlsx'
+    data_filename = filename
+    carerdetails_sheetname = 'Carer Details'
+    careravail_sheetname = 'Carer Availability'
+
+    # Read date of the rota
+    book = openpyxl.load_workbook(data_filename)
+    sheet = book.active
+    sheet = book[careravail_sheetname]
+    date_txt = sheet.cell(row=2, column=1).value
+    date_txt = date_txt.split(' ')[1]
+    date_pd = pd.to_datetime(date_txt, format='%d/%m/%Y').date()
+
+   ### -------------------------------- Client -------------------------------- ###
+
+    clientdetails_sheetname = 'Clients Details'
+    clientcontracts_sheetname = 'Client Contracts'
+
+    # First sheet: Clients Details
[…]

### 2021-12-18 code/CarrotTkinter/tools_and_scripts/tkinter_userinput.py

@@ -1,0 +1,394 @@
+#--------------------#
+# CARROT - CARe ROuting Tool
+# tkinter_userinput.py
+# Tkinter pop-up for users to be able to select parameter values
+# 21/06/2021
+#--------------------#
+
+import os
+from sys import exit
+from tkinter import *
+from tkinter import filedialog
+from tkcalendar import Calendar
+from tkinter import ttk
+
+class UserInputBox:
+    def __init__(self, top):
+        self.areaName = 'Hampshire'
+        self.maxTimeSeconds = 60
+        self.twInterval = 15
+        self.wbCoeff = -1
+        self.maxTravelCoeff = -1
+        self.maxWaitingCoeff = -1
+        self.qualityMeasure = 'default'
+        self.createHtmlWebsite = True
+        self.createPythonPlots = True
+        self.codepointDir = ''
+        self.inputFilename = ''
+        self.dateSelected = 0
+        self.inputBox = Frame(top, width=510, height=420)
+        self.inputBox.pack()
+        self.t_areaName = StringVar()
+        self.t_maxTimeSeconds = IntVar()
+        self.t_twInterval = IntVar()
+        self.t_wbCoeff = DoubleVar()
+        self.t_maxTravelCoeff = DoubleVar() # NEW 01/12/2021
+        self.t_maxWaitingCoeff = DoubleVar() # NEW 01/12/2021
+        # self.t_qualityMeasure = StringVar()
+        self.t_qualityMeasure = IntVar()
+        self.t_createHtmlWebsite = BooleanVar()
+        self.t_createPythonPlots = BooleanVar()
+        self.t_codepointDir = StringVar()
+        self.t_inputFilename = StringVar()
+        self.t_dateSelected = 0
+
+
+        title_label = Label(self.inputBox, text='User Input Parameters')
+        title_label.configure(font=('Helvetica', 12))
+        title_label.place(x=25, y=15)
+
+        area_label = Label(self.inputBox, text='Area name')
+        area_label.configure(font=('Helvetica', 10))
+        area_label.place(x=25, y=45)
+        # area_text = Entry(self.inputBox, textvariable=self.t_areaName, bg="white")
+        area_text = ttk.Combobox(self.inputBox, textvariable=self.t_areaName)
+        area_text['values'] = ['Aldershot', 'Andover', 'Hampshire', 'Monmouth', 'N.Wiltshire', 'Salisbury', 'Swindon']
+        area_text.set('Hampshire')
+        area_text.place(x=180, y=47, width=100, height=17)
[…]

### 2021-12-18 code/CarrotTkinter/tools_and_scripts/week_counter.py

@@ -1,0 +1,67 @@
+#--------------------#
+# CARROT - CARe ROuting Tool
+# week_counter.py
+# Functions to calculate the index value for the day of the week based on 2-week and 8-week schedules
+# 30/07/2021
+#--------------------#
+
+import pandas as pd
+
+def date_in_visit_cycle(given_date, ref_date, week_cycle):
+    """
+    Calculates whether there is a visit due
+    in the week of "given_date", based on a
+    contract that started on "ref_date" and is
+    on a "week_cycle" visit cycle.
+    """
+    week_diff = calculate_week_difference(given_date, ref_date)
+    return week_diff % week_cycle == 0
+
+
+def date_to_mon_9am(given_date):
+    mon = given_date - pd.Timedelta(days=given_date.dayofweek)
+    # print('mon before replace: ', mon)
+    mon.replace(hour=9, minute=0, second=0)
+    return mon
+
+def calculate_week_difference(planning_date, ref_date):
+    # We use Monday 9 am for reference
+    monday1 = date_to_mon_9am(planning_date)
+    # print('monday1: ', monday1)
+    monday2 = date_to_mon_9am(ref_date)
+    # print('monday2: ', monday2)
+
+    days_diff = abs((monday2 - monday1).days)
+    # print('days_diff: ', days_diff)
+    weeks_diff = int(days_diff/7 + 0.001) # Tolerance to make int weeks
+    return weeks_diff
+
+
+def calculate_cycle_day(planning_date, week_period, ref_date=pd.Timestamp("2021-05-31 09:00")):
+    """
+    planning_date  - Pandas timestamp that represents the date of interest
+    week_period - integer containing the week period (2 will return day 0 to 13, 8 will return 0 to 55)
+    ref_date    - Must be a date in what is considered "Week 1"
+    """
+    # print('planning_date: ', planning_date, ' week_period: ', week_period, ' ref_date: ', ref_date)
+    weeks_diff = calculate_week_difference(planning_date, ref_date)
+    # print('weeks_diff: ', weeks_diff)
+    weeks_off_cycle = weeks_diff % week_period
+    # print('weeks_off_cycle: ', weeks_off_cycle)
+    day_of_cycle = weeks_off_cycle*7 + planning_date.dayofweek
+    # print('day_of_cycle: ', day_of_cycle)
+    return day_of_cycle
+
+
+# Example of carer availability: this actually returns the day (0-55) that the planning_date is on.
[…]

### 2021-12-18 code/ExactMethod/2)Model-Code-SingleFile.py

@@ -1,0 +1,288 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+import sys
+from collections import defaultdict
+import gurobipy as gp
+from gurobipy import GRB
+import numpy as np
+import xlrd
+import xlsxwriter
+import matplotlib.pyplot as plt
+
+import read_helper_functions as rhf
+
+# Determine the file to read:
+file_to_read = '45-10-s-2b'
+# inst = rhf.read_ait_h_instance(file_to_read, split=False) # Normal instance
+inst = rhf.read_ait_h_instance(file_to_read, split=True) # Split instance into skills.
+# exit(-1)
+
+# Now you can access this data fields, using a dot:
+# CoutPref - Matrix of the preference scores between patients/carers
+# Demandes - Matrix with skills required by the client
+# DureeDeVisite - Service time per skill -> DureeDeVisite[job][skill]
+# NbClients - Number of patients/clients
+# NbServices - Number of different skills
+# NbVehicles - Number of nurses available
+# Offres - For each nurse, what skills they have. Binary: Offres[nurse][skill]
+# WindowsC - For each job, WindowsC[job][0] is the start of the time window, WindowsC[job][1] the finishing time
+# WindowsD - WindowsD[nurse][0] time the nurse starts working, WindowsD[nurse][1] time the nurse finishes
+# od - od[i,j] as the distance between two patients. row/col 0 represent the nurses location (same for all of them)
+# gap - Gap between two services (for those that require two skills only!) if the gap is 0, it means they are simultaneous
+
+#print('Shape of the OD matrix:')
+#print(inst.od.shape)
+#print('Shape of the preference matrix:')
+#print(inst.CoutPref.shape)
+#print(inst.depl)
+#print(inst.CoutPref)
+
+
+# In[4]:
+
+
+class Customer():
+    def __init__(self,CID,CPref,SDemand,TWinS,TWinE,SGap):
+        self.CID = CID
+        self.CPref = CPref
+        self.SDemand = SDemand
+        self.TWinS = TWinS
+        self.TWinE = TWinE
+        self.SGap = SGap
+    def __str__(self):
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/18-4-s-2a.py

@@ -1,0 +1,49 @@
+NbVehicules=4
+NbClients=18
+NbServices=2
+Demandes=[[1,1],[0,1],[1,1],[1,0],[1,0],[0,1],[0,1],[0,1],[1,0],[1,0],[1,0],[1,0],[0,1],[1,0],[1,0],[0,1],[0,1],[1,0]]
+Offres=[[1,0],[0,1],[1,0],[0,1]]
+depl=[
+23,3,42,43,23,25,4,33,29,29,38,3,39,37,40,24,14,
+23,23,35,36,25,35,20,23,22,17,35,23,31,24,19,7,25,
+3,23,40,41,22,23,4,32,28,28,36,2,38,36,40,24,12,
+42,35,40,2,20,27,39,13,14,18,7,39,4,14,29,28,31,
+43,36,41,2,20,27,40,14,15,19,7,40,6,15,30,29,32,
+23,25,22,20,20,12,21,15,10,15,15,21,18,20,31,20,12,
+25,35,23,27,27,12,24,26,22,27,21,23,27,31,43,31,13,
+4,20,4,39,40,21,24,30,26,26,35,3,36,34,36,21,12,
+33,23,32,13,14,15,26,30,5,6,14,31,9,6,19,16,25,
+29,22,28,14,15,10,22,26,5,7,13,27,11,10,22,15,20,
+29,17,28,18,19,15,27,26,6,7,19,27,14,8,16,10,23,
+38,35,36,7,7,15,21,35,14,13,19,35,8,17,32,28,26,
+3,23,2,39,40,21,23,3,31,27,27,35,37,35,39,23,11,
+39,31,38,4,6,18,27,36,9,11,14,8,37,10,25,24,29,
+37,24,36,14,15,20,31,34,6,10,8,17,35,10,16,17,30,
+40,19,40,29,30,31,43,36,19,22,16,32,39,25,16,16,37,
+24,7,24,28,29,20,31,21,16,15,10,28,23,24,17,16,23,
+14,25,12,31,32,12,13,12,25,20,23,26,11,29,30,37,23,
+21,18,20,22,23,7,18,18,13,9,11,19,19,19,17,26,13,13,
+21,18,20,22,23,7,18,18,13,9,11,19,19,19,17,26,13,13]
+WindowsC=[[170,288],[284,378],[339,497],[447,560],[536,684],[0,29],[38,136],[358,549],[566,667],[0,88],[0,107],[97,206],[165,333],[355,419],[0,72],[4,172],[270,400],[413,492]]
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/18-4-s-2b.py

@@ -1,0 +1,49 @@
+NbVehicules=4
+NbClients=18
+NbServices=2
+Demandes=[[1,0],[1,0],[1,1],[1,1],[0,1],[0,1],[0,1],[1,0],[0,1],[0,1],[0,1],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1]]
+Offres=[[1,0],[1,0],[0,1],[0,1]]
+depl=[
+23,30,4,12,17,14,38,11,16,9,31,37,10,23,20,15,11,
+23,24,25,27,19,31,22,18,28,24,11,26,20,32,9,15,13,
+30,24,33,23,38,26,16,33,20,23,18,9,20,18,31,15,28,
+4,25,33,15,17,16,42,11,19,12,34,41,14,26,22,19,13,
+12,27,23,15,28,5,35,22,5,4,31,31,7,11,28,13,19,
+17,19,38,17,28,31,41,7,32,25,30,43,23,38,12,24,11,
+14,31,26,16,5,31,39,24,7,7,35,34,11,12,32,17,23,
+38,22,16,42,35,41,39,38,33,35,11,10,30,33,30,24,32,
+11,18,33,11,22,7,24,38,25,18,28,39,17,31,12,19,7,
+16,28,20,19,5,32,7,33,25,8,30,28,9,7,30,13,22,
+9,24,23,12,4,25,7,35,18,8,29,31,5,14,25,12,16,
+31,11,18,34,31,30,35,11,28,30,29,18,25,32,20,18,23,
+37,26,9,41,31,43,34,10,39,28,31,18,28,26,34,22,33,
+10,20,20,14,7,23,11,30,17,9,5,25,28,15,21,7,13,
+23,32,18,26,11,38,12,33,31,7,14,32,26,15,35,17,28,
+20,9,31,22,28,12,32,30,12,30,25,20,34,21,35,19,10,
+15,15,15,19,13,24,17,24,19,13,12,18,22,7,17,19,13,
+11,13,28,13,19,11,23,32,7,22,16,23,33,13,28,10,13,
+15,12,18,18,16,21,20,24,16,16,13,17,24,9,21,15,4,10,
+15,12,18,18,16,21,20,24,16,16,13,17,24,9,21,15,4,10]
+WindowsC=[[0,99],[56,119],[305,387],[393,460],[485,616],[0,105],[135,264],[179,299],[297,410],[385,516],[0,75],[2,144],[493,620],[654,715],[0,102],[86,133],[176,247],[399,479]]
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/18-4-s-2c.py

@@ -1,0 +1,49 @@
+NbVehicules=4
+NbClients=18
+NbServices=2
+Demandes=[[1,1],[0,1],[1,1],[1,0],[1,0],[0,1],[0,1],[0,1],[1,0],[1,0],[1,0],[1,0],[0,1],[1,0],[1,0],[0,1],[0,1],[1,0]]
+Offres=[[1,0],[0,1],[1,0],[0,1]]
+depl=[
+23,3,42,43,23,25,4,33,29,29,38,3,39,37,40,24,14,
+23,23,35,36,25,35,20,23,22,17,35,23,31,24,19,7,25,
+3,23,40,41,22,23,4,32,28,28,36,2,38,36,40,24,12,
+42,35,40,2,20,27,39,13,14,18,7,39,4,14,29,28,31,
+43,36,41,2,20,27,40,14,15,19,7,40,6,15,30,29,32,
+23,25,22,20,20,12,21,15,10,15,15,21,18,20,31,20,12,
+25,35,23,27,27,12,24,26,22,27,21,23,27,31,43,31,13,
+4,20,4,39,40,21,24,30,26,26,35,3,36,34,36,21,12,
+33,23,32,13,14,15,26,30,5,6,14,31,9,6,19,16,25,
+29,22,28,14,15,10,22,26,5,7,13,27,11,10,22,15,20,
+29,17,28,18,19,15,27,26,6,7,19,27,14,8,16,10,23,
+38,35,36,7,7,15,21,35,14,13,19,35,8,17,32,28,26,
+3,23,2,39,40,21,23,3,31,27,27,35,37,35,39,23,11,
+39,31,38,4,6,18,27,36,9,11,14,8,37,10,25,24,29,
+37,24,36,14,15,20,31,34,6,10,8,17,35,10,16,17,30,
+40,19,40,29,30,31,43,36,19,22,16,32,39,25,16,16,37,
+24,7,24,28,29,20,31,21,16,15,10,28,23,24,17,16,23,
+14,25,12,31,32,12,13,12,25,20,23,26,11,29,30,37,23,
+21,18,20,22,23,7,18,18,13,9,11,19,19,19,17,26,13,13,
+21,18,20,22,23,7,18,18,13,9,11,19,19,19,17,26,13,13]
+WindowsC=[[45,288],[284,439],[333,497],[373,560],[484,684],[0,88],[38,156],[358,555],[496,667],[0,139],[0,140],[58,231],[165,333],[355,459],[0,132],[4,179],[270,459],[413,575]]
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/18-4-s-2d.py

@@ -1,0 +1,49 @@
+NbVehicules=4
+NbClients=18
+NbServices=2
+Demandes=[[1,0],[1,0],[1,1],[1,1],[0,1],[0,1],[0,1],[1,0],[0,1],[0,1],[0,1],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1]]
+Offres=[[1,0],[1,0],[0,1],[0,1]]
+depl=[
+23,30,4,12,17,14,38,11,16,9,31,37,10,23,20,15,11,
+23,24,25,27,19,31,22,18,28,24,11,26,20,32,9,15,13,
+30,24,33,23,38,26,16,33,20,23,18,9,20,18,31,15,28,
+4,25,33,15,17,16,42,11,19,12,34,41,14,26,22,19,13,
+12,27,23,15,28,5,35,22,5,4,31,31,7,11,28,13,19,
+17,19,38,17,28,31,41,7,32,25,30,43,23,38,12,24,11,
+14,31,26,16,5,31,39,24,7,7,35,34,11,12,32,17,23,
+38,22,16,42,35,41,39,38,33,35,11,10,30,33,30,24,32,
+11,18,33,11,22,7,24,38,25,18,28,39,17,31,12,19,7,
+16,28,20,19,5,32,7,33,25,8,30,28,9,7,30,13,22,
+9,24,23,12,4,25,7,35,18,8,29,31,5,14,25,12,16,
+31,11,18,34,31,30,35,11,28,30,29,18,25,32,20,18,23,
+37,26,9,41,31,43,34,10,39,28,31,18,28,26,34,22,33,
+10,20,20,14,7,23,11,30,17,9,5,25,28,15,21,7,13,
+23,32,18,26,11,38,12,33,31,7,14,32,26,15,35,17,28,
+20,9,31,22,28,12,32,30,12,30,25,20,34,21,35,19,10,
+15,15,15,19,13,24,17,24,19,13,12,18,22,7,17,19,13,
+11,13,28,13,19,11,23,32,7,22,16,23,33,13,28,10,13,
+15,12,18,18,16,21,20,24,16,16,13,17,24,9,21,15,4,10,
+15,12,18,18,16,21,20,24,16,16,13,17,24,9,21,15,4,10]
+WindowsC=[[0,126],[11,139],[242,408],[303,533],[397,645],[0,125],[40,264],[136,299],[268,438],[365,548],[0,105],[2,144],[435,640],[540,726],[0,135],[26,245],[95,289],[272,495]]
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/45-10-m-3.py

@@ -1,0 +1,103 @@
+NbVehicules=10
+NbClients=45
+NbServices=2
+Demandes=[[1,1],[0,1],[0,1],[1,1],[0,1],[1,1],[1,0],[0,1],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[0,1],[1,0],[1,0],[1,0],[1,0],[0,1],[1,0],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[0,1],[1,0],[0,1],[0,1],[1,0],[1,0],[1,0]]
+Offres=[[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1]]
+depl=[
+24,30,31,17,9,18,42,17,14,11,24,25,13,32,37,7,6,28,18,23,25,14,22,9,22,17,7,12,32,36,27,23,31,12,19,26,11,21,4,16,12,24,32,30,
+24,20,38,10,26,15,27,23,35,20,7,31,27,41,23,21,22,8,9,39,22,23,3,32,14,38,27,17,43,23,35,34,42,26,6,22,21,41,24,9,14,35,28,43,
+30,20,23,26,25,12,13,16,33,32,14,18,23,27,9,23,25,13,24,31,8,19,22,33,9,35,35,30,30,7,23,23,30,23,21,6,33,37,32,22,26,24,9,32,
+31,38,23,38,23,24,31,17,24,39,32,7,19,5,29,26,27,33,37,15,17,19,38,29,24,25,38,39,8,27,6,9,9,19,35,18,40,24,35,35,36,9,15,12,
+17,10,26,38,21,16,35,22,30,10,14,32,23,41,31,16,16,17,2,36,25,21,8,25,18,33,18,7,42,31,35,32,41,23,6,25,11,36,15,4,5,34,32,41,
+9,26,25,23,21,15,37,10,10,18,23,17,4,24,33,6,6,27,21,16,19,7,24,9,18,13,15,18,24,32,18,15,23,4,21,20,18,16,13,19,17,16,25,22,
+18,15,12,24,16,15,24,9,24,21,10,17,14,27,19,12,14,12,15,26,9,10,15,23,4,27,24,19,29,18,21,20,28,14,12,10,21,30,20,13,15,21,17,30,
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/45-10-s-3c.py

@@ -1,0 +1,103 @@
+NbVehicules=10
+NbClients=45
+NbServices=2
+Demandes=[[1,1],[0,1],[0,1],[1,1],[0,1],[1,1],[1,0],[0,1],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[0,1],[1,0],[1,0],[1,0],[1,0],[0,1],[1,0],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[0,1],[1,0],[0,1],[0,1],[1,0],[1,0],[1,0]]
+Offres=[[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1]]
+depl=[
+24,30,31,17,9,18,42,17,14,11,24,25,13,32,37,7,6,28,18,23,25,14,22,9,22,17,7,12,32,36,27,23,31,12,19,26,11,21,4,16,12,24,32,30,
+24,20,38,10,26,15,27,23,35,20,7,31,27,41,23,21,22,8,9,39,22,23,3,32,14,38,27,17,43,23,35,34,42,26,6,22,21,41,24,9,14,35,28,43,
+30,20,23,26,25,12,13,16,33,32,14,18,23,27,9,23,25,13,24,31,8,19,22,33,9,35,35,30,30,7,23,23,30,23,21,6,33,37,32,22,26,24,9,32,
+31,38,23,38,23,24,31,17,24,39,32,7,19,5,29,26,27,33,37,15,17,19,38,29,24,25,38,39,8,27,6,9,9,19,35,18,40,24,35,35,36,9,15,12,
+17,10,26,38,21,16,35,22,30,10,14,32,23,41,31,16,16,17,2,36,25,21,8,25,18,33,18,7,42,31,35,32,41,23,6,25,11,36,15,4,5,34,32,41,
+9,26,25,23,21,15,37,10,10,18,23,17,4,24,33,6,6,27,21,16,19,7,24,9,18,13,15,18,24,32,18,15,23,4,21,20,18,16,13,19,17,16,25,22,
+18,15,12,24,16,15,24,9,24,21,10,17,14,27,19,12,14,12,15,26,9,10,15,23,4,27,24,19,29,18,21,20,28,14,12,10,21,30,20,13,15,21,17,30,
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/73-16-l-5.py

@@ -1,0 +1,159 @@
+NbVehicules=16
+NbClients=73
+NbServices=2
+Demandes=[[1,1],[1,0],[0,1],[1,1],[1,0],[0,1],[1,0],[0,1],[1,1],[0,1],[0,1],[1,0],[1,1],[1,0],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[1,1],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1]]
+Offres=[[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1]]
+depl=[
+11,42,28,43,12,10,42,8,21,31,31,41,29,47,32,14,47,9,45,24,41,39,35,9,19,46,26,21,34,13,25,32,26,8,42,33,21,33,38,10,12,43,29,44,13,11,43,9,22,32,32,42,30,48,33,15,48,10,46,25,42,40,36,10,20,47,27,22,35,14,26,33,
+11,32,26,34,5,14,33,4,19,22,21,30,24,38,30,7,37,11,35,24,30,28,29,11,9,37,21,12,25,15,23,22,25,13,34,25,11,22,27,12,10,33,27,35,6,15,34,5,20,23,22,31,25,39,31,8,38,12,36,25,31,29,30,12,10,38,22,13,26,16,24,23,
+42,32,35,11,31,39,13,35,34,16,25,11,26,14,36,33,11,41,6,38,18,11,27,37,27,15,27,30,17,38,35,11,37,40,18,15,24,16,11,43,33,10,36,12,32,40,14,36,35,17,26,12,27,15,37,34,12,42,7,39,19,12,28,38,28,16,28,31,18,39,36,12,
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/73-16-s-3.py

@@ -1,0 +1,159 @@
+NbVehicules=16
+NbClients=73
+NbServices=2
+Demandes=[[1,1],[1,0],[0,1],[1,1],[1,0],[0,1],[1,0],[0,1],[1,1],[0,1],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[1,0],[1,0],[1,0],[0,1],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1]]
+Offres=[[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1],[1,0],[0,1]]
+depl=[
+11,42,28,43,12,10,42,8,21,31,31,41,29,47,32,14,47,9,45,24,41,39,35,9,19,46,26,21,34,13,25,32,26,8,42,33,21,33,38,10,12,43,29,44,13,11,43,9,22,32,32,42,30,48,33,15,48,10,46,25,42,40,36,10,20,47,27,22,35,14,26,33,
+11,32,26,34,5,14,33,4,19,22,21,30,24,38,30,7,37,11,35,24,30,28,29,11,9,37,21,12,25,15,23,22,25,13,34,25,11,22,27,12,10,33,27,35,6,15,34,5,20,23,22,31,25,39,31,8,38,12,36,25,31,29,30,12,10,38,22,13,26,16,24,23,
+42,32,35,11,31,39,13,35,34,16,25,11,26,14,36,33,11,41,6,38,18,11,27,37,27,15,27,30,17,38,35,11,37,40,18,15,24,16,11,43,33,10,36,12,32,40,14,36,35,17,26,12,27,15,37,34,12,42,7,39,19,12,28,38,28,16,28,31,18,39,36,12,
[…]

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/allf.txt.py

@@ -1,0 +1,1 @@
+['18-4-l-2a', '18-4-l-2b', '18-4-l-2c', '18-4-l-2d', '18-4-l-2e', '18-4-m-2a', '18-4-m-2b', '18-4-m-2c', '18-4-m-2d', '18-4-m-2e', '18-4-s-2a', '18-4-s-2b', '18-4-s-2c', '18-4-s-2d', '45-10-l-2a', '45-10-l-2b', '45-10-l-3', '45-10-l-4', '45-10-m-2a', '45-10-m-2b', '45-10-m-3', '45-10-m-4', '45-10-s-2a', '45-10-s-2b', '45-10-s-3a', '45-10-s-3ac', '45-10-s-3b', '45-10-s-3c', '73-16-l-2', '73-16-l-3', '73-16-l-4', '73-16-l-5', '73-16-m-2', '73-16-m-3a', '73-16-m-3b', '73-16-m-3bc', '73-16-s-2a', '73-16-s-2b', '73-16-s-3', '18-4-l-2ac', '18-4-s-2ac', '45-10-l-2ac']

### 2021-12-18 code/ExactMethod/ait _haddadene_et_al_2016/renameall.sh.py

@@ -1,0 +1,4 @@
+for ff in $(ls)
+do
+	cp $ff $ff".py"
+done

### 2021-12-18 code/ExactMethod/coordnetworkx.py

@@ -1,0 +1,21 @@
+import networkx as nx
+import numpy as np 
+import matplotlib.pyplot as plt
+
+fig, ax = plt.subplots()
+
+pos = {0: (4, 2), 1: (2, 3), 2: (1, 3), 3: (3, 1)} 
+X=nx.Graph()
+
+X.add_nodes_from(pos.keys())
+for n, p in iter(pos.items()):
+    X.nodes[n]['pos'] = p
+nx.draw(X, pos)
+
+ax.set_xlim(0,5)
+ax.set_ylim(0,5)
+ax.tick_params(left=True, bottom=True, labelleft=True, labelbottom=True)
+plt.axis('on')
+plt.show()
+
+# nx.draw_networkx_nodes(X,pos,node_size=3000,nodelist=[0,1,2,3],node_color='r')

### 2021-12-18 code/ExactMethod/coords_from_od.py

@@ -1,0 +1,72 @@
+import numpy as np 
+import matplotlib.pyplot as plt
+import read_helper_functions as rhf
+
+def get_od(points):
+    npts = np.max(points.shape)
+    od = np.zeros((npts,npts))
+    for i in range(npts):
+        for j in range(npts):
+            od[i,j] = np.linalg.norm(points[i,:] - points[j,:])
+    
+    return od
+### --- End def get_od --- ###
+
+def get_M(D): 
+    # Create matrix M using values in D (od matrix)
+    n,m = D.shape # n = number of rows in od matrix, m = number of columns in od matrix
+    M = np.zeros((n,m)) # Create empty matrix M of size n x m
+
+    # Fill M matrix using od matrix, M matrix is a Gram Matrix:
+    for i in range(n): 
+        for j in range(m):
+            M[i,j] = (D[0,j]**2 + D[i,0]**2 - D[i,j]**2)/2
+
+    return M
+### --- End def get_M --- ###
+
+def coords_from_M(M):
+    # Eigenvalue decomposition
+    lambd, U = np.linalg.eig(M) # lambd = eigenvalues, U = eigenvector
+    S = np.round(np.diag(lambd), 8) # np.diag(lambd) = create diagonal matrix with eignvalues, round eignvalues to 8dp
+    # S = np.diag(lambd)
+    X = U.dot(np.sqrt(S)) # if M = USU^T (because od is a real symmetric matrix), then the matrix X = U.sqrt(S) contains the position of the points (each row corresponding to one point)
+    return X
+### --- End def coords_from_M --- ###
+
+def coords_from_od(od):
+    M = get_M(od) # Create Gram matrix M
+    return coords_from_M(M) # Get matrix X of position of points (each row corresponds to one point)
+### --- End def coords_from_od --- ###
+
+file_to_read = '18-4-l-2a'
+inst = rhf.read_ait_h_instance(file_to_read)
+nbc = inst.NbClients
+
+# T = np.zeros((nbc+2, nbc+2))
+# T[:nbc+1, :nbc+1] = inst.od
+
+# Complete last column and last row:
+# T[nbc+1, :] = T[0, :] # Last row equal to first row
+# T[:, nbc+1] = T[:, 0] # Last column equal to first column
+
+T = np.zeros((3,3))
+T[0,1] = 3
+T[1,0] = 3
[…]

### 2021-12-18 code/ExactMethod/main.py

@@ -1,0 +1,307 @@
+import sys
+from collections import defaultdict
+import gurobipy as gp
+from gurobipy import GRB
+import numpy as np
+import xlrd
+import xlsxwriter
+import read_helper_functions as rhf
+import networkx_graph as nxg
+import testnetworkx as testnx
+
+# filenames=['18-4-s-2a','18-4-s-2b','18-4-s-2c','18-4-s-2d','18-4-m-2a','18-4-m-2b','18-4-m-2c','18-4-m-2d','18-4-m-2e','18-4-l-2a','18-4-l-2b','18-4-l-2c','18-4-l-2d','18-4-l-2e']
+filenames=['18-4-s-2a']
+# plot_graph = True
+plot_graph = False
+# workbook = xlsxwriter.Workbook('C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/ExactMethod/Values.xlsx')
+# worksheet = workbook.add_worksheet()
+# worksheet.write(0, 0, 'Filename' )
+# worksheet.write(0, 1, 'Objective value' )
+for fl in range(len(filenames)):
+    # Determine the file to read:
+    file_to_read = filenames[fl]
+    inst = rhf.read_ait_h_instance(file_to_read, split=False) # Normal instance
+    # inst = rhf.read_ait_h_instance(file_to_read, split=True) # the instance file is a split file
+    # inst = rhf.split_instance(file_to_read, print_statements=True)
+    # exit(-1)
+    
+    # NbVehicles - Number of nurses available
+    # NbClients - Number of patients/clients
+    # NbServices - Number of different skills
+    # Demandes - NbClients x NbServices, binary matrix with skills required by the client. Demandes[client][skill] = 1 if client requires skill, 0 otherwise
+    # Offres - NbVehicules x NbServices, binary matrix: for each nurse, what skills they have. Binary: Offres[nurse][skill] = 1 if nurse can perform skill, = 0 otherwise
+    # od - NbClients+1 x NbClients+1, distance between patients. row/col 0 represent the nurses location (same for all of them, single depot)
+    # WindowsC - NbClients x 2, for each job, WindowsC[job][0] is the start of the time window, WindowsC[job][1] the finishing time
+    # DureeDeVisite - NbClients x NbServices, service time per skill, DureeDeVisite[client][skill]
[…]

### 2021-12-18 code/ExactMethod/networkx_graph.py

@@ -1,0 +1,136 @@
+import networkx as nx
+import matplotlib.pyplot as plt
+import pandas as pd
+import os
+import sys
+
+def create_graph(file_to_read, num_clients, edge_set, nurse_index=999, is_nurse_list=False):
+
+    fig, ax = plt.subplots()
+    # plt.figure()
+
+    # Figure title
+    if is_nurse_list: # list of specific nurses, add all indices of nurses to the title
+        title_string = str(file_to_read) + '-nurses'  
+        for i in range(len(nurse_index)):
+            title_string += '-' + str(nurse_index[i])
+        plt.title(title_string)
+    else:
+        if nurse_index == 999: # all nurses
+            plt.title(str(file_to_read))
+        elif nurse_index < 999:
+            plt.title(str(file_to_read) + '-nurse' + str(nurse_index)) # one specific nurse
+
+    # Create graph
+    G = nx.MultiDiGraph()
+
+    # Add nodes to graph
+    df = pd.read_csv(r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/ExactMethod/18-4-s-2a_coords.csv') 
+
+    pos = {}
+    client_column = df['n']
+    x_column = df['x']
+    y_column = df['y']
+    min_x = 999
+    max_x = 0
+    min_y = 999
+    max_y = 0
+
+
+    for i in range(len(x_column)):
+        x_coord = df['x'][i]
+        y_coord = df['y'][i]
+        if x_coord > max_x:
+            max_x = x_coord
+        elif x_coord < min_x:
+            min_x = x_coord
+        if y_coord > max_y:
+            max_y = y_coord
+        elif y_coord < min_y:
+            min_y = y_coord
+        pos[df['n'][i]] = x_coord, y_coord
+    
+    print('min_x: ', min_x, ' max_x: ', max_x, ' min_y: ', min_y, ' max_y: ', max_y)
+    # pos = {0: (0,0), 1: (0,1), 2: (0, 2), 3: (0, 3), 4: (0,4), 5: (1,0), 6: (1,1), 7: (1,2), 8: (1,3), 9: (1,4), 10: (2,0), 11: (2,1), 12: (2,2), 13: (2,3), 14: (2,4), 15: (3,0), 16: (3,1), 17: (3,2), 18: (3,3), 19: (3,4) }
+    # G.add_nodes_from(range(0,num_clients))
[…]

### 2021-12-18 code/ExactMethod/networkxexample.py

@@ -1,0 +1,33 @@
+import pandas as pd
+import networkx as nx
+import matplotlib.pyplot as plt
+
+df = pd.DataFrame({'id_emp' : [13524791000109, 12053850000137, 4707821000113, 4707821000114, 1],
+           'name_emp': ['Cristiano', 'Gaúcho', 'Fenômeno','Angelin', 'Souza'],
+           'name_dep': ['Ronaldo','Ronaldo', 'Ronaldo', 'Ronaldo', 'Bruno'],
+           'weight_1': [8,9,10,11,12],
+           'weight_2':[5,6,7,8,9] })
+
+G = nx.MultiDiGraph()
+
+G.add_nodes_from(df['id_emp'], bipartite = 0)
+emp = [v for v in G.nodes if G.nodes[v]['bipartite'] == 0]
+G.add_nodes_from(df['name_dep'], bipartite = 1)
+dep = [v for v in G.nodes if G.nodes[v]['bipartite'] == 1]
+
+G.add_weighted_edges_from(df[['name_dep', 'id_emp', 'weight_1']].values)
+G.add_weighted_edges_from(df[['id_emp', 'name_dep', 'weight_2']].values)
+edge_width = [a[2]['weight']//2 for a in G.edges(data=True)]
+
+plt.figure(figsize=(5,5))
+
+pos = nx.spring_layout(G, k=0.9)
+
+# Here there is the addition:
+
+nx.draw_networkx_nodes(G, pos, nodelist=dep, node_color='#bfbf7f', node_shape="h", node_size=300)
+nx.draw_networkx_nodes(G, pos, nodelist=emp, node_color='red', node_size=300)
+nx.draw_networkx_edges(G, pos, connectionstyle='arc3, rad = 0.3', width=edge_width, alpha=0.2)
+
+plt.axis('off')
+plt.show()

### 2021-12-18 code/ExactMethod/plot.py

@@ -1,0 +1,46 @@
+import pandas as pd
+import numpy as np
+import matplotlib.pyplot as plt
+
+df = pd.read_csv(r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/ExactMethod/18-4-s-2a_coords.csv') 
+
+pos_dict = {}
+client_column = df['n']
+x_column = df['x']
+y_column = df['y']
+
+
+for i in range(len(x_column)):
+    x_coord = df['x'][i]
+    y_coord = df['y'][i]
+    pos_dict[df['n'][i]] = x_coord, y_coord
+
+print('pos dict: ', pos_dict)
+
+exit(-1)
+fig, ax = plt.subplots()
+my_scatterplot = ax.scatter(df['x'], df['y'])
+N = df['n']
+X = df['x']
+Y = df['y']
+ax.set_xlabel('x')
+ax.set_ylabel('y')
+for i, txt in enumerate(N):
+    ax.annotate(txt, (X[i], Y[i]))
+plt.show()
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+

### 2021-12-18 code/ExactMethod/read_data_example.py

@@ -1,0 +1,29 @@
+from read_helper_functions import read_ait_h_instance
+
+
+# Determine the file to read:
+file_to_read = '45-10-m-4'
+inst = read_ait_h_instance(file_to_read)
+
+
+# Now you can access this data fields, using a dot:
+# CoutPref - Matrix of the preference scores between patients/carers
+# Demandes - Matrix with skills required by the client
+# DureeDeVisite - Service time per skill -> DureeDeVisite[job][skill]
+# NbClients - Number of patients/clients
+# NbServices - Number of different skills
+# NbVehicles - Number of nurses available
+# Offres - For each nurse, what skills they have. Binary: Offres[nurse][skill]
+# WindowsC - For each job, WindowsC[job][0] is the start of the time window, WindowsC[job][1] the finishing time
+# WindowsD - WindowsD[nurse][0] time the nurse starts working, WindowsD[nurse][1] time the nurse finishes
+# od - od[i,j] as the distance between two patients. row/col 0 represent the nurses location (same for all of them)
+# gap - Gap between two services (for those that require two skills only!) if the gap is 0, it means they are simultaneous
+
+print('The number of nurses in the instance is', inst.NbVehicules, 'and the number of patients is', inst.NbClients)
+print(len(inst.gap))
+
+print('Shape of the OD matrix:')
+print(inst.od.shape)
+
+print('Shape of the preference matrix:')
+print(inst.CoutPref.shape)

### 2021-12-18 code/ExactMethod/read_helper_functions.py

@@ -1,0 +1,340 @@
+import os, importlib
+import numpy as np
+import copy
+
+def read_ait_h_instance(file_to_read, split=False):
+
+    if split: # split instances being called, so need to use file in 'split_instances' folder and no need to call depl_to_od function
+        folder = 'split_instances'
+        inst = importlib.import_module(folder + '.' + file_to_read)
+        inst.od = np.array(inst.od)
+        inst.CoutPref = np.array(inst.CoutPref)
+    else: # instance is in standard format, not split, call as normal.
+        folder = 'ait _haddadene_et_al_2016'
+        inst = importlib.import_module(folder + '.' + file_to_read)
+        inst.od = depl_to_od(inst)
+        inst.CoutPref = np.array(inst.CoutPref)
+    return inst
+### --- End def read_ait_h_instance --- ###
+
+def depl_to_od(ldi):
+
+    d = []
+    normalRow = ldi.NbClients - 1
+    extRow = ldi.NbClients
+    finalRow = ldi.depl[-1*extRow:]
+    # print('ldi.depl:\n')
+    # print(ldi.depl)
+    for ii in range(ldi.NbClients + 1):
+        if ii == 0:
+            map_i = ldi.NbClients
+            rowSize = extRow
+        else:
+            map_i = ii - 1
+            rowSize = normalRow
+        stLim = rowSize*map_i
+        endLim = rowSize*(map_i + 1)
+        dlRow = copy.copy(ldi.depl[stLim:endLim])
+
+        if ii != 0:
+            # print('Inserting ' + str(map_i) + ' / ' + str(len(finalRow)))
+            dlRow.insert(0,finalRow[map_i])
+            dlRow.insert(map_i + 1, 0)
+        else:
+            dlRow.insert(0, 0)
+
+        d.append(copy.copy(dlRow))
+    return np.array(d)
+### --- End def depl_to_od --- ####
+
+
+def split_instance(file_to_read, print_statements = True):
+
+    folder = 'ait _haddadene_et_al_2016'
+    inst = importlib.import_module(folder + '.' + file_to_read)
+    inst.od = depl_to_od(inst)
+    inst.CoutPref = np.array(inst.CoutPref)
+
+    # Demandes clients x services
+    Demandes1 = []
+    Clients1 = []
+    for i in range(len(inst.Demandes)):
+        if inst.Demandes[i][0] == 1:
[…]

### 2021-12-18 code/ExactMethod/split_instances/18-4-l-2d-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=2
+NbClients=10
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0]]
+od=[[0, 11, 15, 25, 22, 20, 27, 20, 11, 15, 20], [11, 0, 17, 23, 17, 16, 37, 31, 12, 6, 23], [15, 17, 0, 38, 12, 32, 36, 27, 24, 22, 34], [25, 23, 38, 0, 40, 8, 37, 37, 15, 19, 10], [22, 17, 12, 40, 0, 33, 46, 37, 28, 22, 39], [20, 16, 32, 8, 33, 0, 38, 35, 11, 12, 13], [27, 37, 36, 37, 46, 38, 0, 11, 29, 39, 27], [20, 31, 27, 37, 37, 35, 11, 0, 26, 34, 28], [11, 12, 24, 15, 28, 11, 29, 26, 0, 11, 12], [15, 6, 22, 19, 22, 12, 39, 34, 11, 0, 21], [20, 23, 34, 10, 39, 13, 27, 28, 12, 21, 0]]
+WindowsC=[[0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554]]
+DureeDeVisite=[[50, 0], [120, 0], [71, 0], [66, 0], [93, 0], [42, 0], [74, 0], [82, 0], [85, 0], [84, 0]]
+WindowsD=[[0, 554], [0, 554]]
+CoutPref=[[-8.21899, 6.56398], [-4.31381, 3.31036], [-7.94533, 4.9545], [-6.74515, 5.55603], [-5.12531, 3.18814], [-8.50266, 9.10865], [3.91134, 9.46388], [-7.83694, 0.135426], [-9.42898, -4.37351], [-3.8727, 7.82865]]
+gap=[-1, -1, -1, -1, -1, -1, -1, -1, -1, -1]

### 2021-12-18 code/ExactMethod/split_instances/18-4-l-2d-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=2
+NbClients=10
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1]]
+od=[[0, 21, 21, 12, 19, 15, 20, 15, 23, 21, 4], [21, 0, 9, 16, 27, 32, 21, 33, 44, 39, 23], [21, 9, 0, 11, 20, 34, 14, 35, 43, 35, 21], [12, 16, 11, 0, 12, 26, 9, 27, 32, 24, 11], [19, 27, 20, 12, 0, 31, 8, 31, 31, 19, 16], [15, 32, 34, 26, 31, 0, 34, 1, 17, 24, 16], [20, 21, 14, 9, 8, 34, 0, 34, 36, 26, 18], [15, 33, 35, 27, 31, 1, 34, 0, 17, 24, 16], [23, 44, 43, 32, 31, 17, 36, 17, 0, 14, 22], [21, 39, 35, 24, 19, 24, 26, 24, 14, 0, 18], [4, 23, 21, 11, 16, 16, 18, 16, 22, 18, 0]]
+WindowsC=[[0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554], [0, 554]]
+DureeDeVisite=[[0, 68], [0, 180], [0, 24], [0, 64], [0, 41], [0, 74], [0, 85], [0, 17], [0, 46], [0, 83]]
+WindowsD=[[0, 554], [0, 554]]
+CoutPref=[[0.703705, -8.09906], [-9.74309, -0.670695], [-0.712971, -1.74332], [-1.5397, -5.03829], [-9.39871, -7.39046], [1.22136, -1.63518], [-2.50244, 4.38737], [-1.98321, -3.7502], [0.466, -1.84221], [4.04154, 5.3464]]
+gap=[-1, -1, -1, -1, -1, -1, -1, -1, -1, -1]

### 2021-12-18 code/ExactMethod/split_instances/45-10-l-2a-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=5
+NbClients=26
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 25, 16, 20, 16, 22, 16, 18, 22, 21, 20, 20, 20, 19, 18, 21, 26, 19, 18, 15, 3, 13, 22, 22, 15, 16], [23, 0, 47, 27, 26, 33, 42, 28, 26, 43, 10, 12, 31, 19, 41, 33, 39, 34, 14, 8, 27, 21, 17, 43, 4, 34, 14], [25, 47, 0, 33, 33, 25, 10, 31, 36, 16, 44, 45, 33, 39, 11, 23, 23, 33, 43, 42, 28, 27, 36, 13, 47, 21, 40], [16, 27, 33, 0, 35, 9, 34, 3, 4, 21, 31, 17, 5, 34, 23, 33, 15, 42, 15, 27, 31, 18, 11, 23, 24, 13, 13], [20, 26, 33, 35, 0, 35, 25, 35, 37, 38, 17, 32, 39, 9, 32, 11, 39, 9, 32, 18, 7, 17, 29, 37, 28, 33, 30], [16, 33, 25, 9, 35, 0, 28, 7, 12, 13, 35, 25, 8, 36, 15, 31, 7, 40, 23, 32, 30, 19, 17, 15, 31, 5, 20], [22, 42, 10, 34, 25, 28, 0, 32, 37, 22, 37, 42, 35, 31, 16, 14, 28, 23, 40, 36, 20, 23, 34, 19, 43, 24, 38], [16, 28, 31, 3, 35, 7, 32, 0, 6, 19, 31, 19, 5, 33, 21, 32, 13, 41, 17, 28, 30, 18, 12, 21, 25, 11, 15], [18, 26, 36, 4, 37, 12, 37, 6, 0, 24, 31, 15, 6, 34, 26, 36, 18, 43, 13, 27, 33, 20, 10, 26, 
[…]

### 2021-12-18 code/ExactMethod/split_instances/45-10-l-2a-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=5
+NbClients=21
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 13, 20, 16, 7, 14, 21, 21, 16, 11, 18, 12, 19, 27, 19, 11, 29, 23, 17, 19, 16, 27], [13, 0, 30, 10, 19, 26, 25, 27, 27, 23, 30, 5, 32, 20, 16, 23, 23, 19, 20, 19, 21, 19], [20, 30, 0, 35, 14, 20, 37, 35, 4, 21, 16, 30, 11, 33, 38, 18, 32, 27, 14, 38, 12, 33], [16, 10, 35, 0, 22, 25, 18, 21, 31, 22, 31, 6, 34, 29, 7, 23, 32, 28, 28, 10, 29, 28], [7, 19, 14, 22, 0, 10, 24, 22, 10, 9, 12, 18, 13, 31, 25, 7, 31, 26, 16, 24, 15, 30], [14, 26, 20, 25, 10, 0, 19, 17, 16, 4, 7, 23, 12, 40, 26, 4, 41, 35, 25, 23, 24, 40], [21, 27, 35, 21, 22, 17, 0, 4, 31, 14, 23, 23, 28, 46, 17, 17, 48, 43, 36, 13, 36, 45], [21, 10, 35, 6, 25, 30, 21, 0, 31, 22, 31, 6, 34, 29, 7, 23, 32, 28, 28, 10, 29, 28], [16, 27, 4, 31, 10, 16, 33, 31, 0, 17, 13, 27, 9, 33, 34, 14, 32, 27, 14, 34, 12, 32], [11, 23, 21, 22, 9, 4, 17, 14, 17, 0, 10, 20, 15, 38, 22, 3, 39, 34, 25, 20, 24, 38], [18, 30, 16, 31, 12, 7, 26, 23, 13, 10, 0, 28, 6, 42, 32, 9, 42, 37, 25, 30, 23, 42], [12, 5, 30, 6, 18, 23, 21, 23, 
[…]

### 2021-12-18 code/ExactMethod/split_instances/45-10-l-2b-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=5
+NbClients=23
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 22, 17, 19, 20, 9, 29, 3, 17, 17, 25, 8, 11, 21, 17, 20, 24, 15, 13, 7, 22, 22, 17, 17], [22, 0, 20, 38, 10, 26, 27, 23, 35, 7, 23, 21, 22, 3, 32, 38, 23, 35, 34, 26, 21, 41, 9, 14], [17, 20, 0, 23, 26, 25, 13, 16, 33, 14, 9, 23, 8, 22, 33, 35, 7, 23, 23, 23, 33, 37, 22, 26], [19, 38, 23, 0, 38, 23, 31, 17, 24, 32, 29, 26, 17, 38, 29, 25, 27, 6, 9, 19, 40, 24, 35, 36], [20, 10, 26, 38, 0, 21, 35, 22, 30, 14, 31, 16, 25, 8, 25, 33, 31, 35, 32, 23, 11, 36, 4, 5], [9, 26, 25, 23, 21, 0, 37, 10, 10, 23, 33, 6, 19, 24, 9, 13, 32, 18, 15, 4, 18, 16, 19, 17], [29, 27, 13, 31, 35, 37, 0, 28, 45, 22, 5, 35, 20, 30, 45, 47, 6, 33, 34, 35, 43, 49, 32, 37], [3, 23, 16, 17, 22, 10, 28, 0, 18, 18, 24, 11, 9, 23, 18, 20, 23, 14, 11, 7, 24, 22, 19, 19], [17, 35, 33, 24, 30, 10, 45, 18, 0, 32, 41, 15, 26, 34, 7, 4, 40, 19, 16, 11, 25, 7, 28, 25], [17, 7, 14, 32, 14, 23, 22, 18, 32, 0, 18, 19, 16, 8, 30, 35, 18, 30, 29, 23, 22, 38, 10, 15], [25, 23, 9, 29, 31, 33,
[…]

### 2021-12-18 code/ExactMethod/split_instances/45-10-l-3-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=5
+NbClients=24
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 15, 19, 9, 8, 21, 12, 7, 22, 10, 20, 20, 18, 2, 11, 21, 20, 23, 22, 18, 12, 18, 14, 17, 22], [15, 0, 31, 9, 18, 11, 25, 13, 32, 6, 28, 18, 23, 14, 22, 7, 12, 32, 31, 19, 26, 4, 24, 32, 30], [19, 31, 0, 23, 24, 39, 7, 19, 5, 27, 33, 37, 15, 19, 24, 38, 39, 8, 9, 35, 18, 35, 9, 15, 12], [9, 9, 23, 0, 15, 18, 17, 4, 24, 6, 27, 21, 16, 7, 18, 15, 18, 24, 23, 21, 20, 13, 16, 25, 22], [8, 18, 24, 15, 0, 21, 17, 14, 27, 14, 12, 15, 26, 10, 4, 24, 19, 29, 28, 12, 10, 20, 21, 17, 30], [21, 11, 39, 18, 21, 0, 33, 21, 41, 13, 25, 11, 33, 21, 23, 9, 3, 42, 41, 14, 30, 8, 33, 36, 40], [12, 25, 7, 17, 17, 33, 0, 14, 10, 21, 27, 31, 14, 13, 18, 32, 32, 13, 12, 29, 13, 29, 7, 12, 15], [7, 13, 19, 4, 14, 21, 14, 0, 21, 9, 26, 23, 13, 5, 17, 19, 21, 21, 20, 22, 18, 16, 12, 22, 19], [22, 32, 5, 24, 27, 41, 10, 21, 0, 29, 37, 41, 13, 21, 28, 39, 41, 4, 5, 38, 22, 36, 9, 19, 9], [10, 6, 27, 6, 14, 13, 21, 9, 29, 0, 24, 16, 21, 9, 17, 12, 13, 29, 28, 1
[…]

### 2021-12-18 code/ExactMethod/split_instances/45-10-l-3-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=5
+NbClients=24
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 15, 22, 17, 19, 20, 9, 29, 3, 17, 17, 25, 8, 11, 21, 17, 20, 24, 15, 13, 7, 22, 22, 17, 17], [15, 0, 24, 30, 31, 17, 9, 42, 17, 14, 24, 37, 7, 25, 22, 9, 17, 36, 27, 23, 12, 11, 21, 16, 12], [22, 24, 0, 20, 38, 10, 26, 27, 23, 35, 7, 23, 21, 22, 3, 32, 38, 23, 35, 34, 26, 21, 41, 9, 14], [17, 30, 20, 0, 23, 26, 25, 13, 16, 33, 14, 9, 23, 8, 22, 33, 35, 7, 23, 23, 23, 33, 37, 22, 26], [19, 31, 38, 23, 0, 38, 23, 31, 17, 24, 32, 29, 26, 17, 38, 29, 25, 27, 6, 9, 19, 40, 24, 35, 36], [20, 17, 10, 26, 38, 0, 21, 35, 22, 30, 14, 31, 16, 25, 8, 25, 33, 31, 35, 32, 23, 11, 36, 4, 5], [9, 9, 26, 25, 23, 21, 0, 37, 10, 10, 23, 33, 6, 19, 24, 9, 13, 32, 18, 15, 4, 18, 16, 19, 17], [29, 42, 27, 13, 31, 35, 37, 0, 28, 45, 22, 5, 35, 20, 30, 45, 47, 6, 33, 34, 35, 43, 49, 32, 37], [3, 17, 23, 16, 17, 22, 10, 28, 0, 18, 18, 24, 11, 9, 23, 18, 20, 23, 14, 11, 7, 24, 22, 19, 19], [17, 14, 35, 33, 24, 30, 10, 45, 18, 0, 32, 41, 15, 26, 34, 7, 4, 40, 19,
[…]

### 2021-12-18 code/ExactMethod/split_instances/45-10-l-4-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=5
+NbClients=25
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 15, 22, 19, 9, 8, 21, 12, 7, 22, 10, 20, 20, 18, 2, 11, 21, 20, 23, 22, 18, 12, 18, 14, 17, 22], [15, 0, 24, 31, 9, 18, 11, 25, 13, 32, 6, 28, 18, 23, 14, 22, 7, 12, 32, 31, 19, 26, 4, 24, 32, 30], [22, 24, 0, 38, 26, 15, 20, 31, 27, 41, 22, 8, 9, 39, 23, 14, 27, 17, 43, 42, 6, 22, 24, 35, 28, 43], [19, 31, 38, 0, 23, 24, 39, 7, 19, 5, 27, 33, 37, 15, 19, 24, 38, 39, 8, 9, 35, 18, 35, 9, 15, 12], [9, 9, 26, 23, 0, 15, 18, 17, 4, 24, 6, 27, 21, 16, 7, 18, 15, 18, 24, 23, 21, 20, 13, 16, 25, 22], [8, 18, 15, 24, 15, 0, 21, 17, 14, 27, 14, 12, 15, 26, 10, 4, 24, 19, 29, 28, 12, 10, 20, 21, 17, 30], [21, 11, 20, 39, 18, 21, 0, 33, 21, 41, 13, 25, 11, 33, 21, 23, 9, 3, 42, 41, 14, 30, 8, 33, 36, 40], [12, 25, 31, 7, 17, 17, 33, 0, 14, 10, 21, 27, 31, 14, 13, 18, 32, 32, 13, 12, 29, 13, 29, 7, 12, 15], [7, 13, 27, 19, 4, 14, 21, 14, 0, 21, 9, 26, 23, 13, 5, 17, 19, 21, 21, 20, 22, 18, 16, 12, 22, 19], [22, 32, 41, 5, 24, 27, 41, 1
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-l-2-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=39
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 31, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 21, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 42, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 34, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 22, 19, 33, 23, 12, 31], [19, 10, 14, 19, 38, 0, 34, 21,
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-l-2-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=36
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 23, 21, 21, 11, 20, 15, 17, 10, 22, 25, 18, 22, 19, 12, 18, 19, 13, 20, 12, 17, 24, 22, 22, 12, 21, 11, 16, 25, 26, 23, 26, 18, 26, 19, 20, 14], [23, 0, 42, 28, 12, 42, 8, 21, 31, 41, 47, 14, 45, 35, 26, 21, 13, 32, 8, 33, 33, 10, 43, 44, 13, 43, 32, 30, 33, 48, 46, 42, 10, 47, 22, 14, 33], [21, 42, 0, 35, 31, 13, 35, 34, 16, 11, 14, 33, 6, 27, 27, 30, 38, 11, 40, 15, 16, 43, 10, 12, 32, 14, 17, 27, 37, 12, 7, 19, 38, 16, 31, 39, 12], [21, 28, 35, 0, 22, 26, 25, 8, 20, 41, 31, 33, 33, 12, 9, 36, 16, 32, 21, 20, 37, 29, 36, 29, 23, 27, 21, 10, 6, 34, 34, 46, 20, 30, 37, 17, 33], [11, 12, 31, 22, 0, 30, 5, 15, 20, 30, 35, 12, 33, 25, 16, 16, 11, 21, 11, 21, 23, 13, 32, 32, 10, 31, 21, 20, 27, 36, 34, 33, 9, 35, 17, 12, 22], [20, 42, 13, 26, 30, 0, 35, 28, 11, 23, 6, 37, 8, 16, 20, 36, 35, 19, 38, 9, 25, 43, 14, 4, 31, 10, 12, 19, 
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-l-3-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=40
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 15, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 8, 31, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 4, 21, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 25, 42, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 36, 34, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 22, 19, 33, 23, 12, 31], [19,
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-l-5-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=41
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 15, 22, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 8, 31, 41, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 4, 21, 30, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 25, 42, 41, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 36, 34, 22, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 2
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-l-5-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=37
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 23, 21, 21, 11, 20, 15, 17, 10, 22, 25, 18, 22, 25, 19, 12, 18, 19, 13, 20, 12, 17, 24, 22, 22, 12, 21, 11, 16, 25, 26, 23, 26, 18, 26, 19, 20, 14], [23, 0, 42, 28, 12, 42, 8, 21, 31, 41, 47, 14, 45, 41, 35, 26, 21, 13, 32, 8, 33, 33, 10, 43, 44, 13, 43, 32, 30, 33, 48, 46, 42, 10, 47, 22, 14, 33], [21, 42, 0, 35, 31, 13, 35, 34, 16, 11, 14, 33, 6, 18, 27, 27, 30, 38, 11, 40, 15, 16, 43, 10, 12, 32, 14, 17, 27, 37, 12, 7, 19, 38, 16, 31, 39, 12], [21, 28, 35, 0, 22, 26, 25, 8, 20, 41, 31, 33, 33, 45, 12, 9, 36, 16, 32, 21, 20, 37, 29, 36, 29, 23, 27, 21, 10, 6, 34, 34, 46, 20, 30, 37, 17, 33], [11, 12, 31, 22, 0, 30, 5, 15, 20, 30, 35, 12, 33, 32, 25, 16, 16, 11, 21, 11, 21, 23, 13, 32, 32, 10, 31, 21, 20, 27, 36, 34, 33, 9, 35, 17, 12, 22], [20, 42, 13, 26, 30, 0, 35, 28, 11, 23, 6, 37, 8, 30, 16, 20, 36, 35, 19, 38, 9, 
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-m-2-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=39
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 31, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 21, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 42, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 34, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 22, 19, 33, 23, 12, 31], [19, 10, 14, 19, 38, 0, 34, 21, 
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-m-3a-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=40
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 15, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 8, 31, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 4, 21, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 25, 42, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 36, 34, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 22, 19, 33, 23, 12, 31], [19,
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-m-3a-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=36
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 23, 21, 21, 11, 20, 15, 17, 10, 22, 25, 18, 22, 19, 12, 18, 19, 13, 20, 12, 17, 24, 22, 22, 12, 21, 11, 16, 25, 26, 23, 26, 18, 26, 19, 20, 14], [23, 0, 42, 28, 12, 42, 8, 21, 31, 41, 47, 14, 45, 35, 26, 21, 13, 32, 8, 33, 33, 10, 43, 44, 13, 43, 32, 30, 33, 48, 46, 42, 10, 47, 22, 14, 33], [21, 42, 0, 35, 31, 13, 35, 34, 16, 11, 14, 33, 6, 27, 27, 30, 38, 11, 40, 15, 16, 43, 10, 12, 32, 14, 17, 27, 37, 12, 7, 19, 38, 16, 31, 39, 12], [21, 28, 35, 0, 22, 26, 25, 8, 20, 41, 31, 33, 33, 12, 9, 36, 16, 32, 21, 20, 37, 29, 36, 29, 23, 27, 21, 10, 6, 34, 34, 46, 20, 30, 37, 17, 33], [11, 12, 31, 22, 0, 30, 5, 15, 20, 30, 35, 12, 33, 25, 16, 16, 11, 21, 11, 21, 23, 13, 32, 32, 10, 31, 21, 20, 27, 36, 34, 33, 9, 35, 17, 12, 22], [20, 42, 13, 26, 30, 0, 35, 28, 11, 23, 6, 37, 8, 16, 20, 36, 35, 19, 38, 9, 25, 43, 14, 4, 31, 10, 12, 19,
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-m-3b-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=40
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 15, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 8, 31, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 4, 21, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 25, 42, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 36, 34, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 22, 19, 33, 23, 12, 31], [19, 
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-m-3b-skill2.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=36
+NbServices=2
+Demandes=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+Offres=[[0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1], [0, 1]]
+od=[[0, 23, 21, 21, 11, 20, 15, 17, 10, 22, 25, 18, 22, 19, 12, 18, 19, 13, 20, 12, 17, 24, 22, 22, 12, 21, 11, 16, 25, 26, 23, 26, 18, 26, 19, 20, 14], [23, 0, 42, 28, 12, 42, 8, 21, 31, 41, 47, 14, 45, 35, 26, 21, 13, 32, 8, 33, 33, 10, 43, 44, 13, 43, 32, 30, 33, 48, 46, 42, 10, 47, 22, 14, 33], [21, 42, 0, 35, 31, 13, 35, 34, 16, 11, 14, 33, 6, 27, 27, 30, 38, 11, 40, 15, 16, 43, 10, 12, 32, 14, 17, 27, 37, 12, 7, 19, 38, 16, 31, 39, 12], [21, 28, 35, 0, 22, 26, 25, 8, 20, 41, 31, 33, 33, 12, 9, 36, 16, 32, 21, 20, 37, 29, 36, 29, 23, 27, 21, 10, 6, 34, 34, 46, 20, 30, 37, 17, 33], [11, 12, 31, 22, 0, 30, 5, 15, 20, 30, 35, 12, 33, 25, 16, 16, 11, 21, 11, 21, 23, 13, 32, 32, 10, 31, 21, 20, 27, 36, 34, 33, 9, 35, 17, 12, 22], [20, 42, 13, 26, 30, 0, 35, 28, 11, 23, 6, 37, 8, 16, 20, 36, 35, 19, 38, 9, 25, 43, 14, 4, 31, 10, 12, 19,
[…]

### 2021-12-18 code/ExactMethod/split_instances/73-16-s-2b-skill1.py

@@ -1,0 +1,11 @@
+NbVehicules=8
+NbClients=39
+NbServices=2
+Demandes=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+Offres=[[1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0], [1, 0]]
+od=[[0, 23, 13, 21, 21, 19, 22, 15, 24, 25, 23, 22, 25, 21, 17, 13, 25, 13, 19, 21, 22, 12, 19, 14, 22, 20, 16, 18, 23, 23, 26, 19, 24, 23, 22, 20, 14, 13, 14, 20], [23, 0, 11, 28, 43, 10, 31, 29, 32, 47, 9, 24, 41, 39, 9, 19, 46, 34, 25, 26, 42, 21, 38, 12, 29, 11, 9, 22, 32, 42, 48, 15, 10, 25, 40, 36, 20, 27, 35, 26], [13, 11, 0, 26, 34, 14, 21, 24, 30, 37, 11, 24, 30, 28, 11, 9, 37, 25, 23, 25, 34, 11, 27, 10, 27, 15, 5, 20, 22, 31, 39, 8, 12, 25, 29, 30, 10, 22, 26, 24], [21, 28, 26, 0, 28, 19, 42, 9, 5, 33, 34, 5, 45, 39, 19, 32, 29, 19, 4, 3, 23, 31, 38, 27, 10, 20, 26, 9, 43, 42, 32, 34, 35, 6, 40, 13, 33, 10, 20, 5], [21, 43, 34, 28, 0, 38, 34, 20, 29, 5, 44, 32, 29, 21, 36, 32, 5, 11, 30, 30, 8, 29, 21, 35, 29, 39, 37, 31, 35, 23, 6, 39, 45, 33, 22, 19, 33, 23, 12, 31], [19, 10, 14, 19, 38, 0, 34, 21, 
[…]

### 2021-12-18 code/ExactMethod/testeigen.py

@@ -1,0 +1,14 @@
+import numpy as np 
+import matplotlib.pyplot as plt
+
+M = np.zeros((3,3))
+M[1,1] = 9
+M[2,2] = 16
+
+print('M:\n', M)
+
+lambd, U = np.linalg.eig(M)
+
+print('lambd:\n', lambd)
+print('U:\n', U)
+

### 2021-12-18 code/ExactMethod/testnetworkx.py

@@ -1,0 +1,275 @@
+import networkx as nx
+import matplotlib.pyplot as plt
+import pandas as pd
+import os
+import sys
+
+def create_graph(file_to_read, num_clients, edge_set, nurse_index=999, is_nurse_list=False):
+
+    fig, ax = plt.subplots()
+    # plt.figure()
+
+    # Figure title
+    if is_nurse_list: # list of specific nurses, add all indices of nurses to the title
+        title_string = str(file_to_read) + '-nurses'  
+        for i in range(len(nurse_index)):
+            title_string += '-' + str(nurse_index[i])
+        plt.title(title_string)
+    else:
+        if nurse_index == 999: # all nurses
+            plt.title(str(file_to_read))
+        elif nurse_index < 999:
+            plt.title(str(file_to_read) + '-nurse' + str(nurse_index)) # one specific nurse
+
+    # Create graph
+    G = nx.MultiGraph()
+
+    # Add nodes to graph
+    df = pd.read_csv(r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/ExactMethod/18-4-s-2a_coords.csv') 
+
+    pos = {}
+    client_column = df['n']
+    x_column = df['x']
+    y_column = df['y']
+    min_x = 999
+    max_x = 0
+    min_y = 999
+    max_y = 0
+
+
+    for i in range(len(x_column)):
+        x_coord = df['x'][i]
+        y_coord = df['y'][i]
+        if x_coord > max_x:
+            max_x = x_coord
+        elif x_coord < min_x:
+            min_x = x_coord
+        if y_coord > max_y:
+            max_y = y_coord
+        elif y_coord < min_y:
+            min_y = y_coord
+        pos[df['n'][i]] = x_coord, y_coord
+    
+    print('min_x: ', min_x, ' max_x: ', max_x, ' min_y: ', min_y, ' max_y: ', max_y)
+
+    
+    # pos = {0: (0,0), 1: (0,1), 2: (0, 2), 3: (0, 3), 4: (0,4), 5: (1,0), 6: (1,1), 7: (1,2), 8: (1,3), 9: (1,4), 10: (2,0), 11: (2,1), 12: (2,2), 13: (2,3), 14: (2,4), 15: (3,0), 16: (3,1), 17: (3,2), 18: (3,3), 19: (3,4) }
+    # G.add_nodes_from(range(0,num_clients))
+    G.add_nodes_from(pos.keys())
+    for n, p in iter(pos.items()):
[…]

### 2021-12-18 code/ExactMethod/testnetworkx1.py

@@ -1,0 +1,173 @@
+import networkx as nx
+import matplotlib.pyplot as plt
+import pandas as pd
+import os
+import sys
+
+def create_graph(file_to_read, num_clients, edge_set, nurse_index=999, is_nurse_list=False):
+
+    fig, ax = plt.subplots()
+    # plt.figure()
+
+    # Figure title
+    if is_nurse_list: # list of specific nurses, add all indices of nurses to the title
+        title_string = str(file_to_read) + '-nurses'  
+        for i in range(len(nurse_index)):
+            title_string += '-' + str(nurse_index[i])
+        plt.title(title_string)
+    else:
+        if nurse_index == 999: # all nurses
+            plt.title(str(file_to_read))
+        elif nurse_index < 999:
+            plt.title(str(file_to_read) + '-nurse' + str(nurse_index)) # one specific nurse
+
+    # Create graph
+    G = nx.MultiGraph()
+
+    # Add nodes to graph
+    df = pd.read_csv(r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/ExactMethod/18-4-s-2a_coords.csv') 
+
+    pos = {}
+    client_column = df['n']
+    x_column = df['x']
+    y_column = df['y']
+    min_x = 999
+    max_x = 0
+    min_y = 999
+    max_y = 0
+
+
+    for i in range(len(x_column)):
+        x_coord = df['x'][i]
+        y_coord = df['y'][i]
+        if x_coord > max_x:
+            max_x = x_coord
+        elif x_coord < min_x:
+            min_x = x_coord
+        if y_coord > max_y:
+            max_y = y_coord
+        elif y_coord < min_y:
+            min_y = y_coord
+        pos[df['n'][i]] = x_coord, y_coord
+    
+    print('min_x: ', min_x, ' max_x: ', max_x, ' min_y: ', min_y, ' max_y: ', max_y)
+    # pos = {0: (0,0), 1: (0,1), 2: (0, 2), 3: (0, 3), 4: (0,4), 5: (1,0), 6: (1,1), 7: (1,2), 8: (1,3), 9: (1,4), 10: (2,0), 11: (2,1), 12: (2,2), 13: (2,3), 14: (2,4), 15: (3,0), 16: (3,1), 17: (3,2), 18: (3,3), 19: (3,4) }
+    # G.add_nodes_from(range(0,num_clients))
+    G.add_nodes_from(pos.keys())
+
[…]

### 2022-01-25 tex/Paper/main.tex
added pseudocode.

@@
 %\correction{What happens if the next "available" slot is not enough for the job.}
-\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
+%\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
+\noindent The \textsc{CalculateJobTimes} function in \ref{alg:shiftwork} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
+For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job.
+Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e.
+$g_i = 1$.
+Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ and job $j$ would be within a break.
+In both of these cases, the function \textsc{FindValidTime} is used.
+This function scans through the nurse's day to find the next available time period in which the nurse can feasibly perform the job $j$, and returns the start time of this period, $t_j$.
+If the job $j$ is a DS, then \textsc{FindValidTime} seeks a start time that is suitable for the current nurse $i$ and the second nurse on the job.
+
+\begin{algorithm}[h]
+\caption{\textsc{CalculateJobTimes}$(i, r_i)$}
+\label{alg:shiftwork}
+\begin{algorithmic}[1]
+	\State Let $f$ be the current shift for nurse $i$
+	\State $a(i,0) \gets 0$
+	\For{$j \gets 1 \To m_i$}
+		\State $a(i,0) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+		\If{$g_i = 1$}
+			\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
+			\State $z_j \gets \max\{0, (a(i,j) - l_j))\}$
+			\State $t_j \gets a(i,j) + w_{ij}$
+		\ElsIf{$g_i \neq 1 \AAnd \max\{e_j + s_j, a(i,j) + s_j\} > v(\pi_{f}^{i})$}
+			\State $t_j \gets$ \textsc{FindValidTime}$(i, j)$
+			\State $w_{ij} \gets t_j - l_j$
+			\State $z_j \gets t_j - l_j$
+			\State $f \gets$ shift containing time $t_j$
+		\ElsIf{$g_i \neq 1 \AAnd a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f}^{i})$}
+			\State $t_j \gets$ \textsc{FindValidTime}$(i,j)$
+			\State $w_{ij} \gets t_j - v(\pi_{f}^{i})$
+			\State $z_j \gets t_j - l_j$
+			\State $f \gets$ shift containing time $t_j$
+		\EndIf
+	\EndFor
+\end{algorithmic}
+\end{algorithm}
+
 For simplicity, we will consider all nurses' break periods as waiting time, so that the arrival time calculations $a(i,j)$ for nurses $i \in N$ at jobs $j \in J$ (see \eqref{eq:arrivaltime}) remain valid.

### 2022-02-10 tex/Paper/main.tex
added findjobtimes pseudocode.

@@
 
-\noindent however, they may start late. The tardiness with which a job $j$ starts, $z_j$, can be simply calculated using the following:
+\noindent however, they may start late. The tardiness with which a job $j$ starts, $z_j$, can be simply calculated using the following equation:
 \begin{equation}
@@
 
-The arrival time $a(i,j)$ of nurse $i$ at job $j \in r_i$ is calculated as follows:
+The arrival time $a(i,j)$ of nurse $i$ at job $j \in r_i$ is calculated as follows \chng{This is not true with shifts}:
 \begin{equation}
@@
 Note that there are no waiting or service times at the depot; that is, $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 $ for all $i \in N$.
-Furthermore, let $en_{\text{act}}(i)$ be the actual end time of a nurse $i \in N$, that is:
+Furthermore, let $en_{\text{act}}(i)$ be the actual end time of a nurse $i \in N$, that is: \chng{Is this different from $a(i, k_{0}^i)$?}
 \begin{equation}
@@
 Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, where $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$, and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$.
-Then, we enforce the following: \correction{Does this still hold?}
+Then, we enforce the following: \correction{Does this still hold?} \chng{Doesn't hold with shifts}
 %  $i_1 < i_2$, \ie $j_1$ appears earlier in the vector $\mathcal{O}$ than $j_2$. Then, we enforce that:
@@
 In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the end of the shift, and the start time for $j$ is moved forward to the start of $i$'s next shift.
-The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
+The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.\chng{not reflected in (12)}
 	\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time.
@@
 \noindent The \textsc{CalculateJobTimes} function in \ref{alg:shiftwork} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
-For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job.
-Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e.
+For each job $j \in r_i$, the \chng{arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job.
+- key, but not written in (7)} Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e.
 $g_i = 1$.
-Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ and job $j$ would be within a break.
+Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ and job $j$ would be within a break \chng{what about case where there are shifts but job is okay to do?}.
 In both of these cases, the function \textsc{FindValidTime} is used.
@@
 \begin{algorithmic}[1]
-	\State Let $f$ be the current shift for nurse $i$
-	\State $a(i,0) \gets 0$
-	\For{$j \gets 1 \To m_i$}
-		\State $a(i,0) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
-		\If{$g_i = 1$}
+	\State Let $f$ be the initial shift for nurse $i$
+	\State $a(i,k_{0}^i) \gets 0$
+%	\For{$j \gets 1 \To m_i$}
+	\For{$j \in r_i$}
+		\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+		\If{$g_i \neq 1 \AAnd \max\{e_j + s_j, a(i,j) + s_j\} > v(\pi_{f}^{i})$}
+			\State $t_j, f \gets$ \textsc{FindValidTime}$(i, j)$
+			\State $w_{ij} \gets t_j - a(i,j)$
+			\State $z_j \gets t_j - l_j$
+%			\State $f \gets$ shift containing time $t_j$
+		\ElsIf{$g_i \neq 1 \AAnd a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i})$}
+			\State $v \gets v(\pi_{f}^{i})$
+			\State $t_j, f \gets$ \textsc{FindValidTime}$(i,j)$
+			\State $w_{ij} \gets t_j - v$
+			\State $z_j \gets t_j - l_j$
+%			\State $f \gets$ shift containing time $t_j$
+		\Else
 			\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
-			\State $z_j \gets \max\{0, (a(i,j) - l_j))\}$
+			\State $z_j \gets \max\{0, (a(i,j) - l_j)\}$
 			\State $t_j \gets a(i,j) + w_{ij}$
-		\ElsIf{$g_i \neq 1 \AAnd \max\{e_j + s_j, a(i,j) + s_j\} > v(\pi_{f}^{i})$}
-			\State $t_j \gets$ \textsc{FindValidTime}$(i, j)$
-			\State $w_{ij} \gets t_j - l_j$
-			\State $z_j \gets t_j - l_j$
-			\State $f \gets$ shift containing time $t_j$
-		\ElsIf{$g_i \neq 1 \AAnd a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f}^{i})$}
-			\State $t_j \gets$ \textsc{FindValidTime}$(i,j)$
-			\State $w_{ij} \gets t_j - v(\pi_{f}^{i})$
-			\State $z_j \gets t_j - l_j$
-			\State $f \gets$ shift containing time $t_j$
 		\EndIf
 	\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[h]
+\footnotesize
+\caption{\textsc{FindValidTime}$(i, j, f, a(i,j))$}
+\label{alg:findvalidtime}
+\begin{algorithmic}[1]
+	\State $\tilde{f} \gets f$
+	\State $\tilde{a}(i,j) \gets a(i,j)$
+	\State valid $\gets \False$
+	\State DS $\gets \False$
+	\State second $\gets \False$
+	\If{$j$ is a double service}
+		\State DS $\gets \True$
+		\State $\tilde{i} \gets$ second nurse assigned to $j$
+		\If{$i > \tilde{i}$}
+			\State second $\gets \True$
+			\If{$\tilde{a}(i,j) > t_j$} %t_j is the start time for j, already set for 1st nurse earlier
+				% infeasible, exit
+				\State valid $\gets \False$
+				\Break
+			\EndIf
+		\EndIf
+	\EndIf % End of if j is double service
+	\While{$\Not$ valid}
+		\State $\tilde{f} \gets \tilde{f}+1$
+		\State $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
+		\If{DS$= \True$}
+			\If{second $= \False$}
+				\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+					\Continue
+				\Else
+					\State valid $\gets \True$
+					\State $t_j \gets \tilde{a}(i,j)$
+					\State $w_{ij} \gets t_j - a(i,j)$
+					\State $z_j \gets t_j - l_j$
+					\State $f \gets \tilde{f}$
+				\EndIf
+			\ElsIf{second $= \True$}
+				\If{$\tilde{a}(i,j) > t_j$} %t_j is the start time of the job from 1st nurse, already set
+					%infeasible, exit
+					\State valid $\gets \False$
+					\Break
+				\ElsIf{$t_j + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+					\Continue
+				\Else
+					\State valid $\gets \True$
+					%\State $t_j \gets \tilde{a}(i,j)$
+					\State $w_{ij} \gets t_j - a(i,j)$
+					%\State $z_j \gets t_j - l_j$
+					\State $f \gets \tilde{f}$
+				\EndIf
+			\EndIf %End if second is false
+		\ElsIf{DS $= \False$}
+			\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+				\Continue
+			\Else
+				\State valid $\gets \True$
+				\State $t_j \gets \tilde{a}(i,j)$
+				\State $w_{ij} \gets t_j - a(i,j)$
+				\State $z_j \gets t_j - l_j$
+				\State $f \gets \tilde{f}$
+			\EndIf
+		\EndIf %End if DS is True/True
+	\EndWhile
+	\Return either infeasible or starting time $t_j$, shift $f$, waiting time $w_{ij}$ and tardiness $z_j$
 \end{algorithmic}

### 2022-02-16 tex/Paper/main.tex
made changes to pseudocodes and equations, added sets DJ and DS for dependent and double services, where double services are now two separate jobs that need to be started at exactly the same time.

--- 2021-12-18 tex/Paper/main.tex
@@
 \begin{itemize}
-    \item Change all terms `nurse(s)' to `carer(s)' and `patient(s)' to `client(s)'
+    \item Make sure there are no 'carer's or 'clients', should be 'nurse' and patient'.
+    \item All max \{\}
 \end{itemize}
@@
 Each nurse $i$ might have a different starting time, $st(i)$, and end time, $en(i)$, of their working day.
-During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$, a subset of which, $DS \subset J$, are double services; that is, they require two nurses to be present simultaneously.
+During the planning horizon, which is one day in our case, the nurses need to perform a number of jobs $J = \{1,\dots,m\}$.
+A subset of these jobs, $J' \subset J$ will have temporal dependencies between them.
+The set $DJ = \{(j_1, j_2), (j_3,j_4),\dotsc \}$ organises the jobs from $J'$ into pairs that are dependent on one another.
+Further, $DS \subset DJ$ details those jobs that are double services; that is, they require two nurses to be present simultaneously for the jobs to start.
+We define a function $D : J' \to DJ$ such that $D(j)$ is the other job $j'$ paired with job $j$ in $DJ$.
 Note that jobs refer to individual tasks that a nurse needs to perform and are not necessarily identified with unique patients or locations, meaning that the same patient might have more than one job on the list.
+In this work we assume that each job can only be dependent on at most one other job and no further dependencies are allowed, which is in line with current practice.
+
+%\chng{add functions $D : J' \to DJ$ where $D(j)$ returns the other job in the pair with $j$, and $NJ : J \to N$, where $NJ(j)$ returns the nurse assigned to $j$.}
 
@@
 After a schedule is generated, for each job $j \in J$ we denote by $t_j$ its starting time.
-Note that this refers to the time at which the job starts, rather than the time at which the assigned nurse (or nurses) arrive to the job location, which may differ.
+Note that this refers to the time at which the job starts, rather than the time at which the assigned nurse arrive to the job location, which may differ.
+For convenience, let $NJ : J \to N$ define a function which returns the nurse $i \in N$ assigned to any given job $j \in J$.
 
@@
 To model this, we define a function $S_1: N\times J \rightarrow \{0,1\}$, that takes value $1$ if nurse $i$ is skilled to perform job $j$, $S_1(i,j) = 1$ and $0$ otherwise.
-Naturally, for double service jobs no single nurse is skilled to perform them, $S_1(i,j) = 0$ $\forall i \in N, \forall j \in DS$.
-To account for this, we define a second function, $S_2: N\times N\times J \rightarrow \{0,1\}$, that takes value $S_2(i_1,i_2,j) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform job $j$ and $0$ otherwise.
+Naturally, for double service jobs no single nurse is skilled to perform them; that is, $S_1(i,j) = 0 \text{ } \forall \text{ } i \in N \text{ if }\exists \text{ } j' \in J'$ such that $(j, j') \in DS$.
+This might also apply to pairs of dependent jobs in general.
+To account for this, we define a second function, $S_2: N\times N\times DJ \rightarrow \{0,1\}$, that takes value $S_2(i,i',(j, j')) = 1$ if the combination of nurses $i$ and $i'$ is sufficiently skilled to perform the pair of dependent jobs $(j, j')$ and $0$ otherwise.
+
+% that takes value $S_2(i_1,i_2,(j_1, j_2)) = 1$ if the combination of nurses $i_1$ and $i_2$ is sufficiently skilled to perform the pair of dependent jobs $(j_1, j_2)$ and $0$ otherwise.
 
@@
 All individual service jobs must be assigned to a skilled nurse, thus in any feasible solution it must hold that if $j \in r_i$ then $S_1(i,j) = 1$.
-Similarly, all double service jobs must be assigned to a pair of skilled nurses so if $j \in r_i \cap r_{i'}$ then it must hold that $S_2(i, i', j) = 1$.
+Similarly, all double service jobs must be assigned to a pair of skilled nurses; that is, for each $(j, j') \in DS$ such that $j \in r_{i}$ and $j' \in r_{i'}$, it must hold that $S_2(i, i', (j, j')) = 1$.
+
+%that is, for each $(j_1, j_2) \in DS$ such that $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$, it must hold that $S_2(i_1, i_2, (j_1, j_2)) = 1$.
+
+%\chng{Similarly, all double service jobs must be assigned to a pair of skilled nurses so if $j \in r_{i_1} \cap r_{i_2}$ then it must hold that $S_2(i_1, i_2, j) = 1$.
+- dbl service jobs are now duplicated, so now it's $(j_1, j_2)$, not $j$.
+Need to change this sentence.}
 
@@
 We distinguish the following cases:
-\begin{itemize}
+\begin{enumerate}
     \item \textbf{Duplicated}. In this case it is required for all staff members visiting the patient to have all necessary skills to perform the service.
@@
 This is the approach used in \cite{mankowska2014} and \cite{aithaddadene2016}.
-    \item \textbf{Shared-Duplicated}. This case is analogous to \emph{Shared}, but it allows some skills to be required by more than one of the staff members, hence it generalises the previous cases.
-\end{itemize}
-
-\noindent In our DST, the default approach is \emph{Shared-Duplicated}, as it allows for the flexibility that is observed in real operations.
+    \item \textbf{Shared-Duplicated}. This case is analogous to Shared, but it allows some skills to be required by more than one of the staff members, hence it generalises the previous cases.
+\end{enumerate}
+
+\noindent In our DST, the default approach is Shared-Duplicated, as it allows for the flexibility that is observed in real operations.
 The data is introduced as a list of skills and, if some skills need to be present in more than one staff member, they can simply appear duplicated on the list.
-We have, however, also implemented the \emph{Strictly Shared} approach in order to solve the instances from the literature.
+We have, however, also implemented the Strictly Shared approach in order to solve the instances from the literature.
 
@@
 \begin{equation}
-    z_j = \max(0, t_j - l_j). \label{eq:tardiness}
+    z_j = \max\{0, t_j - l_j\}. \label{eq:tardiness}
 \end{equation}
@@
 This can occur in practice, for example, in situations where a second dose of medication must be administered a few hours after the first dose.
-If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$ that depends on the start time of $j_1$, $[e^{(j_1)}_{j_2} + t_{j_1}, l^{(j_1)}_{j_2} + t_{j_1}]$.
+If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$ that depends on the start time of $j_1$, $[t_{j_1} + e_{j_2}^{(j_1)}, t_{j_1} + l_{j_2}^{(j_1)}]$.
 Then, the same time window constraints hold:
@@
 	\begin{align}
-	t_{j_2} &\geq e^{(j_1)}_{j_2} + t_{j_1}, \label{eq:timewindow1}\\[3pt] % & \forall j \in J
-	t_{j_2} &\leq l^{(j_1)}_{j_2} + t_{j_1}. \label{eq:timewindow2} % & \forall j \in J
+	t_{j_2} &\geq t_{j_1} + e^{(j_1)}_{j_2}, \label{eq:timewindow1}\\[3pt] % & \forall j \in J
+	t_{j_2} &\leq t_{j_1} + l^{(j_1)}_{j_2}. \label{eq:timewindow2} % & \forall j \in J
 	\end{align}
@@
 	\begin{align}
-	e^{(j_1)}_{j_2} &= -e^{(j_2)}_{j_1}, \label{eq:earliesttimedependecies}\\[3pt]
-	l^{(j_1)}_{j_2} &= -l^{(j_2)}_{j_1}. \label{eq:latesttimedependencies}
+		e^{(j_1)}_{j_2} &= -l^{(j_2)}_{j_1}, \label{eq:earliesttimedependecies}\\[3pt]
+		l^{(j_1)}_{j_2} &= -e^{(j_2)}_{j_1}. \label{eq:latesttimedependencies}
 	\end{align}
@@
 \begin{equation}
-    \tilde{z}_{j_2} = \max(0, t_{j_2} - (l^{(j_1)}_{j_2} + t_{j_1})). \label{eq:gaptardiness}
-\end{equation}
-
-\noindent Since some jobs have time windows or might need to be synchronised, waiting times may occur.
+    \tilde{z}_{j_2} = \max\{0, t_{j_2} - (t_{j_1} + l_{j_2}^{(j_1)})\}. \label{eq:gaptardiness}
+\end{equation}
+
+\noindent Note that for jobs $(j_1, j_2) \in DS$, $e_{j_2}^{(j_1)}$ and $l_{j_2}^{(j_1)}$ are equal to 0 as both jobs must start at the same time.
+
+Since some jobs have time windows or might need to be synchronised, waiting times may occur.
 We denote by $w_{ij}$ the waiting time of nurse $i$ at job $j$.
@@
 
-The arrival time $a(i,j)$ of nurse $i$ at job $j \in r_i$ is calculated as follows:
-\begin{equation}
-    a(i,j) = st(i) + \sum_{h=0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1}), \label{eq:arrivaltime}
-\end{equation}
-
-\noindent where $j'$ denotes the position of job $j$ in the route $r_i$, i.e.
+%The arrival time $a(i,j)$ of nurse $i$ at job $j \in r_i$ is calculated as follows:
+%\begin{equation}
+%    a(i,j) = st(i) + \sum_{h=0}^{j' - 1} w_{ik^i_h} + s_{k^i_h} + T_i(k^i_h, k^i_{h + 1}), \label{eq:arrivaltime}
+%\end{equation}
+
+%\noindent \chng{where $j'$ denotes the position of job $j$ in the route $r_i$, i.e.
 $j = k^i_{j'} \in r_i$.
-Note that there are no waiting or service times at the depot; that is, $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 $ for all $i \in N$.
+Note that there are no waiting or service times at the depot; that is, $w_{ik^i_0} = 0$ and $s_{k^i_0} = 0 $ for all $i \in N$.}
+
+The arrival time $a(i,j)$ of nurse $i$ at job $j \in r_i$ is calculated recursively as follows:
+\begin{subequations}
+	\begin{align}
+		a(i,k_0^i) &= 0, \label{eq:recursionarrivalbase}\\[3pt]
+		a(i,j) &= a(i,j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j). \label{eq:arrivaltime}
+	\end{align}
+\end{subequations}
+The function \textsc{CalculateJobTimes} in Algorithm~\ref{alg:cjt} below describes the process further.
+Note that there are no waiting or service times at the depot.
 Furthermore, let $en_{\text{act}}(i)$ be the actual end time of a nurse $i \in N$, that is:
 \begin{equation}
-	en_{\text{act}}(i) = a(i,m_i) + w_{ik_{m_i}^i} + s_{k_{m_i}^i}, \label{eq:actualendtime}
+	en_{\text{act}}(i) = a(i,m_i) + w_{im_i} + s_{m_i}, \label{eq:actualendtime}
 \end{equation}
 where $m_i$ is the final job in nurse $i$'s route $r_i$.
+
+%Furthermore, let $en_{\text{act}}(i)$ be the actual end time of a nurse $i \in N$, that is:
+%\begin{equation}
+%	en_{\text{act}}(i) = a(i,m_i) + w_{ik_{m_i}^i} + s_{k_{m_i}^i}, \label{eq:actualendtime}
+%\end{equation}
+%where $m_i$ is the final job in nurse $i$'s route $r_i$.
 
@@
     \begin{equation}
-        w_{i j} \geq t_j - a(i',j) \quad \forall j \in DS \text{ such that } j \in r_{i} \cap r_{i'}. \label{eq:waiting2DS}
+        %w_{i j} \geq t_j - a(i',j) \quad \forall j \in DS \text{ such that } j \in r_{i} \cap r_{i'}. \label{eq:waiting2DS}
+       \chng{ w_{i j} \geq t_j - a(i',j') \text{ } \forall (j,j') \in DS \text{ such that } j \in r_{i} \text{ and } j' \in r_{i'}.} \label{eq:waiting2DS}
     \end{equation}
@@
 
-\noindent Furthermore, we assume that waiting times are never larger than strictly necessary.
+\noindent \chng{Furthermore, we assume that waiting times are never larger than strictly necessary.
 To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based on the order in which they appear in the solution.
 Let $j_1, j_2 \subset J, j_1 \neq j_2$, be two jobs that have a dependency, where $j_1 \in r_{i_1}$ and $j_2 \in r_{i_2}$, and nurse $i_1$ appears earlier than nurse $i_2$ in the vector of nurse order, $\mathcal{O}$.
-Then, we enforce the following: \correction{Does this still hold?}
+Then, we enforce the following:} \correction{Does this still hold?}
 %  $i_1 < i_2$, \ie $j_1$ appears earlier in the vector $\mathcal{O}$ than $j_2$. Then, we enforce that:
@@
 Given these shifts, we must ensure that jobs are assigned to nurses during their actual working time.
-This also involves considering two cases that may occur:
+This also involves considering two cases that may occur: \chng{Need to ref the figures.}
 \begin{enumerate}[label={(\arabic*)}]
@@
 Tardiness as calculated as above.
-	\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period.
-In this case, the waiting time incurred is from the arrival time of nurse $i$ at $j$ to the end of the shift, and the start time for $j$ is moved forward to the start of $i$'s next shift.
-The tardiness is calculated as the time between the end of $j$'s time window, $l_j$, and the new start time of $j$.
-	\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time.
-In this case, no waiting time is incurred and the new start time for job $j$ and tardiness are as above.
+	\chng{\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period.}
+	\chng{\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time.}
 \end{enumerate}
@@
 %\correction{What happens if the next "available" slot is not enough for the job.}
-\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
+%\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
+\noindent The \textsc{CalculateJobTimes} function in Algorithm \ref{alg:cjt} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
+For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job (lines 3--4).
+Line 6 covers the case where the duration of the current job $j$ extends into a break period, whilst line 12 considers the event when the arrival time of nurse $i$ at $j$ would be within a break period.
+In both of these cases, the function \textsc{FindValidTime} in Algorithm~\ref{alg:fvt} is called, which scans through nurse $i$'s day to find a suitable time in which $i$ can perform $j$ (lines 6 and 12).
+If $j$ cannot be feasibly scheduled into nurse $i$'s day, then no feasible solution exists and the algorithm terminates (lines 7--8 and lines 13--14).
+Otherwise, \textsc{FindValidTime} has successfully placed job $j$ into the schedule of nurse $i$, and returns the start time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, and gap tardiness $\tilde{z}_j$ (if $j \in J'$ and gap tardiness is permitted) of job $j$, as well as the current shift $f$ of nurse $i$, and \textsc{CalculateJobTimes} continues on to the next job $j \in r_i$ (lines 9 and 15).
+There are two scenarios in which \textsc{FindValidTime} does not need to be called: (1) if nurse $i$ can feasibly perform job $j$ with the arrival time $a(i,j)$; and (2) if $f$ is the last shift of nurse $i$, i.e.
+$f = g_i$.
+In both instances, the algorithm simply calculates the starting time $t_j$, waiting time $w_{ij}$, and tardiness $z_j$ (lines 16--19).
+
+%\chng{The \textsc{CalculateJobTimes} function in \ref{alg:shiftwork} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
+For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job.
+Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e.
+$g_i = 1$.
+Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ at job $j$ would be within a break.
+In both of these cases, the function \textsc{FindValidTime} is used.
+This function scans through the nurse's day to find the next available time period in which the nurse can feasibly perform the job $j$, and returns the start time of this period, $t_j$.
+If the job $j$ is a DS, then \textsc{FindValidTime} seeks a start time that is suitable for the current nurse $i$ and the second nurse on the job.}
+
+\begin{algorithm}[h]
+\footnotesize
+\caption{\textsc{CalculateJobTimes}$(i, r_i)$}
+\label{alg:cjt}
+\begin{algorithmic}[1]
+	\State Let $f$ be the initial shift for nurse $i$
+	\State $a(i,k_{0}^i) \gets 0$
+%	\For{$j \gets 1 \To m_i$}
+	\For{$j \in r_i$}
+		\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+		\If{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
+			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i,j,f, a(i,j)$)
+			\If{feasible $= \False$}
+				\Break
+			\Else \hspace{0.2mm} \textbf{continue}
+				%\Continue
+			\EndIf
+			%\State $t_j, f \gets$ \textsc{FindValidTime}$(i, j, f, a(i,j))$
+			%\State $w_{ij} \gets t_j - a(i,j)$
+			%\State $z_j \gets t_j - l_j$
+%			\State $f \gets$ shift containing time $t_j$
+		\ElsIf{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
+			\State $v \gets v(\pi_{f}^{i})$
+			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i,j,f,v$)
+			\If{feasible $= \False$}
+				\Break
+			\Else \hspace{0.2mm} \textbf{continue}
+				%\Continue
+			\EndIf
+			%\State $t_j, f \gets$ \textsc{FindValidTime}$(i,j, f, v)$
+			%\State $w_{ij} \gets t_j - v$
+			%\State $z_j \gets t_j - l_j$
+%			\State $f \gets$ shift containing time $t_j$
+		\Else
+			\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
+			\State $z_j \gets \max\{0, (a(i,j) - l_j)\}$
+			\State $t_j \gets a(i,j) + w_{ij}$
+		\EndIf
+	\EndFor
+	\Return either statement of infeasibility or feasible schedule for nurse $i$
+\end{algorithmic}
+\end{algorithm}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{algorithm}[htbp!]
+\scriptsize
+\caption{\textsc{FindValidTime($i, j, f, a(i,j)$)}}
+\label{alg:fvt}
+\begin{algorithmic}[1]
+	\State feasible $\gets \False$, DJ $\gets \False$, DS $\gets \False$, secondN $\gets \False$, gap $\gets \False$
+	\State $t_j \gets \Null$, $w_{ij} \gets \Null$, $z_j \gets \Null$, $\tilde{z}_j \gets \Null$
+	\State $\tilde{f} \gets f$, $\tilde{a}(i,j) \gets a(i,j)$
+	\If{$j \in J'$}
+		\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
+		\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+		\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+			\State secondN $\gets \True$
+		\EndIf
+		\If{$(j, j') \in DS$}
+			\State DS $\gets \True$
+				\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
+					%\State valid $\gets \False$
+					%\Return infeasible
+					\Return feasible $\gets \False$
+				\EndIf % arrival later than latest start time
+		\Else 
+			\State $DJ \gets \True$
+			%\If{secondN $= \False$}
+			%	\State original TWs are used
+			\If{secondN $= \True$} % This nurse is the 2nd one to be assessed, other job has been done
+				%Need to find out if this job j is before or after j' in the schedule
+				%\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+				\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+				%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
+				\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
+					%\State valid $\gets \False$
+					%\Return infeasible
+					\Return feasible $\gets \False$
+				\EndIf
+			\EndIf % i appears later than i' in O
+		\EndIf %(j, j') in DS
+	\EndIf %if j in J'
+	%\While{$\Not$ valid}
+	\While{$\Not$ feasible}
+		\State $\tilde{f} \gets \tilde{f}+1$
+		\State $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
+		%\If{secondN $= \True \AAnd$(DJ $= \True \Or$DS $= \True$)}
+		\If{secondN $= \True$}
+			%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
+			\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$}
+				%\State valid $\gets \False$
+				%\Break
+				%\Return infeasible
+				\Return feasible $\gets \False$
+			%\ElsIf{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \True$}
+			\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \True$}
+				\If{$l_j^{(j')} > t_{j'}$} % job j occurs after job j', gap tardiness allowed
+					\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+						\Continue
+					\Else
+						%\State valid $\gets \True$
+						\State feasible $\gets \True$
+						\State $t_j \gets \tilde{a}(i,j)$
+						\State $w_{ij} \gets t_j - a(i,j)$
+						\State $z_j \gets t_j - l_j$
+						%\State $\tilde{z}_j \gets t_j - l_j^{(j')}$
+						\State $\tilde{z}_j \gets t_j - (t_{j'} + l_j^{(j')})$
+						\State $f \gets \tilde{f}$
+					\EndIf
+				\Else \hspace{0.2mm} \textbf{return} feasible $\gets \False$ % gap tardiness allowed but job is before, so gap is smaller
+					%\State valid $\gets \False$
+					%\Break
+					%\Return infeasible
+				\EndIf
+			%\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}
+			\ElsIf{$\max\{\tilde{a}(i,j), (t_{j'} + e_j^{(j')})\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}
+				\Continue
+			\Else
+				%\State valid $\gets \True$
+				\State feasible $\gets \True$
+				\State $t_j \gets \tilde{a}(i,j)$
+				\State $w_{ij} \gets t_j - a(i,j)$
+				\State $z_j \gets t_j - l_j$
+				\State $f \gets \tilde{f}$
+			\EndIf
+		\ElsIf{secondN $= \False$}
+			\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+				\Continue
+			\Else
+				%\State valid $\gets \True$
+				\State feasible $\gets \True$
+				\State $t_j \gets \tilde{a}(i,j)$
+				\State $w_{ij} \gets t_j - a(i,j)$
+				\State $z_j \gets t_j - l_j$
+				\State $f \gets \tilde{f}$
+			\EndIf
+		\EndIf
+	\EndWhile
+	\Return statement of feasibility, starting time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, gap tardiness $\tilde{z}_j$, and shift $f$
+\end{algorithmic}
+\end{algorithm}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Explanation of FindValidTime function
+The \textsc{FindValidTime} function shown in Algorithm~\ref{alg:fvt} operates as follows: firstly, if $j$ is a dependent job or a double service, the algorithm obtains the paired job $j'$ and the nurse $i'$ assigned to $j'$ (lines 4--6).
+If the schedule for $i'$ has already been set, then $i$ is considered to be the ``second nurse'' (line 7).
+Then, if $j$ is a double service job and the current time $\tilde{a}(i,j)$ is after the start time set for $j'$, no feasible solution can be found and the algorithm terminates (lines 9--12).
+Likewise, if $j$ is a dependent job, and the current time $\tilde{a}(i,j)$ occurs after the latest allowed start time and no gap tardiness is permitted, the algorithm terminates with a statement of infeasibility (lines 13--18).
+
+The main while loop of \textsc{FindValidTime} involves progressing through each shift $\pi_{\tilde{f}}^i \in \Pi_i$ and setting the current time $\tilde{a}(i,j)$ to the start of each shift, $u_(\pi_{\tilde{f}}^{i})$ (lines 20--21).
+Each loop comprises two parts.
+If $i$ is the second nurse assessed, this implies that $j$ is a dependent job.
+The same condition from line 17 is checked, where $l_{j}^{(j')} = 0$ if $j$ is a double service, and again the algorithm terminates if the condition is not met.
+However, if gap tardiness is permitted, this implies that $j$ is not a double service (i.e.
+DJ $= \True$).
+The algorithm evaluates the order of the jobs $j$ and $j'$ and determines whether the job can be performed feasibly (lines 25--36).
+Otherwise, $\tilde{a}(i,j)$ occurs before the latest start time, and the usual checks are performed (lines 37--44).
+Finally, the same assessments occur if $j$ is a paired job and $i$ is the first nurse to be evaluated, or if $j$ is a single job (lines 45--53).
+
 For simplicity, we will consider all nurses' break periods as waiting time, so that the arrival time calculations $a(i,j)$ for nurses $i \in N$ at jobs $j \in J$ (see \eqref{eq:arrivaltime}) remain valid.
@@
 
+\begin{comment}
+\begin{algorithm}[htbp!]
+\footnotesize
+\caption{\textsc{FindValidTime}$(i, j, f, a(i,j))$}
+\label{alg:oldfindvalidtime}
+\begin{algorithmic}[1]
+\State \chng{all $e/l_j^{(i')}$ need to change to $e/l_j^{(j')}$ where $j'$ is the other dep/ds job done by nurse $i'$}
+\State $\tilde{f} \gets f$
+\State $\tilde{a}(i,j) \gets a(i,j)$
+\State valid $\gets \False$
+\State DS $\gets \False$
+\State second $\gets \False$
+\If{$j$ is a double service}
+\State DS $\gets \True$
+\State $i' \gets$ other nurse assigned to $j$
+%\If{$i > \tilde{i}$}
+\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+\State second $\gets \True$
+\If{$\tilde{a}(i,j) > l_j^{(i')}$} % this should be l_{j}^{j'}, where j' is the matching job done by i'
+% infeasible, exit
+\State valid $\gets \False$
+\Return
+\EndIf
+\EndIf
+\EndIf % End of if j is double service
+\While{$\Not$ valid}
+\State $\tilde{f} \gets \tilde{f}+1$
+\State $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
+\If{\chng{DS $= \True$ or j is synchronised}} % rewrite with new variables
+\If{second $= \False$}
+\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+\State valid $\gets \True$
+\State $t_j \gets \tilde{a}(i,j)$
+\State $w_{ij} \gets t_j - a(i,j)$
+\State $z_j \gets t_j - l_j$
+\State $f \gets \tilde{f}$
+\EndIf
+\ElsIf{second $= \True$}
+%\State $i' \gets $ first nurse assigned to the job
+\State $[e_j^{(i')}, l_j^{(i')}] \gets$ dependent time window, based on $t_j$ set by nurse $i'$.
+\If{$\tilde{a}(i,j) > l_j^{(i')}$} %t_j is the start time of the job from 1st nurse, already set
+%infeasible, exit
+\State valid $\gets \False$
+\Break
+\ElsIf{$l_j^{(i')} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+\State valid $\gets \True$
+%\State $t_j \gets \tilde{a}(i,j)$
+\State $w_{ij} \gets t_j - a(i,j)$
+%\State $z_j \gets t_j - l_j$
+\State $f \gets \tilde{f}$
+\EndIf
+\EndIf %End if second is false
+\ElsIf{DS $= \False$}
+\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+\State valid $\gets \True$
+\State $t_j \gets \tilde{a}(i,j)$
+\State $w_{ij} \gets t_j - a(i,j)$
+\State $z_j \gets t_j - l_j$
+\State $f \gets \tilde{f}$
+\EndIf
+\EndIf %End if DS is True/True
+\EndWhile
+\Return either infeasible or starting time $t_j$, shift $f$, waiting time $w_{ij}$ and tardiness $z_j$
+\end{algorithmic}
+\end{algorithm}
+\end{comment}
+
+\begin{comment}
+\begin{algorithm}[htbp!]
+\footnotesize
+\caption{\chng{ROUGH FIND VALID TIME ALG}}
+\label{alg:roughfvt}
+\begin{algorithmic}[1]
+\State $\tilde{f} \gets f$, $\tilde{a}(i,j) \gets a(i,j)$, valid $\gets \False$, 
+\State DS $\gets \False$, DJ $\gets \False$, secondN $\gets \False$, jType $\gets$ NULL
+\State gap $\gets \False$
+\If{$j \in J'$}
+\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
+\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+\State secondN $\gets \True$
+\EndIf
+\If{$(j, j') \in DS$}
+\State DS $\gets \True$
+\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
+\State valid $\gets \False$, infeasible, EXIT
+\EndIf % arrival later than latest start time
+\Else 
+\State $DJ \gets \True$
+\If{secondN $= \False$}
+\State original TWs are used
+\ElsIf{secondN $= \True$} % This nurse is the 2nd one to be assessed, other job has been done
+%Need to find out if this job j is before or after j' in the schedule
+\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_j'$ set by $i'$
+\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
+\State valid $\gets \False$, infeasible, EXIT
+\EndIf
+\EndIf % i appears later than i' in O
+\EndIf %(j, j') in DS
+\EndIf %if j in J'
+\While{$\Not$ valid}
+\State $\tilde{f} \gets \tilde{f}+1$
+\State $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
+\If{DS $= \True$}
+%\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$}
+\If{secondN $= \True$}
+\If{$\tilde{a}(i,j) > t_{j'}$}
+\State valid $\gets \False$, infeasible, EXIT
+\ElsIf{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+\State valid $\gets \True$ etc...
+\EndIf
+\ElsIf{secondN $= \False$}
+\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+\State valid $\gets \True$ etc...
+\EndIf
+\EndIf
+\ElsIf{DJ $= \True$}
+\If{secondN $= \True$}
+\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$} % past TW and no gap tard allowed
+\State valid $\gets \False$, infeasible, EXIT
+\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}%\tilde{a}(i,j) is before or within TW, so check if job can be done in shift
+\Continue
+\Else
+\State valid $\gets \True$ etc...
+\EndIf
+\ElsIf{secondN $= \False$}
+\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+\State valid $\gets \True$ etc...
+\EndIf
+\EndIf
+\EndIf % End if DS/DJ is True
+\EndWhile
+\end{algorithmic}
+\end{algorithm}
+\end{comment}
+
+%		\If{DS $= \True \AAnd$secondN $= \True$}
+%			\If{$\tilde{a}(i,j) > t_{j'}$}
+%				\State valid $\gets \False$, infeasible, EXIT
+%			\ElsIf{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+%				\Continue
+%			\Else
+%				\State valid $\gets \True$ etc...
+%			\EndIf
+%		\ElsIf{DJ $= \True \AAnd$secondN $= \True$}
+%			\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$} % past TW and no gap tard allowed
+%				\State valid $\gets \False$, infeasible, EXIT
+%			\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}%\tilde{a}(i,j) is before or within TW, so check if job can be done in shift
+%				\Continue
+%			\Else
+%				\State valid $\gets \True$ etc...
+%			\EndIf
+%		\EndIf % End if DS/DJ is True
+
 \end{document}

### 2022-02-21 tex/Paper/main.tex
updated main.tex
made changes to CalculateJobTimes pseudocode.

@@
 		\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+		\If{$j \in J'$}
+			\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
+			\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+			\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+				\State secondN $\gets \True$
+			\EndIf
+			\If{$(j, j') \in DS$}
+				\State DS $\gets \True$
+				\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
+					\Return feasible $\gets \False$
+				\EndIf % arrival later than latest start time
+			\Else 
+				\State $DJ \gets \True$
+				\If{secondN $= \True$} % This nurse is 2nd one to be assessed, other job has been done
+				%Need to find out if this job j is before or after j' in the schedule
+					\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+					\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
+						\Return feasible $\gets \False$
+					\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
+						\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
+							\Return feasible $\gets \False$
+						\EndIf
+					\EndIf
+				\EndIf % i appears later than i' in O
+			\EndIf %(j, j') in DS
+		\EndIf %if j in J'
 		\If{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
@@
 			\EndIf
-			%\State $t_j, f \gets$ \textsc{FindValidTime}$(i, j, f, a(i,j))$
-			%\State $w_{ij} \gets t_j - a(i,j)$
-			%\State $z_j \gets t_j - l_j$
-%			\State $f \gets$ shift containing time $t_j$
 		\ElsIf{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
@@
 			\EndIf
-			%\State $t_j, f \gets$ \textsc{FindValidTime}$(i,j, f, v)$
-			%\State $w_{ij} \gets t_j - v$
-			%\State $z_j \gets t_j - l_j$
-%			\State $f \gets$ shift containing time $t_j$
 		\Else
@@
 					\Return feasible $\gets \False$
+				\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
+					\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
+						\Return feasible $\gets \False$
+					\EndIf
 				\EndIf
@@
 			\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \True$}
-				\If{$l_j^{(j')} > t_{j'}$} % job j occurs after job j', gap tardiness allowed
+				\If{$t_{j'} + l_j^{(j')} > t_{j'}$} % job j occurs after job j', gap tardiness allowed
 					\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
@@
     \centering
-    \includegraphics[width=0.7\textwidth]{images/scheme}
+    \includegraphics[width=0.7\textwidth]{figures/scheme}
     \caption{Scheme of the proposed GRASP algorithm.}
@@
 \end{algorithm}
+
+\begin{algorithm}[h]
+\footnotesize
+\caption{\textsc{CalculateJobTimes}$(i, r_i)$}
+\label{alg:cjtold}
+\begin{algorithmic}[1]
+\State Let $f$ be the initial shift for nurse $i$
+\State $a(i,k_{0}^i) \gets 0$
+%	\For{$j \gets 1 \To m_i$}
+\For{$j \in r_i$}
+\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+\If{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
+\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i,j,f, a(i,j)$)
+\If{feasible $= \False$}
+\Break
+\Else \hspace{0.2mm} \textbf{continue}
+%\Continue
+\EndIf
+%\State $t_j, f \gets$ \textsc{FindValidTime}$(i, j, f, a(i,j))$
+%\State $w_{ij} \gets t_j - a(i,j)$
+%\State $z_j \gets t_j - l_j$
+%			\State $f \gets$ shift containing time $t_j$
+\ElsIf{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
+\State $v \gets v(\pi_{f}^{i})$
+\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i,j,f,v$)
+\If{feasible $= \False$}
+\Break
+\Else \hspace{0.2mm} \textbf{continue}
+%\Continue
+\EndIf
+%\State $t_j, f \gets$ \textsc{FindValidTime}$(i,j, f, v)$
+%\State $w_{ij} \gets t_j - v$
+%\State $z_j \gets t_j - l_j$
+%			\State $f \gets$ shift containing time $t_j$
+\Else
+\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
+\State $z_j \gets \max\{0, (a(i,j) - l_j)\}$
+\State $t_j \gets a(i,j) + w_{ij}$
+\EndIf
+\EndFor
+\Return either statement of infeasibility or feasible schedule for nurse $i$
+\end{algorithmic}
+\end{algorithm}
+
 \end{comment}

### 2022-02-25 tex/Paper/main.tex
updated CalculateJobTimes and FindValidTimes pseudocode to make sure all options are covered.
changes to pseudocode, swapped statements around.

@@
 
-\begin{algorithm}[h]
-\footnotesize
+\begin{algorithm}[htbp!]
+\scriptsize
 \caption{\textsc{CalculateJobTimes}$(i, r_i)$}
@@
 \begin{algorithmic}[1]
+	\State feasible $\gets \False$
 	\State Let $f$ be the initial shift for nurse $i$
@@
 		\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+		\State $i' \gets \Null$, $j' \gets \Null$
+		\State DJ $\gets \False$
+		\State DS $\gets \False$
+		\State secondN $\gets \False$
 		\If{$j \in J'$}
@@
 				\State DS $\gets \True$
-				\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
-					\Return feasible $\gets \False$
+				\If{secondN $= \True \AAnd a(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
+					\State feasible $\gets \False$
+					\Break
 				\EndIf % arrival later than latest start time
@@
 					\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-					\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
-						\Return feasible $\gets \False$
-					\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
+					\If{$a(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
+						\State feasible $\gets \False$
+						\Break
+					\ElsIf{$a(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
 						\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
-							\Return feasible $\gets \False$
+							\State feasible $\gets \False$
+							\Break
 						\EndIf
@@
 		\EndIf %if j in J'
-		\If{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
-			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i,j,f, a(i,j)$)
+		\If{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
+			\State $v \gets v(\pi_{f}^{i})$
+			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, v, i', j'$, DS, DJ, secondN)
 			\If{feasible $= \False$}
@@
 			\EndIf
-		\ElsIf{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
-			\State $v \gets v(\pi_{f}^{i})$
-			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i,j,f,v$)
+		\ElsIf{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
+			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN)
 			\If{feasible $= \False$}
@@
 \scriptsize
-\caption{\textsc{FindValidTime($i, j, f, a(i,j)$)}}
+\caption{\textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN, gap)}
 \label{alg:fvt}
 \begin{algorithmic}[1]
-	\State feasible $\gets \False$, DJ $\gets \False$, DS $\gets \False$, secondN $\gets \False$, gap $\gets \False$
+	\State feasible $\gets \False$
 	\State $t_j \gets \Null$, $w_{ij} \gets \Null$, $z_j \gets \Null$, $\tilde{z}_j \gets \Null$
 	\State $\tilde{f} \gets f$, $\tilde{a}(i,j) \gets a(i,j)$
-	\If{$j \in J'$}
-		\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
-		\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
-		\If{$i$ appears later than $i'$ in $\mathcal{O}$}
-			\State secondN $\gets \True$
-		\EndIf
-		\If{$(j, j') \in DS$}
-			\State DS $\gets \True$
-				\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
-					%\State valid $\gets \False$
-					%\Return infeasible
-					\Return feasible $\gets \False$
-				\EndIf % arrival later than latest start time
-		\Else 
-			\State $DJ \gets \True$
-			%\If{secondN $= \False$}
-			%	\State original TWs are used
-			\If{secondN $= \True$} % This nurse is the 2nd one to be assessed, other job has been done
-				%Need to find out if this job j is before or after j' in the schedule
-				%\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+	\If{DJ $= \True \Or$DS $= \True$}
 		\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-				%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
-				\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
-					%\State valid $\gets \False$
-					%\Return infeasible
-					\Return feasible $\gets \False$
-				\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
-					\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
-						\Return feasible $\gets \False$
-					\EndIf
-				\EndIf
-			\EndIf % i appears later than i' in O
-		\EndIf %(j, j') in DS
-	\EndIf %if j in J'
+	\EndIf
+%	\If{$j \in J'$}
+%		\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
+%		\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+%		\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+%			\State secondN $\gets \True$
+%		\EndIf
+%		\If{$(j, j') \in DS$}
+%			\State DS $\gets \True$
+%				\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
+%					%\State valid $\gets \False$
+%					%\Return infeasible
+%					\Return feasible $\gets \False$
+%				\EndIf % arrival later than latest start time
+%		\Else 
+%			\State $DJ \gets \True$
+%			%\If{secondN $= \False$}
+%			%	\State original TWs are used
+%			\If{secondN $= \True$} % This nurse is the 2nd one to be assessed, other job has been done
+%				%Need to find out if this job j is before or after j' in the schedule
+%				%\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+%				\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+%				%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
+%				\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
+%					%\State valid $\gets \False$
+%					%\Return infeasible
+%					\Return feasible $\gets \False$
+%				\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
+%					\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
+%						\Return feasible $\gets \False$
+%					\EndIf
+%				\EndIf
+%			\EndIf % i appears later than i' in O
+%		\EndIf %(j, j') in DS
+%	\EndIf %if j in J'
 	%\While{$\Not$ valid}

### 2022-03-23 tex/Paper/main.tex
made changes to match overleaf.

@@
 \label{sub:preferences}
-\noindent It is not uncommon in practice for district nurses to do their planning using preferences.
+\noindent In practice it is often that district nurses take into account preferences during their planning.
 These might arise from multiple reasons, such as a preserving continuity of care, gender preference, patients and nurses speaking the same language, or previous conflicts between a patient and a nurse.
@@
 	\begin{align}
-		a(i,k_0^i) &= 0, \label{eq:recursionarrivalbase}\\[3pt]
+		a(i,k_0^i) &= st(i), \label{eq:recursionarrivalbase}\\[3pt]
 		a(i,j) &= a(i,j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j). \label{eq:arrivaltime}
@@
         %w_{i j} \geq t_j - a(i',j) \quad \forall j \in DS \text{ such that } j \in r_{i} \cap r_{i'}. \label{eq:waiting2DS}
-       \chng{ w_{i j} \geq t_j - a(i',j') \text{ } \forall (j,j') \in DS \text{ such that } j \in r_{i} \text{ and } j' \in r_{i'}.} \label{eq:waiting2DS}
+       w_{i j} \geq t_j - a(i',j') \text{ } \forall (j,j') \in DS \text{ such that } j \in r_{i} \text{ and } j' \in r_{i'}. \label{eq:waiting2DS}
     \end{equation}
@@
 
-\noindent \chng{Furthermore, we assume that waiting times are never larger than strictly necessary.
+\noindent Furthermore, we assume that waiting times are never larger than strictly necessary.
+To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based on the order in which they appear in the solution.
+Note that this implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
+To mitigate this effect, in Section~\ref{sec:dst} we describe strategies to search different orderings of nurses when solving the problem.
+
+%\noindent \chng{Furthermore, we assume that waiting times are never larger than strictly necessary.
 To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based on the order in which they appear in the solution.
@@
 %  $i_1 < i_2$, \ie $j_1$ appears earlier in the vector $\mathcal{O}$ than $j_2$. Then, we enforce that:
-\begin{align}
-	w_{i_1 j_1} &= \max \{0, e_{j_1} - a(i_1,j_1)\}, \\ 
-		t_{j_1} &= \max \{ a(i_1,j_1), e_{j_1} \}.\label{eq:earliestpossiblestartassumption}
-\end{align}
+%\begin{align}
+%	w_{i_1 j_1} &= \max \{0, e_{j_1} - a(i_1,j_1)\}, \\ 
+%		t_{j_1} &= \max \{ a(i_1,j_1), e_{j_1} \}.\label{eq:earliestpossiblestartassumption}
+%\end{align}
 %\begin{equation}\label{eq:earliestpossiblestartassumption1}
@@
 %\end{equation}
-\noindent Note that this implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
+%\noindent Note that this implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
 To mitigate this effect, in Section~\ref{sec:dst} we describe strategies to search different orderings of nurses when solving the problem.
@@
 Given these shifts, we must ensure that jobs are assigned to nurses during their actual working time.
-This also involves considering two cases that may occur: \chng{Need to ref the figures.}
+This also involves considering two cases that may occur:
 \begin{enumerate}[label={(\arabic*)}]
@@
 Tardiness as calculated as above.
-	\chng{\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period.}
-	\chng{\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time.}
+	\item A nurse $i$ arrives at a job $j$ during a shift, but the duration of $j$ results in the end time of $j$ extending into the nurse's break period.
+	\item A nurse $i$'s arrival time at job $j$ is determined to be during the nurse's unavailable time.
 \end{enumerate}
@@
 \end{figure}
-%\correction{What happens if the next "available" slot is not enough for the job.}
-%\noindent In the event that the next available shift for a nurse $i$ is not long enough to perform job $j$, the job is moved forward again in the nurse's schedule.
-\noindent The \textsc{CalculateJobTimes} function in Algorithm \ref{alg:cjt} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
-For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job (lines 3--4).
-Line 6 covers the case where the duration of the current job $j$ extends into a break period, whilst line 12 considers the event when the arrival time of nurse $i$ at $j$ would be within a break period.
-In both of these cases, the function \textsc{FindValidTime} in Algorithm~\ref{alg:fvt} is called, which scans through nurse $i$'s day to find a suitable time in which $i$ can perform $j$ (lines 6 and 12).
-If $j$ cannot be feasibly scheduled into nurse $i$'s day, then no feasible solution exists and the algorithm terminates (lines 7--8 and lines 13--14).
-Otherwise, \textsc{FindValidTime} has successfully placed job $j$ into the schedule of nurse $i$, and returns the start time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, and gap tardiness $\tilde{z}_j$ (if $j \in J'$ and gap tardiness is permitted) of job $j$, as well as the current shift $f$ of nurse $i$, and \textsc{CalculateJobTimes} continues on to the next job $j \in r_i$ (lines 9 and 15).
-There are two scenarios in which \textsc{FindValidTime} does not need to be called: (1) if nurse $i$ can feasibly perform job $j$ with the arrival time $a(i,j)$; and (2) if $f$ is the last shift of nurse $i$, i.e.
-$f = g_i$.
-In both instances, the algorithm simply calculates the starting time $t_j$, waiting time $w_{ij}$, and tardiness $z_j$ (lines 16--19).
-
-%\chng{The \textsc{CalculateJobTimes} function in \ref{alg:shiftwork} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
-For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job.
-Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e.
-$g_i = 1$.
-Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ at job $j$ would be within a break.
-In both of these cases, the function \textsc{FindValidTime} is used.
-This function scans through the nurse's day to find the next available time period in which the nurse can feasibly perform the job $j$, and returns the start time of this period, $t_j$.
-If the job $j$ is a DS, then \textsc{FindValidTime} seeks a start time that is suitable for the current nurse $i$ and the second nurse on the job.}
+
+\noindent Both of these cases are illustrated in Figure~\ref{fig:itemsaligned}.
+The \textsc{CalculateJobTimes} (CJT) function in Algorithm~\ref{alg:cjt} describes how start times, waiting times, and tardiness are calculated for each nurse $i \in N$.
+For each job $j$ in nurse $i$'s route $r_i$, CJT obtains the arrival time $a(i,j)$ of nurse $i$ at job $j$ using the arrival time, waiting time, and service time of the previous job $j-1$ in the nurse's route, as well as the travel time from $j-1$ to the current job $j$ (lines 3--4).
+The function then determines whether the job is a linked service (i.e.
+a dependent job or a double service) and, if so, whether the other nurse $i'$ performing the linked job has already been assessed (lines 6--12).
+Then, CJT evaluates the current nurse-job assignment under four conditions.
+Firstly, if the nurse's day comprises just one shift -- that is, there are no breaks in the nurse's day -- then the function schedules job $j$ into nurse $i$'s day using the current arrival time $a(i,j)$ (lines 13--18).
+Secondly, if the nurse's day does consists of multiple shifts, and the proposed arrival time $a(i,j)$ occurs within a break period, then CJT calls upon the \textsc{FindValidTime} (FVT) function (see Algorithm~\ref{alg:fvt}) to find a suitable start time for job $j$ (lines 19--20).
+Note that as the proposed arrival time is during an invalid time period, the ``arrival time'' $a(i,j)_{\textup{temp}}$ sent to FVT is in fact the end of the current shift, $v(\pi_{f}^{i})$ (line 20).
+FVT is also called in the third case, where although the arrival time of the nurse at the job is during a valid shift, the duration of the job extends into the nurse's break period (lines 22--23).
+However, if the previous conditions do not hold, then nurse $i$ can perform job $j$ feasibly, and CJT schedules job $j$ into nurse $i$'s route (lines 24--29).
+The function returns a full schedule for nurse $i$.
 
 \begin{algorithm}[htbp!]
-\scriptsize
+% \tiny
 \caption{\textsc{CalculateJobTimes}$(i, r_i)$}
@@
 \begin{algorithmic}[1]
-	\State feasible $\gets \False$
+	%\State feasible $\gets \False$
 	\State Let $f$ be the initial shift for nurse $i$
 	\State $a(i,k_{0}^i) \gets 0$
-%	\For{$j \gets 1 \To m_i$}
-	\For{$j \in r_i$}
+	\State $w_{ik_{0}^i} \gets 0$
+	\State $s_{k_{0}^i} \gets 0$
+%	\For{$j \in r_i$}
+	\ForEach{job $j$ in nurse $i$'s route $r_i$}
 		\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
-		\State $i' \gets \Null$, $j' \gets \Null$
-		\State DJ $\gets \False$
-		\State DS $\gets \False$
-		\State secondN $\gets \False$
-		\If{$j \in J'$}
-			\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
-			\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+		\State $i' \gets \Null$, $j' \gets \Null$, DJ $\gets \False$, DS $\gets \False$ %, secondN $\gets \False$
+		\If{$j \in J'$} % D : J' -> DJ, returns other job in pair with job j, NJ : J -> N, returns the nurse assigned to the job j
+			\State $j' \gets D(j)$, $i' \gets NJ(j')$
 			\If{$i$ appears later than $i'$ in $\mathcal{O}$}
-				\State secondN $\gets \True$
-			\EndIf
 				\If{$(j, j') \in DS$}
 					\State DS $\gets \True$
-				\If{secondN $= \True \AAnd a(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
-					\State feasible $\gets \False$
-					\Break
-				\EndIf % arrival later than latest start time
-			\Else 
-				\State $DJ \gets \True$
-				\If{secondN $= \True$} % This nurse is 2nd one to be assessed, other job has been done
-				%Need to find out if this job j is before or after j' in the schedule
-					\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-					\If{$a(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
-						\State feasible $\gets \False$
-						\Break
-					\ElsIf{$a(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
-						\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
-							\State feasible $\gets \False$
-							\Break
-						\EndIf
-					\EndIf
-				\EndIf % i appears later than i' in O
-			\EndIf %(j, j') in DS
+				\Else \hspace{0.2mm} DJ $\gets \True$
+				\EndIf
+%				\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+				\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+			\EndIf
 		\EndIf %if j in J'
-		\If{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
-			\State $v \gets v(\pi_{f}^{i})$
-			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, v, i', j'$, DS, DJ, secondN)
-			\If{feasible $= \False$}
-				\Break
-			\Else \hspace{0.2mm} \textbf{continue}
-				%\Continue
-			\EndIf
+		% 1. if nurse only has one shift
+		\If{nurse $i$ only has one shift, $g_i = 1$}
+			\State $w_{ij} \gets \max\{0, (e_j \chng{e_j^{(j')}} - a(i,j))\}$
+			\State $t_j \gets a(i,j) + w_{ij}$
+			\State $z_j \gets \max\{0, (t_j - l_j)\}$
+			\If{DJ $= \True \Or$DS $= \True$}
+				\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
+			\EndIf
+		% 2. Else nurse has multiple shifts, arrival is between shifts
+		\ElsIf{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
+			\State $a(i,j)_{\textup{temp}}$ $\gets v(\pi_{f}^{i})$
+			\State $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, a(i,j)_{\textup{temp}}, i', j'$, DS, DJ)
+		% 3. Else arrival is in valid shift but job extends into break
 		\ElsIf{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
-			\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN)
-			\If{feasible $= \False$}
-				\Break
-			\Else \hspace{0.2mm} \textbf{continue}
-				%\Continue
-			\EndIf
-		\Else
-			\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
-			\State $z_j \gets \max\{0, (a(i,j) - l_j)\}$
+			\State $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ)
+		% 4. Job can be performed feasibly, or it is the final shift of nursei 
+		\Else
+			\State $w_{ij} \gets \max\{0, (e_j \chng{e_j^{(j')}} - a(i,j))\}$
 			\State $t_j \gets a(i,j) + w_{ij}$
+			\State $z_j \gets \max\{0, (t_j - l_j)\}$
+			\If{DJ $= \True \Or$DS $= \True$}
+				\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
+			\EndIf
 		\EndIf
 	\EndFor
-	\Return either statement of infeasibility or feasible schedule for nurse $i$
+	\Return feasible schedule for nurse $i$
 \end{algorithmic}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%Explanation of FindValidTime function
+The \textsc{FindValidTime} (FVT) function shown in Algorithm~\ref{alg:fvt} is utilised when a nurse's day consists of multiple shifts.
+The main while loop of FVT involves progressing through each shift $\pi_{\tilde{f}}^i \in \Pi_i$ for the nurse $i$, and setting the current time $\tilde{a}(i,j)$ to the start of each shift $u(\pi_{\tilde{f}}^i)$ (line 7).
+Each loop comprises two parts, depending on whether or not the job $j$ is a linked service.
+In either case, if the duration of the job extends into the nurse's next break period, the function continues onto the next shift $\tilde{f}+1$, provided $\tilde{f}$ is not the last shift of the nurse's day (lines 9 and 18).
+Otherwise, a valid start time has been found for job $j$ (lines 10--16 and 19--25); FVT returns the starting time, waiting time, tardiness and current shift for job $j$ back to the CJT function (line 26).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{algorithm}[htbp!]
-\scriptsize
-\caption{\textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN, gap)}
+% \scriptsize
+\caption{\textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ)}
 \label{alg:fvt}
@@
 	\If{DJ $= \True \Or$DS $= \True$}
-		\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-	\EndIf
-%	\If{$j \in J'$}
-%		\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
-%		\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
-%		\If{$i$ appears later than $i'$ in $\mathcal{O}$}
-%			\State secondN $\gets \True$
+%		\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+		\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+	\EndIf
+	\While{$\Not$ feasible}
+		\State $\tilde{f} \gets \tilde{f}+1$, $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
+		% A. nurse i is the second nurse being assessed, i.e. job is DJ or DS.
+%		\If{secondN $= \True$}
+		\If{DJ $= \True \Or$DS $= \True$}
+			% 1a. currentTime is LATER than latest allowed start time
+%			\If{$\tilde{a}(i,j) > l_j^{(j')}$}
+%				\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+%					\Continue
+%				\Else
+%					\State feasible $\gets \True$
+%					\State $t_j \gets \tilde{a}(i,j)$
+%					\State $w_{ij} \gets \max\{0, (t_j - a(i,j))\}$
+%					\State $z_j \gets \max\{0, (t_j - l_j)\}$
+%					\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
+%					\State $f \gets \tilde{f}$
 %				\EndIf
-%		\If{$(j, j') \in DS$}
-%			\State DS $\gets \True$
-%				\If{secondN $= \True \AAnd \tilde{a}(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
-%					%\State valid $\gets \False$
-%					%\Return infeasible
-%					\Return feasible $\gets \False$
-%				\EndIf % arrival later than latest start time
-%		\Else 
-%			\State $DJ \gets \True$
-%			%\If{secondN $= \False$}
-%			%	\State original TWs are used
-%			\If{secondN $= \True$} % This nurse is the 2nd one to be assessed, other job has been done
-%				%Need to find out if this job j is before or after j' in the schedule
-%				%\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-%				\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-%				%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
-%				\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
-%					%\State valid $\gets \False$
-%					%\Return infeasible
-%					\Return feasible $\gets \False$
-%				\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
-%					\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
-%						\Return feasible $\gets \False$
-%					\EndIf
-%				\EndIf
-%			\EndIf % i appears later than i' in O
-%		\EndIf %(j, j') in DS
-%	\EndIf %if j in J'
-	%\While{$\Not$ valid}
-	\While{$\Not$ feasible}
-		\State $\tilde{f} \gets \tilde{f}+1$
-		\State $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
-		%\If{secondN $= \True \AAnd$(DJ $= \True \Or$DS $= \True$)}
-		\If{secondN $= \True$}
-			%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
-			\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$}
-				%\State valid $\gets \False$
-				%\Break
-				%\Return infeasible
-				\Return feasible $\gets \False$
-			%\ElsIf{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \True$}
-			\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \True$}
-				\If{$t_{j'} + l_j^{(j')} > t_{j'}$} % job j occurs after job j', gap tardiness allowed
-					\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+			% 2a. job cannot be performed feasibly in shift, extends into break
+%			\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+			\If{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$} %\hspace{0.2mm}\textbf{continue}
 				\Continue
-					\Else
-						%\State valid $\gets \True$
+			% 3a. nurse can perform job feasibly in shift OR shift \tilde{f} is the last shift of the day
+			\Else
 				\State feasible $\gets \True$
 				\State $t_j \gets \tilde{a}(i,j)$
-						\State $w_{ij} \gets t_j - a(i,j)$
-						\State $z_j \gets t_j - l_j$
-						%\State $\tilde{z}_j \gets t_j - l_j^{(j')}$
-						\State $\tilde{z}_j \gets t_j - (t_{j'} + l_j^{(j')})$
+				\State $w_{ij} \gets \max\{0, (t_j - a(i,j))\}$
+				\State $z_j \gets \max\{0, (t_j - l_j)\}$
+				\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
 				\State $f \gets \tilde{f}$
 			\EndIf
-				\Else \hspace{0.2mm} \textbf{return} feasible $\gets \False$ % gap tardiness allowed but job is before, so gap is smaller
-					%\State valid $\gets \False$
-					%\Break
-					%\Return infeasible
-				\EndIf
-			%\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}
-			\ElsIf{$\max\{\tilde{a}(i,j), (t_{j'} + e_j^{(j')})\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}
+		% End A. If secondN = True
+		% B. nurse i is the first nurse assessing DJ/DS OR job is not a DJ/DS
+		\Else
+			% 1b. job cannot be performed feasibly in shift, extends into break
+			\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$} %\hspace{0.2mm}\textbf{continue}
 				\Continue
-			\Else
-				%\State valid $\gets \True$
+			% 2b. nurse can perform job feasibly in shift OR \tilde{f} is the last shift of the day
+			\Else
 				\State feasible $\gets \True$
 				\State $t_j \gets \tilde{a}(i,j)$
-				\State $w_{ij} \gets t_j - a(i,j)$
-				\State $z_j \gets t_j - l_j$
+				\State $w_{ij} \gets \max\{0, (t_j - a(i,j))\}$
+				\State $z_j \gets \max\{0, (t_j - l_j)\}$
+				\State $\tilde{z}_j \gets 0$
 				\State $f \gets \tilde{f}$
 			\EndIf
-		\ElsIf{secondN $= \False$}
-			\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
-				\Continue
-			\Else
-				%\State valid $\gets \True$
-				\State feasible $\gets \True$
-				\State $t_j \gets \tilde{a}(i,j)$
-				\State $w_{ij} \gets t_j - a(i,j)$
-				\State $z_j \gets t_j - l_j$
-				\State $f \gets \tilde{f}$
-			\EndIf
 		\EndIf
 	\EndWhile
-	\Return statement of feasibility, starting time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, gap tardiness $\tilde{z}_j$, and shift $f$
+	\Return starting time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, gap tardiness $\tilde{z}_j$, and shift $f$
 \end{algorithmic}
@@
 
-%Explanation of FindValidTime function
-The \textsc{FindValidTime} function shown in Algorithm~\ref{alg:fvt} operates as follows: firstly, if $j$ is a dependent job or a double service, the algorithm obtains the paired job $j'$ and the nurse $i'$ assigned to $j'$ (lines 4--6).
-If the schedule for $i'$ has already been set, then $i$ is considered to be the ``second nurse'' (line 7).
-Then, if $j$ is a double service job and the current time $\tilde{a}(i,j)$ is after the start time set for $j'$, no feasible solution can be found and the algorithm terminates (lines 9--12).
-Likewise, if $j$ is a dependent job, and the current time $\tilde{a}(i,j)$ occurs after the latest allowed start time and no gap tardiness is permitted, the algorithm terminates with a statement of infeasibility (lines 13--18).
-
-The main while loop of \textsc{FindValidTime} involves progressing through each shift $\pi_{\tilde{f}}^i \in \Pi_i$ and setting the current time $\tilde{a}(i,j)$ to the start of each shift, $u_(\pi_{\tilde{f}}^{i})$ (lines 20--21).
-Each loop comprises two parts.
-If $i$ is the second nurse assessed, this implies that $j$ is a dependent job.
-The same condition from line 17 is checked, where $l_{j}^{(j')} = 0$ if $j$ is a double service, and again the algorithm terminates if the condition is not met.
-However, if gap tardiness is permitted, this implies that $j$ is not a double service (i.e.
-DJ $= \True$).
-The algorithm evaluates the order of the jobs $j$ and $j'$ and determines whether the job can be performed feasibly (lines 25--36).
-Otherwise, $\tilde{a}(i,j)$ occurs before the latest start time, and the usual checks are performed (lines 37--44).
-Finally, the same assessments occur if $j$ is a paired job and $i$ is the first nurse to be evaluated, or if $j$ is a single job (lines 45--53).
-
 For simplicity, we will consider all nurses' break periods as waiting time, so that the arrival time calculations $a(i,j)$ for nurses $i \in N$ at jobs $j \in J$ (see \eqref{eq:arrivaltime}) remain valid.
 Then, the waiting times will be corrected by removing all relevant breaks, as shown in \eqref{eq:waitingtimenurse}.
-
-%There may be cases where some nurses have multiple shifts in a single day, e.g.
-a morning shift and an evening shift; these nurses will not be available at all times during the day.
-Therefore, we must ensure that jobs are assigned to nurses during their actual working times.
-This also involves considering the case where although a nurse may be working and available at the time that a job starts, the duration of the job may extend into the nurse's unavailable time, i.e.
-the job ends after the nurse's shift ends.
-In these instances, the job must be moved forward in the nurse's schedule such that the job starts at the beginning of the nurse's next available shift, and tardiness will be incurred.
-For convenience, we denote by $ct(i)$ the actual contracted working time of nurse $i \in N$, where $ct(i) = en(i) - st(i)$ if nurse $i$'s day comprises a single shift starting at $en(i)$ and ending at $st(i)$, and $ct(i) < en(i) - st(i)$ if nurse $i$'s day consists of one or more shifts.
 
@@
 
+\begin{algorithm}[htbp!]
+\scriptsize
+\caption{\textsc{CalculateJobTimes}$(i, r_i)$}
+\label{alg:cjtprev}
+\begin{algorithmic}[1]
+\State feasible $\gets \False$
+\State Let $f$ be the initial shift for nurse $i$
+\State $a(i,k_{0}^i) \gets 0$
+%	\For{$j \gets 1 \To m_i$}
+\For{$j \in r_i$}
+\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+\State $i' \gets \Null$, $j' \gets \Null$
+\State DJ $\gets \False$
+\State DS $\gets \False$
+\State secondN $\gets \False$
+\If{$j \in J'$}
+\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
+\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+\State secondN $\gets \True$
+\EndIf
+\If{$(j, j') \in DS$}
+\State DS $\gets \True$
+\If{secondN $= \True \AAnd a(i,j) > t_{j'}$} %can also use l_j^(j'), latest ST allowed for j based on j'
+\State feasible $\gets \False$
+\Break
+\EndIf % arrival later than latest start time
+\Else 
+\State $DJ \gets \True$
+\If{secondN $= \True$} % This nurse is 2nd one to be assessed, other job has been done
+%Need to find out if this job j is before or after j' in the schedule
+\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+\If{$a(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$} % current time is later than latest allowed start based on start time of j' and no gap tard allowed
+\State feasible $\gets \False$
+\Break
+\ElsIf{$a(i,j) > t_{j'} + l_{j}^{(j')} \AAnd$gap $= \True$}
+\If{$t_{j'} + l_{j}^{(j')} < t_{j'}$} %job j occurs before job j', gap too small
+\State feasible $\gets \False$
+\Break
+\EndIf
+\EndIf
+\EndIf
+\EndIf %(j, j') in DS
+\EndIf %if j in J'
+\If{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
+\State $v \gets v(\pi_{f}^{i})$
+\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, v, i', j'$, DS, DJ, secondN)
+\If{feasible $= \False$}
+\Break
+\Else \hspace{0.2mm} \textbf{continue}
+%\Continue
+\EndIf
+\ElsIf{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
+\State feasible, $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN)
+\If{feasible $= \False$}
+\Break
+\Else \hspace{0.2mm} \textbf{continue}
+%\Continue
+\EndIf
+\Else
+\State $w_{ij} \gets \max\{0, (e_j-a(i,j))\}$
+\State $z_j \gets \max\{0, (a(i,j) - l_j)\}$
+\State $t_j \gets a(i,j) + w_{ij}$
+\EndIf
+\EndFor
+\Return either statement of infeasibility or feasible schedule for nurse $i$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[htbp!]
+\scriptsize
+\caption{\textsc{CalculateJobTimes}$(i, r_i)$}
+\label{alg:cjt08032022}
+\begin{algorithmic}[1]
+\State feasible $\gets \False$
+\State Let $f$ be the initial shift for nurse $i$
+\State $a(i,k_{0}^i) \gets 0$
+\For{$j \in r_i$}
+\State $a(i,j) \gets a(i, j-1) + w_{ij-1} + s_{j-1} + T_i(j-1, j)$
+\State $i' \gets \Null$, $j' \gets \Null$
+\State DJ $\gets \False$, DS $\gets \False$, secondN $\gets \False$
+\If{$j \in J'$}
+\State $j' \gets D(j)$ % D : J' -> DJ, returns other job in pair with job j
+\State $i' \gets NJ(j')$ % NJ : J -> N, returns the nurse assigned to the job j
+\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+\State secondN $\gets \True$
+\EndIf
+\If{$(j, j') \in DS$}
+\State DS $\gets \True$
+\Else 
+\State DJ $\gets \True$
+\If{secondN $= \True$} % This nurse is 2nd one to be assessed, other job has been done
+%Need to find out if this job j is before or after j' in the schedule
+\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+\EndIf
+\EndIf %(j, j') in DS
+\EndIf %if j in J'
+% 1. if nurse only has one shift
+\If{nurse $i$ only has one shift, $g_i = 1$}
+\State $w_{ij} \gets \max\{0, (e_j \chng{e_j^{(j')}} - a(i,j))\}$
+\State $t_j \gets a(i,j) + w_{ij}$
+\State $z_j \gets \max\{0, (t_j - l_j)\}$
+\If{DJ $= \True$}
+\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
+\EndIf
+% 2. Else nurse has multiple shifts, arrival is between shifts
+\ElsIf{$a(i,j) > v(\pi_{f}^{i}) \AAnd a(i,j) < u(\pi_{f+1}^{i}) \AAnd f \neq g_i$}
+\State $v \gets v(\pi_{f}^{i})$
+\State $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, v, i', j'$, DS, DJ, secondN)
+% 3. Else arrival is in valid shift but job extends into break
+\ElsIf{$\max\{a(i,j), e_j\} + s_j > v(\pi_{f}^{i}) \AAnd f \neq g_i$}
+\State $t_j, w_{ij}, z_j, \tilde{z}_j, f \gets$ \textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN)
+% 4. Job can be performed feasibly, or it is the final shift of nursei 
+\Else
+\State $w_{ij} \gets \max\{0, (e_j \chng{e_j^{(j')}} - a(i,j))\}$
+\State $t_j \gets a(i,j) + w_{ij}$
+\State $z_j \gets \max\{0, (t_j - l_j)\}$
+\If{DJ $= \True$}
+\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
+\EndIf
+\EndIf
+\EndFor
+\Return feasible schedule for nurse $i$
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}[htbp!]
+\scriptsize
+\caption{\textsc{FindValidTime}($i, j, f, a(i,j), i', j'$, DS, DJ, secondN, gap)}
+\label{alg:fvtprev}
+\begin{algorithmic}[1]
+\State feasible $\gets \False$
+\State $t_j \gets \Null$, $w_{ij} \gets \Null$, $z_j \gets \Null$, $\tilde{z}_j \gets \Null$
+\State $\tilde{f} \gets f$, $\tilde{a}(i,j) \gets a(i,j)$
+\If{DJ $= \True \Or$DS $= \True$}
+\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+\EndIf
+%\While{$\Not$ valid}
+\While{$\Not$ feasible}
+\State $\tilde{f} \gets \tilde{f}+1$
+\State $\tilde{a}(i,j) \gets u(\pi_{\tilde{f}}^{i})$
+%\If{secondN $= \True \AAnd$(DJ $= \True \Or$DS $= \True$)}
+\If{secondN $= \True$}
+%\If{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \False$}
+\If{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \False$}
+%\State valid $\gets \False$
+%\Break
+%\Return infeasible
+\Return feasible $\gets \False$
+%\ElsIf{$\tilde{a}(i,j) > l_j^{(j')} \AAnd$gap $= \True$}
+\ElsIf{$\tilde{a}(i,j) > t_{j'} + l_j^{(j')} \AAnd$gap $= \True$}
+\If{$t_{j'} + l_j^{(j')} > t_{j'}$} % job j occurs after job j', gap tardiness allowed
+\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+%\State valid $\gets \True$
+\State feasible $\gets \True$
+\State $t_j \gets \tilde{a}(i,j)$
+\State $w_{ij} \gets t_j - a(i,j)$
+\State $z_j \gets t_j - l_j$
+%\State $\tilde{z}_j \gets t_j - l_j^{(j')}$
+\State $\tilde{z}_j \gets t_j - (t_{j'} + l_j^{(j')})$
+\State $f \gets \tilde{f}$
+\EndIf
+\Else \hspace{0.2mm} \textbf{return} feasible $\gets \False$ % gap tardiness allowed but job is before, so gap is smaller
+%\State valid $\gets \False$
+%\Break
+%\Return infeasible
+\EndIf
+%\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}
+\ElsIf{$\max\{\tilde{a}(i,j), (t_{j'} + e_j^{(j')})\} + s_j > v(\pi_{\tilde{f}}^i) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+%\State valid $\gets \True$
+\State feasible $\gets \True$
+\State $t_j \gets \tilde{a}(i,j)$
+\State $w_{ij} \gets t_j - a(i,j)$
+\State $z_j \gets t_j - l_j$
+\State $f \gets \tilde{f}$
+\EndIf
+\ElsIf{secondN $= \False$}
+\If{$\tilde{a}(i,j) + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
+\Continue
+\Else
+%\State valid $\gets \True$
+\State feasible $\gets \True$
+\State $t_j \gets \tilde{a}(i,j)$
+\State $w_{ij} \gets t_j - a(i,j)$
+\State $z_j \gets t_j - l_j$
+\State $f \gets \tilde{f}$
+\EndIf
+\EndIf
+\EndWhile
+\Return statement of feasibility, starting time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, gap tardiness $\tilde{z}_j$, and shift $f$
+\end{algorithmic}
+\end{algorithm}
+
 \end{comment}
@@
 
+%\If{$i$ appears later than $i'$ in $\mathcal{O}$}
+%\State secondN $\gets \True$
+%\EndIf
+%\If{$(j, j') \in DS$}
+%\State DS $\gets \True$
+%\Else 
+%\State DJ $\gets \True$
+%\If{secondN $= \True$} % This nurse is 2nd one to be assessed, other job has been done
+%%Need to find out if this job j is before or after j' in the schedule
+%\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+%\EndIf
+%\EndIf %(j, j') in DS
+
+%\noindent The \textsc{CalculateJobTimes} function in Algorithm \ref{alg:cjt} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
+For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job (lines 3--4).
+Line 6 covers the case where the duration of the current job $j$ extends into a break period, whilst line 12 considers the event when the arrival time of nurse $i$ at $j$ would be within a break period.
+In both of these cases, the function \textsc{FindValidTime} in Algorithm~\ref{alg:fvt} is called, which scans through nurse $i$'s day to find a suitable time in which $i$ can perform $j$ (lines 6 and 12).
+If $j$ cannot be feasibly scheduled into nurse $i$'s day, then no feasible solution exists and the algorithm terminates (lines 7--8 and lines 13--14).
+Otherwise, \textsc{FindValidTime} has successfully placed job $j$ into the schedule of nurse $i$, and returns the start time $t_j$, waiting time $w_{ij}$, tardiness $z_j$, and gap tardiness $\tilde{z}_j$ (if $j \in J'$ and gap tardiness is permitted) of job $j$, as well as the current shift $f$ of nurse $i$, and \textsc{CalculateJobTimes} continues on to the next job $j \in r_i$ (lines 9 and 15).
+There are two scenarios in which \textsc{FindValidTime} does not need to be called: (1) if nurse $i$ can feasibly perform job $j$ with the arrival time $a(i,j)$; and (2) if $f$ is the last shift of nurse $i$, i.e.
+$f = g_i$.
+In both instances, the algorithm simply calculates the starting time $t_j$, waiting time $w_{ij}$, and tardiness $z_j$ (lines 16--19).
+
+%\chng{The \textsc{CalculateJobTimes} function in \ref{alg:shiftwork} describes how start times, waiting times and tardiness are calculated for each nurse $i \in N$.
+For each job $j \in r_i$, the arrival time is evaluated using the arrival time, waiting time, service time and travel time of the previous job.
+Line 1 covers the case where nurse $i$'s day comprises just one shift, i.e.
+$g_i = 1$.
+Line 9 is for the case where the duration of the current job $j$ extends in to a break period, whilst Line 14 considers the event that the arrival time of nurse $i$ at job $j$ would be within a break.
+In both of these cases, the function \textsc{FindValidTime} is used.
+This function scans through the nurse's day to find the next available time period in which the nurse can feasibly perform the job $j$, and returns the start time of this period, $t_j$.
+If the job $j$ is a DS, then \textsc{FindValidTime} seeks a start time that is suitable for the current nurse $i$ and the second nurse on the job.}
+
+%Explanation of FindValidTime function
+%The \textsc{FindValidTime} function shown in Algorithm~\ref{alg:fvt} operates as follows: firstly, if $j$ is a dependent job or a double service, the algorithm obtains the paired job $j'$ and the nurse $i'$ assigned to $j'$ (lines 4--6).
+If the schedule for $i'$ has already been set, then $i$ is considered to be the ``second nurse'' (line 7).
+Then, if $j$ is a double service job and the current time $\tilde{a}(i,j)$ is after the start time set for $j'$, no feasible solution can be found and the algorithm terminates (lines 9--12).
+Likewise, if $j$ is a dependent job, and the current time $\tilde{a}(i,j)$ occurs after the latest allowed start time and no gap tardiness is permitted, the algorithm terminates with a statement of infeasibility (lines 13--18).
+
+%The main while loop of \textsc{FindValidTime} involves progressing through each shift $\pi_{\tilde{f}}^i \in \Pi_i$ and setting the current time $\tilde{a}(i,j)$ to the start of each shift, $u_(\pi_{\tilde{f}}^{i})$ (lines 20--21).
+Each loop comprises two parts.
+If $i$ is the second nurse assessed, this implies that $j$ is a dependent job.
+The same condition from line 17 is checked, where $l_{j}^{(j')} = 0$ if $j$ is a double service, and again the algorithm terminates if the condition is not met.
+However, if gap tardiness is permitted, this implies that $j$ is not a double service (i.e.
+DJ $= \True$).
+The algorithm evaluates the order of the jobs $j$ and $j'$ and determines whether the job can be performed feasibly (lines 25--36).
+Otherwise, $\tilde{a}(i,j)$ occurs before the latest start time, and the usual checks are performed (lines 37--44).
+Finally, the same assessments occur if $j$ is a paired job and $i$ is the first nurse to be evaluated, or if $j$ is a single job (lines 45--53).
+
+%There may be cases where some nurses have multiple shifts in a single day, e.g.
+a morning shift and an evening shift; these nurses will not be available at all times during the day.
+Therefore, we must ensure that jobs are assigned to nurses during their actual working times.
+This also involves considering the case where although a nurse may be working and available at the time that a job starts, the duration of the job may extend into the nurse's unavailable time, i.e.
+the job ends after the nurse's shift ends.
+In these instances, the job must be moved forward in the nurse's schedule such that the job starts at the beginning of the nurse's next available shift, and tardiness will be incurred.
+For convenience, we denote by $ct(i)$ the actual contracted working time of nurse $i \in N$, where $ct(i) = en(i) - st(i)$ if nurse $i$'s day comprises a single shift starting at $en(i)$ and ending at $st(i)$, and $ct(i) < en(i) - st(i)$ if nurse $i$'s day consists of one or more shifts.
+
 \end{document}

### 2022-08-28 tex/Paper/main.tex
new commit

@@
 
+%This can occur in practice, for example, in situations where a second dose of medication must be administered a few hours after the first dose.
+If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$ that depends on the start time of $j_1$, $[t_{j_1} + e_{j_2}^{(j_1)}, t_{j_1} + l_{j_2}^{(j_1)}]$.
+Then, the same time window constraints hold:
 \noindent Note that tardiness here refers only to the difference between the job start and its original time window.
 Jobs that have dependencies or require two nurses may be further restricted temporally.
-This can occur in practice, for example, in situations where a second dose of medication must be administered a few hours after the first dose.
-If two jobs $j_1$ and $j_2$ are dependent, this is modelled by imposing a second time window on $j_2$ that depends on the start time of $j_1$, $[t_{j_1} + e_{j_2}^{(j_1)}, t_{j_1} + l_{j_2}^{(j_1)}]$.
+This can occur in practice, for example, in situations where a set period of time is required between two separate doses of medication administered to a patient.
+Let $j_1$ and $j_2$ be two dependent jobs.
+Then, we model the additional temporal restriction by imposing a second time window of $j_2$ that depends on the start time of $j_1$, $[t_{j_1} + e_{j_2}^{(j_1)}, t_{j_1} + l_{j_2}^{(j_1)}]$, where $e_{j_2}^{(j_1)}$ and $l_{j_2}^{(j_1)}$ are the minimum and maximum time gaps from the start time of $j_1$ to the start time of $j_2$.
 Then, the same time window constraints hold:
@@
 To achieve this, we consider that the starting times of jobs are fixed to their earliest possible time, based on the order in which they appear in the solution.
-Note that this implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
+Note that for dependent jobs $(j_1, j_2) \in DJ$, the first nurse that is scheduled to do one of the two jobs sets the start time for that job, and therefore the other job acquires the additional time window.
+This implies that the order in which nurse routes appear in a solution may have an impact on the solution quality.
 To mitigate this effect, in Section~\ref{sec:dst} we describe strategies to search different orderings of nurses when solving the problem.
@@
 				\EndIf
-%				\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-				\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+				\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+%				\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
 			\EndIf
@@
 		\If{nurse $i$ only has one shift, $g_i = 1$}
-			\State $w_{ij} \gets \max\{0, (e_j \chng{e_j^{(j')}} - a(i,j))\}$
+			\If{DJ $= \True \Or$DS $= \True$}			
+				\State $w_{ij} \gets \max\{0, ((t_{j'} + e_j^{(j')}) - a(i,j))\}$
 				\State $t_j \gets a(i,j) + w_{ij}$
+				\State $\tilde{z}_j \gets \max\{0, (t_j - (t_{j'} + l_j^{(j')}))\}$
+			\Else
+				\State $w_{ij} \gets \max\{0, (e_j - a(i,j))\}$
+				\State $t_j \gets a(i,j) + w_{ij}$
+			\EndIf
 			\State $z_j \gets \max\{0, (t_j - l_j)\}$
-			\If{DJ $= \True \Or$DS $= \True$}
-				\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
-			\EndIf
 		% 2. Else nurse has multiple shifts, arrival is between shifts
@@
 		\Else
-			\State $w_{ij} \gets \max\{0, (e_j \chng{e_j^{(j')}} - a(i,j))\}$
+			\If{DJ $= \True \Or$DS $= \True$}			
+				\State $w_{ij} \gets \max\{0, ((t_{j'} + e_j^{(j')}) - a(i,j))\}$
 				\State $t_j \gets a(i,j) + w_{ij}$
+				\State $\tilde{z}_j \gets \max\{0, (t_j - (t_{j'} + l_j^{(j')}))\}$
+			\Else
+				\State $w_{ij} \gets \max\{0, (e_j - a(i,j))\}$
+				\State $t_j \gets a(i,j) + w_{ij}$
+			\EndIf
 			\State $z_j \gets \max\{0, (t_j - l_j)\}$
-			\If{DJ $= \True \Or$DS $= \True$}
-				\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
-			\EndIf
 		\EndIf
@@
 	\If{DJ $= \True \Or$DS $= \True$}
-%		\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
-		\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+		\State $[t_{j'} + e_j^{(j')}, t_{j'} + l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
+%		\State $[e_j^{(j')}, l_j^{(j')}] \gets$ dependent time window, based on $t_{j'}$ set by $i'$
 	\EndIf
@@
 %			\ElsIf{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$}
-			\If{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$} %\hspace{0.2mm}\textbf{continue}
+%			\If{$\max\{\tilde{a}(i,j), e_j^{(j')}\} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$} %\hspace{0.2mm}\textbf{continue}
+			\If{$\max\{\tilde{a}(i,j), (t_{j'} + e_j^{(j')})\} + s_j > v(\pi_{\tilde{f}}^{i}) \AAnd \tilde{f} \neq g_i$} %\hspace{0.2mm}\textbf{continue}
 				\Continue
@@
 				\State $z_j \gets \max\{0, (t_j - l_j)\}$
-				\State $\tilde{z}_j \gets \max\{0, (t_j - l_j^{(j')})\}$
+				\State $\tilde{z}_j \gets \max\{0, (t_j - (t_{j'} + l_j^{(j')}))\}$
 				\State $f \gets \tilde{f}$

### 2022-08-28 code/Carrot/constructive_wrapper.py

@@ -21,20 +21,20 @@ import tools_and_scripts.week_counter as wc
 def main():
     
     # Get user input variables using tkinter entry box:
-    # area, max_time_seconds, tw_interval, wb_coeff, maxtravel_coeff, maxwaiting_coeff, quality_measure, create_html_website, create_python_plots, codepoint_directory, input_filename, date_selected = tui.get_user_input_variables()
-    area = 'Hampshire'
-    tw_interval = 15
-    wb_balance = 1
-    wb_coeff = -1
-    maxtravel_coeff = -1
-    maxwaiting_coeff = -1
-    quality_measure = 'default' # default = paper
-    max_time_seconds = 60
-    create_html_website = True
-    create_python_plots = True
-    input_filename = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/Carrot/data/abicare/2021.07.15 - UOS - Clients - Carers - Rota - Anonymised.xlsx'
-    codepoint_directory = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/Carrot/data/codepo_gb'
-    date_selected = pd.Timestamp("2021-07-27")
+    area, max_time_seconds, tw_interval, wb_coeff, maxtravel_coeff, maxwaiting_coeff, quality_measure, create_html_website, create_python_plots, codepoint_directory, input_filename, date_selected = tui.get_user_input_variables()
+    # area = 'Hampshire'
+    # tw_interval = 15
+    # wb_balance = 1
+    # wb_coeff = -1
+    # maxtravel_coeff = -1
+    # maxwaiting_coeff = -1
+    # quality_measure = 'default' # default = paper
+    # max_time_seconds = 60
+    # create_html_website = True
+    # create_python_plots = True
+    # input_filename = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/Carrot/data/abicare/2021.07.15 - UOS - Clients - Carers - Rota - Anonymised.xlsx'
+    # codepoint_directory = r'C:/Users/ah4c20/Asyl/PostDoc/SOCIALCARE/code/Carrot/data/codepo_gb'
+    # date_selected = pd.Timestamp("2021-07-27")
 
     # Convert date_selected from datetime_date to Timestamp
     planning_date = pd.Timestamp(date_selected)

### 2022-08-28 code/Carrot/src/GRASP_VNS/constructive.h

@@ -50,15 +50,14 @@ struct INSTANCE {
 	double objWBQuality;
 	double objPaperQuality;
 	double objQuality;
-    double objMaxTravelTime; // NB new 16/11/2021
-    double objMaxWaitingTime; // NB new 16/11/2021
-    double objMaxDiffWorkload; // NB new 16/11/2021
+    double objMaxTravelTime;
+    double objMaxWaitingTime;
+    double objMaxDiffWorkload;
 	float totalPref; // Total preference score
 	int qualityMeasure; // Temp. Allow to choose (Mk, Ait h., etc.) Ideally we want a vector with weights! This should be set ONLY in "InstanceFromPython" (read_instance.c). The value is hardcoded at the moment.
 	float MAX_TIME_SECONDS; // Maximum time limit in seconds
 	int twInterval; // Time window interval
 	bool excludeNurseTravel; // True if excluding the nurse travel time from home to the first job when updating current time in 'set_nurse_time' function, else false.
-	// double **od_cost;
     int*** unavailMatrix; // 50 X 4 X nNurses 3d matrix, col[0] = unavailable shift number, col[1] = start of unavailable time, col[2] = end of unavailable time, col[3] = duration of unavailable time.
     int* nurseUnavail; // 1D array, size = nNurses, for each nurse contains the number of unavailable shifts.
 	int** nurseWorkingTimes; // UPDATED: 2D array, size = nNurses x 5. For each nurse i (row): column[0] = start time, column[1] = finish time, column[2] = length of day for nurse, 
@@ -92,11 +91,6 @@ struct INSTANCE {
     double totalServiceTime; // New 06/11/2021, moving total service time here as it should be fixed rather than calculating it every time in Objective function
     double totalServiceTimeIncDS; // New 06/11/2021, totalServiceTime but also including double time for the double services; this will therefore be the total working time of all nurses.
     double** arrivalTimes; //NEW 28/02/2022, size = nNurses x nJobs, keeps the original arrival time of nurse i at job j.
-	// int startTime;
-	// int ** jobTimeWindow;
-	// int * jobDuration;
-	// int **nurseSkilled;
-
 };
 
[…]

### 2022-08-28 code/Carrot/src/GRASP_VNS/grasp.c

@@ -13,11 +13,6 @@ UoS
 
 int GRASP(struct INSTANCE* ip) {
 
-	// PR01 Perform path relinking between every generated solution and a random solution from the pool.
-	// PR02 Perform path relinking between some new solutions (above a threshold) and all solutions in the pool.
-	// PR03 Perform path relinking at the end of the algorithm, between all pairs of solutions in the pool.
-	// PR04 Do not perform path relinking at all.
-	
 	clock_t start = clock();
 	clock_t end = clock();
 	double eps = 1e-6;
@@ -25,8 +20,8 @@ int GRASP(struct INSTANCE* ip) {
 	double average_gen_quality = 0.0;
 	long double sum_of_qualities = 0.0;
 	int given_quality = ip->qualityMeasure;
-	int solutions_in_pool = (int)(ip->algorithmOptions[8] + eps); // Allocate space for a pool of solutions (aO[8] = 10)
-	int rcl_strategy = (int)(ip->algorithmOptions[10] + eps); //Type of strategy for cut-off value in RCL (used in RandomisedConstructive). (aO[10] = 1 (strategy 1 C=B-delta*(B-W)))
+	int solutions_in_pool = (int)(ip->algorithmOptions[8] + eps);
+	int rcl_strategy = (int)(ip->algorithmOptions[10] + eps);
 
 	double* pool_quality = malloc(solutions_in_pool * sizeof(double));
 	struct INSTANCE** pool = (struct INSTANCE**)malloc(solutions_in_pool * sizeof(struct INSTANCE*));
@@ -35,9 +30,8 @@ int GRASP(struct INSTANCE* ip) {
 		pool_quality[i] = -DBL_MAX;
 	}
 
-	int PR_STRATEGY = (int)(ip->algorithmOptions[9] + eps); // 1, 2, 3 or 4, as above (aO[9] = 1, so PR with one random solution in pool)
-	int PR_DIRECTION_PARAMETER = (int)(ip->algorithmOptions[11] + eps); // 0 forward, 1 backward, 2 forward & backward, 3 randomly chosen (including f&b) (aO[11] = 0, so forward)
-	// int performPathRelinking = (int) ip->algorithmOptions[7] + eps;
+	int PR_STRATEGY = (int)(ip->algorithmOptions[9] + eps);
+	int PR_DIRECTION_PARAMETER = (int)(ip->algorithmOptions[11] + eps);
 
 	if (ip->verbose > 5) {
 		printf("Started GRASP procedure...\n");
[…]

### 2022-08-28 code/Carrot/src/GRASP_VNS/grasp.h

@@ -10,15 +10,11 @@ UoS
 
 int GRASP(struct INSTANCE * ip);
 void RandomisedConstructive(struct INSTANCE * ip, int randomness, double delta, int rclStrategy);
-void GRASPLS(struct INSTANCE * ip, int LS_ITERS);
 void GenerateRCL(double delta, int* rclSeeds, int* rclSize, double* rankValues, int rvSize, double bestRank, double worstRank, int strategy);
 void RCLPick(int ** bestIndices, int * rclSeeds, int rclSize, int * cNurse, int * cJob);
-void BestIndexPick(int ** bestIndices, int GRASP_param, int * cNurse, int * cJob);
 int PickInteger(int max_int);
-void UpdateTopValues(int GRASP_param, int ** bestIndices, double * bestValues, double newValue, int bestNurse, int bestJob);
 void CleanSolutionFromStruct(struct INSTANCE * ip);
 int IdentifyEarlyInsert(struct INSTANCE * ip, struct INSTANCE * guiding, int guidingNurse, int guidingNursePos);
-void BestWorstPoolQualities(double * poolBest, double * poolWorst, struct INSTANCE ** pool, double * poolQuality, int poolSize);
 double SolutionDissimilarity(struct INSTANCE * input1, struct INSTANCE * input2);
 int FindArcDestination(int sourceNurse, int sourceJob, struct INSTANCE * ip);
 void PrintRecalculatePoolContents(struct INSTANCE ** pool, double * poolQuality, int solutionsInPool);
@@ -32,5 +28,3 @@ void NurseAndJobPosition(struct INSTANCE * ip, int job, int * nurse, int * posit
 void NurseAndJobPositionDS(struct INSTANCE * ip, int job, int * nurse1, int * position1, int * nurse2, int * position2);
 void StandardLocalSearch(struct INSTANCE * ip, int MAX_ITERATIONS, double MAX_TIME);
 double StandardLocalSearchTest(struct INSTANCE * ip, int MAX_ITERATIONS, double MAX_TIME, int TEST_ITERATIONS);
-double SolutionDistance(struct INSTANCE * input1, struct INSTANCE * input2);
-// void standard_local_search_two_opt_first(struct INSTANCE * ip, int MAX_ITERATIONS, double MAX_TIME);

### 2022-08-28 code/Carrot/src/GRASP_VNS/read_instance.c

@@ -8,11 +8,6 @@ UoS
 
 #include "constructive.h"
 void OverwriteInstance(struct INSTANCE* ip, struct INSTANCE* oi) {
-	/*
-	Copies the contents of the solution of "oi" into "ip"
-	without creating any new allocations.
-	Only copies the solution, no algorithm options or data
-	*/
 
 	for (int i = 0; i < oi->nNurses; ++i){
         for(int j = 0; j < oi->nJobs; ++j){
@@ -56,7 +51,7 @@ void OverwriteInstance(struct INSTANCE* ip, struct INSTANCE* oi) {
 		ip->violatedTWMK[i] = oi->violatedTWMK[i];
 	}
 
-} //END OF OverwriteInstance function.
+}
 
 struct INSTANCE CopyInstance(struct INSTANCE* original_instance) {
 	int LOCAL_VERBOSE = original_instance->verbose;
@@ -64,9 +59,9 @@ struct INSTANCE CopyInstance(struct INSTANCE* original_instance) {
 	int** solMatrix;
 	int nRows = original_instance->nNurses;
 	int nCols = original_instance->nJobs;
-	solMatrix = malloc(nRows * sizeof(int*)); // Rows
+	solMatrix = malloc(nRows * sizeof(int*));
 	for (int i = 0; i < nRows; i++){
-        solMatrix[i] = malloc(nCols*sizeof(int)); // Cols
+        solMatrix[i] = malloc(nCols*sizeof(int));
     }
 	for (int i = 0; i < nRows; ++i){
         for(int j = 0; j < nCols; ++j){
@@ -81,9 +76,9 @@ struct INSTANCE CopyInstance(struct INSTANCE* original_instance) {
 	double** timeMatrix;
 	nRows = original_instance->nNurses;
 	nCols = original_instance->nJobs;
-	timeMatrix = malloc(nRows * sizeof(double*)); // Rows
+	timeMatrix = malloc(nRows * sizeof(double*));
 	for (int i = 0; i < nRows; i++){
-        timeMatrix[i] = malloc(nCols*sizeof(double)); // Cols
+        timeMatrix[i] = malloc(nCols*sizeof(double));
     }
 	for (int i = 0; i < nRows; ++i){
         for(int j = 0; j < nCols; ++j){
@@ -97,9 +92,9 @@ struct INSTANCE CopyInstance(struct INSTANCE* original_instance) {
 	int** allNurseRoutes;
[…]
